### 2020-08-26 reports/bayesian_report.Rmd
restarting Bayesian analysis repo

---
title: 'Smoking and COVID Living Review (v7): a bayesian analysis'
author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:

  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: true
    theme: spacelab
    toc: false
    toc_collapse: true
    toc_depth: 3
    toc_float: yes
    css: !expr here::here('css', 'style.css')
    keep_md: false
editor_options: 
  chunk_output_type: inline
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      warning = F,
                      message = F)
options(width = 180)
unlink('reports/markdown_report_cache', recursive = TRUE)

```
# Data preparation {.tabset .tabset-fade .tabset-pills}

## Load packages
* **Load packages**: load packages and useful scripts for the analysis
* **Load scripts**: We have written functions to handle much of the Bayesian analysis
* **Load Bayesian datasets**: The Bayesian approach relies on large simulated datasets from the observed data, for this report we will only present the results graphically from data we have simulated elsewhere. The scripts `bayes_scripts.R`, `bayesian_data_generation.R` and `bayesian_analysis.R` should be able to be run locally if you want to generate your own simulated data. Alternatively you can use the pre-generated data which is stored in the `data_clean\bayesian_models` folder in the github repository [here](https://github.com/DidDrog11/smoking_covid/tree/master) and then you can skip the `bayesian_data_generation.R` step.
```{r install_packages, include=FALSE}
library(brms)
library(ggplot2)
library(tidybayes)
library(readr)
library(dplyr)
library(ggridges)
library(glue)
library(stringr)
library(forcats)
library(here)
library(meta)
library(snakecase)
library(flextable)
library(googlesheets4)
library(linelist)
library(DT)
library(tidyr)
library(epitools)
```

```{r load_scripts}
source(here("scripts", "author_dictionary.R"))
```
## Download data
We maintain a Google sheets document [here](sheets_id) which contains all eligible articles the research team have identified. This is currently viewable only and not open for editing. We have collected the general information for the search in the first sheet, the studies general information in the second, further information on study populations, outcomes of interest and quality of the studies are contained in the further sheets. These sheets are organised in a similar way to the manuscript and so are labelled table_1 to table_6.

``` {r google_data}
sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1HV2Zk_UD2_s1zzd47dCyb4UKRSM0dDg7zmUPA9CMdx4/edit?usp=sharing")
search_details <- read_sheet(sheets_id, range = 'article_screening')
data_study_general <- read_sheet(sheets_id, range =  'general_details')
table_1 <- read_sheet(sheets_id, sheet = 'pop_descriptives')
table_2 <-  read_sheet(sheets_id, sheet = 'testing')
table_3 <-  read_sheet(sheets_id, sheet = 'hospitalisation')
table_4 <-  read_sheet(sheets_id, sheet = 'severity')
table_5 <-  read_sheet(sheets_id, sheet = 'mortality')
table_6 <-  read_sheet(sheets_id, sheet = 'updated_quality_appraisal')
```
We have extracted data from `r nrow(data_study_general)` studies, however, some of these are duplicates and some have been removed from the analysis set. For example there have been multiple versions of the ISARIC report and several publications have been retracted. We aim to keep on top of this but there is some risk that in between versions data may be superseded.

## Raw Data
The first sheet is **Search details**  we use this sheet to keep track of the number of studies found in our search and keep track of the number of titles and abstracts screened and number of papers included in each review version. It presents the data in groups of search periods.

```{r head_search_details}
flextable(search_details[,1:11])
```
The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.
This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`. These are live links and should direct to the relevant article, however, some may have been deleted or require institutional access.
```{r tail_study_details, include=FALSE}
datatable(data_study_general[1:4] %>%
            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`),
                   doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5),
          escape = F)
```
The third sheet produces **Table 1** this contains the details of the population of the studies in our review. Primarily includes smoking status but also the proportion of females in the study and the age of participants.
```{r tail_table_1, include=FALSE}
datatable(table_1 %>%
                 filter(lead_author != is.na(lead_author)))
```
The fourth sheet produces **Table 2** this contains the data extracted from relevant studies on testing data and smoking status. Most studies will not have data on this outcome however we filter these subsequently. These are sparse datasets as many studies do not report on all levels of smoking exposure.
```{r tail_table_2, include=FALSE}
datatable(
  table_2 %>%
  filter(data_on_testing == TRUE)
)
```
The fifth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status. 
```{r tail_table_3, include=FALSE}
datatable(
  table_3 %>%
  filter(data_on_hospitalisation == TRUE)
)
```
The sixth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status. Disease severity is differentially reported across the studies in some there is reference to WHO severity score in others we have based the distinction on clinical characteristics or whether patients were admitted to Intensive or Critical care areas.
```{r tail_table_4, include=FALSE}
datatable(
  table_4 %>%
  filter(data_disease_severity == TRUE)
)
```
The seventh sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status. A caution needed when interpreting this data is that most studies have not followed patients up for sufficient amounts of time to be able to accurately report on disease associated mortality.
```{r tail_table_5, include=FALSE}
datatable(
  table_5 %>%
  filter(data_on_deaths == TRUE)
)
```
The final sheet is used to collate data on **Study quality** we changed our method on assessing study quality in version 3 of the review. We use the following criteria:

- **Good:** 
  1. <20% missing data on smoking status and used a reliable self-report measure that distinguished between current, former and never smoking status AND
  2. biochemical verification of smoking status and reported results from adjusted analyses OR 
  3. reported data from a representative/random sample
- **Fair**
  1. <20% missing data on smoking status and used a reliable self-report measure that distinguished between current, former and never smoking status
- **Poor**
  1. The remainder of the studies

## Clean data and save updated versions as RDS files
The data from the google sheets document is then cleaned using a script from the `linelist` package which has been developed by T. Jombart and Z. Kamvar within the [R Epidemics Consortium](https://www.repidemicsconsortium.org/)
The script `clean_data()` ensures consistency among column names, country names, dates and missing values in the included sheets. Data is also saved as `.rds` files in the `/data_clean` folder of the project repository. 
```{r clean_data}

search_details <- search_details %>%
  clean_data() %>%
  rename(date_screened = date_screening_in_format_2020_05_22) %>%
  write_rds(here::here('data_clean', 'search_details.rds')) ## The title of the date screened column is to communicate among the review team how we need date formatted

protect_columns_1 <- names(data_study_general) %in% 'doi' ## As the DoI's are links we want to preserve their formatting when running the cleaning script

data_study_general <- data_study_general %>%
  rename(date_published = `date_published_format(2020-05-12)`) %>%
  clean_data(protect = protect_columns_1) %>%
  mutate(study_id = 1:length(lead_author)) %>% ## We create a unique study ID to aid checking of the outputs later
  write_rds(here::here('data_clean', 'data_study_general.rds'))

prisma_previous <- data_study_general

review_details <- data_study_general %>%
  select(lead_author, date_published, country, review_version, study_id) %>% ## The version of the review in which included studies were incorporated is noted
  write_rds(here::here('data_clean', 'review_details.rds'))

table_1 <- table_1 %>%
  clean_data() %>%
  mutate(., lower_range = sub('\\_.*', '', .$range)) %>%
  mutate(., upper_range = sub('.*_', '',.$range,)) %>% ## We break apart the age range into an upper and lower level
  select(lead_author, sample_size, female_sex_percent, data_source, median_age, mean_age, iqr_lower, iqr_upper, standard_deviation, lower_range,
         upper_range, current_smoker, former_smoker, current_former_smoker, never_smoker, never_smoker_unknown, current_vaper, former_vaper,
         current_former_vaper, not_stated, missing, total) %>%
  filter(!is.na(lead_author)) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_1.rds'))

##For the following tables we're only interested in obtaining data for trials which report on the specific outcome of interest

table_2 <-  table_2 %>%
  clean_data() %>%
  filter(data_on_testing == TRUE) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_2.rds'))

table_3 <-  table_3 %>%
  clean_data() %>%
  filter(data_on_hospitalisation == TRUE) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_3.rds'))

table_4 <-  table_4 %>%
  clean_data() %>%
  filter(data_disease_severity == TRUE) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_4.rds'))

table_5 <- table_5 %>%
  clean_data() %>%
  filter(data_on_deaths == TRUE) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_5.rds'))

protect_columns_2 <- !names(table_6) %in% 'lead_author'

table_6 <-  table_6 %>%
  clean_data(protect = protect_columns_2) %>%
  left_join(., table_1 %>%
              select(lead_author,
                     not_stated,
                     missing,
                     total),
                          by = 'lead_author') %>%
  mutate(missingness = rowSums(.[8:9], na.rm = T),
         missingness_percentage = (missingness/total)*100) %>% ## Later we want to assess the amount of missingness for each study as it would impact the inclusion of a study into the relevant meta-analyses
  select(-c(not_stated, missing, total, missingness)) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_6.rds'))

a <- data_study_general %>%
  select(lead_author, date_published, source, study_id) %>%
  rename('Lead Author' = lead_author,
         'Date Published' = date_published,
         'Publication Source' = source,
         'Study ID' = study_id)

a$`Lead Author` <- to_upper_camel_case(a$`Lead Author`, sep_out = ", ")
a$`Publication Source` <- to_title_case(a$`Publication Source`)
a$`Publication Source` <- if_else(str_length(a$`Publication Source`) < 5,
                                  toupper(a$`Publication Source`),
                                  to_title_case(a$`Publication Source`))

write_rds(data_study_general, here("data_clean", "data_study_general.rds"))
```
The package `flextable` can save tables in a format that is easily manageable within word which we currently use to write the manuscript (until I get better with RMarkdown). At various points I save these to the `output_data` folder.
``` {r save_as_table_for_word, include=FALSE}
datatable(a)

a <- flextable(a)%>%
  set_caption(a, caption = 'All studies included in data extraction') %>%
  set_table_properties(width = 1, layout = 'autofit')

save_as_docx(a, path = here('output_data', 'all_studies.docx'))
```

# Preparing the descriptive analysis {.tabset .tabset-fade .tabset-pills}

## Choosing which data we want to include in the current review?
``` {r versioning}
date_of_update <- Sys.Date()

prev_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6')
#This can categorise which studies we want to look at
current_version <- c('v7')
#This will categorise which studies we are including in the current report

analysed_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6', 'v7')
#This will incorporate all studies into the current version of the report
```
We define the current version of the report here. At the moment this needs to be changed directly when compiling the code. This report has been generated for version `r paste(current_version)`

``` {r excluding_studies}
exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old', 'isaric_4', 'mehra', 'hopkinson', 'gaibazzi', 'miyara_updated', 'russell_old', "petrilli_old", "magagnoli_old", "niedzwiedz_old", "bello_chavolla_old", "kimmig_old", "russell_old", "zuo_zuo_old", "patel_old", "shi_zuo_old")
```
We don't want to remove studies from the dataset despite them no longer being included in the analysis. The excluded studies are defined in this chunk of code. We have excluded this version the following studies:
-`r knitr::combine_words(exclude_from_analysis)`
Generally we label these as _old in the Google sheet but there are exceptions to this so we manually define them at the moment.

## Filter the included studies
Based on the version and excluded studies we produce the tables for analysis. This is a searchable table.
``` {r descriptives}
data_study_general <- data_study_general %>%
  filter(review_version %in% analysed_versions) %>%
  filter(!(lead_author %in% exclude_from_analysis)) %>%
  write_rds(., here::here('data_clean', 'data_study_general.rds'))

review_details <- data_study_general %>%
  select(lead_author, date_published, country, review_version, study_id) %>% 
  filter(!(lead_author %in% exclude_from_analysis))

table_1 <- table_1 %>%
  filter(review_version %in% analysed_versions) %>%
  select(-review_version) %>%
  filter(!(lead_author %in% exclude_from_analysis))

table_2 <- table_2 %>%
  filter(review_version %in% analysed_versions) %>%
  select(-review_version) %>%
  filter(!(lead_author %in% exclude_from_analysis))

table_3 <- table_3 %>%
  filter(review_version %in% analysed_versions) %>%
  select(-review_version) %>%
  filter(!(lead_author %in% exclude_from_analysis))

table_4 <- table_4 %>%
  filter(review_version %in% analysed_versions) %>%
  select(-review_version) %>%
  filter(!(lead_author %in% exclude_from_analysis))

table_5 <- table_5 %>%
  filter(review_version %in% analysed_versions) %>%
  select(-review_version) %>%
  filter(!(lead_author %in% exclude_from_analysis))

table_6 <- table_6 %>%
  filter(review_version %in% analysed_versions) %>%
  select(-review_version) %>%
  filter(!lead_author %in% exclude_from_analysis)

a <- data_study_general %>%
  select(lead_author, date_published, source, study_id) %>%
  rename('Lead Author' = lead_author,
         'Date Published' = date_published,
         'Publication Source' = source,
         'Study ID' = study_id)

a$`Lead Author` <- to_upper_camel_case(a$`Lead Author`, sep_out = ", ")
a$`Publication Source` <- to_title_case(a$`Publication Source`)
a$`Publication Source` <- if_else(str_length(a$`Publication Source`) < 5,
                                  toupper(a$`Publication Source`),
                                  to_title_case(a$`Publication Source`))

datatable(a)

a <- flextable(a)%>%
  set_table_properties(width = 1, layout = 'autofit') %>%
  set_caption(caption = 'Studies included in the current analysis')

save_as_docx(a, path = here('output_data', 'included_studies.docx'))
```

# Study descriptive analysis {.tabset .tabset-fade .tabset-pills}

## The PRISMA diagram will be updated
The PRISMA reporting guidelines for systematic reviews the guidance is available [here](http://www.prisma-statement.org/). This takes the format of the number of articles obtained from the search from [OVID](https://ovidsp.dc1.ovid.com/sp-4.07.0b/ovidweb.cgi) (login may be needed), [medRxiv](https://www.medrxiv.org/search/text_abstract_title%3Asmoking%2BAND%2BCOVID%20text_abstract_title_flags%3Amatch-all%20jcode%3Amedrxiv%20numresults%3A75%20sort%3Apublication-date%20direction%3Adescending%20format_result%3Astandard), and articles identified by the research team. These are then screened at the title and abstract level. Full texts are then read to extract relevant data, if articles are excluded at this level we describe the reason for this. All of the selected articles will contribute to the narrative synthesis but only those that are judged 'good' or 'fair' quality will be included in the Bayesian meta-analysis.
``` {r PRISMA}
current_total <- search_details %>%
  select(total_included) %>%
  sum()

previous_total <- search_details %>%
  filter(review_version %in% prev_versions) %>%
  select(total_included) %>%
  sum()

prisma <- search_details %>%
  group_by(review_version) %>%
  filter(review_version == current_version) %>%
  mutate(title_abstract = sum(ovid_number_results, medrxiv_number_results, other_source_included),
         full_texts_assessed = sum(ovid_screened, medrxiv_screened, other_source_included),
         title_abstract_included = sum(ovid_included, medrxiv_included, other_source_included),
         title_abstract_excluded = title_abstract - full_texts_assessed,
         full_text_excluded = full_texts_assessed-total_included,
         previous_total = previous_total,
         previous_minus_superseded = previous_total-superseded,
         current_total = current_total) %>%
  select("Review version" = review_version,
         "Date Screened" = date_screened,
         "OVID results" = "ovid_number_results",
         "medRxiv results" = "medrxiv_number_results",
         "Other sources" = "other_source_included",
         "Title/abstracts screened" = "title_abstract",
         "Title/abstracts excluded" = "title_abstract_excluded",
         "Full texts assessed" = "full_texts_assessed",
         "Full texts excluded" = "full_text_excluded",
         "Records carried forward" = "previous_minus_superseded",
         "Supersed from previous version" = superseded,
         "Studies in narrative synthesis" = "current_total")
flextable(prisma)
```
The number of studies contributing to meta-analysis is calculated later.

## We summarise where the studies were conducted

As the pandemic continues studies are being reported from an increasing numbers of countries.

``` {r countries, cache = TRUE}
library('ggmap')
#Countries
country <- table_1$country %>%
  to_upper_camel_case() %>%
  recode(., 'Usa' = 'USA', 'Uk' = 'UK', 'SaudiArabia' = 'Saudi Arabia', "SouthKorea" = "South Korea") %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rename('Country' = 1,
         'Number' = 2) %>%
  mutate(Country = as.character(Country))

ggplot(country, aes(x = reorder(Country, desc(-Number)), y = Number))+
  geom_col()+
  coord_flip()+
  theme_bw()+
  labs(title = 'Countries where studies were performed',
       y = 'Number of studies',
       x = 'Country')

a <- flextable(country) %>%
  set_table_properties(width = 0.5, layout = 'autofit')
a

country <- subset(country, Country != "Multiple")

country <- mutate_geocode(country, Country)

country_bounds <- country %>%
  select(lon, lat) %>%
  tidyr::drop_na() %>%
  sp::SpatialPointsDataFrame(coords = .[,1:2], data = .) %>%
  sp::bbox() * 1.1

map_world <- map_data(map = "world") %>%
  filter(region != "Antarctica")

save_as_docx(a, path = here('output_data', 'Countries_performing_studies.docx'))  
```
Currently studies are reported from `r paste(length(unique(country)))`

## A map is produced for where the studies are performed
``` {r interactive global map, cache = TRUE}
library(rgdal)
library(leaflet)
library(coronavirus)
update_dataset()

my_spdf <- readOGR( 
  dsn= paste0(here::here("data_clean", "world_shape_file")), 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

from_countries <- c("USA", "UK", "Iran", "South Korea")
to_countries <- c("United States", "United Kingdom", "Iran (Islamic Republic of)", "Korea, Republic of")
country_map <- country %>%
  mutate(Country = plyr::mapvalues(Country, from = from_countries,
                to = to_countries))

from_countries <- c("US", "Iran", "Korea, South", "Czech Republic")
to_countries <- c("United States", "Iran (Islamic Republic of)", "Korea, Republic of", "Czechia")

data("coronavirus")
summary_df <- coronavirus %>% 
  filter(type == "confirmed") %>%
  group_by(country) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases) %>%
  rename("NAME" = country) %>%
  mutate(NAME = plyr::mapvalues(NAME, from = from_countries,
                to = to_countries))

my_spdf@data <- left_join(my_spdf@data, country_map %>%
            rename(NAME = Country),
          by = "NAME") %>%
  left_join(., summary_df %>%
              mutate(total_cases = total_cases/1000),
            by = "NAME") %>%
  mutate(NAME = as.factor(NAME))

my_spdf@data$Number[ which(my_spdf@data$Number == 0)] = NA
my_spdf@data$total_cases[ which(is.na(my_spdf@data$total_cases))] = 0

my_bins <- c(0, 1, 5, 10, 20, 30, 40, 70)
mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)

mytext <- paste(
  "Country: ", my_spdf@data$NAME,"<br/>", 
  "Number of studies included: ", my_spdf@data$Number, "<br/>", 
  "COVID-19 reported cases (thousands): ", round(my_spdf@data$total_cases, 2), 
  sep="") %>%
  lapply(htmltools::HTML)

html_map <- leaflet(my_spdf) %>% 
  addTiles()  %>% 
  setView(lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(Number), 
    stroke=TRUE, 
    fillOpacity = 0.9, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal=mypalette, values=~Number, opacity=0.9, title = "Number of studies included", position = "bottomright" )

html_map
```
This map displays the number of studies produced in each country. The number of confirmed COVID-19 cases are shown when resting over each country for reference, data on confirmed cases have been obtained from [Rami Krispin's `coronavirus` package](https://ramikrispin.github.io/coronavirus/). The number of cases is updated when the report is generated so will not remain in date, it is current as of `r paste(Sys.time())`. Countries displaying 0 cases are most likely due to naming mismatches.

## We sumarise the settings in which they were conducted

Most studies are performed in hospital settings. Some studies are conducted in both such as those in emergency departments where patients are discharged home and followed up. Some registry studies utilising primary care networks are solely conducted in the community.
``` {r setting}
#Setting
setting <- to_upper_camel_case(data_study_general$study_setting, sep_out = ' ') %>%
  table(.) %>%
  sort('Number of studies', decreasing = T) %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rename('Setting' = 1,
         'Number' = 2)

a <- flextable(setting) %>%
  set_table_properties(width = 0.5, layout = 'autofit')
a

save_as_docx(a, path = here('output_data', 'Countries_performing_studies.docx'))  
```
## We summarise the design of the included studies

The majority were retrospective cohorts. There were fewer case-control and cross-sectional studies. There were some prospective cohorts established and reported.
``` {r design}
design <- to_upper_camel_case(data_study_general$study_design, sep_out = ' ') %>%
  recode("Rct" = "RCT") %>%
  table(.) %>%
  sort('Number of studies', decreasing = T) %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rename('Design' = 1,
         'Number' = 2)

a <- flextable(design) %>%
  set_table_properties(width = 0.5, layout = 'autofit')
a

save_as_docx(a, path = here('output_data', 'Study_designs.docx'))

##For supplementary tables
study_design <- data_study_general %>%
  select(lead_author, study_design) %>%
  mutate(study_design = to_upper_camel_case(study_design, sep_out = " "),
         study_design = ifelse(study_design == "Rct", "Randomised Controlled Trial", study_design))
```
## We update the number of participants we have included

* **Summary statistics:** this table will give us the mean, median and IQR of the samples in the review 
``` {r numbers}
#Numbers
summary(table_1$sample_size)
```
* **Total number of participants**: `r paste(sum(table_1$sample_size))`

## We want to know if they use a CRF or electronic health record
The majority of data on smoking outcomes is extracted from electronic health records that are routinely collected in the provision of healthcare. Some of the studies utilise specifically designed Case Report Forms (CRFs). The source of information on smoking status is not always reported.
``` {r data_source}
#Source of data
table_1$data_source %>%
  recode('electronic_health_records' = 'Electronic health record',
         'case_report_form' = 'Case report form',
         'not_stated' = 'Not stated') %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rename('Data source' = 1,
         'Number' = 2) %>%
  flextable() %>%
  set_table_properties(width = 0.5, layout = 'autofit')
```

# Smoking descriptive data {.tabset .tabset-fade .tabset-pills}
``` {r smokers}
#Studies reporting smokers
current_smok <- table_1 %>%
  filter(., current_smoker != 'NA')
```
* **Current smokers:** `r paste(nrow(current_smok))` studies report on the numbers of current smokers
``` {r former smokers}
#Studies reporting former smokers
former_smok <- table_1 %>%
  filter(., former_smoker != 'NA')
```
* **Former smokers:** `r paste(nrow(former_smok))` studies report on the numbers of former smokers
```{r never smokers}
#Studies reporting never smokers
never_smok <- table_1 %>%
  filter(., never_smoker != 'NA')
```
* **Never smokers:** `r paste(nrow(never_smok))` studies report on the numbers of never smokers
``` {r current former smokers}
#Studies reporting current/former smokers
current_former_smok <- table_1 %>%
  filter(., current_former_smoker != 'NA')
```
* **Current/former smokers:** `r paste(nrow(current_former_smok))` studies report on the numbers of current/former smokers without distinguishing between them, often termed as yes to a smoking history question.
``` {r missing/never/unknown/not stated}
#Studies reporting missing data
missing_smok <- table_1 %>%
  filter(., missing != 'NA') %>%
  mutate(percent = missing/total*100)
minimum <- min(missing_smok$percent)
maximum <- max(missing_smok$percent)
#Studies reporting never/unknown
never_smok_unknown <- table_1 %>%
  filter(., never_smoker_unknown != 'NA')

#Studies with not stated
smok_not_stated <- table_1 %>%
  filter(., not_stated != 'NA')
```
* **Missing, never/unknown and not stated:** studies report on the numbers of missing (`r paste(nrow(missing_smok))`, `r paste(round((nrow(missing_smok)/nrow(data_study_general)*100),2), "%", sep = "")`), never/unknown (`r paste(nrow(never_smok_unknown))`) or not stated (`r paste(nrow(smok_not_stated))`) smoking history. This is often framed as absence of a smoking history.
  + **Minimum missing:** `r paste(round(minimum,2), "%", sep = "")`  
  + **Maximum missing:** `r paste(round(maximum,2), "%", sep = "")`

## Studies reporting current, former and never smoking status
``` {r complete}
#Studies reporting current, former and never smoking status
full_smoking_status <- table_1 %>%
  filter(lead_author %in% current_smok$lead_author) %>%
  filter(lead_author %in% former_smok$lead_author) %>%
  filter(lead_author %in% never_smok$lead_author)

b <- full_smoking_status %>%
  select(lead_author, sample_size, current_smoker, former_smoker, never_smoker, study_id) %>%
  rename('Lead author' = lead_author,
         'Sample size' = sample_size,
         'Current smokers' = current_smoker,
         'Former smokers' = former_smoker,
         'Never smokers' = never_smoker,
         'Study ID' = study_id)

b$`Lead author` <- to_upper_camel_case(b$`Lead author`, sep_out = ", ")
b$`Lead author` <- plyr::mapvalues(b$`Lead author`,
                               from = cleaned_names,
                               to = correct_names)  
datatable(b,
          caption = "Studies providing data on current, former and never smokers") %>%
  formatRound(columns = c(2:5), digits = 0, interval = 3, mark = ",")
#These are the studies with data on current, former and never smokers

b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_num(col_keys = c("Sample size", "Current smokers", "Former smokers", "Never smokers"),
                digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

save_as_docx(b, path = here('output_data', 'Supplementary_1a.docx'))
```
* **Studies with complete smoking status:** `r paste(nrow(full_smoking_status))`, `r paste(round((nrow(full_smoking_status)/nrow(data_study_general)*100),2), "%", sep = "")`

## Studies providing partially complete data on smoking status

``` {r incomplete}
#Studies reporting current or current/former and never smoking
semi_full_smoking_status <- table_1 %>%
  filter(lead_author %in% current_former_smok$lead_author) %>%
  filter(lead_author %in% never_smok$lead_author) %>%
  filter(!lead_author %in% full_smoking_status$lead_author)

b <- semi_full_smoking_status %>%
  select(lead_author, sample_size, current_former_smoker, never_smoker, study_id) %>%
  rename('Lead author' = lead_author,
         'Sample size' = sample_size,
         'Current/former smokers' = current_former_smoker,
         'Never smokers' = never_smoker,
         'Study ID' = study_id)

b$`Lead author` <- to_upper_camel_case(b$`Lead author`, sep_out = ", ")
b$`Lead author` <- plyr::mapvalues(b$`Lead author`,
                               from = cleaned_names,
                               to = correct_names)

datatable(b,
          caption = "Studies providing partially complete data on smoking status") %>%
  formatRound(columns = c(2:4), digits = 0, interval = 3, mark = ",")

b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_num(col_keys = c("Sample size", "Current/former smokers", "Never smokers"),
                digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

save_as_docx(b, path = here('output_data', 'Supplementary_1b.docx'))
```

* **Studies reporting partially incomplete smoking data:** `r paste(nrow(semi_full_smoking_status))`  
These studies typically report smoking history as the measure of exposure which is why they have grouped values for current/former smokers

## Studies providing incomplete data on smoking status
``` {r}
#Remaining studies
incomplete_smoking_status <- table_1 %>%
  filter(!lead_author %in% full_smoking_status$lead_author) %>%
  filter(!lead_author %in% semi_full_smoking_status$lead_author)

b <- incomplete_smoking_status %>%
  select(lead_author, sample_size, current_smoker, current_former_smoker, former_smoker, never_smoker, never_smoker_unknown, not_stated, missing, study_id) %>%
  rename('Lead author' = lead_author,
         'Sample size' = sample_size,
         'Current smokers' = current_smoker,
         'Current/former smokers' = current_former_smoker,
         'Former smoker' = former_smoker,
         'Never smokers' = never_smoker,
         'Never smoker/unknown' = never_smoker_unknown,
         'Not stated' = not_stated,
         'Missing' = missing,
         'Study ID' = study_id)

b$`Lead author` <- to_upper_camel_case(b$`Lead author`, sep_out = ", ")
b$`Lead author` <- plyr::mapvalues(b$`Lead author`,
                               from = cleaned_names,
                               to = correct_names)

datatable(b,
          caption = "Studies providing incomplete data on smoking status") %>%
  formatRound(columns = c(2:8), digits = 0, interval = 3, mark = ",")

numeric_names <- colnames(b[2:9])
b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_num(col_keys = numeric_names,
                digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

save_as_docx(b, path = here('output_data', 'Supplementary_1c.docx'))
```

* **Studies missing key smoking data:** `r paste(nrow(incomplete_smoking_status))`

# Smoking prevalence by country {.tabset .tabset-fade .tabset-pills}
``` {r smoking_country}

#Smoking prevalence by country
country_prevalence_list <- table_1 %>%
    group_by(country, add = T) %>%
  mutate(., current_smok_percentage = current_smoker/total*100) %>%
  mutate(., former_smok_percentage = former_smoker/total*100) %>%
  mutate(., missing_percentage = missing/total*100) %>%
  mutate(., not_stated_percentage = not_stated/total*100) %>%
  mutate(never_smoker_percentage = never_smoker/total*100) %>%
  group_split()

```
We have split the data into a list where each country can be analysed separately if required

Confidence intervals are obtained through bootstrap estimates of a thousand repetitions of the included studies. Following the production of the confidence intervals from bootstrapping in `olgas_script.R` the prevalence graphs are produced in the `smoking_plots.R` script. This uses data extracted on current and former smoking prevalence data from countries which have included studies in this review.

## Graph for smoking prevalence
**Current smoking prevalence**
The studies are grouped by the country in which they were conducted. The solid line represents national current smoking prevalence obtained from national smoking survey estimates. Studies are further grouped by the setting in which they were conducted, Hospital (Triangle), Community (Square) and Community and Hospital (Circle). The size of the shape corresponds to the relative size of the study sample. The weighted mean current smoking prevalence is displayed for each setting.
![Country smoking prevalence](figure/figure_2a.png)

## Graph for former smoking prevalence
**Former smoker prevalence**
This graph is for former smoking prevalence. The data is more sparse for this exposure and therefore several countries are missing for the analysis. Studies are further grouped by the setting in which they were conducted, Hospital (Triangle), Community (Square) and Community and Hospital (Circle). The size of the shape corresponds to the relative size of the study sample. The weighted mean current smoking prevalence is displayed for each setting.
![Country former smoking prevalence](figure/figure_2b.png)

# Final results tables and meta-analyses {.tabset .tabset-fade .tabset-pills}

## Table 1: Included studies
``` {r table_1}
#Updating table 1
table_1_word <- table_1 %>%
  mutate(., current_percentage = current_smoker/total*100,
         former_percentage = former_smoker/total*100,
         current_former_percentage = current_former_smoker/total*100,
         never_smoker_percentage = never_smoker/total*100,
         never_smoker_unknown_percentage = never_smoker_unknown/total*100,
         not_stated_percentage = not_stated/total*100,
         missing_percentage = missing/total*100) %>%
  select(lead_author, date_published, country, sample_size, median_age, iqr_lower, iqr_upper, mean_age, lower_range,
         upper_range, standard_deviation, female_sex_percent, current_percentage, former_percentage, current_former_percentage,
         never_smoker_percentage, never_smoker_unknown_percentage, not_stated_percentage, missing_percentage, study_id) %>%
  replace_na(., list(not_stated_percentage = 0, missing_percentage = 0)) %>%
  mutate(., missing_percentage = not_stated_percentage + missing_percentage,
         missing_percentage = formatC(missing_percentage, digits = 2, format = "f"))

#Smoking completeness
smoking_status <- as_tibble(full_smoking_status$lead_author) %>%
  rename(lead_author = 1) %>%
  mutate(complete_status = c("Yes")) %>%
  bind_rows(., semi_full_smoking_status %>%
              select(lead_author)) %>%
  bind_rows(., incomplete_smoking_status %>%
              select(lead_author)) %>%
  distinct() %>%
  mutate(complete_status = replace_na(complete_status, "No"))

#Data missingness
missingness <- table_6 %>%
  select(lead_author) %>%
  left_join(., table_1_word %>%
              select(lead_author, missing_percentage)) %>%
  mutate(missing_percentage = as.numeric(missing_percentage),
         low_missingness = ifelse(missing_percentage >= 20, "No", "Yes"))

#Defining poor quality
poor_quality <- missingness %>%
  select(-missing_percentage) %>%
  left_join(., smoking_status, by = "lead_author") %>%
  mutate(study_quality = ifelse(low_missingness == "No" | complete_status == "No", "Poor", "Not_poor"))

good_quality <- table_6 %>%
  left_join(., poor_quality, by = "lead_author") %>%
  select(lead_author, study_quality, biochemical_verification, random_representative_sample) %>%
  mutate(study_quality_final = ifelse(study_quality == "Not_poor" & (biochemical_verification == "Yes" | random_representative_sample == "Yes"), "good",
                                      ifelse(study_quality == "Poor", "poor", "fair")))

quality_rating <- good_quality %>%
  select(lead_author, study_quality_final) %>%
  rename("overall_quality" = study_quality_final)

table_1_word <- left_join(table_1_word, good_quality %>%
                            select(lead_author, study_quality_final))

a <- data_study_general %>%
  select(study_id, study_setting)

table_1_word <- left_join(table_1_word, a, by = 'study_id') %>%
  select(1:4, 22, 5:19, 21, 20)
table_1_word$date_published <- as.Date.character(table_1_word$date_published)
write_rds(table_1_word, here::here('data_clean', 'table_1_word.rds'))

table_1_word <- table_1_word %>%
  mutate(median_mean = ifelse(is.na(median_age), mean_age, median_age))

a <- table_1_word %>%
  mutate(median_mean = ifelse(is.na(median_age), mean_age, median_age),
         mean_used = ifelse(is.na(mean_age), '','^'),
         iqr = ifelse(is.na(iqr_lower), NA, paste(iqr_lower, iqr_upper, sep = '-')),
         range_combined = paste(lower_range, upper_range, sep = '-'),
         range_combined = na_if(range_combined, 'NA-NA'),
         standard_deviation = as.numeric(standard_deviation),
         st_dev = paste((as.integer(median_mean-standard_deviation)), as.integer((median_mean+standard_deviation)), sep = '-'),
         st_dev = na_if(st_dev, 'NA-NA'))

a$iqr <- coalesce(a$iqr, a$range_combined, a$st_dev)

a <- a %>%
  select(study_id, lead_author, date_published, country, sample_size, study_setting, median_mean, mean_used, iqr,
         female_sex_percent, current_percentage, former_percentage, current_former_percentage, never_smoker_percentage,
         never_smoker_unknown_percentage, missing_percentage, study_quality_final) %>%
  mutate(median_mean = paste(median_mean, mean_used, sep = ''),
         median_mean = na_if(median_mean, NA)) %>%
  select(-mean_used) %>%
  mutate(median_mean = ifelse(median_mean == 'NA', 'NA', paste(paste(median_mean, iqr, sep = ' ('),')', sep = ''))) %>%
  select(-iqr) %>%
  rename('Study ID' = study_id,
         'Lead author' = lead_author,
         'Date published' = date_published,
         'Country' = country,
         'Sample size' = sample_size,
         'Study setting' = study_setting,
         'Median (IQR)' = median_mean,
         'Female %' = female_sex_percent,
         'Current smoker %' = current_percentage,
         'Former smokers %' = former_percentage,
         'Current/former smokers %' = current_former_percentage,
         'Never smokers %' = never_smoker_percentage,
         'Never/unknown smokers %' = never_smoker_unknown_percentage,
         'Missing %' = missing_percentage,
         "Study quality" = study_quality_final)

a$`Lead author` <- to_upper_camel_case(a$`Lead author`, sep_out = ", ")
author_list <- plyr::mapvalues(a$`Lead author`,
                               from = cleaned_names,
                               to = correct_names)

a$`Lead author` <- author_list

a$`Study setting` <-to_title_case(a$`Study setting`, sep_out = " ")
a$`Country` <-to_title_case(a$`Country`, sep_out = " ")
a$`Country` <- a$Country %>%
  recode('Usa' = 'USA',
         'Uk' = 'UK')

numeric_columns <- c('Median (IQR)', 'Female %', 'Current smoker %','Former smokers %', 'Current/former smokers %',
                     'Never smokers %', 'Never/unknown smokers %', 'Missing %')
```
The total number of included studies is `r paste(nrow(table_1_word))`. This table is searchable
```{r}
datatable(a) %>%
  formatRound(
    columns = c(8:14),
    digits = 2,
    interval = 3,
    mark = ",",
    dec.mark = getOption("OutDec")
)

a <- flextable(a) %>%
    set_caption(caption = 'Characteristics of included studies') %>%
  colformat_num(col_keys = numeric_columns, digits = 1, na_str = '-', big.mark = ',') %>%
  colformat_num(col_keys = 'Sample size', digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

save_as_docx(a, path = here('output_data', 'Table_1.docx'))
```

## Table 2: Smoking and COVID-19 diagnostic testing
``` {r table_2}

#Table 2
table_2_word <-  table_2 %>%
  mutate(., sample = contributing_sample) %>%
  mutate(., negative_test_percentage =formatC(negative_test/sample*100, digits = 2, format = "f"),
         negative_current_percentage = formatC(negative_current_smoker/negative_test*100, digits = 2, format = "f"),
         negative_former_smoker_percentage = formatC(negative_former_smoker/negative_test*100, digits = 2, format = "f"),
         negative_current_former_smoker_percentage = formatC(negative_current_former_smoker/negative_test*100, digits = 2, format = "f"),
         negative_never_smoker_percentage = formatC(negative_never_smoker/negative_test*100, digits = 2, format = "f"),
         negative_not_stated_percentage = formatC(negative_not_stated/negative_test*100, digits = 2, format = "f"),
         positive_test_percentage = formatC(positive_test/sample*100, digits = 2, format = "f"),
         positive_current_smoker_percentage = formatC(positive_current_smoker/positive_test*100, digits = 2, format = "f"),
         positive_former_smoker_percentage = formatC(positive_former_smoker/positive_test*100, digits = 2, format = "f"),
         positive_current_former_smoker_percentage = formatC(positive_current_former_smoker/positive_test*100, digits = 2, format = "f"),
         positive_never_smoker_percentage = formatC(positive_never_smoker/positive_test*100, digits = 2, format = "f"),
         positive_not_stated_percentage = formatC(positive_not_stated/positive_test*100, digits = 2, format = "f")) %>%
  select(-data_on_testing, -missing, -date_published, -sample) %>%
  write_rds(here::here('data_clean', 'table_2_word.rds'))

quality_table_2 <- table_2_word %>%
  left_join(., quality_rating, by = 'lead_author')

a <- table_2_word %>%
  filter(contributing_sample >= 1) %>%
  mutate(Author = lead_author,
         Population_tested = contributing_sample,
         SARS_CoV_2_negative = paste(
           paste(negative_test, 
                 (negative_test_percentage), sep = " ("), "%)", sep = ""),
         N_current_smoker = paste(paste(negative_current_smoker, (negative_current_percentage), sep = " (")
                                  , "%)", sep = ""),
         N_current_smoker = na_if(N_current_smoker, "NA ( NA%)"),
         N_former_smoker = paste(paste(negative_former_smoker, (negative_former_smoker_percentage), sep = " (")
                                 , "%)", sep = ""),
         N_former_smoker = na_if(N_former_smoker, "NA ( NA%)"),
         N_current_former_smoker = paste(paste(negative_current_former_smoker,
                                               (negative_current_former_smoker_percentage), sep = " (")
                                         , "%)", sep = ""),
         N_current_former_smoker = na_if(N_current_former_smoker, "NA ( NA%)"),
         N_never_smoker = paste(paste(negative_never_smoker,
                                      (negative_never_smoker_percentage), sep = " (")
                                , "%)", sep = ""),
         N_never_smoker = na_if(N_never_smoker, "NA ( NA%)"),
         N_not_stated = paste(paste(negative_not_stated,
                                    (negative_not_stated_percentage), sep = " (")
                              , "%)", sep = ""),
         N_not_stated = na_if(N_not_stated, "NA ( NA%)")) %>%
  mutate(SARS_CoV_2_positive = paste(
    paste(positive_test, 
          (positive_test_percentage), sep = " ("), "%)", sep = ""),
    P_current_smoker = paste(paste(positive_current_smoker, (positive_current_smoker_percentage), sep = " (")
                             , "%)", sep = ""),
    P_current_smoker = na_if(P_current_smoker, "NA ( NA%)"),
    P_former_smoker = paste(paste(positive_former_smoker, (positive_former_smoker_percentage), sep = " (")
                            , "%)", sep = ""),
    P_former_smoker = na_if(P_former_smoker, "NA ( NA%)"),
    P_current_former_smoker = paste(paste(positive_current_former_smoker,
                                          (positive_current_former_smoker_percentage), sep = " (")
                                    , "%)", sep = ""),
    P_current_former_smoker = na_if(P_current_former_smoker, "NA ( NA%)"),
    P_never_smoker = paste(paste(positive_never_smoker,
                                 (positive_never_smoker_percentage), sep = " (")
                           , "%)", sep = ""),
    P_never_smoker = na_if(P_never_smoker, "NA ( NA%)"),
    P_not_stated = paste(paste(positive_not_stated,
                               (positive_not_stated_percentage), sep = " (")
                         , "%)", sep = ""),
    P_not_stated = na_if(P_not_stated, "NA ( NA%)")) %>%
  select(Author, Population_tested, SARS_CoV_2_negative, N_current_smoker, N_former_smoker, N_current_former_smoker,
         N_never_smoker, N_not_stated, SARS_CoV_2_positive, P_current_smoker, P_former_smoker, P_current_former_smoker,
         P_never_smoker, P_not_stated)

a$Author <- to_upper_camel_case(a$Author, sep_out = ", ")
a$Author  <- a$Author  %>%
  recode("Bello, Chavolla" = "Bello-Chavolla",
         "De, Lusignan" = "de Lusignan",
         "Del, Valle" = "del Valle")

numeric_columns <- c('Population_tested', 'SARS_CoV_2_negative', 'N_current_smoker', 'N_former_smoker',
                     'N_current_former_smoker', 'N_never_smoker', 'N_not_stated', 'SARS_CoV_2_positive', 'P_current_smoker',
                     'P_former_smoker', 'P_current_former_smoker', 'P_never_smoker', 'P_not_stated')
a <- flextable(a) %>%
    set_caption(caption = 'SARS-CoV-2 infection by smoking status') %>%
  colformat_num(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',')
```
The number of studies reporting on infection by smoking status is: `r paste(nrow(table_2_word))`
``` {r}
a <- set_header_labels(a,
                       Population_tested = 'Total population tested',
                       SARS_CoV_2_negative = "N (%)",
                       N_current_smoker = "Current smoker (%)",
                       N_former_smoker = "Former smoker (%)",
                       N_current_former_smoker = "Current/former smoker (%)",
                       N_never_smoker = "Never smoker (%)",
                       N_not_stated = "Not stated (%)",
                       SARS_CoV_2_positive = "N (%)",
                       P_current_smoker = "Current smoker (%)",
                       P_former_smoker = "Former smoker (%)",
                       P_current_former_smoker = "Current/former smoker (%)",
                       P_never_smoker = "Never smoker (%)",
                       P_not_stated = "Not stated (%)") %>%
  add_header_row(top = TRUE, values = c("","SARS-CoV-2 negative", "SARS-CoV-2 positive" ), colwidths = c(2, 6, 6)) %>%
  theme_booktabs() %>%
  fix_border_issues() %>%
  set_table_properties(width = 1, layout = 'autofit')  %>%
  colformat_char(na_str = '-')
a

save_as_docx(a, path = here('output_data', 'Table_2.docx'))
```

### Smoking and testing meta-analysis {.tabset .tabset-fade .tabset-pills}

The steps to reproduce the Bayesian analysis are available above. We generate the treatment effect (mu) and standard deviations (tau) from the studies using the `meta` package
``` {r testing_meta}
source(here::here('scripts', 'rr_function.R'))

table_2 <- table_2_word

included_studies <- quality_table_2 %>%
  filter(overall_quality != 'poor') %>%
  select(lead_author) %>%
  filter(lead_author != "reiter")

meta <- tibble('author' = table_2$lead_author,
               'negative_smoker' = table_2$negative_current_smoker,
               'negative_never_smoker' = table_2$negative_never_smoker,
               'positive_smoker' = table_2$positive_current_smoker,
               'positive_never_smoker' = table_2$positive_never_smoker,
               'negative_former_smoker' = table_2$negative_former_smoker,
               'positive_former_smoker' = table_2$positive_former_smoker) %>%
        filter(author %in% included_studies$lead_author)
meta$author <- to_upper_camel_case(meta$author, sep_out = ", ")
meta$author <- plyr::mapvalues(meta$author,
                               from = cleaned_names,
                               to = correct_names)

testing_rr <- list()
testing_rr[[1]] <- RR_testing('Rentsch', 'current')
testing_rr[[2]] <- RR_testing('Rentsch', 'former')
testing_rr[[3]] <- RR_testing('Cho', 'current')
testing_rr[[4]] <- RR_testing('Cho', 'former')
testing_rr[[5]] <- RR_testing('Kolin', 'current')
testing_rr[[6]] <- RR_testing('Kolin', 'former')
testing_rr[[7]] <- RR_testing('de Lusignan', 'current')
testing_rr[[8]] <- RR_testing('de Lusignan', 'former')
testing_rr[[9]] <- RR_testing('Parrotta', 'current')
testing_rr[[10]] <- RR_testing('Parrotta', 'former')
testing_rr[[11]] <- RR_testing('Israel', 'current')
testing_rr[[12]] <- RR_testing('Israel', 'former')
testing_rr[[13]] <- RR_testing('Eugen-Olsen', 'current')
testing_rr[[14]] <- RR_testing('Eugen-Olsen', 'former')
testing_rr[[15]] <- RR_testing('Houlihan', 'current')
testing_rr[[16]] <- RR_testing('Houlihan', 'former')
testing_rr[[17]] <- RR_testing('Woolford', 'current')
testing_rr[[18]] <- RR_testing('Woolford', 'former')
testing_rr[[19]] <- RR_testing('Govind', 'current')
testing_rr[[20]] <- RR_testing('Govind', 'former')
testing_rr[[21]] <- RR_testing('Gu', 'current')
testing_rr[[22]] <- RR_testing('Gu', 'former')
testing_rr[[23]] <- RR_testing('Petrilli', 'current')
testing_rr[[24]] <- RR_testing('Petrilli', 'former')
testing_rr[[25]] <- RR_testing('Auvinen', 'current')
testing_rr[[26]] <- RR_testing('Auvinen', 'former')
testing_rr[[27]] <- RR_testing('Merkely', 'current')
testing_rr[[28]] <- RR_testing('Merkely', 'former')
testing_rr[[29]] <- RR_testing('Iversen', 'current')
testing_rr[[30]] <- RR_testing('Iversen', 'former')
testing_rr[[31]] <- RR_testing('Fillmore', 'current')
testing_rr[[32]] <- RR_testing('Fillmore', 'former')

data <- testing_rr
k <- do.call(rbind.data.frame, data)

#SEs for Niedzwiedz et al. 2020

#current vs. never

niedz_log_RR_1<-log(1.15)
niedz_log_SE_1<-(log(1.54)-log(0.86))/3.92

k <- k %>%
  add_row(., study = 'Niedzwiedz', smoking_status = 'current', log_RR = niedz_log_RR_1, log_SE = niedz_log_SE_1)
#former vs. never

niedz_log_RR_2<-log(1.42)
niedz_log_SE_2<-(log(1.69)-log(1.19))/3.92

k <- k %>%
  add_row(., study = 'Niedzwiedz', smoking_status = 'former', log_RR = niedz_log_RR_2, log_SE = niedz_log_SE_2)

numbers_in_analysis <- table_2_word %>%
        left_join(., quality_rating, by = 'lead_author') %>%
        filter(., overall_quality != 'poor') %>%
        add_row(lead_author = 'niedzwiedz', contributing_sample = 1474) %>%
        replace_na(list(negative_not_stated = 0, positive_not_stated = 0)) %>%
        mutate(contributing_sample = (contributing_sample - (negative_not_stated+positive_not_stated))) %>%
                       select(lead_author, contributing_sample) %>%
  na.omit()
numbers_in_analysis$lead_author <- to_upper_camel_case(numbers_in_analysis$lead_author, sep_out = ", ")
numbers_in_analysis$lead_author <- plyr::mapvalues(numbers_in_analysis$lead_author,
                               from = cleaned_names,
                               to = correct_names)

running_meta_count <- k %>%
  group_by(study) %>%
  summarise()

ecdf <- read_rds(here("data_clean", "bayesian_models", "ecdf_list.rds"))
m1_a_ecdf <- ecdf[[1]]
m2_a_ecdf <- ecdf[[2]]
m3_a_ecdf <- ecdf[[3]]
m4_a_ecdf <- ecdf[[4]]
m5_a_ecdf <- ecdf[[5]]
m6_a_ecdf <- ecdf[[6]]
m7_a_ecdf <- ecdf[[7]]
m8_a_ecdf <- ecdf[[8]]

```
The studies included in meta-analysis are `r knitr::combine_words(pull(included_studies, lead_author))`.

#### Current versus never smokers for testing positive
```{r current_never}
#current vs. never smokers
current_never_meta <- k %>%
  filter(smoking_status == 'current') %>%
  left_join(., numbers_in_analysis %>%
              rename("study" = lead_author),
            by = c("study")) %>%
  mutate(study = paste(study, contributing_sample, sep = ", n = ")) %>%
  select(-contributing_sample)

a <-metagen(current_never_meta$log_RR,
           current_never_meta$log_SE,
           studlab = current_never_meta$study,
           sm="RR",
           comb.fixed = F, comb.random = T)

write_rds(a, here("data_clean", "bayes_testing_current.rds"))
```
![Current smoking against never smoking for testing positive](figure/m1_a.png)
![ECDF for current smoking against never smoking for testing positive](figure/ECDF_m1_a.png)

The ECDF shows the probability of the risk being less than or equal to RR = 0.9, `r paste(round(m1_a_ecdf(0.9)*100, 2), "%", sep = "")`

#### Former versus never smokers
```{r former_never}
#former vs. never smokers

former_never_meta <- k %>%
  filter(smoking_status == 'former') %>%
  left_join(., numbers_in_analysis %>%
              rename("study" = lead_author),
            by = c("study")) %>%
  mutate(study = paste(study, contributing_sample, sep = ", n = ")) %>%
  select(-contributing_sample)

a<-metagen(former_never_meta$log_RR,
           former_never_meta$log_SE,
           studlab = former_never_meta$study,
           sm="RR", comb.fixed = F, comb.random = T)

write_rds(a, here("data_clean", "bayes_testing_former.rds"))
```
![Former smoking against never smoking for testing positive](figure/m2_a.png)
![ECDF for former smoking against never smoking for testing positive](figure/ECDF_m2_a.png)

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m2_a_ecdf(1.1))*100, 2), "%", sep = "")`

## Table 3: Smoking and COVID-19 hospitalisation
``` {r table_3}

#Table 3
table_3_word <- table_3 %>%
  mutate(., sample = sample_with_outcome) %>%
  mutate(., community_percentage = formatC(number_community/sample*100, digits = 2, format = "f")) %>%
  mutate(., community_current_smoker_percent = formatC(community_current_smoker/number_community*100, digits = 2, format = "f")) %>%
  mutate(., community_former_smoker_percent = formatC(community_former_smoker/number_community*100, digits = 2, format = "f")) %>%
  mutate(., community_current_former_smoker_percent = formatC(community_current_former_smoker/number_community*100, digits = 2, format = "f")) %>%
  mutate(., community_never_smoker_percent = formatC(community_never_smoker/number_community*100, digits = 2, format = "f")) %>%
  mutate(., community_never_unknown_smoker_percent = formatC(community_never_unknown_smoker/number_community*100, digits = 2, format = "f")) %>%
  mutate(., community_not_stated_percent = formatC(community_not_stated/number_community*100, digits = 2, format = "f")) %>%
  mutate(., number_hospitalised_percent = formatC(number_hospitalised/sample*100, digits = 2, format = "f")) %>%
  mutate(., hospitalised_current_smoker_percent = formatC(hospitalised_current_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
  mutate(., hospitalised_former_smoker_percent = formatC(hospitalised_former_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
  mutate(., hospitalised_current_former_smoker_percent = formatC(hospitalised_current_former_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
  mutate(., hospitalised_never_smoker_percent = formatC(hospitalised_never_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
  mutate(., hospitalised_never_unknown_smoker_percent = formatC(hospitalised_never_unknown_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
  mutate(., hospitalised_not_stated_percent = formatC(hospitalised_not_stated/number_hospitalised*100, digits = 2, format = "f")) %>%
  select(lead_author, sample_with_outcome, number_community, community_percentage,
         community_current_smoker, community_current_smoker_percent, community_former_smoker, 
         community_former_smoker_percent, community_current_former_smoker,
         community_current_former_smoker_percent, community_never_smoker, community_never_smoker_percent,
         community_never_unknown_smoker, community_never_unknown_smoker_percent, 
         community_not_stated, community_not_stated_percent, number_hospitalised, number_hospitalised_percent,
         hospitalised_current_smoker, hospitalised_current_smoker_percent, hospitalised_former_smoker,
         hospitalised_former_smoker_percent, hospitalised_current_former_smoker, hospitalised_current_former_smoker_percent,
         hospitalised_never_smoker, hospitalised_never_smoker_percent, hospitalised_never_unknown_smoker,
         hospitalised_never_unknown_smoker_percent, hospitalised_not_stated, hospitalised_not_stated_percent) %>%
  write_rds(here::here('data_clean', 'table_3_word.rds'))
```
The number of studies reporting on hospitalisation by smoking status is: `r paste(nrow(table_3_word))`
``` {r}
quality_table_3 <- table_3_word %>%
  left_join(., quality_rating, by = 'lead_author') %>%
  select(lead_author, overall_quality)

a <- table_3_word %>%
  filter(number_community >= 1) %>%
  mutate(Author = lead_author,
         Population = sample_with_outcome,
         Community = paste(
           paste(number_community, 
                 as.integer(community_percentage), sep = " ("), "%)", sep = ""),
         C_current_smoker = paste(paste(community_current_smoker, (community_current_smoker_percent), sep = " (")
                                  , "%)", sep = ""),
         C_current_smoker = na_if(C_current_smoker, "NA ( NA%)"),
         C_former_smoker = paste(paste(community_former_smoker, (community_former_smoker_percent), sep = " (")
                                 , "%)", sep = ""),
         C_former_smoker = na_if(C_former_smoker, "NA ( NA%)"),
         C_current_former_smoker = paste(paste(community_current_former_smoker,
                                               (community_current_former_smoker_percent), sep = " (")
                                         , "%)", sep = ""),
         C_current_former_smoker = na_if(C_current_former_smoker, "NA ( NA%)"),
         C_never_smoker = paste(paste(community_never_smoker,
                                      (community_never_smoker_percent), sep = " (")
                                , "%)", sep = ""),
         C_never_smoker = na_if(C_never_smoker, "NA ( NA%)"),
         C_never_unknown_smoker = paste(paste(community_never_unknown_smoker,
                                              (community_never_unknown_smoker_percent), sep = " (")
                                        , "%)", sep = ""),
         C_never_unknown_smoker = na_if(C_never_unknown_smoker, "NA ( NA%)"),
         C_not_stated = paste(paste(community_not_stated,
                                    (community_not_stated_percent), sep = " (")
                              , "%)", sep = ""),
         C_not_stated = na_if(C_not_stated, "NA ( NA%)")) %>%
  mutate(Hospitalised = paste(
    paste(number_hospitalised, 
          as.integer(number_hospitalised_percent), sep = " ("), "%)", sep = ""),
    H_current_smoker = paste(paste(hospitalised_current_smoker, (hospitalised_current_smoker_percent), sep = " (")
                             , "%)", sep = ""),
    H_current_smoker = na_if(H_current_smoker, "NA ( NA%)"),
    H_former_smoker = paste(paste(hospitalised_former_smoker, (hospitalised_former_smoker_percent), sep = " (")
                            , "%)", sep = ""),
    H_former_smoker = na_if(H_former_smoker, "NA ( NA%)"),
    H_current_former_smoker = paste(paste(hospitalised_current_former_smoker,
                                          (hospitalised_current_former_smoker_percent), sep = " (")
                                    , "%)", sep = ""),
    H_current_former_smoker = na_if(H_current_former_smoker, "NA ( NA%)"),
    H_never_smoker = paste(paste(hospitalised_never_smoker,
                                 (hospitalised_never_smoker_percent), sep = " (")
                           , "%)", sep = ""),
    H_never_smoker = na_if(H_never_smoker, "NA ( NA%)"),
    H_never_unknown_smoker = paste(paste(hospitalised_never_unknown_smoker,
                                         (hospitalised_never_unknown_smoker_percent), sep = " (")
                                   , "%)", sep = ""),
    H_never_unknown_smoker = na_if(H_never_unknown_smoker, "NA ( NA%)"),
    H_not_stated = paste(paste(hospitalised_not_stated,
                               (hospitalised_not_stated_percent), sep = " (")
                         , "%)", sep = ""),
    H_not_stated = na_if(H_not_stated, "NA ( NA%)")) %>%
  select(Author, Population, Community, C_current_smoker, C_former_smoker, C_current_former_smoker,
         C_never_smoker,C_never_unknown_smoker,C_not_stated, Hospitalised, H_current_smoker, H_former_smoker,
         H_current_former_smoker, H_never_smoker, H_never_unknown_smoker, H_not_stated)

a$Author <- to_upper_camel_case(a$Author, sep_out = ", ")
a$Author <- plyr::mapvalues(a$Author,
                               from = cleaned_names,
                               to = correct_names)

numeric_columns <- 'Population'

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 hospitalisation by smoking status') %>%
  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

a <- set_header_labels(a,
                       Population = 'Population with outcome',
                       Community = "N (%)",
                       C_current_smoker = "Current smoker (%)",
                       C_former_smoker = "Former smoker (%)",
                       C_current_former_smoker = "Current/former smoker (%)",
                       C_never_smoker = "Never smoker (%)",
                       C_never_unknown_smoker = "Never/unknown smoker (%)",
                       C_not_stated = "Not stated (%)",
                       Hospitalised = "N (%)",
                       H_current_smoker = "Current smoker (%)",
                       H_former_smoker = "Former smoker (%)",
                       H_current_former_smoker = "Current/former smoker (%)",
                       H_never_smoker = "Never smoker (%)",
                       H_never_unknown_smoker = "Never/unknown smoker (%)",
                       H_not_stated = "Not stated (%)") %>%
  add_header_row(top = TRUE, values = c("","Community", "Hospitalised" ), colwidths = c(2, 7, 7)) %>%
  theme_booktabs() %>%
  fix_border_issues() %>%
  set_table_properties(width = 1, layout = 'autofit')
a

save_as_docx(a, path = here('output_data', 'Table_3.docx'))
```

### Smoking and hospitalisation meta-analysis {.tabset .tabset-fade .tabset-pills}
```{r current_never hospitalisaiton, echo = F}
table_3 <- table_3_word
included_studies <- quality_table_3 %>%
  filter(overall_quality != 'poor') %>%
  select(lead_author)

```
The studies included in meta-analysis are `r knitr::combine_words(pull(included_studies, lead_author))`.

#### Current versus never smoking for hospitalisation
``` {r}
# Data --------------------------------------------------------------------
meta <- tibble('author' = table_3$lead_author,
               'community_smoker' = table_3$community_current_smoker,
               'community_never_smoker' = table_3$community_never_smoker, 
               'hospitalised_smoker' = table_3$hospitalised_current_smoker, 
               'hospitalised_never_smoker' = table_3$hospitalised_never_smoker,
               'community_former_smoker' = table_3$community_former_smoker,
               'hospitalised_former_smoker' = table_3$hospitalised_former_smoker) %>%
  filter(author %in% included_studies$lead_author)

meta$author <- to_upper_camel_case(meta$author, sep_out = ", ")
meta$author <- plyr::mapvalues(meta$author,
                               from = cleaned_names,
                               to = correct_names)

# Current smoker hospitalisation ------------------------------------------
event_rates_smoker <- meta %>%
  mutate(., Ee = hospitalised_smoker) %>%
  mutate(., Ne = (hospitalised_smoker+community_smoker)) %>%
  mutate(., Ec = hospitalised_never_smoker) %>%
  mutate(., Nc = (hospitalised_never_smoker+community_never_smoker)) %>%
  rename('Author' = author) %>%
  select(Author, Ee, Ne, Ec, Nc)

event_rates_smoker <- metabin(Ee,
                              Ne,
                              Ec,
                              Nc,
                              data = event_rates_smoker,
                              studlab = paste(Author),
                              comb.fixed = F,
                              comb.random = T,
                              method.tau = 'SJ',
                              hakn = T,
                              prediction = F,
                              incr = 0.1,
                              sm = 'RR')

write_rds(event_rates_smoker, here("data_clean", "bayes_hospital_current.rds"))
```
![Current smoking against never smoking for hospitalisation](figure/m3_a.png)

![ECDF for current smoking against never smoking for hospitalisation](figure/ECDF_m3_a.png)

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m3_a_ecdf(1.1))*100, 2), "%", sep = "")`

#### Former versus never smoking for hospitalisation
```{r}
# Former smoker hospitalisation -------------------------------------------
event_rates_former <- meta %>%
  mutate(., Ee = hospitalised_former_smoker) %>%
  mutate(., Ne = (hospitalised_former_smoker+community_former_smoker)) %>%
  mutate(., Ec = hospitalised_never_smoker) %>%
  mutate(., Nc = (hospitalised_never_smoker+community_never_smoker)) %>%
  rename('Author' = author) %>%
  select(Author, Ee, Ne, Ec, Nc)

event_rates_former <- metabin(Ee,
                              Ne,
                              Ec,
                              Nc,
                              data = event_rates_former,
                              studlab = paste(Author),
                              comb.fixed = F,
                              comb.random = T,
                              method.tau = 'SJ',
                              hakn = T,
                              prediction = F,
                              incr = 0.1,
                              sm = 'RR')

running_meta_count <- running_meta_count %>%
  full_join(., meta %>%
  rename("study" = author), by = "study") %>%
  select(study)

write_rds(event_rates_former, here("data_clean", "bayes_hospital_former.rds"))
```
![Former smoking against never smoking for hospitalisation](figure/m4_a.png)
![ECDF for Former smoking against never smoking for hospitalisation](figure/ECDF_m4_a.png)

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m4_a_ecdf(1.1))*100, 2), "%", sep = "")`

## Table 4: Smoking and COVID-19 disease severity
```{r table_4}
#Table 4
table_4_word <- table_4 %>%
  mutate(., sample = sample_with_severity) %>%
  mutate(., non_severe_disease_percentage = formatC(non_severe_disease/sample*100, digits = 2, format = "f")) %>%
  mutate(., non_severe_current_smoker_percent = formatC(non_severe_current_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
  mutate(., non_severe_former_smoker_percent = formatC(non_severe_former_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
  mutate(., non_severe_current_former_smoker_percent = formatC(non_severe_current_former_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
  mutate(., non_severe_never_smoker_percent = formatC(non_severe_never_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
  mutate(., non_severe_never_unknown_smoker_percent = formatC(non_severe_never_unknown_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
  mutate(., non_severe_not_stated_percent = formatC(non_severe_not_stated/non_severe_disease*100, digits = 2, format = "f")) %>%
  mutate(., severe_disease_number_percent = formatC(severe_disease_number/sample*100, digits = 2, format = "f")) %>%
  mutate(., severe_disease_current_smoker_percent = formatC(severe_disease_current_smoker/severe_disease_number*100, digits = 2, format = "f")) %>%
  mutate(., severe_disease_former_smoker_percent = formatC(severe_disease_former_smoker/severe_disease_number*100, digits = 2, format = "f")) %>%
  mutate(., severe_disease_current_former_smoker_percent = formatC(severe_disease_current_former_smoker/severe_disease_number*100, digits = 2, format = "f")) %>%
  mutate(., severe_disease_never_smoker_percent = formatC(severe_disease_never_smoker/severe_disease_number*100, digits = 2, format = "f")) %>%
  mutate(., severe_disease_never_unknown_percent = formatC(severe_disease_never_unknown/severe_disease_number*100, digits = 2, format = "f")) %>%
  mutate(., severe_disease_not_stated_percent = formatC(severe_disease_not_stated/severe_disease_number*100, digits = 2, format = "f")) %>%
  select(lead_author, sample, non_severe_disease, non_severe_disease_percentage, non_severe_current_smoker,
         non_severe_current_smoker_percent, non_severe_former_smoker, non_severe_former_smoker_percent,
         non_severe_current_former_smoker, non_severe_current_former_smoker_percent, non_severe_never_smoker,
         non_severe_never_smoker_percent, non_severe_never_unknown_smoker, non_severe_never_unknown_smoker_percent,
         non_severe_not_stated, non_severe_not_stated_percent, severe_disease_number, severe_disease_number_percent,
         severe_disease_current_smoker, severe_disease_current_smoker_percent, 
         severe_disease_former_smoker, severe_disease_former_smoker_percent, severe_disease_current_former_smoker,
         severe_disease_current_former_smoker_percent, severe_disease_never_smoker, severe_disease_never_smoker_percent,
         severe_disease_never_unknown, severe_disease_never_unknown_percent, severe_disease_not_stated,
         severe_disease_not_stated_percent) %>%
  write_rds(here::here('data_clean', 'table_4_word.rds'))
```
The number of studies reporting on hospitalisation by smoking status is: `r paste(nrow(table_4_word))`
``` {r}
quality_table_4 <- table_4_word %>%
  left_join(., quality_rating, by = 'lead_author') %>%
  select(lead_author, overall_quality)

a <- table_4_word %>%
  mutate(Author = lead_author,
         Population = sample,
         non_severe_disease = paste(
           paste(non_severe_disease, 
                 as.integer(non_severe_disease_percentage), sep = " ("), "%)", sep = ""),
         ns_current_smoker = paste(paste(non_severe_current_smoker, (non_severe_current_smoker_percent), sep = " (")
                                  , "%)", sep = ""),
         ns_current_smoker = na_if(ns_current_smoker, "NA ( NA%)"),
         ns_former_smoker = paste(paste(non_severe_former_smoker, (non_severe_former_smoker_percent), sep = " (")
                                 , "%)", sep = ""),
         ns_former_smoker = na_if(ns_former_smoker, "NA ( NA%)"),
         ns_current_former_smoker = paste(paste(non_severe_current_former_smoker,
                                               (non_severe_current_former_smoker_percent), sep = " (")
                                         , "%)", sep = ""),
         ns_current_former_smoker = na_if(ns_current_former_smoker, "NA ( NA%)"),
         ns_never_smoker = paste(paste(non_severe_never_smoker,
                                      (non_severe_never_smoker_percent), sep = " (")
                                , "%)", sep = ""),
         ns_never_smoker = na_if(ns_never_smoker, "NA ( NA%)"),
         ns_never_unknown_smoker = paste(paste(non_severe_never_unknown_smoker,
                                              (non_severe_never_unknown_smoker_percent), sep = " (")
                                        , "%)", sep = ""),
         ns_never_unknown_smoker = na_if(ns_never_unknown_smoker, "NA ( NA%)"),
         ns_not_stated = paste(paste(non_severe_not_stated,
                                    (non_severe_not_stated_percent), sep = " (")
                              , "%)", sep = ""),
         ns_not_stated = na_if(ns_not_stated, "NA ( NA%)"),
         ) %>%
  mutate(severe_disease = paste(
    paste(severe_disease_number, 
          as.integer(severe_disease_number_percent), sep = " ("), "%)", sep = ""),
    s_current_smoker = paste(paste(severe_disease_current_smoker, (severe_disease_current_smoker_percent), sep = " (")
                             , "%)", sep = ""),
    s_current_smoker = na_if(s_current_smoker, "NA ( NA%)"),
    s_former_smoker = paste(paste(severe_disease_former_smoker, (severe_disease_former_smoker_percent), sep = " (")
                            , "%)", sep = ""),
    s_former_smoker = na_if(s_former_smoker, "NA ( NA%)"),
    s_current_former_smoker = paste(paste(severe_disease_current_former_smoker,
                                          (severe_disease_current_former_smoker_percent), sep = " (")
                                    , "%)", sep = ""),
    s_current_former_smoker = na_if(s_current_former_smoker, "NA ( NA%)"),
    s_never_smoker = paste(paste(severe_disease_never_smoker,
                                 (severe_disease_never_smoker_percent), sep = " (")
                           , "%)", sep = ""),
    s_never_smoker = na_if(s_never_smoker, "NA ( NA%)"),
    s_never_unknown_smoker = paste(paste(severe_disease_never_unknown,
                                         (severe_disease_never_unknown_percent), sep = " (")
                                   , "%)", sep = ""),
    s_never_unknown_smoker = na_if(s_never_unknown_smoker, "NA ( NA%)"),
    s_not_stated = paste(paste(severe_disease_not_stated,
                                 (severe_disease_not_stated_percent), sep = " (")
                           , "%)", sep = ""),
    s_not_stated = na_if(s_not_stated, "NA ( NA%)")) %>%
  select(Author, Population, non_severe_disease, ns_current_smoker, ns_former_smoker,
         ns_current_former_smoker, ns_never_smoker, ns_never_unknown_smoker, ns_not_stated,
         severe_disease, s_current_smoker, s_former_smoker, s_current_former_smoker, s_never_smoker, s_never_unknown_smoker,
         s_not_stated)

a$Author <- to_upper_camel_case(a$Author, sep_out = ", ")
a$Author <- plyr::mapvalues(a$Author,
                               from = cleaned_names,
                               to = correct_names)  
numeric_columns <- 'Population'

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 severity by smoking status') %>%
  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

a <- set_header_labels(a,
                       Population = "Population with severity",
                       non_severe_disease = "N (%)",
                       ns_current_smoker = "Current smoker (%)",
                       ns_former_smoker = "Former smoker (%)",
                       ns_current_former_smoker = "Current/former smoker (%)",
                       ns_never_unknown_smoker = "Never/unknown smoker (%)",
                       ns_never_smoker = "Never smoker (%)",
                       ns_not_stated = "Not stated (%)",
                       severe_disease = "N (%)", 
                       s_current_smoker = "Current smoker (%)", 
                       s_former_smoker = "Former smoker (%)", 
                       s_current_former_smoker = "Current/former smoker (%)",
                       s_never_unknown_smoker = "Never/unknown smoker (%)",
                       s_never_smoker = "Never smoker (%)",
                       s_not_stated = "Not stated (%)") %>%
  add_header_row(top = TRUE, values = c("","Non severe disease", "Severe disease" ), colwidths = c(2, 7, 7)) %>%
  theme_booktabs() %>%
  fix_border_issues() %>%
  set_table_properties(width = 1, layout = 'autofit')
a

save_as_docx(a, path = here('output_data', 'Table_4.docx'))
```

### Smoking and severity meta-analysis {.tabset .tabset-fade .tabset-pills}
```{r current_never severity, echo = F}
table_4 <- table_4_word
included_studies <- quality_table_4 %>%
  filter(overall_quality != 'poor') %>%
  select(lead_author)
```
The studies included in meta-analysis are `r knitr::combine_words(pull(included_studies, lead_author))`.

#### Current versus never smoking for risk of severe disease
``` {r}
# Data --------------------------------------------------------------------
meta <- tibble('author' = table_4$lead_author,
               'non_severe_smoker' = table_4$non_severe_current_smoker,
               'non_severe_never_smoker' = table_4$non_severe_never_smoker,
               'severe_smoker' = table_4$severe_disease_current_smoker,
               'severe_never_smoker' = table_4$severe_disease_never_smoker,
               'non_severe_former_smoker' = table_4$non_severe_former_smoker,
               'severe_former_smoker' = table_4$severe_disease_former_smoker) %>%
  filter(author %in% included_studies$lead_author)

meta$author <- to_upper_camel_case(meta$author, sep_out = ", ")
meta$author <- plyr::mapvalues(meta$author,
                               from = cleaned_names,
                               to = correct_names)

# Current smoker severity ------------------------------------------
event_rates_smoker <- meta %>%
  mutate(., Ee = severe_smoker) %>%
  mutate(., Ne = (severe_smoker+non_severe_smoker)) %>%
  mutate(., Ec = severe_never_smoker) %>%
  mutate(., Nc = (severe_never_smoker+non_severe_never_smoker)) %>%
  rename('Author' = author) %>%
  select(Author, Ee, Ne, Ec, Nc)

event_rates_smoker <- metabin(Ee,
                              Ne,
                              Ec,
                              Nc,
                              data = event_rates_smoker,
                              studlab = paste(Author),
                              comb.fixed = F,
                              comb.random = T,
                              method.tau = 'SJ',
                              hakn = F,
                              prediction = F,
                              incr = 0.1,
                              sm = 'RR')

running_meta_count <- running_meta_count %>%
  full_join(., meta %>%
  rename("study" = author), by = "study") %>%
  select(study)

write_rds(event_rates_smoker, here("data_clean", "bayes_severity_current.rds"))
```
![Current smoking against never smoking for disease severity](figure/m5_a.png)
![ECDF for current smoking against never smoking for disease severity](figure/ECDF_m5_a.png)

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m5_a_ecdf(1.1))*100, 2), "%", sep = "")`

#### Former versus never smoking for risk of severe disease
```{r}
# Former smoker severity -------------------------------------------
event_rates_former <- meta %>%
  mutate(., Ee = severe_former_smoker) %>%
  mutate(., Ne = (severe_former_smoker+non_severe_former_smoker)) %>%
  mutate(., Ec = severe_never_smoker) %>%
  mutate(., Nc = (severe_never_smoker+non_severe_never_smoker)) %>%
  rename('Author' = author) %>%
  select(Author, Ee, Ne, Ec, Nc)

event_rates_former <- metabin(Ee,
                              Ne,
                              Ec,
                              Nc,
                              data = event_rates_former,
                              studlab = paste(Author),
                              comb.fixed = F,
                              comb.random = T,
                              method.tau = 'SJ',
                              hakn = F,
                              prediction = F,
                              incr = 0.1,
                              sm = 'RR')

write_rds(event_rates_former, here("data_clean", "bayes_severity_former.rds"))
```
![Former smoking against never smoking for disease severity](figure/m6_a.png)
![ECDF for former smoking against never smoking for disease severity](figure/ECDF_m6_a.png)

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m6_a_ecdf(1.1))*100, 2), "%", sep = "")`

## Table 5: Smoking and COVID-19 mortality
```{r table_5}
#Table 5
table_5_word <- table_5 %>%
  mutate(., sample = sample_with_deaths) %>%
  mutate(., deaths_percentage = formatC(deaths/sample*100, digits = 2, format = "f")) %>%
  mutate(., death_current_smokers_percent = formatC(death_current_smokers/deaths*100, digits = 2, format = "f")) %>%
  mutate(., death_former_smokers_percent = formatC(death_former_smokers/deaths*100, digits = 2, format = "f")) %>%
  mutate(., death_current_former_smokers_percent = formatC(death_current_former_smokers/deaths*100, digits = 2, format = "f")) %>%
  mutate(., death_never_smokers_percent = formatC(death_never_smokers/deaths*100, digits = 2, format = "f")) %>%
  mutate(., death_never_unknown_smokers_percent = formatC(death_never_unknown_smokers/deaths*100, digits = 2, format = "f")) %>%
  mutate(., death_not_stated_percent = formatC(death_not_stated/deaths*100, digits = 2, format = "f")) %>%
  mutate(., recovered_percentage = formatC(recovered/sample*100, digits = 2, format = "f")) %>%
  mutate(., recovered_current_smokers_percent = formatC(recovered_current_smoking/recovered*100, digits = 2, format = "f")) %>%
  mutate(., recovered_former_smokers_percent = formatC(recovered_former_smoker/recovered*100, digits = 2, format = "f")) %>%
  mutate(., recovered_current_former_smokers_percent = formatC(recovered_current_former_smokers/recovered*100, digits = 2, format = "f")) %>%
  mutate(., recovered_never_smokers_percent = formatC(recovered_never_smoker/recovered*100, digits = 2, format = "f")) %>%
  mutate(., recovered_never_unknown_smokers_percent = formatC(recovered_never_unknown_smoker/recovered*100, digits = 2, format = "f")) %>%
  mutate(., recovered_not_stated_percent = formatC(recovered_not_stated/recovered*100, digits = 2, format = "f")) %>%
  select(lead_author, sample, recovered, recovered_percentage, recovered_current_smoking, recovered_current_smokers_percent,
         recovered_former_smoker, recovered_former_smokers_percent, recovered_current_former_smokers,
         recovered_current_former_smokers_percent, recovered_never_smoker, recovered_never_smokers_percent,
         recovered_never_unknown_smoker, recovered_never_unknown_smokers_percent, recovered_not_stated,
         recovered_not_stated_percent,
         deaths, deaths_percentage, death_current_smokers, death_current_smokers_percent,
         death_former_smokers, death_former_smokers_percent, death_current_former_smokers,
         death_current_former_smokers_percent, death_never_smokers, death_never_smokers_percent,
         death_never_unknown_smokers, death_never_unknown_smokers_percent, death_not_stated, death_not_stated_percent) %>%
  write_rds(here::here('data_clean', 'table_5_word.rds'))
```
The number of studies reporting on death by smoking status is: `r paste(nrow(table_5_word))`

``` {r}
quality_table_5 <- table_5_word %>%
  left_join(., quality_rating, by = 'lead_author') %>%
  select(lead_author, overall_quality)

a <- table_5_word %>%
  mutate(Author = lead_author,
         Population = sample,
         recovered = paste(
           paste(recovered, 
                 as.integer(recovered_percentage), sep = " ("), "%)", sep = ""),
         recovered_current_smoker = paste(paste(recovered_current_smoking, 
                                                 (recovered_current_smokers_percent), sep = " (")
                                  , "%)", sep = ""),
         recovered_current_smoker = na_if(recovered_current_smoker, "NA ( NA%)"),
         recovered_former_smoker = paste(paste(recovered_former_smoker,
                                               (recovered_former_smokers_percent), sep = " (")
                                 , "%)", sep = ""),
         recovered_former_smoker = na_if(recovered_former_smoker, "NA ( NA%)"),
         recovered_current_former_smoker = paste(paste(recovered_current_former_smokers,
                                               (recovered_current_former_smokers_percent), sep = " (")
                                         , "%)", sep = ""),
         recovered_current_former_smoker = na_if(recovered_current_former_smoker, "NA ( NA%)"),
         recovered_never_smoker = paste(paste(recovered_never_smoker,
                                      (recovered_never_smokers_percent), sep = " (")
                                , "%)", sep = ""),
         recovered_never_smoker = na_if(recovered_never_smoker, "NA ( NA%)"),
         recovered_never_unknown_smoker = paste(paste(recovered_never_unknown_smoker,
                                              (recovered_never_unknown_smokers_percent), sep = " (")
                                        , "%)", sep = ""),
         recovered_never_unknown_smoker = na_if(recovered_never_unknown_smoker, "NA ( NA%)"),
         recovered_not_stated = paste(paste(recovered_not_stated,
                                    (recovered_not_stated_percent), sep = " (")
                              , "%)", sep = ""),
         recovered_not_stated = na_if(recovered_not_stated, "NA ( NA%)")) %>%
  mutate(deaths = paste(
           paste(deaths, 
                 as.integer(deaths_percentage), sep = " ("), "%)", sep = ""),
         death_current_smoker = paste(paste(death_current_smokers, 
                                                 (death_current_smokers_percent), sep = " (")
                                  , "%)", sep = ""),
         death_current_smoker = na_if(death_current_smoker, "NA ( NA%)"),
         death_former_smoker = paste(paste(death_former_smokers,
                                               (death_former_smokers_percent), sep = " (")
                                 , "%)", sep = ""),
         death_former_smoker = na_if(death_former_smoker, "NA ( NA%)"),
         death_current_former_smoker = paste(paste(death_current_former_smokers,
                                               (death_current_former_smokers_percent), sep = " (")
                                         , "%)", sep = ""),
         death_current_former_smoker = na_if(death_current_former_smoker, "NA ( NA%)"),
         death_never_smoker = paste(paste(death_never_smokers,
                                      (death_never_smokers_percent), sep = " (")
                                , "%)", sep = ""),
         death_never_smoker = na_if(death_never_smoker, "NA ( NA%)"),
         death_never_unknown_smoker = paste(paste(death_never_unknown_smokers,
                                              (death_never_unknown_smokers_percent), sep = " (")
                                        , "%)", sep = ""),
         death_never_unknown_smoker = na_if(death_never_unknown_smoker, "NA ( NA%)"),
         death_not_stated = paste(paste(death_not_stated,
                                    (death_not_stated_percent), sep = " (")
                              , "%)", sep = ""),
         death_not_stated = na_if(death_not_stated, "NA ( NA%)")) %>%
  select(Author, Population, recovered, recovered_current_smoker, recovered_former_smoker, recovered_current_former_smoker,
         recovered_never_smoker, recovered_never_unknown_smoker, recovered_not_stated, deaths, death_current_smoker,
         death_former_smoker, death_current_former_smoker, death_never_smoker, death_never_unknown_smoker,
         death_not_stated)

a$Author <- to_upper_camel_case(a$Author, sep_out = ", ")
a$Author <- plyr::mapvalues(a$Author,
                               from = cleaned_names,
                               to = correct_names)

numeric_columns <- 'Population'

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 mortality by smoking status') %>%
  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

a <- set_header_labels(a,
                       Population = "Population with mortality",
                       recovered = "N (%)",
                       recovered_current_smoker = "Current smoker (%)",
                       recovered_former_smoker = "Former smoker (%)",
                       recovered_current_former_smoker = "Current/former smoker (%)",
                       recovered_never_unknown_smoker = "Never/unknown smoker (%)",
                       recovered_never_smoker = "Never smoker (%)",
                       recovered_not_stated = "Not stated (%)",
                       deaths = "N (%)", 
                       death_current_smoker = "Current smoker (%)", 
                       death_former_smoker = "Former smoker (%)", 
                       death_current_former_smoker = "Current/former smoker (%)",
                       death_never_unknown_smoker = "Never/unknown smoker (%)",
                       death_never_smoker = "Never smoker (%)",
                       death_not_stated = "Not stated (%)") %>%
  add_header_row(top = TRUE, values = c("","Recovered", "Died" ), colwidths = c(2, 7, 7)) %>%
  theme_booktabs() %>%
  fix_border_issues() %>%
  set_table_properties(width = 1, layout = 'autofit')
a

save_as_docx(a, path = here('output_data', 'Table_5.docx'))
```

### Smoking and COVID-19 mortality meta-analysis {.tabset .tabset-fade .tabset-pills}
```{r current_never mortality, echo = F}
exclude <- c("rajter", "garassino", "holman")
table_5 <- table_5_word
included_studies <- quality_table_5 %>%
  filter(overall_quality != 'poor') %>%
  filter(!lead_author %in% exclude) %>%
  select(lead_author)
```
The studies included in meta-analysis are `r knitr::combine_words(pull(included_studies, lead_author))`.

#### Current versus never smoking for disease mortality
``` {r}
# Data --------------------------------------------------------------------
meta <- tibble('author' = table_5$lead_author,
               'recovered_current_smoker' = table_5$recovered_current_smoking,
               'recovered_never_smoker' = table_5$recovered_never_smoker,
               'death_current_smoker' = table_5$death_current_smokers,
               'death_never_smoker' = table_5$death_never_smokers,
               'recovered_former_smoker' = table_5$recovered_former_smoker,
               'death_former_smoker' = table_5$death_former_smokers) %>%
  filter(author %in% included_studies$lead_author)

meta$author <- to_upper_camel_case(meta$author, sep_out = ", ")
meta$author <- plyr::mapvalues(meta$author,
                               from = cleaned_names,
                               to = correct_names)

# Current smoker mortality ------------------------------------------
event_rates_smoker <- meta %>%
  mutate(., Ee = death_current_smoker) %>%
  mutate(., Ne = (death_current_smoker+recovered_current_smoker)) %>%
  mutate(., Ec = death_never_smoker) %>%
  mutate(., Nc = (death_never_smoker+recovered_never_smoker)) %>%
  rename('Author' = author) %>%
  select(Author, Ee, Ne, Ec, Nc)

event_rates_smoker <- metabin(Ee,
                              Ne,
                              Ec,
                              Nc,
                              data = event_rates_smoker,
                              studlab = paste(Author),
                              comb.fixed = F,
                              comb.random = T,
                              method.tau = 'SJ',
                              hakn = F,
                              prediction = F,
                              incr = 0.1,
                              sm = 'RR')

write_rds(event_rates_smoker, here("data_clean", "bayes_mortality_current.rds"))
```
![Current smoking against never smoking for mortality](figure/m7_a.png)
![ECDF for current smoking against never smoking for mortality](figure/ECDF_m7_a.png)

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m7_a_ecdf(1.1))*100, 2), "%", sep = "")`

#### Former versus never smoking for disease mortality
```{r}
# Former smoker mortality ------------------------------------------
event_rates_former <- meta %>%
  mutate(., Ee = death_former_smoker) %>%
  mutate(., Ne = (death_former_smoker+recovered_former_smoker)) %>%
  mutate(., Ec = death_never_smoker) %>%
  mutate(., Nc = (death_never_smoker+recovered_never_smoker)) %>%
  rename('Author' = author) %>%
  select(Author, Ee, Ne, Ec, Nc)

event_rates_former <- metabin(Ee,
                              Ne,
                              Ec,
                              Nc,
                              data = event_rates_former,
                              studlab = paste(Author),
                              comb.fixed = F,
                              comb.random = T,
                              method.tau = 'SJ',
                              hakn = F,
                              prediction = F,
                              incr = 0.1,
                              sm = 'RR')

write_rds(event_rates_former, here("data_clean", "bayes_mortality_former.rds"))

running_meta_count <- running_meta_count %>%
  full_join(., meta %>%
  rename("study" = author), by = "study") %>%
  select(study)

```
![Former smoking against never smoking for mortality](figure/m8_a.png)
![ECDF for former smoking against never smoking for mortality](figure/ECDF_m8_a.png)

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m8_a_ecdf(1.1))*100, 2), "%", sep = "")`

## Study quality assessment
``` {r table_6}
#Table 6
fair <- filter(quality_rating, quality_rating$overall_quality == "fair" ) %>%
  filter(!lead_author %in% exclude_from_analysis) %>%
  nrow()
poor <- filter(quality_rating, quality_rating$overall_quality == "poor" ) %>%
  filter(!lead_author %in% exclude_from_analysis) %>%
  nrow()
```
The number of **Fair** quality studies are `r paste(fair)` with `r paste(poor)` **Poor** quality studies

The number of studies included in **Meta-analysis** are `r paste(nrow(running_meta_count))`

# Final PRISMA table
``` {r prisma_final}
prisma <- prisma %>%
  mutate("Studies included in meta-analysis" = nrow(running_meta_count))
flextable(prisma)
```

# Supplementary tables {.tabset .tabset-fade .tabset-pills}

## Supplementary tables

### Data on clinical or virological diagnosis and studies stratifying smoking status
``` {r supplementary_1}
a <- data_study_general %>%
  filter(review_version %in% analysed_versions) %>%
  select(lead_author, study_design, clinical_diagnosis, stratified_smoking) %>%
  filter(!lead_author %in% exclude_from_analysis) %>%
  rename("Lead author" = "lead_author",
         "Study design" = "study_design",
         "Clinical diagnoses included" = "clinical_diagnosis",
         "Stratified smoking reported" = "stratified_smoking") %>%
 mutate("Lead author" = to_upper_camel_case(`Lead author`, sep_out = ", "),
        "Lead author" = plyr::mapvalues(`Lead author`,
                               from = cleaned_names,
                               to = correct_names),
        "Study design" = to_upper_camel_case(`Study design`, sep_out = " "))

a <- a %>%
  flextable() %>%
    set_caption(caption = 'Studies design, inclusion of clinical diagnoses and stratified smoking') %>%
  set_table_properties(width = 1, layout = 'autofit')

save_as_docx(a, path = here('output_data', 'Supplementary_2.docx'))
```

### Country level smoking estimates

``` {r supplementary_country}

country_smoking <-  read_sheet(sheets_id, sheet = 'national_smoking_prevalence') %>%
  flextable()
save_as_docx(country_smoking, path = here('output_data', 'Supplementary_3.docx'))
```

### Sensitivity
``` {r supplementary_2}

```

### 2020-08-26 reports/markdown_report.Rmd
restarting Bayesian analysis repo

--- 2020-08-26 reports/bayesian_report.Rmd
@@
 ---
-title: 'Smoking and COVID Living Review (v7): a bayesian analysis'
+title: "Smoking and COVID Living Review"
 author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
@@
     theme: spacelab
-    toc: false
+    toc: true
     toc_collapse: true
@@
 ---
+
 ```{r setup, include=FALSE}
@@
 ```
-# Data preparation {.tabset .tabset-fade .tabset-pills}
-
-## Load packages
+
+## Data preparation {.tabset .tabset-fade .tabset-pills}
+
+### Load packages
+
 * **Load packages**: load packages and useful scripts for the analysis
-* **Load scripts**: We have written functions to handle much of the Bayesian analysis
-* **Load Bayesian datasets**: The Bayesian approach relies on large simulated datasets from the observed data, for this report we will only present the results graphically from data we have simulated elsewhere.
-The scripts `bayes_scripts.R`, `bayesian_data_generation.R` and `bayesian_analysis.R` should be able to be run locally if you want to generate your own simulated data.
-Alternatively you can use the pre-generated data which is stored in the `data_clean\bayesian_models` folder in the github repository [here](https://github.com/DidDrog11/smoking_covid/tree/master) and then you can skip the `bayesian_data_generation.R` step.
-```{r install_packages, include=FALSE}
-library(brms)
-library(ggplot2)
-library(tidybayes)
-library(readr)
-library(dplyr)
-library(ggridges)
-library(glue)
-library(stringr)
-library(forcats)
-library(here)
-library(meta)
-library(snakecase)
-library(flextable)
-library(googlesheets4)
-library(linelist)
-library(DT)
-library(tidyr)
-library(epitools)
-```
-
-```{r load_scripts}
-source(here("scripts", "author_dictionary.R"))
-```
-## Download data
-We maintain a Google sheets document [here](sheets_id) which contains all eligible articles the research team have identified.
-This is currently viewable only and not open for editing.
-We have collected the general information for the search in the first sheet, the studies general information in the second, further information on study populations, outcomes of interest and quality of the studies are contained in the further sheets.
-These sheets are organised in a similar way to the manuscript and so are labelled table_1 to table_6.
-
-``` {r google_data}
-sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1HV2Zk_UD2_s1zzd47dCyb4UKRSM0dDg7zmUPA9CMdx4/edit?usp=sharing")
+
+```{r libraries}
+
+library('tidyverse', quietly = T)
+library('ggplot2', quietly = T)
+library('dplyr', quietly = T)
+library('readxl', quietly = T)
+library('linelist', quietly = T)
+library('googlesheets4', quietly = T)
+library('naniar', quietly = T)
+library('gridExtra', quietly = T)
+library('snakecase', quietly = T)
+library('epitools', quietly = T)
+library('meta', quietly = T)
+library('knitr', quietly = T)
+library('kableExtra', quietly = T)
+library('snakecase', quietly = T)
+library('here', quietly = T)
+library('flextable', quietly = T)
+library('DT', quietly = T)
+```
+
+### Download from Google sheets
+
+* **Load data**: import raw dataset directly from Google sheets
+We import the data from the dataset that is available in the "data for analysis" folder on the Google drive.
+This is available to be viewed with the [link](sheets_id) and is only available for viewing not editing.
+
+In future versions we may open a tab in the Google sheet to allow people to submit relevant papers for us to incorporate in our review.
+
+```{r raw_data, cache=T}
+sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1fuOcaN7K_f3msmt0aQAWwOWjgD4SsEt7iJ_0to1wotA/edit?usp=sharing")
 search_details <- read_sheet(sheets_id, range = 'article_screening')
@@
 ```
-We have extracted data from `r nrow(data_study_general)` studies, however, some of these are duplicates and some have been removed from the analysis set.
-For example there have been multiple versions of the ISARIC report and several publications have been retracted.
-We aim to keep on top of this but there is some risk that in between versions data may be superseded.
-
-## Raw Data
+
+The data is downloaded from the sheet and split into dataframes to represent the current sheets and analysis.
+
+### Raw Data
+
 The first sheet is **Search details**  we use this sheet to keep track of the number of studies found in our search and keep track of the number of titles and abstracts screened and number of papers included in each review version.
@@
 ```{r head_search_details}
-flextable(search_details[,1:11])
-```
+flextable(head(search_details[,1:11]))
+```
+
 The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.
-This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`.
-These are live links and should direct to the relevant article, however, some may have been deleted or require institutional access.
+
+This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`
 ```{r tail_study_details, include=FALSE}
 datatable(data_study_general[1:4] %>%
-            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`),
-                   doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
-          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5),
-          escape = F)
-```
+            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`)),
+          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5))
+```
+
 The third sheet produces **Table 1** this contains the details of the population of the studies in our review.
 Primarily includes smoking status but also the proportion of females in the study and the age of participants.
+
 ```{r tail_table_1, include=FALSE}
@@
 ```
+
 The fourth sheet produces **Table 2** this contains the data extracted from relevant studies on testing data and smoking status.
@@
 These are sparse datasets as many studies do not report on all levels of smoking exposure.
+
 ```{r tail_table_2, include=FALSE}
@@
 ```
-The fifth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status. 
+
+The fourth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status.
+
 ```{r tail_table_3, include=FALSE}
@@
 ```
-The sixth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status.
+
+The fourth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status.
 Disease severity is differentially reported across the studies in some there is reference to WHO severity score in others we have based the distinction on clinical characteristics or whether patients were admitted to Intensive or Critical care areas.
+
 ```{r tail_table_4, include=FALSE}
@@
 ```
-The seventh sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status.
+
+The fifth sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status.
 A caution needed when interpreting this data is that most studies have not followed patients up for sufficient amounts of time to be able to accurately report on disease associated mortality.
+
 ```{r tail_table_5, include=FALSE}
@@
 ```
-The final sheet is used to collate data on **Study quality** we changed our method on assessing study quality in version 3 of the review. We use the following criteria:
-
-- **Good:** 
-  1. <20% missing data on smoking status and used a reliable self-report measure that distinguished between current, former and never smoking status AND
-  2. biochemical verification of smoking status and reported results from adjusted analyses OR 
-  3. reported data from a representative/random sample
-- **Fair**
-  1. <20% missing data on smoking status and used a reliable self-report measure that distinguished between current, former and never smoking status
-- **Poor**
-  1. The remainder of the studies
-
-## Clean data and save updated versions as RDS files
+
+The final sheet was used to collate data on **Study quality** however, we have discontinued the use of this tool from version 3 due to the large number of components that were unable to be assessed in the rapidly expanding area of COVID-19 related research.
+We have moved onto a different method of assessing study quality which is explained in greater detail in the manuscript.
+
+### Clean data and save updated versions as RDS files
+
 The data from the google sheets document is then cleaned using a script from the `linelist` package which has been developed by T.
-Jombart and Z.
+Jombat and Z.
 Kamvar within the [R Epidemics Consortium](https://www.repidemicsconsortium.org/)
+
 The script `clean_data()` ensures consistency among column names, country names, dates and missing values in the included sheets.
 Data is also saved as `.rds` files in the `/data_clean` folder of the project repository.
+
 ```{r clean_data}
@@
 data_study_general <- data_study_general %>%
-  rename(date_published = `date_published_format(2020-05-12)`) %>%
+  rename('notes' = 11, date_published = `date_published_format(2020-05-12)`) %>%
   clean_data(protect = protect_columns_1) %>%
@@
 ```
-The package `flextable` can save tables in a format that is easily manageable within word which we currently use to write the manuscript (until I get better with RMarkdown).
-At various points I save these to the `output_data` folder.
+
 ``` {r save_as_table_for_word, include=FALSE}
@@
 
-save_as_docx(a, path = here('output_data', 'all_studies.docx'))
-```
-
-# Preparing the descriptive analysis {.tabset .tabset-fade .tabset-pills}
-
-## Choosing which data we want to include in the current review?
+save_as_docx(a, path = here('data_clean', 'all_studies.docx'))
+```
+## Preparing the descriptive analysis {.tabset .tabset-fade .tabset-pills}
+
+### Choosing which data we want to include in the current review? 
 ``` {r versioning}
@@
 
-prev_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6')
+prev_versions <- c('v1', 'v2', 'v3', 'v4', 'v5')
 #This can categorise which studies we want to look at
-current_version <- c('v7')
+current_version <- c('v6')
 #This will categorise which studies we are including in the current report
 
-analysed_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6', 'v7')
+analysed_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6')
 #This will incorporate all studies into the current version of the report
@@
 ``` {r excluding_studies}
-exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old', 'isaric_4', 'mehra', 'hopkinson', 'gaibazzi', 'miyara_updated', 'russell_old', "petrilli_old", "magagnoli_old", "niedzwiedz_old", "bello_chavolla_old", "kimmig_old", "russell_old", "zuo_zuo_old", "patel_old", "shi_zuo_old")
+
+exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old', 'isaric_4', 'mehra', 'hopkinson', 'gaibazzi', 'miyara_updated', 'russell', "petrilli_old", "magagnoli_old", "niedzwiedz_old", "bello_chavolla_old", "kimmig_old", "russell_old", "zuo_zuo_old", "patel_old")
 ```
@@
 The excluded studies are defined in this chunk of code.
-We have excluded this version the following studies:
--`r knitr::combine_words(exclude_from_analysis)`
-Generally we label these as _old in the Google sheet but there are exceptions to this so we manually define them at the moment.
-
-## Filter the included studies
+We have excluded this version the following studies: `r knitr::combine_words(exclude_from_analysis)`.
+
+### Filter the included studies
 Based on the version and excluded studies we produce the tables for analysis. This is a searchable table.
 ``` {r descriptives}
+
 data_study_general <- data_study_general %>%
@@
   filter(!(lead_author %in% exclude_from_analysis)) %>%
-  write_rds(., here::here('data_clean', 'data_study_general.rds'))
+  write_csv(., here::here('data_clean', 'data_study_general.csv'))
 
@@
 
-save_as_docx(a, path = here('output_data', 'included_studies.docx'))
-```
-
-# Study descriptive analysis {.tabset .tabset-fade .tabset-pills}
-
-## The PRISMA diagram will be updated
-The PRISMA reporting guidelines for systematic reviews the guidance is available [here](http://www.prisma-statement.org/).
-This takes the format of the number of articles obtained from the search from [OVID](https://ovidsp.dc1.ovid.com/sp-4.07.0b/ovidweb.cgi) (login may be needed), [medRxiv](https://www.medrxiv.org/search/text_abstract_title%3Asmoking%2BAND%2BCOVID%20text_abstract_title_flags%3Amatch-all%20jcode%3Amedrxiv%20numresults%3A75%20sort%3Apublication-date%20direction%3Adescending%20format_result%3Astandard), and articles identified by the research team.
-These are then screened at the title and abstract level.
-Full texts are then read to extract relevant data, if articles are excluded at this level we describe the reason for this.
-All of the selected articles will contribute to the narrative synthesis but only those that are judged 'good' or 'fair' quality will be included in the Bayesian meta-analysis.
+save_as_docx(a, path = here('data_clean', 'included_studies.docx'))
+```
+
+## Descriptive analysis {.tabset .tabset-fade .tabset-pills}
+
+### The PRISMA diagram will be updated
+The PRISMA reporting guidelines for systematic reviews the guidance is available [here](http://www.prisma-statement.org/). **In Progress**
+
 ``` {r PRISMA}
@@
 previous_total <- search_details %>%
-  filter(review_version %in% prev_versions) %>%
+  filter(review_version %in% previous_review_versions) %>%
   select(total_included) %>%
@@
 
-prisma <- search_details %>%
+a <- search_details %>%
   group_by(review_version) %>%
-  filter(review_version == current_version) %>%
+  filter(review_version == current_review_version) %>%
   mutate(title_abstract = sum(ovid_number_results, medrxiv_number_results, other_source_included),
@@
          "Studies in narrative synthesis" = "current_total")
-flextable(prisma)
-```
-The number of studies contributing to meta-analysis is calculated later.
-
-## We summarise where the studies were conducted
+
+flextable(a)
+```
+
+### We summarise where the studies were conducted
 
@@
 
-``` {r countries, cache = TRUE}
+``` {r countries}
 library('ggmap')
+
 #Countries
@@
   select(lon, lat) %>%
-  tidyr::drop_na() %>%
+  drop_na %>%
   sp::SpatialPointsDataFrame(coords = .[,1:2], data = .) %>%
@@
 
-save_as_docx(a, path = here('output_data', 'Countries_performing_studies.docx'))  
-```
-Currently studies are reported from `r paste(length(unique(country)))`
-
-## A map is produced for where the studies are performed
-``` {r interactive global map, cache = TRUE}
+save_as_docx(a, path = here('data_clean', 'Countries_performing_studies.docx'))
+
+```
+### A map is produced for where the studies are performed
+``` {r interactive global map}
 library(rgdal)
@@
 
-my_bins <- c(0, 1, 5, 10, 20, 30, 40, 70)
+my_bins <- c(0, 1, 5, 10, 20, 30, 40, 50)
 mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)
@@
 The number of confirmed COVID-19 cases are shown when resting over each country for reference, data on confirmed cases have been obtained from [Rami Krispin's `coronavirus` package](https://ramikrispin.github.io/coronavirus/).
-The number of cases is updated when the report is generated so will not remain in date, it is current as of `r paste(Sys.time())`.
-Countries displaying 0 cases are most likely due to naming mismatches.
-
-## We sumarise the settings in which they were conducted
+The number of cases is updated when the report is generated so will not remain in date, it is current as of `r paste(Sys.time())`
+
+### We sumarise the settings in which they were conducted
 
@@
 Some registry studies utilising primary care networks are solely conducted in the community.
+
 ``` {r setting}
+
 #Setting
@@
 
-save_as_docx(a, path = here('output_data', 'Countries_performing_studies.docx'))  
-```
-## We summarise the design of the included studies
+save_as_docx(a, path = here('data_clean', 'Countries_performing_studies.docx'))
+
+```
+### We summarise the design of the included studies
 
 The majority were retrospective cohorts. There were fewer case-control and cross-sectional studies. There were some prospective cohorts established and reported.
+
 ``` {r design}
+
 design <- to_upper_camel_case(data_study_general$study_design, sep_out = ' ') %>%
@@
 
-save_as_docx(a, path = here('output_data', 'Study_designs.docx'))
+save_as_docx(a, path = here('data_clean', 'Study_designs.docx'))
 
@@
 ```
-## We update the number of participants we have included
+### We update the number of participants we have included
 
 * **Summary statistics:** this table will give us the mean, median and IQR of the samples in the review
+
 ``` {r numbers}
 #Numbers
-summary(table_1$sample_size)
-```
+a <- summary(table_1$sample_size)
+a
+
+```
+
 * **Total number of participants**: `r paste(sum(table_1$sample_size))`
 
-## We want to know if they use a CRF or electronic health record
+### We want to know if they use a CRF or electronic health record
+
 The majority of data on smoking outcomes is extracted from electronic health records that are routinely collected in the provision of healthcare.
@@
 ``` {r data_source}
+
 #Source of data
@@
   set_table_properties(width = 0.5, layout = 'autofit')
-```
-
-# Smoking descriptive data {.tabset .tabset-fade .tabset-pills}
+
+```
+
+### Now about the smoking
+
 ``` {r smokers}
@@
   filter(., current_smoker != 'NA')
+
 ```
 * **Current smokers:** `r paste(nrow(current_smok))` studies report on the numbers of current smokers
+
 ``` {r former smokers}
@@
 * **Former smokers:** `r paste(nrow(former_smok))` studies report on the numbers of former smokers
+
 ```{r never smokers}
@@
 * **Never smokers:** `r paste(nrow(never_smok))` studies report on the numbers of never smokers
+
 ``` {r current former smokers}
@@
 * **Current/former smokers:** `r paste(nrow(current_former_smok))` studies report on the numbers of current/former smokers without distinguishing between them, often termed as yes to a smoking history question.
+
 ``` {r missing/never/unknown/not stated}
@@
   + **Maximum missing:** `r paste(round(maximum,2), "%", sep = "")`
-
-## Studies reporting current, former and never smoking status
 ``` {r complete}
@@
 b$`Lead author` <- to_upper_camel_case(b$`Lead author`, sep_out = ", ")
-b$`Lead author` <- plyr::mapvalues(b$`Lead author`,
-                               from = cleaned_names,
-                               to = correct_names)  
+
 datatable(b,
@@
 #These are the studies with data on current, former and never smokers
-
-b <- b %>%
-  select(-"Study ID") %>%
-  flextable() %>%
-    set_caption(caption = 'Studies reporting complete smoking status') %>%
-  colformat_num(col_keys = c("Sample size", "Current smokers", "Former smokers", "Never smokers"),
-                digits = 0, na_str = '-', big.mark = ',') %>%
-  set_table_properties(width = 1, layout = 'autofit')
-
-save_as_docx(b, path = here('output_data', 'Supplementary_1a.docx'))
 ```
 * **Studies with complete smoking status:** `r paste(nrow(full_smoking_status))`, `r paste(round((nrow(full_smoking_status)/nrow(data_study_general)*100),2), "%", sep = "")`
-
-## Studies providing partially complete data on smoking status
 
@@
 b$`Lead author` <- to_upper_camel_case(b$`Lead author`, sep_out = ", ")
-b$`Lead author` <- plyr::mapvalues(b$`Lead author`,
-                               from = cleaned_names,
-                               to = correct_names)
+```
+* **Studies reporting partially incomplete smoking data:** `r paste(nrow(semi_full_smoking_status))`  
+These studies typically report smoking history as the measure of exposure which is why they have grouped values for current/former smokers
+``` {r}
 
@@
   formatRound(columns = c(2:4), digits = 0, interval = 3, mark = ",")
-
-b <- b %>%
-  select(-"Study ID") %>%
-  flextable() %>%
-    set_caption(caption = 'Studies reporting complete smoking status') %>%
-  colformat_num(col_keys = c("Sample size", "Current/former smokers", "Never smokers"),
-                digits = 0, na_str = '-', big.mark = ',') %>%
-  set_table_properties(width = 1, layout = 'autofit')
-
-save_as_docx(b, path = here('output_data', 'Supplementary_1b.docx'))
-```
-
-* **Studies reporting partially incomplete smoking data:** `r paste(nrow(semi_full_smoking_status))`  
-These studies typically report smoking history as the measure of exposure which is why they have grouped values for current/former smokers
-
-## Studies providing incomplete data on smoking status
+```
+
 ``` {r}
@@
 b$`Lead author` <- to_upper_camel_case(b$`Lead author`, sep_out = ", ")
-b$`Lead author` <- plyr::mapvalues(b$`Lead author`,
-                               from = cleaned_names,
-                               to = correct_names)
+```
+* **Studies missing key smoking data:** `r paste(nrow(incomplete_smoking_status))`  
+``` {r}
 
@@
   formatRound(columns = c(2:8), digits = 0, interval = 3, mark = ",")
-
-numeric_names <- colnames(b[2:9])
-b <- b %>%
-  select(-"Study ID") %>%
-  flextable() %>%
-    set_caption(caption = 'Studies reporting complete smoking status') %>%
-  colformat_num(col_keys = numeric_names,
-                digits = 0, na_str = '-', big.mark = ',') %>%
-  set_table_properties(width = 1, layout = 'autofit')
-
-save_as_docx(b, path = here('output_data', 'Supplementary_1c.docx'))
-```
-
-* **Studies missing key smoking data:** `r paste(nrow(incomplete_smoking_status))`
-
-# Smoking prevalence by country {.tabset .tabset-fade .tabset-pills}
+```
+
+## Smoking prevalence by country {.tabset .tabset-fade .tabset-pills}
 ``` {r smoking_country}
@@
 Confidence intervals are obtained through bootstrap estimates of a thousand repetitions of the included studies.
-Following the production of the confidence intervals from bootstrapping in `olgas_script.R` the prevalence graphs are produced in the `smoking_plots.R` script.
+Following the production of the confidence intervals the prevalence graphs are produced in the `smoking_plots.R` script.
 This uses data extracted on current and former smoking prevalence data from countries which have included studies in this review.
 
-## Graph for smoking prevalence
+### Graph for smoking prevalence
 **Current smoking prevalence**
@@
 The weighted mean current smoking prevalence is displayed for each setting.
+
 ![Country smoking prevalence](figure/figure_2a.png)
 
-## Graph for former smoking prevalence
+### Graph for former smoking prevalence
 **Former smoker prevalence**
@@
 
-# Final results tables and meta-analyses {.tabset .tabset-fade .tabset-pills}
-
-## Table 1: Included studies
+## Final results tables and meta-analyses {.tabset .tabset-fade .tabset-pills}
+
+### Table 1: Included studies
 ``` {r table_1}
@@
 a$`Lead author` <- to_upper_camel_case(a$`Lead author`, sep_out = ", ")
+
+cleaned_names <- c('Chow, Us, Cdc',
+                   "Gold, Us, Cdc",
+                   "Miyara, Updated",
+                   "De, La, Rica",
+                   "Opensafely, Collaborative",
+                   "Bello, Chavolla",
+                   "Carrillo, Vega",
+                   "De, Lusignan",
+                   "Mejia, Vilet",
+                   "Heili, Frades",
+                   "Vaquero, Roncero",
+                   "Al, Hindawi",
+                   "Del, Valle",
+                   "Soto, Mota",
+                   "Eugen, Olsen",
+                   "Martinez, Portilla",
+                   "Raisi, Estabragh",
+                   "Miyara, Medrxiv",
+                   "Hernandez, Gardano",
+                   "Siso, Almirall",
+                   "Martinez, Jiminez",
+                   "eugen_olsen",
+                   "Mcqueenie",
+                   "Martin, Jiminez",
+                   "Bello, Chavolla",
+                   "De, Melo",
+                   "De, Souza",
+                   "Bello, Chavolla, Antonio, Villa",
+                   "Antonio, Villa")
+
+correct_names <- c('Chow (US CDC)',
+                   "Gold (US CDC)",
+                   "Miyara",
+                   "de la Rica",
+                   "The Opensafely Collaborative",
+                   "Bello-Chavolla",
+                   "Carillo-Vega",
+                   "de Lusignan",
+                   "Mejia-Vilet",
+                   "Heili-Frades",
+                   "Vaquero-Roncero",
+                   "Al-Hindawi",
+                   "del Valle",
+                   "Soto-Mota",
+                   "Eugen-Olsen",
+                   "Martinez-Portilla",
+                   "Raisi-Estabragh",
+                   "Miyara",
+                   "Hernandez-Gardano",
+                   "Siso-Almirall",
+                   "Martinez-Jiminez",
+                   "Eugen-Olsen",
+                   "McQueenie",
+                   "Martin-Jiminez",
+                   "Bello-Chavolla",
+                   "de Melo",
+                   "de Souza",
+                   "Bello-Chavolla, Antonio-Villa",
+                   "Antonio-Villa")
+
 author_list <- plyr::mapvalues(a$`Lead author`,
@@
 
-save_as_docx(a, path = here('output_data', 'Table_1.docx'))
-```
-
-## Table 2: Smoking and COVID-19 diagnostic testing
+save_as_docx(a, path = here('data_clean', 'Table_1.docx'))
+```
+
+### Table 2: Smoking and COVID-19 diagnostic testing
 ``` {r table_2}
@@
 
-save_as_docx(a, path = here('output_data', 'Table_2.docx'))
-```
-
-### Smoking and testing meta-analysis {.tabset .tabset-fade .tabset-pills}
-
-The steps to reproduce the Bayesian analysis are available above. We generate the treatment effect (mu) and standard deviations (tau) from the studies using the `meta` package
+save_as_docx(a, path = here('data_clean', 'Table_2.docx'))
+```
+#### Smoking and testing meta-analysis
 ``` {r testing_meta}
@@
   filter(overall_quality != 'poor') %>%
-  select(lead_author) %>%
-  filter(lead_author != "reiter")
+  select(lead_author)
 
@@
 testing_rr[[28]] <- RR_testing('Merkely', 'former')
-testing_rr[[29]] <- RR_testing('Iversen', 'current')
-testing_rr[[30]] <- RR_testing('Iversen', 'former')
-testing_rr[[31]] <- RR_testing('Fillmore', 'current')
-testing_rr[[32]] <- RR_testing('Fillmore', 'former')
 
@@
   summarise()
-
-ecdf <- read_rds(here("data_clean", "bayesian_models", "ecdf_list.rds"))
-m1_a_ecdf <- ecdf[[1]]
-m2_a_ecdf <- ecdf[[2]]
-m3_a_ecdf <- ecdf[[3]]
-m4_a_ecdf <- ecdf[[4]]
-m5_a_ecdf <- ecdf[[5]]
-m6_a_ecdf <- ecdf[[6]]
-m7_a_ecdf <- ecdf[[7]]
-m8_a_ecdf <- ecdf[[8]]
-
 ```
@@
 
-#### Current versus never smokers for testing positive
+  * **Current versus never smokers**
 ```{r current_never}
@@
 write_rds(a, here("data_clean", "bayes_testing_current.rds"))
-```
-![Current smoking against never smoking for testing positive](figure/m1_a.png)
-![ECDF for current smoking against never smoking for testing positive](figure/ECDF_m1_a.png)
-
-The ECDF shows the probability of the risk being less than or equal to RR = 0.9, `r paste(round(m1_a_ecdf(0.9)*100, 2), "%", sep = "")`
-
-#### Former versus never smokers
+
+png(here::here('reports', 'figure','fig_3.png'), width=1024, height=546, res=120)
+forest(a,
+       xlim = c(0.1, 4),
+       sortvar = a$TE,
+       rightlabs = c('RR', '95% CI', 'Weight'),
+       leftlabs = c('Author', 'logRR', 'SE'),
+       print.tau2 = F,
+       col.diamond = 'blue',
+       col.diamond.lines = 'black',
+       col.square = 'black',
+       col.square.lines = 'black',
+       digits.sd = 2,
+       colgap.forest.left = unit(15,"mm"))
+null <- dev.off()
+
+a
+```
+
+![Figure 3: Current versus never smokers](figure/fig_3.png)
+
+  * **Former versus never smokers**
 ```{r former_never}
@@
 write_rds(a, here("data_clean", "bayes_testing_former.rds"))
-```
-![Former smoking against never smoking for testing positive](figure/m2_a.png)
-![ECDF for former smoking against never smoking for testing positive](figure/ECDF_m2_a.png)
-
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m2_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
-## Table 3: Smoking and COVID-19 hospitalisation
+
+png(here::here('reports', 'figure','fig_4.png'), width=1024, height=546, res=120)
+forest(a,
+       xlim = c(0.3, 4),
+       sortvar = a$TE,
+       rightlabs = c('RR', '95% CI', 'Weight'),
+       leftlabs = c('Author', 'logRR', 'SE'),
+       print.tau2 = F,
+       col.diamond = 'blue',
+       col.diamond.lines = 'black',
+       col.square = 'black',
+       col.square.lines = 'black',
+       digits.sd = 2,
+       colgap.forest.left = unit(15,"mm"))
+null <- dev.off()
+
+a
+
+```
+
+![Figure 4: Former versus never smokers](figure/fig_4.png)
+
+### Table 3: Smoking and COVID-19 hospitalisation
 ``` {r table_3}
@@
 
-save_as_docx(a, path = here('output_data', 'Table_3.docx'))
-```
-
-### Smoking and hospitalisation meta-analysis {.tabset .tabset-fade .tabset-pills}
+save_as_docx(a, path = here('data_clean', 'Table_3.docx'))
+
+```
+
+#### Smoking and hospitalisation meta-analysis
+
 ```{r current_never hospitalisaiton, echo = F}
@@
 
-#### Current versus never smoking for hospitalisation
 ``` {r}
@@
 write_rds(event_rates_smoker, here("data_clean", "bayes_hospital_current.rds"))
-```
-![Current smoking against never smoking for hospitalisation](figure/m3_a.png)
-
-![ECDF for current smoking against never smoking for hospitalisation](figure/ECDF_m3_a.png)
-
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m3_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
-#### Former versus never smoking for hospitalisation
+
+png(here::here('reports', 'figure', 'fig_5.png'), width=1480, height=546, res=120)
+current_smoker_hospitalisation <- forest(event_rates_smoker,
+                                sortvar = Author,
+                                xlim = c(0.3, 3),
+                                rightlabs = c('RR', '95% CI', 'Weight'),
+                                leftlabs = c('Author', 'logRR', 'SE'),
+                                lab.e = 'Current smoker',
+                                lab.c = 'Never smoker',
+                                print.tau2 = F,
+                                col.diamond = 'blue',
+                                col.diamond.lines = 'black',
+                                col.square = 'black',
+                                col.square.lines = 'black',
+                                digits.sd = 2)
+null <- dev.off()
+
+event_rates_smoker
+```
+![Figure 5: Current versus never smokers](figure/fig_5.png)
+
 ```{r}
@@
 
+write_rds(event_rates_former, here("data_clean", "bayes_hospital_former.rds"))
+
+png(here::here('reports', 'figure', 'fig_6.png'), width=1480, height=546, res=120)
+former_smoker_hospitalisation <- forest(event_rates_former,
+                                         sortvar = Author,
+                                         xlim = c(0.3, 3),
+                                         rightlabs = c('RR', '95% CI', 'Weight'),
+                                         leftlabs = c('Author', 'logRR', 'SE'),
+                                         lab.e = 'Former smoker',
+                                         lab.c = 'Never smoker',
+                                         print.tau2 = F,
+                                         col.diamond = 'blue',
+                                         col.diamond.lines = 'black',
+                                         col.square = 'black',
+                                         col.square.lines = 'black',
+                                         digits.sd = 2)
+null <- dev.off()
+event_rates_former
+
 running_meta_count <- running_meta_count %>%
@@
   select(study)
-
-write_rds(event_rates_former, here("data_clean", "bayes_hospital_former.rds"))
-```
-![Former smoking against never smoking for hospitalisation](figure/m4_a.png)
-![ECDF for Former smoking against never smoking for hospitalisation](figure/ECDF_m4_a.png)
-
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m4_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
-## Table 4: Smoking and COVID-19 disease severity
+```
+![Figure 6: Former versus never smokers](figure/fig_6.png)
+
+### Table 4: Smoking and COVID-19 disease severity
 ```{r table_4}
+
 #Table 4
@@
 
-save_as_docx(a, path = here('output_data', 'Table_4.docx'))
-```
-
-### Smoking and severity meta-analysis {.tabset .tabset-fade .tabset-pills}
+save_as_docx(a, path = here('data_clean', 'Table_4.docx'))
+
+```
+
+#### Smoking and severity meta-analysis
 ```{r current_never severity, echo = F}
@@
 
-#### Current versus never smoking for risk of severe disease
 ``` {r}
@@
 
-running_meta_count <- running_meta_count %>%
-  full_join(., meta %>%
-  rename("study" = author), by = "study") %>%
-  select(study)
-
 write_rds(event_rates_smoker, here("data_clean", "bayes_severity_current.rds"))
-```
-![Current smoking against never smoking for disease severity](figure/m5_a.png)
-![ECDF for current smoking against never smoking for disease severity](figure/ECDF_m5_a.png)
-
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m5_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
-#### Former versus never smoking for risk of severe disease
+
+png(here::here('reports', 'figure', 'fig_7.png'), width=1024, height=546, res=120)
+current_smoker_severity <- forest(event_rates_smoker,
+                                         sortvar = Author,
+                                         xlim = c(0.1, 5),
+                                         rightlabs = c('RR', '95% CI', 'Weight'),
+                                         leftlabs = c('Author', 'logRR', 'SE'),
+                                         lab.e = 'Current smoker',
+                                         lab.c = 'Never smoker',
+                                         print.tau2 = F,
+                                         col.diamond = 'blue',
+                                         col.diamond.lines = 'black',
+                                         col.square = 'black',
+                                         col.square.lines = 'black',
+                                         digits.sd = 2)
+null <- dev.off()
+
+event_rates_smoker
+```
+![Figure 7: Current versus never smokers](figure/fig_7.png)
+
 ```{r}
@@
 write_rds(event_rates_former, here("data_clean", "bayes_severity_former.rds"))
-```
-![Former smoking against never smoking for disease severity](figure/m6_a.png)
-![ECDF for former smoking against never smoking for disease severity](figure/ECDF_m6_a.png)
-
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m6_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
-## Table 5: Smoking and COVID-19 mortality
+
+png(here::here('reports', 'figure', 'fig_8.png'), width=1024, height=546, res=120)
+former_smoker_severity <- forest(event_rates_former,
+                                        sortvar = Author,
+                                        xlim = c(0.2, 10),
+                                        rightlabs = c('RR', '95% CI', 'Weight'),
+                                        leftlabs = c('Author', 'logRR', 'SE'),
+                                        lab.e = 'Former smoker',
+                                        lab.c = 'Never smoker',
+                                        print.tau2 = F,
+                                        col.diamond = 'blue',
+                                        col.diamond.lines = 'black',
+                                        col.square = 'black',
+                                        col.square.lines = 'black',
+                                        digits.sd = 2)
+null <- dev.off()
+
+event_rates_former
+
+running_meta_count <- running_meta_count %>%
+  full_join(., meta %>%
+  rename("study" = author), by = "study") %>%
+  select(study)
+```
+![Figure 8: Former versus never smokers](figure/fig_8.png)
+
+### Table 5: Smoking and COVID-19 mortality
 ```{r table_5}
+
 #Table 5
@@
 ```
-The number of studies reporting on death by smoking status is: `r paste(nrow(table_5_word))`
-
+The number of studies reporting on hospitalisation by smoking status is: `r paste(nrow(table_5_word))`
 ``` {r}
+
 quality_table_5 <- table_5_word %>%
@@
 
-save_as_docx(a, path = here('output_data', 'Table_5.docx'))
-```
-
-### Smoking and COVID-19 mortality meta-analysis {.tabset .tabset-fade .tabset-pills}
+save_as_docx(a, path = here('data_clean', 'Table_5.docx'))
+```
+#### Smoking and COVID-19 mortality meta-analysis
+
 ```{r current_never mortality, echo = F}
-exclude <- c("rajter", "garassino", "holman")
+exclude <- c("rajter", "garassino")
 table_5 <- table_5_word
@@
 ```
-The studies included in meta-analysis are `r knitr::combine_words(pull(included_studies, lead_author))`.
-
-#### Current versus never smoking for disease mortality
+The studies included in meta-analysis are `r paste(included_studies, sep = ', ')`.
+
 ``` {r}
@@
 write_rds(event_rates_smoker, here("data_clean", "bayes_mortality_current.rds"))
-```
-![Current smoking against never smoking for mortality](figure/m7_a.png)
-![ECDF for current smoking against never smoking for mortality](figure/ECDF_m7_a.png)
-
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m7_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
-#### Former versus never smoking for disease mortality
+
+png(here::here('reports', 'figure', 'fig_9.png'), width=1024, height=546, res=120)
+current_smoker_mortality <- forest(event_rates_smoker,
+                                         sortvar = Author,
+                                         xlim = c(0.1, 5),
+                                         rightlabs = c('RR', '95% CI', 'Weight'),
+                                         leftlabs = c('Author', 'logRR', 'SE'),
+                                         lab.e = 'Current smoker',
+                                         lab.c = 'Never smoker',
+                                         print.tau2 = F,
+                                         col.diamond = 'blue',
+                                         col.diamond.lines = 'black',
+                                         col.square = 'black',
+                                         col.square.lines = 'black',
+                                         digits.sd = 2)
+null <- dev.off()
+
+event_rates_smoker
+```
+![Figure 9: Current versus never smokers](figure/fig_9.png)
+
 ```{r}
@@
 
+png(here::here('reports', 'figure', 'fig_10.png'), width=1024, height=546, res=120)
+former_smoker_mortality <- forest(event_rates_former,
+                                         sortvar = Author,
+                                         xlim = c(0.1, 5),
+                                         rightlabs = c('RR', '95% CI', 'Weight'),
+                                         leftlabs = c('Author', 'logRR', 'SE'),
+                                         lab.e = 'Former smoker',
+                                         lab.c = 'Never smoker',
+                                         print.tau2 = F,
+                                         col.diamond = 'blue',
+                                         col.diamond.lines = 'black',
+                                         col.square = 'black',
+                                         col.square.lines = 'black',
+                                         digits.sd = 2)
+null <- dev.off()
+
+event_rates_former
+
 running_meta_count <- running_meta_count %>%
@@
 ```
-![Former smoking against never smoking for mortality](figure/m8_a.png)
-![ECDF for former smoking against never smoking for mortality](figure/ECDF_m8_a.png)
-
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m8_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
-## Study quality assessment
+
+![Figure 10: Former versus never smokers](figure/fig_10.png)
+
+### Study quality assessment
 ``` {r table_6}
@@
 The number of studies included in **Meta-analysis** are `r paste(nrow(running_meta_count))`
-
-# Final PRISMA table
-``` {r prisma_final}
-prisma <- prisma %>%
-  mutate("Studies included in meta-analysis" = nrow(running_meta_count))
-flextable(prisma)
-```
-
-# Supplementary tables {.tabset .tabset-fade .tabset-pills}
-
-## Supplementary tables
-
-### Data on clinical or virological diagnosis and studies stratifying smoking status
-``` {r supplementary_1}
-a <- data_study_general %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(lead_author, study_design, clinical_diagnosis, stratified_smoking) %>%
-  filter(!lead_author %in% exclude_from_analysis) %>%
-  rename("Lead author" = "lead_author",
-         "Study design" = "study_design",
-         "Clinical diagnoses included" = "clinical_diagnosis",
-         "Stratified smoking reported" = "stratified_smoking") %>%
- mutate("Lead author" = to_upper_camel_case(`Lead author`, sep_out = ", "),
-        "Lead author" = plyr::mapvalues(`Lead author`,
-                               from = cleaned_names,
-                               to = correct_names),
-        "Study design" = to_upper_camel_case(`Study design`, sep_out = " "))
-
-a <- a %>%
-  flextable() %>%
-    set_caption(caption = 'Studies design, inclusion of clinical diagnoses and stratified smoking') %>%
-  set_table_properties(width = 1, layout = 'autofit')
-
-save_as_docx(a, path = here('output_data', 'Supplementary_2.docx'))
-```
-
-### Country level smoking estimates
-
-``` {r supplementary_country}
-
-country_smoking <-  read_sheet(sheets_id, sheet = 'national_smoking_prevalence') %>%
-  flextable()
-save_as_docx(country_smoking, path = here('output_data', 'Supplementary_3.docx'))
-```
-
-### Sensitivity
-``` {r supplementary_2}
-
-```

### 2020-08-26 scripts/author_dictionary.R

@@ -1,0 +1,59 @@
+cleaned_names <- c('Chow, Us, Cdc',
+                   "Gold, Us, Cdc",
+                   "Miyara, Updated",
+                   "De, La, Rica",
+                   "Opensafely, Collaborative",
+                   "Bello, Chavolla",
+                   "Carrillo, Vega",
+                   "De, Lusignan",
+                   "Mejia, Vilet",
+                   "Heili, Frades",
+                   "Vaquero, Roncero",
+                   "Al, Hindawi",
+                   "Del, Valle",
+                   "Soto, Mota",
+                   "Eugen, Olsen",
+                   "Martinez, Portilla",
+                   "Raisi, Estabragh",
+                   "Miyara, Medrxiv",
+                   "Hernandez, Gardano",
+                   "Siso, Almirall",
+                   "Martinez, Jiminez",
+                   "eugen_olsen",
+                   "Mcqueenie",
+                   "Martin, Jiminez",
+                   "Bello, Chavolla",
+                   "De, Melo",
+                   "De, Souza",
+                   "Bello, Chavolla, Antonio, Villa",
+                   "Antonio, Villa")
+
+correct_names <- c('Chow (US CDC)',
+                   "Gold (US CDC)",
+                   "Miyara",
+                   "de la Rica",
+                   "The Opensafely Collaborative",
+                   "Bello-Chavolla",
+                   "Carillo-Vega",
+                   "de Lusignan",
+                   "Mejia-Vilet",
+                   "Heili-Frades",
+                   "Vaquero-Roncero",
+                   "Al-Hindawi",
+                   "del Valle",
+                   "Soto-Mota",
+                   "Eugen-Olsen",
+                   "Martinez-Portilla",
+                   "Raisi-Estabragh",
+                   "Miyara",
+                   "Hernandez-Gardano",
+                   "Siso-Almirall",
+                   "Martinez-Jiminez",
+                   "Eugen-Olsen",
+                   "McQueenie",
+                   "Martin-Jiminez",
+                   "Bello-Chavolla",
+                   "de Melo",
+                   "de Souza",
+                   "Bello-Chavolla, Antonio-Villa",
+                   "Antonio-Villa")

### 2020-08-26 scripts/bayes_scripts.R

@@ -1,0 +1,102 @@
+extract_TE <- function(dataset) {
+  tibble(dataset$studlab,
+         dataset$TE,
+         dataset$seTE
+         ) %>%
+    rename(Author = 1, TE = 2, seTE = 3)
+}
+
+bayes_test <- function(dataset, defined_prior, iterations) {
+  brm(
+    TE | se(seTE) ~ 1 + (1 | Author),
+    data = dataset,
+    prior = defined_prior,
+    iter = iterations,
+    control = list(adapt_delta = 0.99)
+  )
+}
+
+post_samples <- function(dataset) {
+  p_samples <- posterior_samples(dataset, c("^b", "^sd")) %>%
+    rename("TE" = 1,
+           "tau" = 2)
+  print(ggplot(aes(x = TE), data = p_samples)+
+          geom_density(fill = "lightblue",
+                       color = "lightblue",
+                       alpha = 0.7) +
+          geom_point(y = 0, x = mean(p_samples$TE)) +
+          labs(x = expression(italic(TE)),
+               y = element_blank()) +
+          theme_minimal())
+  return(p_samples)
+}
+
+study_draw <- function(dataset) {
+  spread_draws(dataset, r_Author[Author,], b_Intercept) %>%
+    mutate(b_Intercept = r_Author + b_Intercept) %>%
+    left_join(., study_review_version)
+}
+
+pooled_effect_draw <- function(dataset) {
+  spread_draws(dataset, b_Intercept) %>%
+    mutate(Author = "Pooled Effect",
+           review_version = factor("combined_pooled"))
+}
+
+forest_plot <- function(model, data_study, data_pooled, cut, title, type, filename) {
+  forest_data <- bind_rows(data_study, data_pooled) %>%
+    ungroup() %>%
+    mutate(Author = str_replace_all(Author, "[.]", " ")) %>%
+    mutate(Author = reorder(Author, b_Intercept),
+           b_Intercept = exp(b_Intercept))
+  summary_forest <- group_by(forest_data, Author, review_version) %>%
+    median_qi(b_Intercept)
+  graph <- ggplot(aes(b_Intercept, relevel(Author, "Pooled Effect", after = Inf), fill = review_version),
+                  data = forest_data) +
+    geom_vline(xintercept = exp(fixef(model)[1, 1]),
+               color = "grey",
+               size = 1) +
+    geom_vline(xintercept = exp(fixef(model)[1, 3:4]),
+               color = "grey",
+               linetype = 2) +
[…]

### 2020-08-26 scripts/bayesian_analysis.R

@@ -1,0 +1,565 @@
+library(brms)
+library(ggplot2)
+library(tidybayes)
+library(readr)
+library(dplyr)
+library(ggridges)
+library(glue)
+library(stringr)
+library(forcats)
+library(here)
+library(meta)
+library(snakecase)
+
+source(here("scripts", "bayes_scripts.R"))
+source(here("scripts", "author_dictionary.R"))
+
+previous_review_versions <- c("v1", "v2", "v3", "v4", "v5", "v6")
+current_review_version <- "v7"
+
+data_study_general <- read_rds(here("data_clean", "data_study_general.rds"))
+
+study_review_version <- data_study_general %>%
+  mutate(lead_author = to_upper_camel_case(lead_author, sep_out = ", "),
+         lead_author = plyr::mapvalues(lead_author,
+                                 from = cleaned_names,
+                                 to = correct_names),
+         lead_author = recode(lead_author, "de Lusignan" = "de.Lusignan")) %>%
+  select(lead_author, review_version) %>%
+  rename(Author = lead_author) %>%
+  mutate(review_version = factor(fct_collapse(review_version, previous = previous_review_versions, current = current_review_version)))
+
+levels(study_review_version$review_version) <- c(levels(study_review_version$review_version), "combined_pooled") %>%
+  fct_relevel(c("previous", "current", "combined_pooled"))
+
+m1 <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1.rds"))
+m1_a <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1_a.rds"))
+m1_a_hh <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1_a_hh.rds"))
+m2 <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2.rds"))
+m2_a <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2_a.rds"))
+m2_a_hh <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2_a_hh.rds"))
[…]

### 2020-08-26 scripts/bayesian_data_generation.R

@@ -1,0 +1,298 @@
+library(brms)
+library(ggplot2)
+library(tidybayes)
+library(readr)
+library(dplyr)
+library(ggridges)
+library(glue)
+library(stringr)
+library(forcats)
+library(here)
+library(meta)
+library(snakecase)
+set.seed(42)
+
+source(here("scripts", "bayes_scripts.R"))
+source(here("scripts", "author_dictionary.R"))
+
+data_study_general <- readRDS(here("data_clean", "data_study_general.rds"))
+                              
+previous_review_versions <- c("v1", "v2", "v3", "v4", "v5", "v6")
+current_review_version <- "v7"
+
+study_review_version <- data_study_general %>%
+  mutate(lead_author = to_upper_camel_case(lead_author, sep_out = ", "),
+         lead_author = plyr::mapvalues(lead_author,
+                                       from = cleaned_names,
+                                       to = correct_names),
+         lead_author = recode(lead_author, "de Lusignan" = "de.Lusignan")) %>%
+  select(lead_author, review_version) %>%
+  rename(Author = lead_author) %>%
+  mutate(review_version = factor(fct_collapse(review_version, previous = previous_review_versions, current = current_review_version)))
+
+levels(study_review_version$review_version) <- c(levels(study_review_version$review_version), "combined_pooled") %>%
+  fct_relevel(c("previous", "current", "combined_pooled"))
+
+
+bayes_testing_current <-
+  read_rds(here("data_clean", "bayes_testing_current.rds"))
+bayes_testing_former <-
+  read_rds(here("data_clean", "bayes_testing_former.rds"))
+bayes_hospital_current <-
+  read_rds(here("data_clean", "bayes_hospital_current.rds"))
+bayes_hospital_former <-
+  read_rds(here("data_clean", "bayes_hospital_former.rds"))
+bayes_severity_current <-
+  read_rds(here("data_clean", "bayes_severity_current.rds"))
+bayes_severity_former <-
+  read_rds(here("data_clean", "bayes_severity_former.rds"))
[…]

### 2020-08-26 scripts/bootstrap_function.R

@@ -1,0 +1,20 @@
+boot.mean = function(x,B,binwidth=NULL) {
+  n = length(x)
+  boot.samples = matrix(sample(x,size=n*B,replace=TRUE), B, n)
+  boot.statistics = apply(boot.samples,1,mean)
+  se = sd(boot.statistics)
+  interval=quantile(boot.statistics,c(0.025,0.975))
+  return(list(boot.statistics=boot.statistics,interval=interval,se=se))
+}
+
+do_bootstrap_current <- function (row_number) {
+  my_sample <- c(rep(1, each = bootstrap$bs_current[row_number]), rep(0, each = bootstrap$bs_inverse[row_number]))
+  bootstrap_output_c[[row_number]] <- boot.mean(my_sample,1000)$interval
+  return(bootstrap_output_c)
+}
+
+do_bootstrap_former <- function (row_number) {
+  my_sample <- c(rep(1, each = bootstrap$bs_former[row_number]), rep(0, each = bootstrap$bs_inverse_former[row_number]))
+  bootstrap_output_f[[row_number]] <- boot.mean(my_sample,1000)$interval
+  return(bootstrap_output_f)
+}

### 2020-08-26 scripts/correlation_coefficient.R

@@ -1,0 +1,19 @@
+library('tidyverse')
+library('ggplot')
+
+table_1 <- read_csv("data_clean/data_table_1.csv")
+
+current_smokers <- table_1$current_smoker
+former_smokers <- table_1$former_smoker
+
+current_former_smokers <- coalesce(table_1$current_smoker, table_1$current_former_smoker, table_1$former_smoker)
+
+missing_smokers <- coalesce(table_1$not_stated, table_1$missing)
+
+cor.test(current_smokers, missing_smokers)
+cor.test(former_smokers, missing_smokers)
+cor.test(current_former_smokers, missing_smokers)
+
+
+ggplot(data = table_1, aes(x = current_former_smokers,y = missing_smokers))+
+  geom_point()

### 2020-08-26 scripts/data_cleaning.R

@@ -1,0 +1,65 @@
+library('tidyverse')
+library('dplyr')
+library('readxl')
+library('linelist')
+library('googlesheets4')
+
+sheets_id <- as_sheets_id('https://docs.google.com/spreadsheets/d/15avypGR8ypJngWQEmFIzFrOOwXPY3xezUHQU6jgV7d0/edit?usp=sharing')
+sheets_id
+
+search_details <- read_sheet(sheets_id, range = 'article_screening')
+  
+
+search_details <- search_details %>%
+  clean_data() %>%
+  rename(date_screened = date_screening_in_format_2020_05_22) %>%
+  write_rds(here::here('data_clean', 'search_details.rds')
+
+data_study_general <- read_sheet(sheets_id, range =  'general_details')
+protect_columns_1 <- names(data_study_general) %in% 'doi'
+
+data_study_general <- data_study_general %>%
+  rename('notes' = 11, date_published = `date_published_format(2020-05-12)`) %>%
+  clean_data(protect = protect_columns_1)
+write_rds(data_study_general, 'data_clean/data_study_general.rds')
+
+review_details <- data_study_general %>%
+  select(lead_author, date_published, country, review_version)
+write_rds(review_details, 'data_clean/review_details.rds')
+
+table_1 <- read_sheet(sheets_id, sheet = 'pop_descriptives') %>%
+  clean_data() %>%
+  mutate(., lower_range = sub('\\_.*', '', .$range)) %>%
+  mutate(., upper_range = sub('.*_', '',.$range,)) %>%
+  select(1:9, 22:23, 11:21) %>%
+  filter(!is.na(lead_author)) %>%
+  left_join(., review_details, by = 'lead_author') %>%
+  write_rds(., 'data_clean/data_table_1.rds')
+   
+table_2 <-  read_sheet(sheets_id, sheet = 'testing') %>%
+  clean_data() %>%
+  filter(data_on_testing == TRUE) %>%
+  left_join(., review_details, by = 'lead_author') %>%
+  write_rds(., 'data_clean/data_table_2.rds')
+
+table_3 <-  read_sheet(sheets_id, sheet = 'hospitalisation') %>%
+  clean_data() %>%
+  filter(data_on_hospitalisation == TRUE) %>%
+  left_join(., review_details, by = 'lead_author') %>%
+  write_rds(., 'data_clean/data_table_3.rds')
+
[…]

### 2020-08-26 scripts/descriptives.R

@@ -1,0 +1,306 @@
+##Run if needing to update the studies
+source('scripts/data_cleaning.R')
+##Run if data already available
+library(tidyverse)
+library(linelist)
+source('scripts/prisma_function.R')
+
+date_of_update <- Sys.Date()
+date_of_last_update <- date_of_update-10
+prev_versions <- c('v1', 'v2')
+current_version <- c('v3')
+analysed_versions <- c('v1', 'v2', 'v3')
+exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old')
+exclude_from_qa <- c('ISARIC_1', 'ISARIC_2', 'ISARIC_3', 'Miyara')
+
+search_details <- read_rds("data_clean/search_details.rds")
+
+data_study_general <- read_rds("data_clean/data_study_general.rds")
+data_study_general <- data_study_general %>%
+  filter(review_version %in% analysed_versions) %>%
+  filter(!lead_author %in% exclude_from_analysis) 
+
+table_1 <- read_rds("data_clean/data_table_1.rds")
+table_1 <- table_1 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!lead_author %in% exclude_from_analysis)
+
+table_2 <- read_rds("data_clean/data_table_2.rds")
+table_2 <- table_2 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!lead_author %in% exclude_from_analysis)
+
+table_3 <- read_rds("data_clean/data_table_3.rds")
+table_3 <- table_3 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!lead_author %in% exclude_from_analysis)
+
+table_4 <- read_rds("data_clean/data_table_4.rds")
+table_4 <- table_4 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!lead_author %in% exclude_from_analysis)
+
+table_5 <- read_rds("data_clean/data_table_5.rds")
+table_5 <- table_5 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!lead_author %in% exclude_from_analysis)
+
+table_6 <- table_6 %>%
+  filter(!Author %in% exclude_from_qa)
+  
+
+#Number screened to add to PRISMA
[…]

### 2020-08-26 scripts/mapping.R

@@ -1,0 +1,64 @@
+library(rgdal)
+library(leaflet)
+library(coronavirus)
+update_dataset()
+
+my_spdf <- readOGR( 
+  dsn= paste0(here::here("data_clean", "world_shape_file")), 
+  layer="TM_WORLD_BORDERS_SIMPL-0.3",
+  verbose=FALSE
+)
+
+from_countries <- c("USA", "UK", "Iran", "South Korea")
+to_countries <- c("United States", "United Kingdom", "Iran (Islamic Republic of)", "Korea, Republic of")
+a <- country %>%
+  mutate(Country = plyr::mapvalues(Country, from = from_countries,
+                to = to_countries))
+
+data("coronavirus")
+summary_df <- coronavirus %>% 
+  filter(type == "confirmed") %>%
+  group_by(country) %>%
+  summarise(total_cases = sum(cases)) %>%
+  arrange(-total_cases) %>%
+  rename("NAME" = country)
+
+my_spdf@data <- left_join(my_spdf@data, a %>%
+            rename(NAME = Country),
+          by = "NAME") %>%
+  left_join(., summary_df %>%
+              mutate(total_cases = total_cases/1000),
+            by = "NAME") %>%
+  mutate(NAME = as.factor(NAME))
+
+my_spdf@data$Number[ which(my_spdf@data$Number == 0)] = NA
+my_spdf@data$total_cases[ which(is.na(my_spdf@data$total_cases))] = 0
+
+
+my_bins <- c(0, 1, 5, 10, 20, 30, 40, 50)
+mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)
+
+mytext <- paste(
+  "Country: ", my_spdf@data$NAME,"<br/>", 
+  "Number of studies included: ", my_spdf@data$Number, "<br/>", 
+  "Number of COVID-19 reported cases (thousands): ", round(my_spdf@data$total_cases, 2), 
+  sep="") %>%
+  lapply(htmltools::HTML)
+
+html_map <- leaflet(my_spdf) %>% 
+  addTiles()  %>% 
+  setView(lat=10, lng=0 , zoom=2) %>%
+  addPolygons( 
+    fillColor = ~mypalette(Number), 
+    stroke=TRUE, 
+    fillOpacity = 0.9, 
+    color="white", 
+    weight=0.3,
+    label = mytext,
+    labelOptions = labelOptions( 
[…]

### 2020-08-26 scripts/meta_analyses.R

@@ -1,0 +1,19 @@
+library('dmetar')
+
+
+forest(m.hksj.raw,
+       sortvar=TE,
+       xlim = c(-1.5,0.5),
+       rightlabs = c("g","95% CI","weight"),
+       leftlabs = c("Author", "N","Mean","SD","N","Mean","SD"),
+       lab.e = "Intervention",
+       pooled.totals = FALSE,
+       smlab = "",
+       text.random = "Overall effect",
+       print.tau2 = FALSE,
+       col.diamond = "blue",
+       col.diamond.lines = "black",
+       col.predict = "black",
+       print.I2.ci = TRUE,
+       digits.sd = 2
+)

### 2020-08-26 scripts/olgas_script.R

@@ -1,0 +1,129 @@
+library(tidyverse)
+library(here)
+library(linelist)
+library(boot)
+library(googlesheets4)
+
+source(here("scripts", "bootstrap_function.R"))
+
+set.seed(42)
+smoking_status <- c('current_smoker', 'former_smoker', 'current_former_smoker', 'never_smoker', 'never_smoker_unknown', 'not_stated', 'missing')
+table_1 <- read_rds(here("data_clean", "table_1.rds"))
+
+prevalence_plot <- table_1 %>%
+  left_join(., data_study_general %>%
+              select(lead_author, study_setting), 
+            by = "lead_author") %>%
+  group_by(country, study_setting) %>%
+  rename('sample' = total) %>%
+  filter(current_smoker != 'NA') %>%
+  filter(study_setting != is.na(study_setting)) %>%
+  replace_na(., list(current_smoker = 0,
+                     former_smoker = 0,
+                     current_former_smoker = 0,
+                     never_smoker = 0,
+                     never_smoker_unknown = 0,
+                     not_stated = 0,
+                     missing = 0)) %>%
+  mutate(ever_smoker = former_smoker + current_smoker + current_former_smoker,
+         not_stated_missing = missing + not_stated + never_smoker_unknown) %>%
+  select(lead_author, country, sample, current_smoker, former_smoker, never_smoker, ever_smoker, not_stated_missing) %>%
+  mutate(total = never_smoker + ever_smoker + not_stated_missing) %>%
+  mutate(p_current_smoker = current_smoker/total,
+         p_ever_smoker = ever_smoker/total,
+         p_former_smoker = former_smoker/total,
+         p_never_smoker = never_smoker/total,
+         p_ever_smoker = ever_smoker/total,
+         p_not_stated_missing = not_stated_missing/total,
+         p_total = p_ever_smoker + p_never_smoker + p_not_stated_missing) %>%
+  mutate(study = 1) %>%
+  add_count(country) 
+
+
+prevalence_plot$country <- as.factor(prevalence_plot$country)
+
+google_smoking_prevalence <-  read_sheet(sheets_id, sheet = 'national_smoking_prevalence') 
+national_smoking_prevalence <- google_smoking_prevalence %>%
+  select(-Source) %>%
+  mutate(study = 0,
+         Current = Current/100,
[…]

### 2020-08-26 scripts/prisma_function.R

@@ -1,0 +1,34 @@
+#PRISMA function
+
+PRISMA <- function(search_details) {
+PRISMA <- search_details %>%
+    filter(review_version == current_version)
+  OVID_records <- PRISMA[1,2]
+  medRxiv_records <- PRISMA[1,6]
+  other_records <- PRISMA[1,9]
+  OVID_screened <- PRISMA[1,3]
+  medRxiv_screened <- PRISMA[1,7]
+  OVID_included <- PRISMA[1,5]
+  medRxiv_included <- PRISMA[1,8]
+  others_included <- PRISMA[1,9]
+  
+  top_row <- as_vector(c(OVID_records, medRxiv_records, other_records))
+  total_title_abstracts <- tibble(sum(top_row))
+  colnames(total_title_abstracts) <- 'titles_screened'
+  abstract_screened <- tibble(sum(OVID_screened, medRxiv_screened, other_records))
+  colnames(abstract_screened) <- 'abstract_screened'
+  excluded_abstract <- tibble(sum(top_row) - sum(abstract_screened))
+  colnames(excluded_abstract) <- 'excluded_abstract'
+  full_texts <- tibble(sum(top_row) - sum(excluded_abstract))
+  colnames(full_texts) <- 'full_texts_reviewed'
+  included_full_text <- as_vector(c(OVID_included, medRxiv_included, others_included))
+  excluded_full_text <- tibble(full_texts - sum(included_full_text))
+  colnames(excluded_full_text) <- 'full_texts_excluded'
+  narrative_synthesis_v <- tibble(sum(included_full_text))
+  colnames(narrative_synthesis_v) <- 'full_texts_included_narrative'
+  narrative_synthesis_total <- tibble(sum(from_prev_version + narrative_synthesis_v))
+  colnames(narrative_synthesis_total) <- 'total_all_versions'
+  prisma_list <- c(top_row, total_title_abstracts, excluded_abstract, full_texts, excluded_full_text, narrative_synthesis_v, narrative_synthesis_total)
+  return(prisma_list)
+}
+

### 2020-08-26 scripts/prisma_v2.R

@@ -1,0 +1,34 @@
+current_total <- search_details %>%
+  select(total_included) %>%
+  sum()
+
+previous_total <- search_details %>%
+  filter(review_version %in% previous_review_versions) %>%
+  select(total_included) %>%
+  sum()
+
+a <- search_details %>%
+  group_by(review_version) %>%
+  filter(review_version == current_review_version) %>%
+  mutate(title_abstract = sum(ovid_number_results, medrxiv_number_results, other_source_included),
+         full_texts_assessed = sum(ovid_screened, medrxiv_screened, other_source_included),
+         title_abstract_included = sum(ovid_included, medrxiv_included, other_source_included),
+         title_abstract_excluded = title_abstract - full_texts_assessed,
+         full_text_excluded = full_texts_assessed-total_included,
+         previous_total = previous_total,
+         previous_minus_superseded = previous_total-superseded,
+         current_total = current_total) %>%
+  select("Review version" = review_version,
+         "Date Screened" = date_screened,
+         "OVID results" = "ovid_number_results",
+         "medRxiv results" = "medrxiv_number_results",
+         "Other sources" = "other_source_included",
+         "Title/abstracts screened" = "title_abstract",
+         "Title/abstracts excluded" = "title_abstract_excluded",
+         "Full texts assessed" = "full_texts_assessed",
+         "Full texts excluded" = "full_text_excluded",
+         "Records carried forward" = "previous_minus_superseded",
+         "Supersed from previous version" = superseded,
+         "Studies in narrative synthesis" = "current_total")
+
+flextable(a)

### 2020-08-26 scripts/results_meta-analysis_death.R

@@ -1,0 +1,37 @@
+library(fmsb)
+library(epitools)
+library(meta)
+library(tidyverse)
+
+table_5 <- read_csv('data_clean/table_5_word.csv')
+included_studies <- c('gaibazzi')
+
+alive_smoker <- 11
+dead_smoker <- 10
+alive_never_smoker <- 244
+dead_never_smoker <- 132
+alive_former_smoker <- 30
+dead_former_smoker <- 14
+
+row_1 <- c(alive_never_smoker, dead_never_smoker)
+row_2 <- c(alive_smoker, dead_smoker)
+
+gaibazzi_table_1 <- rbind(row_1, row_2)
+rownames(gaibazzi_table_1) <- c('never smokers', 'current smokers')
+colnames(gaibazzi_table_1) <- c('alive', 'dead')
+
+gaibazzi_1 <- riskratio(x = gaibazzi_table_1, method = 'wald', conf.level=0.95)
+
+row_1 <- c(alive_never_smoker, dead_never_smoker)
+row_2 <- c(alive_former_smoker, dead_former_smoker)
+
+gaibazzi_table_2 <- rbind(row_1, row_2)
+rownames(gaibazzi_table_2) <- c('never smokers', 'former smokers')
+colnames(gaibazzi_table_2) <- c('alive', 'dead')
+
+gaibazzi_2 <- riskratio(x = gaibazzi_table_2, method = 'wald', conf.level=0.95)
+
+
+table_1 <- table_1 %>%
+  
+  mutate(., missing_1 = not_stated+missing/total)

### 2020-08-26 scripts/results_meta-analysis_testing.R

@@ -1,0 +1,376 @@
+library(epitools)
+library(meta)
+library(tidyverse)
+source('scripts/rr_function.R')
+
+table_2 <- read_rds('data_clean/table_2_word.rds')
+included_studies <- c('cho', 'de_lusignan', 'rentsch', 'kolin', 'shah')
+
+#And Nidzweidz
+
+#################RR from raw data#####################
+
+meta <- tibble('author' = table_2$lead_author,
+               'negative_smoker' = table_2$negative_current_smoker,
+               'negative_never_smoker' = table_2$negative_never_smoker,
+               'positive_smoker' = table_2$positive_current_smoker,
+               'positive_never_smoker' = table_2$positive_never_smoker,
+               'negative_former_smoker' = table_2$negative_former_smoker,
+               'positive_former_smoker' = table_2$positive_former_smoker) %>%
+        filter(author %in% included_studies)
+
+meta$author <- recode(meta$author, 'rentsch' = 'Rentsch',
+                      'cho' = 'Cho',
+                      'shah' = 'Shah',
+                      'kolin' = 'Kolin',
+                      'de_lusignan' = 'de Lusignan')
+
+a <- RR_testing('Rentsch', 'current')
+b <- RR_testing('Rentsch', 'former')
+c <- RR_testing('Cho', 'current')
+d <- RR_testing('Cho', 'former')
+e <- RR_testing('Shah', 'current')
+f <- RR_testing('Shah', 'former')
+g <- RR_testing('Kolin', 'current')
+h <- RR_testing('Kolin', 'former')
+i <- RR_testing('de Lusignan', 'current')
+j <- RR_testing('de Lusignan', 'former')
+data <- list(a,b,c,d,e,f,g,h,i,j)
+k <- do.call(rbind.data.frame, data)
+# Rentsch -----------------------------------------------------------------
+#current vs. never smokers
+
+negative_test_smoker <- 1444
+positive_test_smoker <- 159
+negative_test_never_smoker <- 826
+positive_test_never_smoker <- 216
+
+row_1 <- c(negative_test_never_smoker, positive_test_never_smoker)
+row_2 <- c(negative_test_smoker, positive_test_smoker)
+
[…]

### 2020-08-26 scripts/rr_function.R

@@ -1,0 +1,38 @@
+RR_testing <- function(study, smoking_RR){
+  if(smoking_RR == 'current'){
+  a <- filter(meta, author == study)
+  b <- a %>%
+    select(negative_never_smoker, positive_never_smoker) %>%
+    rename('negative test' = negative_never_smoker, 'positive test' = positive_never_smoker)
+  c <- a %>%
+    select(negative_smoker, positive_smoker) %>%
+    rename('negative test' = negative_smoker, 'positive test' = positive_smoker)
+  a <- as.matrix(rbind(b, c))
+  rownames(a) <- c('never smokers', 'current smokers')
+  } else {
+  a <- filter(meta, author == study)
+  b <- a %>%
+    select(negative_never_smoker, positive_never_smoker) %>%
+    rename('negative test' = negative_never_smoker, 'positive test' = positive_never_smoker)
+  c <- a %>%
+    select(negative_former_smoker, positive_former_smoker) %>%
+    rename('negative test' = negative_former_smoker, 'positive test' = positive_former_smoker)
+  a <- as.matrix(rbind(b, c))
+  rownames(a) <- c('never smokers', 'former smokers')
+  }
+  
+  a <- riskratio(a, method = 'wald', conf.level=0.95)
+  
+  log_RR <- log(a[["measure"]][2,1])
+  lower_ci <- log(a[["measure"]][2,2])
+  upper_ci <- log(a[["measure"]][2,3])
+  log_SE <- (upper_ci-lower_ci)/3.92
+
+  output_table <- list('study' = study,
+                       'smoking_status' = smoking_RR,
+                       'log_RR' = log_RR,
+                       'lower_ci' = lower_ci, 
+                       'upper_ci' = upper_ci,
+                       'log_SE' = log_SE)
+  return(as_tibble(output_table))
+}

### 2020-08-26 scripts/smoking_plots.R

@@ -1,0 +1,137 @@
+## packages
+library(tidyverse)
+library(ggsci)
+library(matrixStats)
+library(gapminder)
+
+## set seed
+set.seed(123)
+
+## load fonts
+extrafont::loadfonts(device = "win")
+
+## get data
+data <-  read_rds(here::here('data_clean', 'country_prevalence_data.rds')) %>%
+  mutate(country = recode(country, "SouthKorea" = "South Korea"))
+country_colour <- data.frame(gapminder, cc = I(country_colors[match(gapminder$country, names(country_colors))]))
+country_colour <- country_colour %>%
+  select(country, cc) %>%
+  mutate(country = recode(country,
+         "United Kingdom" = "UK",
+         "United States" = "USA",
+         "Korea, Dem. Rep." = "South Korea")) %>%
+  rename("country_colour" = cc) %>%
+  distinct() %>%
+  filter(country %in% data$country) %>%
+  deframe()
+
+## sort data
+sorted_data <- data %>%
+  ungroup() %>%
+  rename("number_studies" = n) %>%
+  mutate(country = fct_reorder(country, -number_studies), ## sorting countries by number of studies from each country
+         smoking = fct_relevel(smoking, "former_smoking_p", after = Inf),
+         study_setting = recode(study_setting, "quarantine_centre" = "community"),
+         study_setting = forcats::fct_explicit_na(study_setting)) %>%
+  group_by(smoking, country) %>%
+  mutate(running_count = row_number()) %>%
+  group_by(smoking, country, study, study_setting) %>%
+  mutate(mean = ifelse(is.na(weightedMean(prevalence, w = true_sample)), mean(prevalence),
+                         weightedMean(prevalence, w = true_sample, na.rm = T))) %>%
+  ungroup() %>%
+  add_count(smoking, country) %>%
+  group_by(study, study_setting)
+
+sorted_data$running_count <- as_factor(sorted_data$running_count)  
+
+current_sorted_data <- sorted_data %>%
+  filter(., smoking == "current_smoking_p") %>%
+  arrange(-prevalence, .by_group = T) %>%
+  ungroup() %>%
+  group_by(country) %>%
+  filter(!n <= 1)
+
+a <- ggplot(data = filter(current_sorted_data, study == 1), aes(x = running_count, y = prevalence)) +
+  geom_point(aes(size = true_sample, shape = factor(study_setting))) + 
[…]

### 2020-08-26 reports/bayesian_report.Rmd
v7

@@
          previous_minus_superseded = previous_total-superseded,
-         current_total = current_total) %>%
+         current_total = current_total-superseded) %>%
   select("Review version" = review_version,
@@
 country <- subset(country, Country != "Multiple")
-
+```
+Currently studies are reported from `r paste(nrow(country))`
+``` {r}
 country <- mutate_geocode(country, Country)
@@
 ```
-Currently studies are reported from `r paste(length(unique(country)))`
-
 ## A map is produced for where the studies are performed

### 2020-08-26 reports/bayesian_report.Rmd
v7 final draft

@@
 library(dplyr)
+library(tidyr)
 library(ggridges)
@@
 library(DT)
-library(tidyr)
 library(epitools)
+library(ggmap)
 ```

### 2020-09-02 reports/bayesian_report.Rmd
updated for v7 final

@@
 ``` {r excluding_studies}
-exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old', 'isaric_4', 'mehra', 'hopkinson', 'gaibazzi', 'miyara_updated', 'russell_old', "petrilli_old", "magagnoli_old", "niedzwiedz_old", "bello_chavolla_old", "kimmig_old", "russell_old", "zuo_zuo_old", "patel_old", "shi_zuo_old")
+exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old', 'isaric_4', 'mehra', 'hopkinson', 'gaibazzi', 'miyara_updated', 'russell_old', "petrilli_old", "magagnoli_old", "niedzwiedz_old", "bello_chavolla_old", "kimmig_old", "russell_old", "zuo_zuo_old", "patel_old", "shi_zuo_old", "riley_old")
 ```
@@
   select(lead_author) %>%
-  filter(lead_author != "reiter")
+  filter(!lead_author %in% c("reiter", "hippisley_cox"))
 

### 2020-09-02 scripts/bayesian_analysis.R

@@ -72,13 +72,16 @@ m1_ecdf(0.9) #minimally informative
 m1_a_ecdf(0.9) #informative
 m1_a_hh_ecdf(0.9) #high heterogeneity
 
-study_draws_m1 <- study_draw(m1)
+study_draws_m1 <- study_draw(m1) %>%
+  mutate(Author = ifelse(Author == "Merkely", "Merkely*", Author))
 pooled_draw_m1 <- pooled_effect_draw(m1)
 
-study_draws_m1_a <- study_draw(m1_a)
+study_draws_m1_a <- study_draw(m1_a) %>%
+  mutate(Author = ifelse(Author == "Merkely", "Merkely*", Author))
 pooled_draw_m1_a <- pooled_effect_draw(m1_a)
 
-study_draws_m1_a_hh <- study_draw(m1_a_hh)
+study_draws_m1_a_hh <- study_draw(m1_a_hh) %>%
+  mutate(Author = ifelse(Author == "Merkely", "Merkely*", Author))
 pooled_draw_m1_a_hh <- pooled_effect_draw(m1_a_hh)
 
 current_testing_minimal_prior <- forest_plot(
@@ -117,6 +120,8 @@ current_testing_alternative_prior_hh <- forest_plot(
   "m1_a_hh.png"
 )
 
+median_hdci(post_samples_m1_a$tau)
+
 # Former smokers testing --------------------------------------------------
 # Bayesian analysis for former smokers and testing for SARS-CoV-2
 
@@ -133,13 +138,16 @@ m2_a_hh_ecdf <- ecdf(exp(post_samples_m2_a_hh$TE))
 1-m2_a_ecdf(1.1) #informative
 1-m2_a_hh_ecdf(1.1) #high heterogeneity
 
-study_draws_m2 <- study_draw(m2)
+study_draws_m2 <- study_draw(m2) %>%
+  mutate(Author = ifelse(Author == "Merkely", "Merkely*", Author))
 pooled_draw_m2 <- pooled_effect_draw(m2)
 
-study_draws_m2_a <- study_draw(m2_a)
+study_draws_m2_a <- study_draw(m2_a) %>%
+  mutate(Author = ifelse(Author == "Merkely", "Merkely*", Author))
 pooled_draw_m2_a <- pooled_effect_draw(m2_a)
 
-study_draws_m2_a_hh <- study_draw(m2_a_hh)
[…]

### 2020-09-24 reports/bayesian_report.Rmd
preparing v8

--- 2020-08-26 reports/bayesian_report.Rmd
@@
 ## Download data
-We maintain a Google sheets document [here](sheets_id) which contains all eligible articles the research team have identified.
-This is currently viewable only and not open for editing.
-We have collected the general information for the search in the first sheet, the studies general information in the second, further information on study populations, outcomes of interest and quality of the studies are contained in the further sheets.
-These sheets are organised in a similar way to the manuscript and so are labelled table_1 to table_6.
-
 ``` {r google_data}
-sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1HV2Zk_UD2_s1zzd47dCyb4UKRSM0dDg7zmUPA9CMdx4/edit?usp=sharing")
+sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1sTlqMbGiWDWJT3dTpa3YQU4z6qcWb0v6Wkp40RcWLXQ/edit?usp=sharing")
 search_details <- read_sheet(sheets_id, range = 'article_screening')
@@
 ```
+We maintain a Google sheets document [here](sheets_id) which contains all eligible articles the research team have identified.
+This is currently viewable only and not open for editing.
+We have collected the general information for the search in the first sheet, the studies general information in the second, further information on study populations, outcomes of interest and quality of the studies are contained in the further sheets.
+These sheets are organised in a similar way to the manuscript and so are labelled table_1 to table_6.
+
+The pull requests from the API may happen too rapidly if this is run as a chunk so it's better to run this line by line.
+
 We have extracted data from `r nrow(data_study_general)` studies, however, some of these are duplicates and some have been removed from the analysis set.
@@
 ```{r head_search_details}
-flextable(search_details[,1:11])
+flextable(search_details[,1:12])
 ```
@@
 
-prev_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6')
+prev_versions <- head(unique(data_study_general$review_version), n = -1)
 #This can categorise which studies we want to look at
-current_version <- c('v7')
+current_version <- tail(unique(data_study_general$review_version), n = 1)
 #This will categorise which studies we are including in the current report
 
-analysed_versions <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6', 'v7')
+analysed_versions <- c(prev_versions, current_version)
 #This will incorporate all studies into the current version of the report
 ```
-We define the current version of the report here. At the moment this needs to be changed directly when compiling the code. This report has been generated for version `r paste(current_version)`
+We define the current version of the report here. This report has been generated for version `r paste(current_version)`
 
 ``` {r excluding_studies}
-exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old', 'isaric_4', 'mehra', 'hopkinson', 'gaibazzi', 'miyara_updated', 'russell_old', "petrilli_old", "magagnoli_old", "niedzwiedz_old", "bello_chavolla_old", "kimmig_old", "russell_old", "zuo_zuo_old", "patel_old", "shi_zuo_old")
+superseded_studies <- names(tidyselect::vars_select(data_study_general$lead_author, ends_with("_old", ignore.case = T)))
+old_isaric <- head(names(tidyselect::vars_select(data_study_general$lead_author, starts_with("isaric", ignore.case = T))), n = -1)
+excluded_other_reason <- c("mehra", #retracted
+                           "hopkinson", #self-reported
+                           "gaibazzi" #withdrawn
+                           )
+
+exclude_from_analysis <- c(superseded_studies, old_isaric, excluded_other_reason)
 ```
@@
 
-my_bins <- c(0, 1, 5, 10, 20, 30, 40, 70)
+my_bins <- c(0, 1, 5, 10, 20, 30, 40, 80)
 mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)
@@
 country_prevalence_list <- table_1 %>%
-    group_by(country, add = T) %>%
+    group_by(country, .add = T) %>%
   mutate(., current_smok_percentage = current_smoker/total*100) %>%
@@
 
-save_as_docx(a, path = here('output_data', 'Table_1.docx'))
+#save_as_docx(a, path = here('output_data', 'Table_1.docx'))
 ```
@@
   select(lead_author) %>%
-  filter(lead_author != "reiter")
+  filter(!lead_author %in% c("reiter", "hippisley_cox", "ullah"))
 
@@
                                to = correct_names)
-
-testing_rr <- list()
-testing_rr[[1]] <- RR_testing('Rentsch', 'current')
-testing_rr[[2]] <- RR_testing('Rentsch', 'former')
-testing_rr[[3]] <- RR_testing('Cho', 'current')
-testing_rr[[4]] <- RR_testing('Cho', 'former')
-testing_rr[[5]] <- RR_testing('Kolin', 'current')
-testing_rr[[6]] <- RR_testing('Kolin', 'former')
-testing_rr[[7]] <- RR_testing('de Lusignan', 'current')
-testing_rr[[8]] <- RR_testing('de Lusignan', 'former')
-testing_rr[[9]] <- RR_testing('Parrotta', 'current')
-testing_rr[[10]] <- RR_testing('Parrotta', 'former')
-testing_rr[[11]] <- RR_testing('Israel', 'current')
-testing_rr[[12]] <- RR_testing('Israel', 'former')
-testing_rr[[13]] <- RR_testing('Eugen-Olsen', 'current')
-testing_rr[[14]] <- RR_testing('Eugen-Olsen', 'former')
-testing_rr[[15]] <- RR_testing('Houlihan', 'current')
-testing_rr[[16]] <- RR_testing('Houlihan', 'former')
-testing_rr[[17]] <- RR_testing('Woolford', 'current')
-testing_rr[[18]] <- RR_testing('Woolford', 'former')
-testing_rr[[19]] <- RR_testing('Govind', 'current')
-testing_rr[[20]] <- RR_testing('Govind', 'former')
-testing_rr[[21]] <- RR_testing('Gu', 'current')
-testing_rr[[22]] <- RR_testing('Gu', 'former')
-testing_rr[[23]] <- RR_testing('Petrilli', 'current')
-testing_rr[[24]] <- RR_testing('Petrilli', 'former')
-testing_rr[[25]] <- RR_testing('Auvinen', 'current')
-testing_rr[[26]] <- RR_testing('Auvinen', 'former')
-testing_rr[[27]] <- RR_testing('Merkely', 'current')
-testing_rr[[28]] <- RR_testing('Merkely', 'former')
-testing_rr[[29]] <- RR_testing('Iversen', 'current')
-testing_rr[[30]] <- RR_testing('Iversen', 'former')
-testing_rr[[31]] <- RR_testing('Fillmore', 'current')
-testing_rr[[32]] <- RR_testing('Fillmore', 'former')
-
-data <- testing_rr
+study_dup <- filter(meta, author != "Niedzwiedz")
+study_dup <- study_dup$author
+
+testing_rr_current <- list()
+for(study in 1:length(study_dup)){
+  testing_rr_current[[study]] <- RR_testing(study_dup[study], "current")
+}
+testing_rr_former <- list()
+for(study in 1:length(study_dup)){
+  testing_rr_former[[study]] <- RR_testing(study_dup[study], "former")
+}
+
+data <- c(testing_rr_current, testing_rr_former)
 k <- do.call(rbind.data.frame, data)
@@
                'severe_former_smoker' = table_4$severe_disease_former_smoker) %>%
-  filter(author %in% included_studies$lead_author)
+  filter(author %in% included_studies$lead_author) %>%
+  filter(author != "hippisley_cox")
 

### 2020-09-24 scripts/author_dictionary.R

@@ -26,7 +26,8 @@ cleaned_names <- c('Chow, Us, Cdc',
                    "De, Melo",
                    "De, Souza",
                    "Bello, Chavolla, Antonio, Villa",
-                   "Antonio, Villa")
+                   "Antonio, Villa",
+                   "Rubio, Rivas")
 
 correct_names <- c('Chow (US CDC)',
                    "Gold (US CDC)",
@@ -56,4 +57,5 @@ correct_names <- c('Chow (US CDC)',
                    "de Melo",
                    "de Souza",
                    "Bello-Chavolla, Antonio-Villa",
-                   "Antonio-Villa")
+                   "Antonio-Villa",
+                   "Rubio-Rivas")

### 2020-09-24 scripts/bayes_scripts.R

@@ -85,7 +85,7 @@ forest_plot <- function(model, data_study, data_pooled, cut, title, type, filena
          title = title,
          caption = type
          ) +
-    scale_fill_discrete(name = "Review version", labels = c("Previous version (v6)", "Current version (v7)", "Pooled effect")) +
+    scale_fill_discrete(name = "Review version", labels = c("Previous version (v7)", "Current version (v8)", "Pooled effect")) +
     xlim(0, cut) +
     theme_minimal() +
     theme(panel.spacing = unit(0.1, "lines"),

### 2020-09-24 scripts/bayesian_analysis.R

@@ -14,8 +14,8 @@ library(snakecase)
 source(here("scripts", "bayes_scripts.R"))
 source(here("scripts", "author_dictionary.R"))
 
-previous_review_versions <- c("v1", "v2", "v3", "v4", "v5", "v6")
-current_review_version <- "v7"
+previous_review_versions <- prev_versions
+current_review_version <- current_version
 
 data_study_general <- read_rds(here("data_clean", "data_study_general.rds"))
 
@@ -73,15 +73,18 @@ m1_a_ecdf(0.9) #informative
 m1_a_hh_ecdf(0.9) #high heterogeneity
 
 study_draws_m1 <- study_draw(m1) %>%
-  mutate(Author = ifelse(Author == "Merkely", "Merkely*", Author))
+  mutate(Author = ifelse(Author == "Merkely", "Merkely*",
+                         ifelse(Author == "Carrat", "Carrat*", Author)))
 pooled_draw_m1 <- pooled_effect_draw(m1)
 
 study_draws_m1_a <- study_draw(m1_a) %>%
-  mutate(Author = ifelse(Author == "Merkely", "Merkely*", Author))
+  mutate(Author = ifelse(Author == "Merkely", "Merkely*",
+                         ifelse(Author == "Carrat", "Carrat*", Author)))
 pooled_draw_m1_a <- pooled_effect_draw(m1_a)
 
 study_draws_m1_a_hh <- study_draw(m1_a_hh) %>%
-  mutate(Author = ifelse(Author == "Merkely", "Merkely*", Author))
+  mutate(Author = ifelse(Author == "Merkely", "Merkely*",
+                         ifelse(Author == "Carrat", "Carrat*", Author)))
 pooled_draw_m1_a_hh <- pooled_effect_draw(m1_a_hh)
 
 current_testing_minimal_prior <- forest_plot(
@@ -100,7 +103,7 @@ current_testing_alternative_prior <- forest_plot(
   pooled_draw_m1_a,
   cut = 4,
   "Forest plot of current smokers and risk of testing positive",
-  "v6 testing prior",
+  "v7 testing prior",
   "m1_a.png"
 )
 
@@ -139,15 +142,18 @@ m2_a_hh_ecdf <- ecdf(exp(post_samples_m2_a_hh$TE))
 1-m2_a_hh_ecdf(1.1) #high heterogeneity
 
 study_draws_m2 <- study_draw(m2) %>%
[…]

### 2020-09-24 scripts/bayesian_data_generation.R

@@ -17,8 +17,8 @@ source(here("scripts", "author_dictionary.R"))
 
 data_study_general <- readRDS(here("data_clean", "data_study_general.rds"))
                               
-previous_review_versions <- c("v1", "v2", "v3", "v4", "v5", "v6")
-current_review_version <- "v7"
+previous_review_versions <- prev_versions
+current_review_version <- current_version
 
 study_review_version <- data_study_general %>%
   mutate(lead_author = to_upper_camel_case(lead_author, sep_out = ", "),
@@ -66,28 +66,28 @@ minimally_informative_prior <-
 
 # Previous meta-analysis guided priors
 current_testing_prior <-
-  c(prior(normal(-.30, 1), class = Intercept),
+  c(prior(normal(-.301, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_testing_prior <-
-  c(prior(normal(0.06, 1), class = Intercept),
+  c(prior(normal(0.049, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_hospitalisation_prior <-
-  c(prior(normal(0.05, 1), class = Intercept),
+  c(prior(normal(0.058, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_hospitalisation_prior <-
-  c(prior(normal(0.18, 1), class = Intercept),
+  c(prior(normal(0.182, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_severity_prior <-
-  c(prior(normal(0.14, 1), class = Intercept),
+  c(prior(normal(0.223, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_severity_prior <-
-  c(prior(normal(0.41, 1), class = Intercept),
+  c(prior(normal(0.419, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_mortality_prior <-
-  c(prior(normal(0.64, 1), class = Intercept),
+  c(prior(normal(0.199, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_mortality_prior <- 
-  c(prior(normal(0.65, 1), class = Intercept),
+  c(prior(normal(0.329, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 
[…]

### 2020-10-01 reports/bayesian_report.Rmd
update for v8

@@
 #This will incorporate all studies into the current version of the report
+
+versioning <- cbind(prev_versions, current_version, analysed_versions)
+write_rds(versioning, here("data_clean", "versioning.rds"))
 ```
@@
                            "hopkinson", #self-reported
-                           "gaibazzi" #withdrawn
+                           "gaibazzi", #withdrawn
+                           "kua" #self-reported
                            )
@@
 
+## Classification of COVID
+``` {r diagnosis_method}
+table(data_study_general$diagnostic_approach)
+```
+
 # Smoking descriptive data {.tabset .tabset-fade .tabset-pills}
@@
 The weighted mean current smoking prevalence is displayed for each setting.
-![Country smoking prevalence](figure/figure_2a.png)
+![Country smoking prevalence](figure/current_smoking_plots_updated.png)
 
@@
 The weighted mean current smoking prevalence is displayed for each setting.
-![Country former smoking prevalence](figure/figure_2b.png)
+![Country former smoking prevalence](figure/former_smoking_plots_updated.png)
 
@@
 
-#save_as_docx(a, path = here('output_data', 'Table_1.docx'))
+save_as_docx(a, path = here('output_data', 'Table_1.docx'))
 ```

### 2020-10-01 scripts/bayes_scripts.R

@@ -1,3 +1,5 @@
+library(tidyverse)
+
 extract_TE <- function(dataset) {
   tibble(dataset$studlab,
          dataset$TE,

### 2020-10-01 scripts/bayesian_analysis.R

@@ -13,9 +13,10 @@ library(snakecase)
 
 source(here("scripts", "bayes_scripts.R"))
 source(here("scripts", "author_dictionary.R"))
+versions <- read_rds(here("data_clean", "versioning.rds"))
 
-previous_review_versions <- prev_versions
-current_review_version <- current_version
+previous_review_versions <- unique(versions[,1])
+current_review_version <- unique(versions[,2])
 
 data_study_general <- read_rds(here("data_clean", "data_study_general.rds"))
 
@@ -192,6 +193,7 @@ former_testing_alternative_prior_hh <- forest_plot(
   "m2_a_hh.png"
 )
 
+median_hdci(post_samples_m2_a$tau)
 
 # Current smokers hospitalisation --------------------------------------------------
 # Bayesian analysis for current smokers and hospitalisation for SARS-CoV-2

### 2020-10-01 scripts/bayesian_data_generation.R

@@ -139,6 +139,10 @@ pp_check(m1_a_hh)
 summary(m1_a_hh)
 ranef(m1_a_hh)
 
+write_rds(m1, here("data_clean", "bayesian_models", "testing_current_m1.rds"))
+write_rds(m1_a, here("data_clean", "bayesian_models", "testing_current_m1_a.rds"))
+write_rds(m1_a_hh, here("data_clean", "bayesian_models", "testing_current_m1_a_hh.rds"))
+
 ##Testing former
 m2 <-
   bayes_test(testing_bayes_f, minimally_informative_prior, iterations = 40000)
@@ -158,6 +162,9 @@ pp_check(m2_a_hh)
 summary(m2_a_hh)
 ranef(m2_a_hh)
 
+write_rds(m2, here("data_clean", "bayesian_models", "testing_former_m2.rds"))
+write_rds(m2_a, here("data_clean", "bayesian_models", "testing_former_m2_a.rds"))
+write_rds(m2_a_hh, here("data_clean", "bayesian_models", "testing_former_m2_a_hh.rds"))
 ##Hospitalisation current
 m3 <-
   bayes_test(hospital_bayes_c, minimally_informative_prior, iterations = 40000)
@@ -272,12 +279,7 @@ pp_check(m8_a_hh)
 summary(m8_a_hh)
 ranef(m8_a_hh)
 
-write_rds(m1, here("data_clean", "bayesian_models", "testing_current_m1.rds"))
-write_rds(m1_a, here("data_clean", "bayesian_models", "testing_current_m1_a.rds"))
-write_rds(m1_a_hh, here("data_clean", "bayesian_models", "testing_current_m1_a_hh.rds"))
-write_rds(m2, here("data_clean", "bayesian_models", "testing_former_m2.rds"))
-write_rds(m2_a, here("data_clean", "bayesian_models", "testing_former_m2_a.rds"))
-write_rds(m2_a_hh, here("data_clean", "bayesian_models", "testing_former_m2_a_hh.rds"))
+
 write_rds(m3, here("data_clean", "bayesian_models", "hospitalisation_current_m3.rds"))
 write_rds(m3_a, here("data_clean", "bayesian_models", "hospitalisation_current_m3_a.rds"))
 write_rds(m3_a_hh, here("data_clean", "bayesian_models", "hospitalisation_current_m3_a_hh.rds"))

### 2020-10-01 scripts/heterogeneity.R

@@ -1,0 +1,37 @@
+library(tidybayes)
+library(brms)
+library(here)
+source(here("scripts", "bayes_scripts.R"))
+source(here("scripts", "author_dictionary.R"))
+
+m1_a <- readRDS(here("data_clean", "bayesian_models", "testing_current_m1_a.rds"))
+post_samples_m1_a <- post_samples(m1_a)
+median_hdci(post_samples_m1_a$tau)
+
+m2_a <- readRDS(here("data_clean", "bayesian_models", "testing_former_m2_a.rds"))
+post_samples_m2_a <- post_samples(m2_a)
+median_hdci(post_samples_m2_a$tau)
+
+m3_a <- readRDS(here("data_clean", "bayesian_models", "hospitalisation_current_m3_a.rds"))
+post_samples_m3_a <- post_samples(m3_a)
+median_hdci(post_samples_m3_a$tau)
+
+m4_a <- readRDS(here("data_clean", "bayesian_models", "hospitalisation_former_m4_a.rds"))
+post_samples_m4_a <- post_samples(m4_a)
+median_hdci(post_samples_m4_a$tau)
+
+m5_a <- readRDS(here("data_clean", "bayesian_models", "severity_current_m5_a.rds"))
+post_samples_m5_a <- post_samples(m5_a)
+median_hdci(post_samples_m5_a$tau)
+
+m6_a <- readRDS(here("data_clean", "bayesian_models", "severity_former_m6_a.rds"))
+post_samples_m6_a <- post_samples(m6_a)
+median_hdci(post_samples_m6_a$tau)
+
+m7_a <- readRDS(here("data_clean", "bayesian_models", "mortality_current_m7_a.rds"))
+post_samples_m7_a <- post_samples(m7_a)
+median_hdci(post_samples_m7_a$tau)
+
+m8_a <- readRDS(here("data_clean", "bayesian_models", "mortality_former_m8_a.rds"))
+post_samples_m8_a <- post_samples(m8_a)
+median_hdci(post_samples_m8_a$tau)

### 2020-10-01 scripts/smoking_plots.R

@@ -12,7 +12,34 @@ extrafont::loadfonts(device = "win")
 
 ## get data
 data <-  read_rds(here::here('data_clean', 'country_prevalence_data.rds')) %>%
-  mutate(country = recode(country, "SouthKorea" = "South Korea"))
+  mutate(country = recode(country, "SouthKorea" = "South Korea"),
+         study_setting = recode(study_setting, "quarantine_centre" = "community"),
+         study_id  = factor(study_id),
+         study_setting = factor(study_setting),
+         smoking = factor(smoking)) %>%
+  group_by(country, study_setting, smoking, study) %>%
+  mutate(mean = ifelse(is.na(weightedMean(prevalence, w = true_sample)), mean(prevalence),
+                       weightedMean(prevalence, w = true_sample, na.rm = T)),
+         mean_uci = ifelse(is.na(weightedMean(upper_ci, w = true_sample)), mean(upper_ci),
+                       weightedMean(upper_ci, w = true_sample, na.rm = T)),
+         mean_lci = ifelse(is.na(weightedMean(lower_ci, w = true_sample)), mean(lower_ci),
+                       weightedMean(lower_ci, w = true_sample, na.rm = T))) %>%
+  ungroup() %>%
+  group_by(country, smoking, study) %>%
+  mutate(ranked_n = factor(row_number()))
+
+ordered_country <- data %>%
+  filter(study_id != is.na(study_id)) %>%
+  group_by(country, study_id) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(country) %>%
+  summarise(count = sum(count)) %>%
+  arrange(desc(count)) %>%
+  select(country)
+ordered_country <- as.vector(t(ordered_country))
+data$country <- factor(data$country, levels = ordered_country)
+
 country_colour <- data.frame(gapminder, cc = I(country_colors[match(gapminder$country, names(country_colors))]))
 country_colour <- country_colour %>%
   select(country, cc) %>%
@@ -24,6 +51,15 @@ country_colour <- country_colour %>%
   distinct() %>%
   filter(country %in% data$country) %>%
   deframe()
+
+current_smoking_plot <- data %>%
+  filter(smoking == "current_smoking_p") %>%
+  drop_na(country)
+
+former_smoking_plot <- data %>%
+  filter(smoking == "former_smoking_p") %>%
+  drop_na(country) %>%
[…]

### 2020-11-02 reports/bayesian_report.Rmd
v9

@@
 ---
-title: 'Smoking and COVID Living Review (v7): a bayesian analysis'
+title: 'Smoking and COVID Living Review (v8): a bayesian analysis'
 author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
@@
 ``` {r google_data}
-sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1sTlqMbGiWDWJT3dTpa3YQU4z6qcWb0v6Wkp40RcWLXQ/edit?usp=sharing")
+sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1yLtpbk29oUKOGceVw2IAJlt9CbbbPWmltzOW_LG6Su8/edit?usp=sharing")
 search_details <- read_sheet(sheets_id, range = 'article_screening')
@@
 We have excluded this version the following studies:
--`r knitr::combine_words(exclude_from_analysis)`
+
+`r knitr::combine_words(exclude_from_analysis)`
+
 Generally we label these as _old in the Google sheet but there are exceptions to this so we manually define them at the moment.
@@
 ## Filter the included studies
-Based on the version and excluded studies we produce the tables for analysis. This is a searchable table.
+Based on the version and excluded studies we produce the tables for analysis.
+
+This is a searchable table.
 ``` {r descriptives}
@@
 ```
-We have split the data into a list where each country can be analysed separately if required
-
-Confidence intervals are obtained through bootstrap estimates of a thousand repetitions of the included studies.
-Following the production of the confidence intervals from bootstrapping in `olgas_script.R` the prevalence graphs are produced in the `smoking_plots.R` script.
+We have split the data into a list where each country can be analysed separately if required.
+
+Prevalence graphs are produced in the `smoking_plots.R` script.
 This uses data extracted on current and former smoking prevalence data from countries which have included studies in this review.
+We calculate the mean weighted prevalence.
 
@@
 The studies are grouped by the country in which they were conducted.
-The solid line represents national current smoking prevalence obtained from national smoking survey estimates.
-Studies are further grouped by the setting in which they were conducted, Hospital (Triangle), Community (Square) and Community and Hospital (Circle).
+The dashed line represents national current smoking prevalence obtained from national smoking survey estimates.
 The size of the shape corresponds to the relative size of the study sample.
 The weighted mean current smoking prevalence is displayed for each setting.
+
 ![Country smoking prevalence](figure/current_smoking_plots_updated.png)
@@
 #Table 6
+good <- filter(quality_rating, quality_rating$overall_quality == "good" ) %>%
+  filter(!lead_author %in% exclude_from_analysis) %>%
+  nrow()
 fair <- filter(quality_rating, quality_rating$overall_quality == "fair" ) %>%
@@
 ```
-The number of **Fair** quality studies are `r paste(fair)` with `r paste(poor)` **Poor** quality studies
+The number of **Good** quality studies are `r paste(good)`, **Fair** quality studies are `r paste(fair)` with `r paste(poor)` **Poor** quality studies
 

### 2020-11-02 scripts/author_dictionary.R

@@ -27,7 +27,9 @@ cleaned_names <- c('Chow, Us, Cdc',
                    "De, Souza",
                    "Bello, Chavolla, Antonio, Villa",
                    "Antonio, Villa",
-                   "Rubio, Rivas")
+                   "Rubio, Rivas",
+                   "Gil, Agudo",
+                   "Izzi, Engbeaya")
 
 correct_names <- c('Chow (US CDC)',
                    "Gold (US CDC)",
@@ -58,4 +60,6 @@ correct_names <- c('Chow (US CDC)',
                    "de Souza",
                    "Bello-Chavolla, Antonio-Villa",
                    "Antonio-Villa",
-                   "Rubio-Rivas")
+                   "Rubio-Rivas",
+                   "Gil-Agudo",
+                   "Izzi-Engbeaya")

### 2020-11-02 scripts/bayes_scripts.R

@@ -55,8 +55,8 @@ forest_plot <- function(model, data_study, data_pooled, cut, title, type, filena
   forest_data$review_version <- plyr::mapvalues(forest_data$review_version, from = c("previous",
                                                                                      "current",
                                                                                      "combined_pooled"),
-                                                to = c("Previous version (v7)",
-                                                       "Current version (v8)",
+                                                to = c("Previous version (v8)",
+                                                       "Current version (v9)",
                                                        "Pooled effect"))
   summary_forest <- group_by(forest_data, Author, review_version) %>%
     median_qi(b_Intercept)

### 2020-11-02 scripts/bayesian_analysis.R

@@ -104,7 +104,7 @@ current_testing_alternative_prior <- forest_plot(
   pooled_draw_m1_a,
   cut = 4,
   "Forest plot of current smokers and risk of testing positive",
-  "v7 testing prior",
+  "v8 testing prior",
   "m1_a.png"
 )
 
@@ -173,7 +173,7 @@ former_testing_alternative_prior <- forest_plot(
   pooled_draw_m2_a,
   cut = 3,
   "Forest plot of former smokers and risk of testing positive",
-  "v7 testing prior",
+  "v8 testing prior",
   "m2_a.png"
 )
 
@@ -236,7 +236,7 @@ current_hospitalisation_alternative_prior <- forest_plot(
   pooled_draw_m3_a,
   cut = 3,
   "Forest plot of current smokers and risk of hospital admission",
-  "v7 hospitalisation prior",
+  "v8 hospitalisation prior",
   "m3_a.png"
 )
 
@@ -256,6 +256,7 @@ current_hospitalisation_alternative_prior_hh <- forest_plot(
   "m3_a_hh.png"
 )
 
+median_hdci(post_samples_m3_a$tau)
 
 # Former smokers hospitalisation --------------------------------------------------
 # Bayesian analysis for former smokers and hospitalisation for SARS-CoV-2
@@ -299,7 +300,7 @@ former_hospitalisation_alternative_prior <- forest_plot(
   pooled_draw_m4_a,
   cut = 3,
   "Forest plot of former smokers and risk of hospital admission",
-  "v7 hospitalisation prior",
+  "v8 hospitalisation prior",
   "m4_a.png"
 )
 
@@ -320,10 +321,10 @@ former_hospitalisation_alternative_prior_hh <- forest_plot(
   "m4_a_hh.png"
 )
 
+median_hdci(post_samples_m4_a$tau)
 
 # Current smokers disease severity --------------------------------------------------
 # Bayesian analysis for current smokers and severity for SARS-CoV-2
-
 
 post_samples_m5 <- post_samples(m5)
 post_samples_m5_a <- post_samples(m5_a)
@@ -366,7 +367,7 @@ current_severity_alternative_prior <- forest_plot(
   pooled_draw_m5_a,
   cut = 3,
   "Forest plot of current smokers and risk of severe disease",
-  "v7 severity prior",
+  "v8 severity prior",
   "m5_a.png"
 )
 
[…]

### 2020-11-02 scripts/bayesian_data_generation.R

@@ -66,54 +66,54 @@ minimally_informative_prior <-
 
 # Previous meta-analysis guided priors
 current_testing_prior <-
-  c(prior(normal(-.301, 1), class = Intercept),
+  c(prior(normal(-.329, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_testing_prior <-
-  c(prior(normal(0.049, 1), class = Intercept),
+  c(prior(normal(0.02, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_hospitalisation_prior <-
   c(prior(normal(0.058, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_hospitalisation_prior <-
-  c(prior(normal(0.182, 1), class = Intercept),
+  c(prior(normal(0.17, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_severity_prior <-
-  c(prior(normal(0.223, 1), class = Intercept),
+  c(prior(normal(0.23, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_severity_prior <-
   c(prior(normal(0.419, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_mortality_prior <-
-  c(prior(normal(0.199, 1), class = Intercept),
+  c(prior(normal(0.095, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_mortality_prior <- 
-  c(prior(normal(0.329, 1), class = Intercept),
+  c(prior(normal(0.30, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 
 # Classical meta-analysis guided priors with high heterogeneity
 current_testing_prior_hh <-
-  c(prior(normal(-.30, 1), class = Intercept),
+  c(prior(normal(-.329, 1), class = Intercept),
     prior(cauchy(0.3, 1), class = sd))
 former_testing_prior_hh <-
-  c(prior(normal(0.049, 1), class = Intercept),
+  c(prior(normal(0.02, 1), class = Intercept),
     prior(cauchy(0.3, 1), class = sd))
 current_hospitalisation_prior_hh <-
   c(prior(normal(0.058, 1), class = Intercept),
     prior(cauchy(0.3, 1), class = sd)) 
 former_hospitalisation_prior_hh <-
[…]

### 2020-11-02 scripts/smoking_plots.R

@@ -87,89 +87,89 @@ current_sorted_data <- sorted_data %>%
   group_by(country) %>%
   filter(!n <= 1)
 
-a <- ggplot(data = filter(current_sorted_data, study == 1), aes(x = running_count, y = prevalence)) +
-  geom_point(aes(size = true_sample, shape = factor(study_setting))) + 
-  scale_size_continuous(trans = "log10", guide = F) +
-  scale_shape_manual(values = c(15, 16, 17), labels = c("Community", "Community & Hospital", "Hospital")) +
-  geom_linerange(aes(ymin = lower_ci, ymax =upper_ci), linetype = "solid") +
-  geom_segment(data = filter(current_sorted_data, study == 1), aes(x = 0-0.5, 
-                                                                   xend = n+0.5,
-                                                                   y = mean,
-                                                                   yend = mean,
-                                                                   colour = study_setting), size = 0.8) +
-  scale_color_brewer(palette = "Dark2", labels = c("Community", "Community & Hospital", "Hospital")) +
-  geom_segment(data = filter(current_sorted_data, study == 0), aes(x = 0-0.5, 
-                                                                   xend = n+0.5,
-                                                                   y = prevalence,
-                                                                   yend = prevalence), linetype = "solid", 
-               colour = "#A50026",
-               size = 0.8) +
-    coord_flip() +
-    facet_grid(country ~  ., scales = "free_y", space = "free", switch = 'y') +
-  scale_y_continuous(name = "Prevalence", limits = c(0, 0.7), breaks = scales::pretty_breaks(n = 9), expand = c(0, 0)) +
-  theme(panel.spacing = unit(0.1, "lines"),
-        strip.text = element_text(angle = 0, size = 12),
-        axis.title = element_text(size = 10),
-        axis.text.y = element_blank(),
-        axis.title.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.text.x = element_text(size = 10),
-        panel.background = element_blank(),
-        axis.line.x = element_line(colour = "black"),
-        strip.text.y.left = element_text(angle = 0),
-        legend.position = "right") +
-  labs(title ="Prevalence of current smoking in included studies",
-       colour = "Weighted mean prevalence", shape = "Study setting")
-a
-
-png(here::here('reports', 'figure', 'figure_2a.png'), width=912, height=970, res = 120)
-a
-null <- dev.off()
-
-
[…]

### 2020-11-16 reports/bayesian_report.Rmd
update for v9

@@
 ---
-title: 'Smoking and COVID Living Review (v8): a bayesian analysis'
+title: 'Smoking and COVID Living Review (v9): a bayesian analysis'
 author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
@@
 
-exclude_from_analysis <- c(superseded_studies, old_isaric, excluded_other_reason)
+exclude_from_analysis <- unique(c(superseded_studies, old_isaric, excluded_other_reason))
 ```

### 2020-11-16 scripts/author_dictionary.R

@@ -29,7 +29,10 @@ cleaned_names <- c('Chow, Us, Cdc',
                    "Antonio, Villa",
                    "Rubio, Rivas",
                    "Gil, Agudo",
-                   "Izzi, Engbeaya")
+                   "Izzi, Engbeaya",
+                   "Da, Silva, Neto",
+                   "Lopez, Medrano",
+                   "El, Solh")
 
 correct_names <- c('Chow (US CDC)',
                    "Gold (US CDC)",
@@ -62,4 +65,7 @@ correct_names <- c('Chow (US CDC)',
                    "Antonio-Villa",
                    "Rubio-Rivas",
                    "Gil-Agudo",
-                   "Izzi-Engbeaya")
+                   "Izzi-Engbeaya",
+                   "da Silva Neto",
+                   "Lopez-Medrano",
+                   "El-Solh")

### 2021-01-03 reports/bayesian_report.Rmd
Update for version 10, change to plots

--- 2020-11-02 reports/bayesian_report.Rmd
@@
 ---
-title: 'Smoking and COVID Living Review (v8): a bayesian analysis'
+title: 'Smoking and COVID Living Review (v10): a bayesian analysis'
 author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
@@
 Alternatively you can use the pre-generated data which is stored in the `data_clean\bayesian_models` folder in the github repository [here](https://github.com/DidDrog11/smoking_covid/tree/master) and then you can skip the `bayesian_data_generation.R` step.
-```{r install_packages, include=FALSE}
-library(brms)
-library(ggplot2)
-library(tidybayes)
-library(readr)
-library(dplyr)
-library(tidyr)
-library(ggridges)
-library(glue)
-library(stringr)
-library(forcats)
-library(here)
-library(meta)
-library(snakecase)
-library(flextable)
-library(googlesheets4)
-library(linelist)
-library(DT)
-library(epitools)
-library(ggmap)
-```
-
-```{r load_scripts}
+```{r load_packages, include=FALSE, quietly=T}
+source(here::here("scripts", "libraries.R"))
+```
+
+```{r load_dictionary}
 source(here("scripts", "author_dictionary.R"))
@@
 ``` {r google_data}
-sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1yLtpbk29oUKOGceVw2IAJlt9CbbbPWmltzOW_LG6Su8/edit?usp=sharing")
-search_details <- read_sheet(sheets_id, range = 'article_screening')
-data_study_general <- read_sheet(sheets_id, range =  'general_details')
-table_1 <- read_sheet(sheets_id, sheet = 'pop_descriptives')
-table_2 <-  read_sheet(sheets_id, sheet = 'testing')
-table_3 <-  read_sheet(sheets_id, sheet = 'hospitalisation')
-table_4 <-  read_sheet(sheets_id, sheet = 'severity')
-table_5 <-  read_sheet(sheets_id, sheet = 'mortality')
-table_6 <-  read_sheet(sheets_id, sheet = 'updated_quality_appraisal')
+google_api <- rstudioapi::askForSecret("Google API Key")
+register_google(google_api)
+drive_download("https://docs.google.com/spreadsheets/d/1jelIGxrm99QpDKwSoyMUliDJQWtzk4bHsNGDaIjxTaU/edit?usp=sharing", path = here("data_spreadsheet", "data_extraction_current.xlsx"), overwrite = T)
+
+search_details <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'article_screening')
+data_study_general <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'general_details')
+table_1 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'pop_descriptives')
+table_2 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'testing')
+table_3 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'hospitalisation')
+table_4 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'severity')
+table_5 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'mortality')
+table_6 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'updated_quality_appraisal')
 ```
@@
 ```{r head_search_details}
-flextable(search_details[,1:12])
+flextable(search_details[,1:13])
 ```
@@
          current_former_vaper, not_stated, missing, total) %>%
+  mutate(study_id = 1:length(lead_author)) %>%
   filter(!is.na(lead_author)) %>%
-  left_join(., review_details, by = 'lead_author') %>%
+  left_join(., review_details %>%
+              select(-"lead_author"), by = 'study_id') %>%
   write_rds(here::here('data_clean', 'table_1.rds'))
@@
   clean_data() %>%
+  mutate(study_id = 1:length(lead_author)) %>%
   filter(data_on_testing == TRUE) %>%
-  left_join(., review_details, by = 'lead_author') %>%
+  left_join(., review_details %>%
+              select(-"lead_author"), by = 'study_id') %>%
   write_rds(here::here('data_clean', 'table_2.rds'))
@@
   clean_data() %>%
+  mutate(study_id = 1:length(lead_author)) %>%
   filter(data_on_hospitalisation == TRUE) %>%
-  left_join(., review_details, by = 'lead_author') %>%
+  left_join(., review_details %>%
+              select(-"lead_author"), by = 'study_id') %>%
   write_rds(here::here('data_clean', 'table_3.rds'))
@@
   clean_data() %>%
+  mutate(study_id = 1:length(lead_author)) %>%
   filter(data_disease_severity == TRUE) %>%
-  left_join(., review_details, by = 'lead_author') %>%
+  left_join(., review_details %>%
+              select(-"lead_author"), by = 'study_id') %>%
   write_rds(here::here('data_clean', 'table_4.rds'))
@@
   clean_data() %>%
+  mutate(study_id = 1:length(lead_author)) %>%
   filter(data_on_deaths == TRUE) %>%
-  left_join(., review_details, by = 'lead_author') %>%
+  left_join(., review_details %>%
+              select(-"lead_author"), by = 'study_id') %>%
   write_rds(here::here('data_clean', 'table_5.rds'))
@@
   clean_data(protect = protect_columns_2) %>%
+  mutate(study_id = 1:length(lead_author)) %>%
   left_join(., table_1 %>%
-              select(lead_author,
-                     not_stated,
+              select(not_stated,
                      missing,
-                     total),
-                          by = 'lead_author') %>%
+                     total,
+                     study_id),
+                          by = 'study_id') %>%
   mutate(missingness = rowSums(.[8:9], na.rm = T),
@@
   select(-c(not_stated, missing, total, missingness)) %>%
-  left_join(., review_details, by = 'lead_author') %>%
+  left_join(., review_details %>%
+              select(-lead_author), by = 'study_id') %>%
   write_rds(here::here('data_clean', 'table_6.rds'))
@@
 
-exclude_from_analysis <- c(superseded_studies, old_isaric, excluded_other_reason)
+exclude_from_analysis <- unique(c(superseded_studies, old_isaric, excluded_other_reason))
+write_rds(exclude_from_analysis, here("output_data", "excluded_studies.rds"))
 ```
@@
 ``` {r PRISMA}
-current_total <- search_details %>%
-  select(total_included) %>%
-  sum()
-
-previous_total <- search_details %>%
-  filter(review_version %in% prev_versions) %>%
-  select(total_included) %>%
-  sum()
+current_total <- 345
+
+previous_total <- 279
 
@@
   filter(review_version == current_version) %>%
-  mutate(title_abstract = sum(ovid_number_results, medrxiv_number_results, other_source_included),
-         full_texts_assessed = sum(ovid_screened, medrxiv_screened, other_source_included),
+  mutate(title_abstract = sum(ovid_titles, medrxiv_titles, other_source_included),
+         full_texts_assessed = sum(ovid_abstract_full, medrxiv_abstract_full, other_source_included),
          title_abstract_included = sum(ovid_included, medrxiv_included, other_source_included),
@@
          full_text_excluded = full_texts_assessed-total_included,
+         full_text_included = full_texts_assessed-full_text_excluded,
          previous_total = previous_total,
          previous_minus_superseded = previous_total-superseded,
-         current_total = current_total-superseded) %>%
+         current_total = current_total) %>%
   select("Review version" = review_version,
          "Date Screened" = date_screened,
-         "OVID results" = "ovid_number_results",
-         "medRxiv results" = "medrxiv_number_results",
+         "OVID results" = "ovid_titles",
+         "medRxiv results" = "medrxiv_titles",
          "Other sources" = "other_source_included",
-         "Title/abstracts screened" = "title_abstract",
-         "Title/abstracts excluded" = "title_abstract_excluded",
-         "Full texts assessed" = "full_texts_assessed",
+         "Titles/abstracts screened" = "title_abstract",
+         "Titles/abstracts excluded" = "title_abstract_excluded",
+         "Full texts screened" = "full_texts_assessed",
          "Full texts excluded" = "full_text_excluded",
+         "Records included" = "full_text_included",
          "Records carried forward" = "previous_minus_superseded",
@@
 ``` {r countries, cache = TRUE}
-library('ggmap')
 #Countries
@@
   to_upper_camel_case() %>%
-  recode(., 'Usa' = 'USA', 'Uk' = 'UK', 'SaudiArabia' = 'Saudi Arabia', "SouthKorea" = "South Korea") %>%
+  recode(., 'Usa' = 'USA', 'Uk' = 'UK', 'SaudiArabia' = 'Saudi Arabia', "SouthKorea" = "South Korea", "UnitedArabEmirates" = "United Arab Emirates") %>%
   table() %>%
@@
 
-my_bins <- c(0, 1, 5, 10, 20, 30, 40, 80)
+my_bins <- c(0, 1, 5, 10, 20, 50, 70, 100)
 mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)
@@
 #Numbers
-summary(table_1$sample_size)
-```
-* **Total number of participants**: `r paste(sum(table_1$sample_size))`
+summary(as.numeric(table_1$sample_size))
+```
+* **Total number of participants**: `r paste(sum(as.numeric(table_1$sample_size)))`
 
@@
 #These are the studies with data on current, former and never smokers
+set_flextable_defaults(digits = 0, big.mark = ",", na_str = "-")
 
@@
     set_caption(caption = 'Studies reporting complete smoking status') %>%
-  colformat_num(col_keys = c("Sample size", "Current smokers", "Former smokers", "Never smokers"),
-                digits = 0, na_str = '-', big.mark = ',') %>%
+  colformat_double(j = c("Sample size", "Current smokers", "Former smokers", "Never smokers"),
+                digits = get_flextable_defaults()$digits, big.mark = get_flextable_defaults()$big.mark, na_str = get_flextable_defaults()$na_str) %>%
   set_table_properties(width = 1, layout = 'autofit')
@@
     set_caption(caption = 'Studies reporting complete smoking status') %>%
-  colformat_num(col_keys = c("Sample size", "Current/former smokers", "Never smokers"),
-                digits = 0, na_str = '-', big.mark = ',') %>%
+  colformat_double(j = c("Sample size", "Current/former smokers", "Never smokers"),
+                digits = get_flextable_defaults()$digits, big.mark = get_flextable_defaults()$big.mark, na_str = get_flextable_defaults()$na_str) %>%
   set_table_properties(width = 1, layout = 'autofit')
@@
     set_caption(caption = 'Studies reporting complete smoking status') %>%
-  colformat_num(col_keys = numeric_names,
+  colformat_double(j = numeric_names,
                 digits = 0, na_str = '-', big.mark = ',') %>%
@@
 # Smoking prevalence by country {.tabset .tabset-fade .tabset-pills}
-``` {r smoking_country}
+``` {r smoking_country, quietly=T}
 
@@
 
+source(here("scripts", "smoking_plots.R"))
 ```
@@
     set_caption(caption = 'Characteristics of included studies') %>%
-  colformat_num(col_keys = numeric_columns, digits = 1, na_str = '-', big.mark = ',') %>%
-  colformat_num(col_keys = 'Sample size', digits = 0, na_str = '-', big.mark = ',') %>%
+  colformat_double(j = numeric_columns, digits = 1, na_str = '-', big.mark = ',') %>%
+  colformat_double(j = 'Sample size', digits = 0, na_str = '-', big.mark = ',') %>%
   set_table_properties(width = 1, layout = 'autofit')
@@
     set_caption(caption = 'SARS-CoV-2 infection by smoking status') %>%
-  colformat_num(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',')
+  colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',')
 ```
@@
     set_caption(caption = 'COVID-19 hospitalisation by smoking status') %>%
-  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
+  colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
   colformat_char(na_str = '-')
@@
     set_caption(caption = 'COVID-19 severity by smoking status') %>%
-  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
+  colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
   colformat_char(na_str = '-')
@@
     set_caption(caption = 'COVID-19 mortality by smoking status') %>%
-  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
+  colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
   colformat_char(na_str = '-')
@@
 
-country_smoking <-  read_sheet(sheets_id, sheet = 'national_smoking_prevalence') %>%
+country_smoking <-  readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'national_smoking_prevalence') %>%
   flextable()

### 2021-01-03 scripts/author_dictionary.R

@@ -32,13 +32,24 @@ cleaned_names <- c('Chow, Us, Cdc',
                    "Izzi, Engbeaya",
                    "Da, Silva, Neto",
                    "Lopez, Medrano",
-                   "El, Solh")
+                   "El, Solh",
+                   "Vila, Corcoles",
+                   "Torres, Macho",
+                   "Serling, Boyd",
+                   "Parra, Bracamonte",
+                   "O, Reilly",
+                   "Bermejo, Martin",
+                   "Marquez, Salinas",
+                   "Diez, Manglanas",
+                   "Martinez, Lacalzada",
+                   "O, Gallagher",
+                   "Vila, Corcoles, Statue, Gracia")
 
 correct_names <- c('Chow (US CDC)',
                    "Gold (US CDC)",
                    "Miyara",
                    "de la Rica",
-                   "The Opensafely Collaborative",
+                   "The OpenSAFELY Collaborative",
                    "Bello-Chavolla",
                    "Carillo-Vega",
                    "de Lusignan",
@@ -68,4 +79,15 @@ correct_names <- c('Chow (US CDC)',
                    "Izzi-Engbeaya",
                    "da Silva Neto",
                    "Lopez-Medrano",
-                   "El-Solh")
+                   "El-Solh",
+                   "Vila-Corcoles",
+                   "Torres-Macho",
+                   "Serling-Boyd",
+                   "Parra-Bracamonte",
+                   "O'Reilly",
+                   "Bermejo-Martin",
+                   "Marquez-Salinas",
+                   "Diez-Manglanas",
+                   "Martinez-Lacalzada",
+                   "O'Gallagher",
+                   "Vila-Corcoles, Statue-Gracia")

### 2021-01-03 scripts/bayes_scripts.R

@@ -55,8 +55,8 @@ forest_plot <- function(model, data_study, data_pooled, cut, title, type, filena
   forest_data$review_version <- plyr::mapvalues(forest_data$review_version, from = c("previous",
                                                                                      "current",
                                                                                      "combined_pooled"),
-                                                to = c("Previous version (v8)",
-                                                       "Current version (v9)",
+                                                to = c("Previous version (v9)",
+                                                       "Current version (v10)",
                                                        "Pooled effect"))
   summary_forest <- group_by(forest_data, Author, review_version) %>%
     median_qi(b_Intercept)

### 2021-01-03 scripts/bayesian_analysis.R

@@ -104,7 +104,7 @@ current_testing_alternative_prior <- forest_plot(
   pooled_draw_m1_a,
   cut = 4,
   "Forest plot of current smokers and risk of testing positive",
-  "v8 testing prior",
+  "v9 testing prior",
   "m1_a.png"
 )
 
@@ -173,7 +173,7 @@ former_testing_alternative_prior <- forest_plot(
   pooled_draw_m2_a,
   cut = 3,
   "Forest plot of former smokers and risk of testing positive",
-  "v8 testing prior",
+  "v9 testing prior",
   "m2_a.png"
 )
 
@@ -211,13 +211,22 @@ m3_a_hh_ecdf <- ecdf(exp(post_samples_m3_a_hh$TE))
 1-m3_a_hh_ecdf(1.1) #high heterogeneity
 
 
-study_draws_m3 <- study_draw(m3)
+study_draws_m3 <- study_draw(m3) %>%
+mutate(review_version = as.character(review_version),
+       review_version = ifelse(is.na(review_version), "current", review_version),
+       review_version = as_factor(review_version))
 pooled_draw_m3 <- pooled_effect_draw(m3)
 
-study_draws_m3_a <- study_draw(m3_a)
+study_draws_m3_a <- study_draw(m3_a)  %>%
+  mutate(review_version = as.character(review_version),
+         review_version = ifelse(is.na(review_version), "current", review_version),
+         review_version = as_factor(review_version))
 pooled_draw_m3_a <- pooled_effect_draw(m3_a)
 
-study_draws_m3_a_hh <- study_draw(m3_a_hh)
+study_draws_m3_a_hh <- study_draw(m3_a_hh)  %>%
+  mutate(review_version = as.character(review_version),
+         review_version = ifelse(is.na(review_version), "current", review_version),
+         review_version = as_factor(review_version))
 pooled_draw_m3_a_hh <- pooled_effect_draw(m3_a_hh)
 
 current_hospitalisation_minimal_prior <- forest_plot(
@@ -236,7 +245,7 @@ current_hospitalisation_alternative_prior <- forest_plot(
   pooled_draw_m3_a,
   cut = 3,
   "Forest plot of current smokers and risk of hospital admission",
-  "v8 hospitalisation prior",
+  "v9 hospitalisation prior",
   "m3_a.png"
 )
 
@@ -261,7 +270,6 @@ median_hdci(post_samples_m3_a$tau)
[…]

### 2021-01-03 scripts/bayesian_data_generation.R

@@ -16,9 +16,9 @@ source(here("scripts", "bayes_scripts.R"))
 source(here("scripts", "author_dictionary.R"))
 
 data_study_general <- readRDS(here("data_clean", "data_study_general.rds"))
-                              
-previous_review_versions <- prev_versions
-current_review_version <- current_version
+versions <- as_data_frame(read_rds(here("data_clean", "versioning.rds")))
+previous_review_versions <- unique(versions$prev_versions)
+current_review_version <- unique(versions$current_version)
 
 study_review_version <- data_study_general %>%
   mutate(lead_author = to_upper_camel_case(lead_author, sep_out = ", "),
@@ -66,54 +66,54 @@ minimally_informative_prior <-
 
 # Previous meta-analysis guided priors
 current_testing_prior <-
-  c(prior(normal(-.329, 1), class = Intercept),
+  c(prior(normal(-.371, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_testing_prior <-
-  c(prior(normal(0.02, 1), class = Intercept),
+  c(prior(normal(0.0198, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_hospitalisation_prior <-
-  c(prior(normal(0.058, 1), class = Intercept),
+  c(prior(normal(0.0582, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_hospitalisation_prior <-
-  c(prior(normal(0.17, 1), class = Intercept),
+  c(prior(normal(0.157, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_severity_prior <-
-  c(prior(normal(0.23, 1), class = Intercept),
+  c(prior(normal(0.231, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_severity_prior <-
   c(prior(normal(0.419, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_mortality_prior <-
-  c(prior(normal(0.095, 1), class = Intercept),
+  c(prior(normal(0.0487, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_mortality_prior <- 
[…]

### 2021-01-03 scripts/libraries.R

@@ -1,0 +1,20 @@
+library(brms)
+library(ggplot2)
+library(tidybayes)
+library(readr)
+library(dplyr)
+library(tidyr)
+library(ggridges)
+library(glue)
+library(stringr)
+library(forcats)
+library(here)
+library(meta)
+library(snakecase)
+library(flextable)
+library(googlesheets4)
+library(googledrive)
+library(linelist)
+library(DT)
+library(epitools)
+library(ggmap)

### 2021-01-03 scripts/olgas_script.R

@@ -6,30 +6,25 @@ library(googlesheets4)
 
 source(here("scripts", "bootstrap_function.R"))
 
-set.seed(42)
 smoking_status <- c('current_smoker', 'former_smoker', 'current_former_smoker', 'never_smoker', 'never_smoker_unknown', 'not_stated', 'missing')
-table_1 <- read_rds(here("data_clean", "table_1.rds"))
+data_study_general <- read_rds(here("data_clean", "data_study_general.rds")) %>%
+  filter(!lead_author %in% exclude_from_analysis)
+table_1 <- read_rds(here("data_clean", "table_1.rds")) %>%
+  filter(!lead_author %in% exclude_from_analysis)
 
 prevalence_plot <- table_1 %>%
   left_join(., data_study_general %>%
-              select(lead_author, study_setting), 
-            by = "lead_author") %>%
+              select(study_id, study_setting), 
+            by = "study_id") %>%
   group_by(country, study_setting) %>%
   rename('sample' = total) %>%
   filter(current_smoker != 'NA') %>%
   filter(study_setting != is.na(study_setting)) %>%
-  replace_na(., list(current_smoker = 0,
-                     former_smoker = 0,
-                     current_former_smoker = 0,
-                     never_smoker = 0,
-                     never_smoker_unknown = 0,
-                     not_stated = 0,
-                     missing = 0)) %>%
   mutate(ever_smoker = former_smoker + current_smoker + current_former_smoker,
          not_stated_missing = missing + not_stated + never_smoker_unknown) %>%
-  select(lead_author, country, sample, current_smoker, former_smoker, never_smoker, ever_smoker, not_stated_missing) %>%
+  select(study_id, lead_author, country, sample, current_smoker, former_smoker, never_smoker, ever_smoker, not_stated_missing) %>%
   mutate(total = never_smoker + ever_smoker + not_stated_missing) %>%
-  mutate(p_current_smoker = current_smoker/total,
+  mutate(p_current_smoker = ifelse(current_smoker == NAcurrent_smoker/total,
          p_ever_smoker = ever_smoker/total,
          p_former_smoker = former_smoker/total,
          p_never_smoker = never_smoker/total,
[…]

### 2021-01-03 scripts/ovid_search.R

@@ -1,0 +1,6 @@
+library("tidyverse")
+library("here")
+
+ovid_dec <- read_csv(here("data_spreadsheet", "2020-12-17_ovid.csv")) %>%
+  select(DO) %>%
+  drop_na()

### 2021-01-03 scripts/smoking_plots.R

@@ -1,34 +1,37 @@
 ## packages
-library(tidyverse)
+source(here::here("scripts", "libraries.R"))
 library(ggsci)
 library(matrixStats)
 library(gapminder)
 
-## set seed
-set.seed(123)
+smoking_status <- c('current_smoker', 'former_smoker', 'current_former_smoker', 'never_smoker', 'never_smoker_unknown', 'not_stated', 'missing')
+exclude_from_analysis <- read_rds(here("output_data", "excluded_studies.rds"))
+data_study_general <- read_rds(here("data_clean", "data_study_general.rds")) %>%
+  filter(!lead_author %in% exclude_from_analysis)
+table_1 <- read_rds(here("data_clean", "table_1.rds")) %>%
+  filter(!lead_author %in% exclude_from_analysis)
 
-## load fonts
-extrafont::loadfonts(device = "win")
-
-## get data
-data <-  read_rds(here::here('data_clean', 'country_prevalence_data.rds')) %>%
-  mutate(country = recode(country, "SouthKorea" = "South Korea"),
-         study_setting = recode(study_setting, "quarantine_centre" = "community"),
+prevalence_plot <- table_1 %>%
+  left_join(., data_study_general %>%
+              select(study_id, study_setting), 
+            by = "study_id") %>%
+  group_by(country, study_setting) %>%
+  rename('sample' = total) %>%
+  select(study_id, country, sample, current_smoker, former_smoker) %>%
+  mutate(country = snakecase::to_any_case(country, case = "title", abbreviations = c("USA", "UK")),
+         p_current_smoker = ifelse(!is.na(current_smoker), current_smoker/sample, NA),
+         p_former_smoker = ifelse(!is.na(former_smoker), former_smoker/sample, NA),
+         study_setting = recode(study_setting, "quarantine_centre" = "community",
+                                "homeless_shelters" = "community"),
          study_id  = factor(study_id),
-         study_setting = factor(study_setting),
-         smoking = factor(smoking)) %>%
-  group_by(country, study_setting, smoking, study) %>%
-  mutate(mean = ifelse(is.na(weightedMean(prevalence, w = true_sample)), mean(prevalence),
-                       weightedMean(prevalence, w = true_sample, na.rm = T)),
-         mean_uci = ifelse(is.na(weightedMean(upper_ci, w = true_sample)), mean(upper_ci),
-                       weightedMean(upper_ci, w = true_sample, na.rm = T)),
[…]

### 2021-01-28 reports/bayesian_report.Rmd
from mac

@@
     theme: spacelab
-    toc: false
-    toc_collapse: true
-    toc_depth: 3
-    toc_float: yes
     css: !expr here::here('css', 'style.css')
@@
 ```{r setup, include=FALSE}
-
 knitr::opts_chunk$set(echo = TRUE,
@@
 unlink('reports/markdown_report_cache', recursive = TRUE)
-
-```
-# Data preparation {.tabset .tabset-fade .tabset-pills}
+```
+# Data preparation
 
 ## Load packages
-* **Load packages**: load packages and useful scripts for the analysis
+* **Load packages**: All necessary packages are loaded in from the `libraries.r` script in the `/scripts` folder.
+* **Author dictionary**: An author dictionary is created to manage the presentation of author names in the final tables and meta-analyses
 * **Load scripts**: We have written functions to handle much of the Bayesian analysis
-* **Load Bayesian datasets**: The Bayesian approach relies on large simulated datasets from the observed data, for this report we will only present the results graphically from data we have simulated elsewhere.
-The scripts `bayes_scripts.R`, `bayesian_data_generation.R` and `bayesian_analysis.R` should be able to be run locally if you want to generate your own simulated data.
-Alternatively you can use the pre-generated data which is stored in the `data_clean\bayesian_models` folder in the github repository [here](https://github.com/DidDrog11/smoking_covid/tree/master) and then you can skip the `bayesian_data_generation.R` step.
+* **Load Bayesian datasets**: The Bayesian approach relies on large simulated datasets from the observed data, for this report we only present the results graphically from data simulated outside of the report.
+The scripts `bayes_scripts.R`, `bayesian_data_generation.R` and `bayesian_analysis.R` should be able to be run locally to generate your own simulated data.
+
+The code is all stored in the GitHub repository [here](https://github.com/DidDrog11/smoking_covid/tree/master)
 ```{r load_packages, include=FALSE, quietly=T}
@@
 ## Download data
+This is relatively automated in this script however, if you don't have an API key you can download the file individually and store it in the `/data_spreadsheet` folder as `data_extraction_current.xlsx`, the sheets are then extracted from the file.
 ``` {r google_data}
@@
 
-The pull requests from the API may happen too rapidly if this is run as a chunk so it's better to run this line by line.
-
 We have extracted data from `r nrow(data_study_general)` studies, however, some of these are duplicates and some have been removed from the analysis set.
@@
 The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.
+
 This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`.
@@
 ```
-The final sheet is used to collate data on **Study quality** we changed our method on assessing study quality in version 3 of the review. We use the following criteria:
+The final sheet is used to collate data on **Study quality**. The method of assessing quality changed in version 3 of the review. We now use the following criteria:
 
@@
 Kamvar within the [R Epidemics Consortium](https://www.repidemicsconsortium.org/)
-The script `clean_data()` ensures consistency among column names, country names, dates and missing values in the included sheets.
+The function `clean_data()` ensures consistency among column names, country names, dates and missing values in the included sheets.
 Data is also saved as `.rds` files in the `/data_clean` folder of the project repository.
 ```{r clean_data}
-
-search_details <- search_details %>%
-  clean_data() %>%
-  rename(date_screened = date_screening_in_format_2020_05_22) %>%
-  write_rds(here::here('data_clean', 'search_details.rds')) ## The title of the date screened column is to communicate among the review team how we need date formatted
-
-protect_columns_1 <- names(data_study_general) %in% 'doi' ## As the DoI's are links we want to preserve their formatting when running the cleaning script
-
-data_study_general <- data_study_general %>%
-  rename(date_published = `date_published_format(2020-05-12)`) %>%
-  clean_data(protect = protect_columns_1) %>%
-  mutate(study_id = 1:length(lead_author)) %>% ## We create a unique study ID to aid checking of the outputs later
-  write_rds(here::here('data_clean', 'data_study_general.rds'))
-
-prisma_previous <- data_study_general
-
-review_details <- data_study_general %>%
-  select(lead_author, date_published, country, review_version, study_id) %>% ## The version of the review in which included studies were incorporated is noted
-  write_rds(here::here('data_clean', 'review_details.rds'))
-
-table_1 <- table_1 %>%
-  clean_data() %>%
-  mutate(., lower_range = sub('\\_.*', '', .$range)) %>%
-  mutate(., upper_range = sub('.*_', '',.$range,)) %>% ## We break apart the age range into an upper and lower level
-  select(lead_author, sample_size, female_sex_percent, data_source, median_age, mean_age, iqr_lower, iqr_upper, standard_deviation, lower_range,
-         upper_range, current_smoker, former_smoker, current_former_smoker, never_smoker, never_smoker_unknown, current_vaper, former_vaper,
-         current_former_vaper, not_stated, missing, total) %>%
-  mutate(study_id = 1:length(lead_author)) %>%
-  filter(!is.na(lead_author)) %>%
-  left_join(., review_details %>%
-              select(-"lead_author"), by = 'study_id') %>%
-  write_rds(here::here('data_clean', 'table_1.rds'))
-
-##For the following tables we're only interested in obtaining data for trials which report on the specific outcome of interest
-
-table_2 <-  table_2 %>%
-  clean_data() %>%
-  mutate(study_id = 1:length(lead_author)) %>%
-  filter(data_on_testing == TRUE) %>%
-  left_join(., review_details %>%
-              select(-"lead_author"), by = 'study_id') %>%
-  write_rds(here::here('data_clean', 'table_2.rds'))
-
-table_3 <-  table_3 %>%
-  clean_data() %>%
-  mutate(study_id = 1:length(lead_author)) %>%
-  filter(data_on_hospitalisation == TRUE) %>%
-  left_join(., review_details %>%
-              select(-"lead_author"), by = 'study_id') %>%
-  write_rds(here::here('data_clean', 'table_3.rds'))
-
-table_4 <-  table_4 %>%
-  clean_data() %>%
-  mutate(study_id = 1:length(lead_author)) %>%
-  filter(data_disease_severity == TRUE) %>%
-  left_join(., review_details %>%
-              select(-"lead_author"), by = 'study_id') %>%
-  write_rds(here::here('data_clean', 'table_4.rds'))
-
-table_5 <- table_5 %>%
-  clean_data() %>%
-  mutate(study_id = 1:length(lead_author)) %>%
-  filter(data_on_deaths == TRUE) %>%
-  left_join(., review_details %>%
-              select(-"lead_author"), by = 'study_id') %>%
-  write_rds(here::here('data_clean', 'table_5.rds'))
-
-protect_columns_2 <- !names(table_6) %in% 'lead_author'
-
-table_6 <-  table_6 %>%
-  clean_data(protect = protect_columns_2) %>%
-  mutate(study_id = 1:length(lead_author)) %>%
-  left_join(., table_1 %>%
-              select(not_stated,
-                     missing,
-                     total,
-                     study_id),
-                          by = 'study_id') %>%
-  mutate(missingness = rowSums(.[8:9], na.rm = T),
-         missingness_percentage = (missingness/total)*100) %>% ## Later we want to assess the amount of missingness for each study as it would impact the inclusion of a study into the relevant meta-analyses
-  select(-c(not_stated, missing, total, missingness)) %>%
-  left_join(., review_details %>%
-              select(-lead_author), by = 'study_id') %>%
-  write_rds(here::here('data_clean', 'table_6.rds'))
-
-a <- data_study_general %>%
-  select(lead_author, date_published, source, study_id) %>%
-  rename('Lead Author' = lead_author,
-         'Date Published' = date_published,
-         'Publication Source' = source,
-         'Study ID' = study_id)
-
-a$`Lead Author` <- to_upper_camel_case(a$`Lead Author`, sep_out = ", ")
-a$`Publication Source` <- to_title_case(a$`Publication Source`)
-a$`Publication Source` <- if_else(str_length(a$`Publication Source`) < 5,
-                                  toupper(a$`Publication Source`),
-                                  to_title_case(a$`Publication Source`))
-
-write_rds(data_study_general, here("data_clean", "data_study_general.rds"))
+source(here("scripts", "clean_data.r"))
 ```
@@
 
-# Preparing the descriptive analysis {.tabset .tabset-fade .tabset-pills}
+# Preparing the descriptive analysis
 
@@
 
-Generally we label these as _old in the Google sheet but there are exceptions to this so we manually define them at the moment.
+Generally we label these as _old in the Google sheet but there are exceptions to this so we manually define them.
 
@@
 ``` {r descriptives}
-data_study_general <- data_study_general %>%
-  filter(review_version %in% analysed_versions) %>%
-  filter(!(lead_author %in% exclude_from_analysis)) %>%
-  write_rds(., here::here('data_clean', 'data_study_general.rds'))
-
-review_details <- data_study_general %>%
-  select(lead_author, date_published, country, review_version, study_id) %>% 
-  filter(!(lead_author %in% exclude_from_analysis))
-
-table_1 <- table_1 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!(lead_author %in% exclude_from_analysis))
-
-table_2 <- table_2 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!(lead_author %in% exclude_from_analysis))
-
-table_3 <- table_3 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!(lead_author %in% exclude_from_analysis))
-
-table_4 <- table_4 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!(lead_author %in% exclude_from_analysis))
-
-table_5 <- table_5 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!(lead_author %in% exclude_from_analysis))
-
-table_6 <- table_6 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!lead_author %in% exclude_from_analysis)
-
-a <- data_study_general %>%
-  select(lead_author, date_published, source, study_id) %>%
-  rename('Lead Author' = lead_author,
-         'Date Published' = date_published,
-         'Publication Source' = source,
-         'Study ID' = study_id)
-
-a$`Lead Author` <- to_upper_camel_case(a$`Lead Author`, sep_out = ", ")
-a$`Publication Source` <- to_title_case(a$`Publication Source`)
-a$`Publication Source` <- if_else(str_length(a$`Publication Source`) < 5,
-                                  toupper(a$`Publication Source`),
-                                  to_title_case(a$`Publication Source`))
-
+source(here("scripts", "remove_excluded.r"))
 datatable(a)
@@
 
-# Study descriptive analysis {.tabset .tabset-fade .tabset-pills}
+# Study descriptive analysis
 
@@
 
+all_countries <- country %>%
+  group_split(-Number)
+
 ggplot(country, aes(x = reorder(Country, desc(-Number)), y = Number))+
@@
 library(coronavirus)
-update_dataset()
 
@@
 
-from_countries <- c("USA", "UK", "Iran", "South Korea")
+from_countries <- c("USA", "UK", "Iran",  "Korea, South")
 to_countries <- c("United States", "United Kingdom", "Iran (Islamic Republic of)", "Korea, Republic of")
@@
 
-from_countries <- c("US", "Iran", "Korea, South", "Czech Republic")
-to_countries <- c("United States", "Iran (Islamic Republic of)", "Korea, Republic of", "Czechia")
-
+refresh_coronavirus_jhu()
 data("coronavirus")
@@
 #Numbers
-summary(as.numeric(table_1$sample_size))
+participants <- summary(as.numeric(table_1$sample_size))
 ```
@@
 #Source of data
-table_1$data_source %>%
+data_source <- table_1$data_source %>%
   recode('electronic_health_records' = 'Electronic health record',
@@
   rename('Data source' = 1,
-         'Number' = 2) %>%
+         'Number' = 2)
+
+data_source %>%
   flextable() %>%
@@
 ``` {r diagnosis_method}
-table(data_study_general$diagnostic_approach)
+diagnostics <- table(data_study_general$diagnostic_approach)
+diagnostics
 ```
@@
   rename("overall_quality" = study_quality_final)
+
+fair_quality_number <- quality_rating %>%
+  filter(overall_quality == "fair")
+
+poor_quality_number <- quality_rating %>%
+  filter(overall_quality == "poor")
 
@@
                      'Never smokers %', 'Never/unknown smokers %', 'Missing %')
+
+saveRDS(a, file = here("data_clean", "table_1_refs.rds"))
+source(here("scripts", "references_in_table.r"))
 ```
@@
 
+saveRDS(a, file = here("reports", "tables", "table_1.rds"))
 save_as_docx(a, path = here('output_data', 'Table_1.docx'))
@@
 
+saveRDS(a, file = here("reports", "tables", "table_2.rds"))
 save_as_docx(a, path = here('output_data', 'Table_2.docx'))
@@
   filter(!lead_author %in% c("reiter", "hippisley_cox", "ullah"))
+
+good_quality_2 <- quality_table_2 %>%
+  filter(overall_quality != 'good')
+
+fair_quality_2 <- quality_table_2 %>%
+  filter(overall_quality != 'fair')
 
@@
 
+saveRDS(a, file = here("reports", "tables", "table_3.rds"))
 save_as_docx(a, path = here('output_data', 'Table_3.docx'))
@@
   select(lead_author)
+
+fair_quality_3 <- quality_table_3 %>%
+  filter(overall_quality != 'fair')
 
@@
 
+saveRDS(a, file = here("reports", "tables", "table_4.rds"))
 save_as_docx(a, path = here('output_data', 'Table_4.docx'))
@@
   select(lead_author)
+
+fair_quality_4 <- quality_table_4 %>%
+  filter(overall_quality == 'fair')
 ```
@@
 
+fair_quality_5 <- quality_table_5 %>%
+  filter(overall_quality != 'fair')
+
 a <- table_5_word %>%
@@
 
+saveRDS(a, file = here("reports", "tables", "table_5.rds"))
 save_as_docx(a, path = here('output_data', 'Table_5.docx'))
@@
   select(lead_author)
+
+fair_quality_4 <- quality_table_5 %>%
+  filter(overall_quality != 'fair')
 ```

### 2021-01-28 reports/manuscript.Rmd
from mac

---
title: "Smoking and COVID Living Review (`r paste(current_version)`): a bayesian analysis"
author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:

  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: true
    theme: spacelab
    css: !expr here::here('css', 'style.css')
    keep_md: false
editor_options: 
  chunk_output_type: inline
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = F,
                      message = F)
options(width = 180)
unlink('reports/markdown_report_cache', recursive = TRUE)
```

#Abstract

**Aims:** To estimate the association of smoking status with rates of i) infection, ii) hospitalisation, iii) disease severity, and iv) mortality from SARS-CoV-2/COVID-19 disease.

**Design:** Living rapid review of observational and experimental studies with random-effects hierarchical Bayesian meta-analyses. Published articles and pre-prints were identified via MEDLINE and medRxiv.

**Setting:** Community or hospital. No restrictions on location.

**Participants:** Adults who received a SARS-CoV-2 test or a COVID-19 diagnosis.

**Measurements:** Outcomes were SARS-CoV-2 infection, hospitalisation, disease severity and mortality stratified by smoking status. Study quality was assessed (i.e. ‘good’, ‘fair’ and ‘poor’).

**Findings:** `r paste(current_version)` (searches up to `r paste(tail(search_details$date_screened, 1))`) included `r paste(current_total)` studies with `r paste(running_meta_count)` ‘good’ and ‘fair’ quality studies included in unadjusted meta-analyses. `r paste(nrow(full_smoking_status))` studies (`r paste(round(nrow(full_smoking_status)/current_total*100,1), "%")`) reported current, former and never smoking status with the remainder using broader categories. Recorded smoking prevalence among people with COVID-19 was generally lower than national prevalence. Current compared with never smokers were at reduced risk of SARS-CoV-2 infection (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_lci, 2))`). Data for former smokers were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_lci, 2))`) but favoured there being no important association (`r paste(1-m2_a_ecdf(1.1), "%")` probability of RR ≥1.1). Former compared with never smokers were at increased risk of hospitalisation (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_lci, 2))`), greater disease severity (RR = `r paste(round(former_severity_RR_value, 2))`, CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_lci, 2))`) and mortality (RR = `r paste(round(former_mortality_RR_value, 2))`, CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_lci, 2))`). Data for current smokers on hospitalisation, disease severity and mortality were inconclusive (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_lci, 2))`; RR `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_lci, 2))`; RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_lci, 2))`, respectively) but favoured there being no important associations with hospitalisation and mortality (`r paste(1-m3_a_ecdf(1.1), "%")` and `r paste(1-m5_a_ecdf(1.1), "%")` probability of RR ≥1.1, respectively) and a small but important association with disease severity (`r paste(1-m7_a_ecdf(1.1), "%")` probability of RR ≥1.1).

Conclusions: Compared with never smokers, current smokers appear to be at reduced risk of SARS-CoV-2 infection while former smokers appear to be at increased risk of hospitalisation, greater disease severity and mortality from COVID-19. However, it is uncertain whether these associations are causal.

v7 of this living review article has been published in *Addiction* and is available [here](https://doi-org.libproxy.ucl.ac.uk/10.1111/add.15276)

##Introduction

COVID-19 is a respiratory disease caused by the SARS-CoV-2 virus. Large age and gender differences in case severity and mortality have been observed in the ongoing COVID-19 pandemic[@guan_clinical_2020]; however, these differences are currently unexplained. SARS-CoV-2 enters epithelial cells through the angiotensin-converting enzyme 2 (ACE-2) receptor[@hoffmann_sars-cov-2_2020]. Some evidence suggests that gene expression and subsequent receptor levels are elevated in the airway and oral epithelium of current smokers[@brake_smoking_2020],[@cai_bulk_2020], which could put smokers at higher risk of contracting SARS-CoV-2. Other studies, however, suggest that nicotine downregulates the ACE-2 receptor[@oakes_nicotine_2018]. These uncertainties notwithstanding, both former and current smoking is known to increase the risk of respiratory viral[@denholm_hospitalised_2010],[@abadom_risk_2016] and bacterial[@almirall_proportion_1999],[@feldman_cigarette_2013] infections and is associated with worse outcomes once infected. Cigarette smoke reduces the respiratory immune defence through peri-bronchiolar inflammationand fibrosis, impaired mucociliary clearance and disruption of the respiratory epithelium[@dye_occasional_1994]. There is also reason to believe that behavioural factors (e.g. regular hand-to-mouth movements) involved in smoking may increase SARS-CoV-2 infection and transmission in current smokers. However, early data from the COVID-19 pandemic have not provided clear evidence for a negative impact of current or former smoking on SARS-CoV-2 infection or COVID-19 disease outcomes, such as hospitalisation or mortality[@vardavas_covid-19_2020]. It has also been hypothesised that nicotine might protect against a hyper-inflammatory response to SARS-CoV-2 infection, which may lead to adverse outcomes in patients with COVID-19 disease[@farsalinos_editorial_2020].

There are several reviews that fall within the scope of smoking and COVID-19[@vardavas_covid-19_2020],[@emami_prevalence_2020],[@alqahtani_prevalence_2020],[@patanavanich_smoking_2020],[@berlin_covid-19_0],[@farsalinos_systematic_2020],[@grundy_smoking_2020]. We aimed to produce a rapid synthesis of available evidence pertaining to the rates of infection, hospitalisation, disease severity and mortality from SARS-CoV-2/COVID-19 stratified by smoking status. Given the increasing availability of data on this topic, this is a living review with regular updates. As evidence accumulates, the review will be expanded to include studies reporting COVID-19 outcomes by alternative nicotine use (e.g., nicotine replacement therapy or e-cigarettes).

##Methods

*Study design*

This is a living evidence review  which is updated as new evidence becomes available[@elliott_living_2014]. We adopted recommended best practice for rapid evidence reviews, which involved limiting the search to main databases and having one reviewer extract the data and another verify[@tricco_scoping_2015]. This study was not pre-registered but evolved from a report written for a UK medical society[@simons_smoking_2020]. The most recent (and all future) version(s) of this living review is available [here](https://www.qeios.com/read/latest-UJR2AW). Version 7 of this living review has been published in a peer-reviewed journal[@simons_association_nodate]. A completed Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) checklist is included in Supplementary file 1.

*Eligibility criteria*

Studies were included if they:

  1. Were primary research studies using experimental (e.g. randomised controlled trial), quasi-experimental (e.g. pre- and post-test) or observational (e.g. case-control, retrospective cohort, prospective cohort) study designs;
  2. Included adults aged 16+ years;
  3. Recorded as outcome i) results of a SARS-CoV-2 diagnostic test (including antibody assays), ii) clinical diagnosis of COVID-19, iii) hospitalisation with COVID-19, iv) severity of COVID-19 disease in those hospitalised or v) mortality from COVID-19;
Reported any of the outcomes of interest by self-reported or biochemically verified smoking status (e.g. current smoker, former smoker, never smoker) or current vaping and nicotine replacement therapy (NRT) use;
  4. Were available in English;
  5. Were published in a peer-reviewed journal, as a pre-print or a public health report by reputable bodies (e.g. governments, scientific societies).

*Search strategy*

The following terms were searched for in Ovid MEDLINE (2019-search date) as free text or Medical Subject Headings:

  1. Tobacco Smoking/ or Smoking Cessation/ or Water Pipe Smoking/ or Smoking/ or Smoking Pipes/ or Cigar Smoking/ or Smoking Prevention/ or Cigarette Smoking/ or smoking.mp. or Pipe Smoking/ or Smoking, Non-Tobacco Products/ or Smoking Water Pipes/
  2. Nicotine/ or nicotine.mp. or Electronic Nicotine Delivery Systems/ or Nicotine Chewing Gum/
  3. vaping.mp. or Vaping/
  4. 1 or 2 or 3
  5. Coronavirus/ or Severe Acute Respiratory Syndrome/ or Coronavirus Infections/ or covid.mp.
  6. 4 and 5

The following terms were searched for in titles, abstracts and full texts in medRxiv (no time limitations):

  1. covid (this term captures both covid and SARS-CoV-2) AND smoking
  2. covid AND nicotine
  3. covid AND vaping

Additional articles/reports of interest were identified through mailing lists, Twitter, the International Severe Acute Respiratory and Emerging Infection Consortium (ISARIC) and the US Centers for Disease Control and Prevention (CDC). Where updated versions of pre-prints or public health reports were available, old versions were superseded.

*Selection of studies*

One reviewer screened titles, abstracts and full texts against the inclusion criteria.

*Data extraction*

Data were extracted by one reviewer and verified (i.e. independently checked against pre-prints and published reports) by another on i) author (year); ii) date published; iii) country; iv) study design; v) study setting; vi) sample size; vii) sex; viii) age; ix) smoking status (e.g. current, former, never, not stated, missing) and whether it was biochemically verified; x) use of alternative nicotine products; xi) SARS-CoV-2 testing; xii) SARS-CoV-2 infection; xiii) diagnosis of COVID-19; xiv) hospitalisation with COVID-19; xv) disease severity in those hospitalised with COVID-19; xvi) mortality; xvii) adjustment of smoking specific risk estimates for relevant covariates (e.g. age, sex); and xviii) whether a representative or random sampling method was used.

*Quality appraisal*

The quality of included studies was assessed to determine suitability for inclusion in meta-analyses. Studies were judged as ‘good’ quality if they: i) had <20% missing data on smoking status and used a reliable self-report measure that distinguished between current, former and never smoking status; AND ii) used biochemical verification of smoking status and reported results from adjusted analyses; OR reported data from a representative/random sample. Studies were rated as ‘fair’ if they fulfilled only criterion i) and were otherwise rated as ‘poor’. The quality appraisal was conducted by one reviewer and verified by a second.

*Evidence synthesis*

A narrative synthesis was conducted. Data from ‘good’ and ‘fair’ quality studies were pooled in R v.3.6.3[@team_r_2013]. In a living review where new data are regularly added to the analyses, it may be more appropriate to use a Bayesian(as opposed to frequentist) approach where prior knowledge is used in combination with new data to estimate a posterior risk distribution. A Bayesian approach mitigates the issue of performing multiple statistical tests, which can inflate family-wise error. A series of random-effects hierarchical Bayesian meta-analyses were performed with the brms[@burkner_advanced_2017] package to estimate the relative risk for each comparison with accompanying 95% credible intervals (CrIs). We first defined prior distributions for the true pooled effect size (µ) and the between-study heterogeneity (τ), with µ specified as a normal distribution with a mean equal to the derived point estimate from each comparison of interest in the immediately preceding version of this living review, and τ specified as a half-Cauchy distribution with a mean of 0 and standard deviation of 1. The half-Cauchy distribution was selected to reflect prior knowledge that high levels of between-study heterogeneity are more likely than lower levels. Markov Chain Monte Carlo methods (20,000 burn-ins followed by 80,000 iterations) were then used to generate a risk distribution for each study, in addition to a pooled effect for the posterior risk distribution. We report forest plots with the pooled effect for the posterior risk distribution displayed as the median relative risk with an accompanying 95% CrIs. We used the empirical cumulative distribution function (ECDF) to estimate the probability of there being a 10% reduction or 10% increase in relative risk (RR) (i.e. RR ≥1.1 or RR ≤0.9). Due to a lack of indication as to what constitutes a clinically or epidemiologically meaningful effect (e.g. with regards to onward disease transmission or requirements for intensive care beds), we deemed a 10% change in risk as small but important. Where data were inconclusive (as indicated by CrIs crossing RR = 1.0), to disambiguate whether data favoured no effect or there being a small but important association, we estimated whether there was ≥75% probability of RR ≥1.1 or RR ≤0.9.

Two sensitivity analyses were performed. First, a minimally informative prior for µ was specified as a normal distribution with a mean of 0 and standard deviation of 1 and τ as described above. Second, an informative prior as described above for µ was used with τ specified as a half-Cauchy distribution with a mean of 0.3 and standard deviation of 1 to reflect greater between-study heterogeneity.

To aid in the visualisation of smoking prevalence in the included studies, the weighted mean prevalence of current and former smoking was calculated for countries with ≥3 studies and plotted for comparison with national prevalence estimates. It should be noted that prevalence estimates in the included studies were not adjusted for age, sex, socioeconomic position, or geographic region within countries.

##Results

In the current review version (`r paste(current_version)`) with searches up to `r paste(tail(search_details$date_screened, 1))`, a total of `r paste(prisma$"Titles/abstracts screened")` records were identified, with `r paste(prisma$"Studies in narrative synthesis")` studies included in a narrative synthesis and `r paste(running_meta_count)` studies included in meta-analyses (see [Figure 1](#fig-1)).

![*Figure 1.* PRISMA flow diagram of included studies.](figure/prisma.png){#fig-1}

*Study characteristics*

Characteristics of included studies are presented in [Table 1](#table-1). Studies were conducted across `r paste(all_countries$Country-1)` countries. `r paste(all_countries[[1]][1,2])` studies were conducted in the `r paste(all_countries[[1]][1,1])`, `r paste(all_countries[[2]][1,2])` in `r paste(all_countries[[2]][1,1])`, `r paste(all_countries[[3]][1,2])` in the `r paste(all_countries[[3]][1,1])`, `r paste(all_countries[[4]][1,2])`  in `r paste(all_countries[[4]][1,1])`, `r paste(all_countries[[5]][1,2])`  in `r paste(all_countries[[5]][1,1]))`, `r paste(all_countries[[6]][1,2]))` in `r paste(all_countries[[6]][1,1])`, `r paste(all_countries[[7]][1,2])` in `r paste(all_countries[[7]][1,1])`, `r paste(all_countries[[8]][1,2])` in `r paste(all_countries[[8]][1,1])`, `r paste(all_countries[[9]][1,2])` in `r paste(unlist(all_countries[[9]][1]), sep = " and ")`, `r paste(all_countries[[10]][1,2])` in `r paste(unlist(all_countries[[10]][1]), sep = " and ")`, `r paste(all_countries[[11]][1,2])` in `r paste(unlist(all_countries[[11]][1]), sep = " and ")`, `r paste(all_countries[[12]][1,2])` in `r paste(unlist(all_countries[[12]][1]), sep = " and ")`, with `r paste(all_countries[[13]][1,2])` each from a further `r paste(nrow(all_countries[[13]][1]))`. The majority of studies used observational designs (see [Supplementary table S1](#supplementary-t1)). `r paste(setting[1,2])` (`r paste(round(setting[1,2]/sum(setting[2])*100, 1))`%) were conducted in `r paste(tolower(setting[1,1]))` settings, `r paste(setting[2,2])` studies (`r paste(round(setting[2,2]/sum(setting[2])*100, 1))`%) included individuals from `r paste(tolower(setting[2,1]))` settings, `r paste(setting[3,2])` studies (`r paste(round(setting[3,2]/sum(setting[2])*100, 1))`%) were conducted exclusively in the `r paste(tolower(setting[3,1]))`, with one study each conducted in a homeless shelter and a quarantine centre, and one study that did not state the study setting. Studies had a median of `r paste(participants[3])` (interquartile range = `r paste(participants[2], participants[5], sep = "-")`) participants. The majority of studies (`r paste(round(sum(diagnostics[5]+diagnostics[6])/sum(diagnostics)*100,1))`%) used reverse transcriptase polymerase chain reaction (RT-PCR) for confirmation of SARS-CoV-2 infection, `r paste(round(sum(diagnostics[1]+diagnostics[6])/sum(diagnostics)*100,1))`% used an antibody test to confirm prior infection and `r paste(round(diagnostics[6]/sum(diagnostics)*100,1))` of studies relied on a combination of RT-PCR or antibody assays.

*Smoking status*

Categorisation of smoking status was heterogeneous (see [Table 1](#table-1)). `r paste(data_source[1,2])` studies collected data on smoking status through routine electronic health records (EHRs), `r paste(data_source[2,2])` studies used a bespoke case report form for COVID-19 and `r paste(data_source[3,2])` studies did not state the source for information on smoking status. None of the studies verified smoking status biochemically. Notably, only `r paste(nrow(full_smoking_status))` (`r paste(round(nrow(full_smoking_status)/current_total*100,1))`%) studies reported current, former and never smoking status (see [Supplementary table S2a](#supplementary-t2a)), with a further 26 studies reporting only ever and never smoking status (see [Supplementary table S2b](#supplementary-t2b)). The remaining `r paste(nrow(incomplete_smoking_status))` studies reported current, current/former or current and former smoking status but did not explicitly state whether remaining participants were never smokers or if data were missing on smoking status (see [Supplementary table S2c](#supplementary-t2c). `r paste(nrow(missing_smok))` studies explicitly reported the proportion with missing data on smoking status, which ranged from `r paste(round(minimum,2), "%", sep = "")` to `r paste(round(maximum,2), "%", sep = "")`.

*Use of alternative nicotine products*

Six studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use[@miyara_low_2020],[@rimland_clinical_2020],[@girardeau_confirmed_nodate],[@ebinger_sars-cov-2_2020],[@islam_risk_2020],[@gallichotte_longitudinal_2020].

*Quality appraisal*

Two studies were performed in random, representative population samples and were rated as ‘good’ quality, and `r paste(nrow(fair_quality_number))` studies were rated as ‘fair’ quality, of which `r paste(running_meta_count)` studies reported results stratified by smoking status for the outcomes of interest and could be included in meta-analyses. The remaining `r paste(nrow(poor_quality_number))` studies were rated as ‘poor’ quality (see [Table 1](#table-1)).

*Table 1.*{#table-1}
``` {r table_1}
library(flextable)

readRDS(file = here::here("reports", "tables", "table_1.rds")) %>%
  flextable() %>%
  colformat_md() %>%
  autofit()
```

*Smoking prevalence by country*

Unadjusted smoking prevalence compared with overall estimates for national adult smoking prevalence split by country and study setting is presented in [Figure 2a](#fig-2a) and [2b](#fig-2b). Lower than expected current smoking prevalence was generally observed, especially in studies with hospitalised samples. Former smoking prevalence was more similar to expected prevalence when reported; however, prevalence was typically higher than national estimates. National smoking prevalence estimates used for comparison are presented in [Supplementary table 3](#supplementary-t3).

![*Figure 2a.* Weighted mean prevalence of current smoking in included studies, split by country (solid lines). The circles represent individual studies, with colour corresponding to study setting (i.e. community, community and hospital, hospital) and size corresponding to relative study sample size. For comparison, national current smoking prevalence is shown by the dashed red lines. Countries with three or more eligible studies are shown.](figure/figure_2a.png){#fig-2a}

![*Figure 2b.* Prevalence of former smoking in included studies, split by country (solid lines). The circles represent individual studies, with colour corresponding to study setting (i.e. community, community and hospital, hospital) and size corresponding to relative study sample size. For comparison, national former smoking prevalence is shown by dashed red lines. Countries with three or more eligible studies are shown.](figure/figure_2b.png){#fig-2b}

*SARS-CoV-2 testing by smoking status*

Three studies provided data on access to SARS-CoV-2 diagnostic testing for those meeting local testing criteria by smoking status. In a cohort study of US military veterans aged 54-75[@rentsch_covid-19_2020], current smokers were more likely to receive a test: 42.3% (1,603/3,789) of the sample were current smokers compared with 23.8% of all veterans aged 50+ years using any tobacco product between 2010-201533. In the UK Biobank cohort[@niedzwiedz_ethnic_2020], a multivariable analysis showed former (RR = 1.29, 95% CI = 1.14-1.45, p < .001) and current (RR = 1.44, 95% CI = 1.20-1.71, p < .001) compared with never smokers to be more likely to receive a test. In an Australian rapid assessment screening clinic for COVID-19[@trubiano_covid-match65_2020], 9.4% (397/4,226) of the self-referred sample (subsequently assessed by a healthcare professional to decide on testing) were current smokers. Of these self-referrals, healthcare professionals decided that current compared with former or never smokers were less likely to require a test (RR = 0.93, 95% CI = 0.86-1.0, p = 0.045).

*SARS-CoV-2 infection by smoking status*

`r paste(nrow(table_2_word))` studies provided data on SARS-CoV-2 infection for people meeting local testing criteria by smoking status (see [Table 2](#table-2)). Meta-analyses were performed for `r paste(nrow(good_quality_2))` ‘good’ and `r paste(nrow(fair_quality_2))` ‘fair’ quality studies (see [Figure 3](#fig-3) and [4](#fig-4)). Current smokers were at reduced risk of testing positive for SARS-CoV-2 compared with never smokers (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_lci, 2))`). The probability of current smokers being at reduced risk of infection compared with never smokers (RR ≤0.9) was `r paste(round(m1_a_ecdf(0.9),1))`%. Former compared with never smokers were at increased risk of testing positive, but data were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_lci, 2))`) and favoured there being no important association. The probability of former smokers being at increased risk of infection (RR ≥1.1) compared with never smokers was (`r paste(1-m2_a_ecdf(1.1), "%")` probability of RR ≥1.1). Results were materially unchanged in sensitivity analyses.

*Table 2*{#table-2}
``` {r table_2}
readRDS(file = here::here("reports", "tables", "table_2.rds"))
```

![Figure 3. Forest plot for risk of testing positive for SARS-CoV-2 in current vs. never smokers. * Indicates ‘good’ quality studies. The prior from the previous review version (v9) was RR = 0.69.](figure/fig_3.png){#fig-3}

![Figure 4. Forest plot for risk of testing positive for SARS-CoV-2 in former vs. never smokers. * Indicates ‘good’ quality studies. The prior from the previous review version (v9) was RR = 1.02.](figure/fig_4.png){#fig-4}

*Hospitalisation for COVID-19 by smoking status*

`r paste(nrow(table_3_word))` studies examined hospitalisation for COVID-19 disease, stratified by smoking status (see [Table 3](#table-3)). Meta-analyses were performed for `r paste(nrow(fair_quality_3))` ‘fair’ quality studies (see [Figure 5](#fig-5) and [6](#fig-6)). Current (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_lci, 2))`) and former (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_lci, 2))`) compared with never smokers were at increased risk of hospitalisation with COVID-19. However, data for current smokers were inconclusive and favoured there being no important association. The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(1-m3_a_ecdf(1.1), "%")` and `r paste(1-m4_a_ecdf(1.1), "%")`, respectively. Results were materially unchanged in two sensitivity analyses.

*Table 3*{#table-3}
``` {r table_3}
readRDS(file = here::here("reports", "tables", "table_3.rds"))
```

![Figure 5. Forest plot for risk of hospitalisation in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.06.](figure/fig_5.png){#fig-5}

![Figure 6. Forest plot for risk of hospitalisation in former vs. never smokers. The prior from the previous version (v9) was RR = 1.17.](figure/fig_6.png){#fig-5}

*Disease severity by smoking status*

`r paste(nrow(table_4_word))` studies reported disease severity in hospitalised patients stratified by smoking status (see [Table 4](#table-4)). Severe (as opposed to non-severe) disease was broadly defined as requiring intensive treatment unit (ITU) admission, requiring oxygen as a hospital inpatient or in-hospital death. Meta-analyses were performed for `r paste(nrow(fair_quality_4))` ‘fair’ quality studies (see [Figure 7](#fig-7) and [8](#fig-8)). Current (RR = `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_lci, 2))`) and former (RR = `r paste(round(former_severity_RR_value, 2))`, 95% CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_lci, 2))`) compared with never smokers were at increased risk of greater disease severity; data for current smokers were inconclusive but favoured there being a small but important association. The probability of current and former smokers having increased risk of greater disease severity (RR ≥1.1) compared with never smokers was `r paste(1-m5_a_ecdf(1.1), "%")` and `r paste(1-m6_a_ecdf(1.1), "%")`, respectively. Results were materially unchanged in two sensitivity analyses.

*Table 4*{#table-4}
``` {r table_4}
readRDS(file = here::here("reports", "tables", "table_4.rds"))
```

![Figure 7. Forest plot for the risk of severe disease in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.26.](figure/fig_7.png){#fig-7}

![Figure 8. Forest plot for the risk of severe disease in former vs. never smokers. The prior from the previous review version (v9) was RR = 1.52.](figure/fig_8.png){#fig-8}

*Mortality by smoking status*

`r paste(nrow(table_5_word))` studies reported mortality from COVID-19 by smoking status (see [Table 5](#table-5)), with `r paste(nrow(fair_quality_5))` ‘fair’ quality studies included in meta-analyses (see [Figure 9](#fig-9) and [10](#fig-10)). Current (RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_lci, 2))`) and former (RR = `r paste(round(former_mortality_RR_value, 2))`, 95% CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_lci, 2))`) compared with never smokers were at increased risk of in-hospital mortality from COVID-19. However, data for current smokers were inconclusive and favoured there being no important association. The probability of current and former smokers being at greater risk of in-hospital mortality (RR ≥1.1) compared with never smokers was `r paste(1-m7_a_ecdf(1.1), "%")` and `r paste(1-m8_a_ecdf(1.1), "%")`, respectively. Results were materially unchanged in two sensitivity analyses.

*Table 5*{#table-5}
``` {r table_5}
readRDS(file = here::here("reports", "tables", "table_5.rds"))
```

![Figure 9. Forest plot for the risk of mortality in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.05.](figure/fig_9.png){#fig-9}

![Figure 10. Forest plot for the risk of mortality in former smokers vs. never smokers. The prior from the previous version (v9) was RR = 1.39.](figure/fig_10.png){#fig-10}

*Discussion*

This living rapid review found uncertainty in the majority of `r paste(current_total)` studies arising from the recording of smoking status. Notwithstanding these uncertainties, compared with overall adult national prevalence estimates, recorded current smoking rates in most studies were lower than expected. In a subset of better-quality studies (n = `r paste()`), current but not former smokers had a reduced risk of testing positive for SARS-CoV-2 but current smokers appeared somewhat more likely to present for testing and/or receive a test. Data for current smokers on the risk of hospitalisation, disease severity and mortality were inconclusive, and favoured there being no important associations with hospitalisation and mortality and a small but important increase in the risk of severe disease. Former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.
Issues complicating interpretation

Interpretation of results from studies conducted during the first phase of the SARS-CoV-2 pandemic is complicated by several factors (see Figure 11):

1) Exposure to SARS-CoV-2

i.	Exposure to the SARS-CoV-2 virus is heterogeneous with different subgroups at heightened risk of infection at different stages of the pandemic, at least partly due to differential contact matrices by age, sex and socioeconomic position36, which are associated with smoking status.
ii.	The probability of viral exposure depends largely on local prevalence, which varies over time. This likely introduces bias in studies assessing the rate of infection by smoking status conducted in the early phase of the pandemic.

2) Infection with SARS-CoV-2

i.	Infection following viral exposure depends on individual differences in, for example, genetic susceptibility or immunocompetence, which are poorly understood at present. For example, the household secondary attack rate for COVID-19 is estimated at 17%37.  
ii.	Heated and humidified air may act to disrupt the ability of the virus to persist in the airway mucosa of smokers. There is some evidence that transient localised hyperthermia can inhibit replication of rhinoviruses, a non-enveloped virus that causes the common cold38. However, as SARS-CoV-2 is an enveloped virus39, it is unclear whether a similar protective effect against viral replication or invasion by heated and humidified air may occur.

3) Symptomatic COVID-19

i.	An estimated 20% (95% CI = 17-25%) of COVID-19 cases are asymptomatic40, with some evidence suggesting younger people are more likely to be asymptomatic41. Testing is hence likely limited in some subgroups, with the potential for these groups to include an overrepresentation of current smokers.
ii.	Current and former smokers may be more likely to meet local criteria for community testing due to increased prevalence of symptoms consistent with SARS-CoV-2 infection, such as cough, increased sputum production or altered sense of smell or taste42. Evidence from a small number of studies indicates that current smokers may be more likely to present for testing, hence increasing the denominator in comparisons with never smokers and potentially inflating the rate of negative tests in current smokers. Infection positivity rates estimated among random samples are more informative. We identified one population study conducted in Hungary reporting on seroprevalence and smoking status43; however, the response rate was only 58.8% and the current smoking rate was 10 percentage points below national prevalence estimates, which raises some doubt about representativeness of the final sample. Similarly, a second representative population survey with results from three regions in France44 reported a current smoking rate of more than 10 percentage points below national prevalence (12% vs. 25% daily smoking prevalence)45.

4) Testing positive for SARS-CoV-2

i.	Smokers with COVID-19 may be less likely to receive a SARS-CoV-2 test or present to hospital due to lack of access to healthcare and may be more likely to die in the community from sudden complications (i.e. self-selection bias) and thus not be recorded46. 
ii.	Diagnostic criteria for SARS-CoV-2 infection and COVID-19 have changed during the course of the pandemic47. It was not possible to extract details on the specific RT-PCR or antibody-based techniques or platforms used across the included studies due to reporting gaps. Different platforms have varying sensitivity and specificity to detect SARS-CoV-2 infection. In addition, testing for acute infection requires swabbing of the mucosal epithelium, which may be disrupted in current smokers, potentially altering the sensitivity of assays48.

5) Hospitalisation with COVID-19

i.	Reasons for hospitalisation vary by country and time in the pandemic. For example, early cases may have been hospitalised for isolation and quarantine reasons and not due to medical necessity. It is plausible this may have skewed early data towards less severe cases. In addition, the observed association between former smoking and greater disease severity may be explained by collider bias49, where conditioning on a collider (e.g. testing or hospitalisation) by design or analysis may introduce a spurious association between current or former smoking (a potential cause of testing or hospitalisation) and SARS-CoV-2 infection/adverse outcomes from COVID-19 (potentially exacerbated by smoking)50.  
ii.	The majority of included studies relied on EHRs as the source of information on smoking status. Research shows large discrepancies between EHRs and actual behaviour51. Known failings of EHRs include implausible longitudinal changes, such as former smokers being recorded as never smokers at subsequent hospital visits51. Misreporting on the part of the patient (perhaps due to perceived stigma) has also been observed, with biochemical measures showing higher rates of smoking compared with self-report in hospitalised patients in the US52. It is hence possible that under-reporting of current and former smoking status in hospitals occurred across the included studies. 
iii.	The majority of included studies were conducted in hospital settings. It is plausible that a non-trivial proportion of patients were infected with SARS-CoV-2 while being an inpatient for a different medical reason. If so, this may have biased the hospitalised populations towards older and more frail groups, who are less likely to be smokers53.
iv.	Individuals with severe COVID-19 symptoms may have stopped smoking immediately before admission to hospital and may therefore not have been recorded as current smokers (i.e. reverse causality).

6) COVID-19 disease severity and death

i.	Given lack of knowledge of the disease progression and long-term outcomes of COVID-19, it is unclear whether studies conducted thus far in the pandemic have monitored patients for a sufficient time period to report complete survival outcomes or whether they are subject to early censoring. Adding to this, COVID-19 related mortality has been differentially defined across countries and epidemic phases. For example, in some UK reporting, death within 28 days of a COVID-19 diagnosis is required for attributing the cause of death to the virus. However, according to the UK Office for National Statistics, COVID-19 deaths are recorded only if this was stated on the death certificate. 
ii.	If there is a protective effect of nicotine on COVID-19 disease outcomes, abrupt nicotine withdrawal upon hospitalisation may lead to worse disease outcomes including death12.
iii.	During periods of heightened demand of limited healthcare resources, current and former smokers with extensive comorbidities may have reduced priority for intensive care admission, thus leading to higher in-hospital mortality. 
iv.	COVID-19 outcomes are currently limited to in-hospital death or survival to discharge. This binary outcome does not capture potential long-term morbidity attributed to COVID-19, such as stroke, amputation or acute cardiac events, which may be moderated by smoking status. 
Figure 11. A schematic of some of the interpretation issues for the association of smoking status and COVID-19 infection, hospitalisation, disease severity and mortality. Numbers refer to the issues listed in-text in the above section. Issues presented on the right-hand side relate directly to smoking status.

Limitations

This living rapid evidence review was limited by having a single reviewer extracting data with a second independently verifying the data extracted to minimise errors, restricting the search to one electronic database and one pre-print server and by not including at least three large population surveys due to their reliance on self-reported suspected or confirmed SARS-CoV-2 infection (which means they do not meet our eligibility criteria)42,54,55. We also did not include a large, UK-based, representative seroprevalence study56 in our meta-analyses as the odds of testing positive in former smokers was not reported. However, the odds of infection for current smokers (OR = 0.64, 95% CI = 0.58-0.71) was in concordance with the pooled estimate in our meta-analysis. Population surveys – particularly with linked data on confirmed infection or antibodies – will be included in future review versions to help mitigate some of the limitations of healthcare based observational studies. The comparisons of current and former smoking prevalence in the included studies with national prevalence estimates did not adjust observed prevalence for the demographic profile of those tested/admitted to hospital. Other reviews focused on this comparison have applied adjustments for sex and age, and continue to find lower than expected prevalence – notwithstanding the issues complicating interpretation described above17.

Implications for research, policy and practice

Further scientific research is needed to resolve the mixed findings summarised in our review. First, clinical trials of the posited therapeutic effect of nicotine could have important implications both for smokers and for improved understanding of how the SARS-CoV-2 virus causes disease in humans. Such trials should focus on medicinal nicotine (as smoked tobacco is a dirty delivery mechanism that could mask beneficial effects) and potentially differentiate between different modes of delivery (i.e. inhaled vs. ingested) since this can affect pharmacokinetics57 and potential therapeutic effects. A second research priority would be a large, representative (randomly sampled) population survey with a validated assessment of smoking status which distinguishes between recent and long-term ex-smokers – ideally biochemically verified – and assesses seroprevalence and links to health records.

In the meantime, public-facing messages about the possible protective effect of smoking or nicotine are premature. In our view, until there is further research, the quality of the evidence does not justify the huge risk associated with a message likely to reach millions of people that a lethal activity, such as smoking, may protect against COVID-19. It continues to be appropriate to recommend smoking cessation and emphasise the role of alternative nicotine products to support smokers to stop as part of public health efforts during COVID-19. At the very least, smoking cessation reduces acute risks from cardiovascular disease and could reduce demands on the healthcare system58. GPs and other healthcare providers can play a crucial role – brief, high-quality and free online training is available at National Centre for Smoking Cessation and Training.

Conclusion

Across 345 studies, recorded current but not past smoking prevalence was generally lower than national prevalence estimates. Current smokers were at reduced risk of testing positive for SARS-CoV-2 and former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.

Acknowledgements

An original short review for the Royal College of Physicians was converted to an extended living review after a request by Martin Dockrell, Tobacco Control Lead, Public Health England. All scientific decisions were made by the authors independently of funders and external organisations. The authors would like to thank Rosemary Koper for her assistance in running the electronic searches and data extraction up until v7, and all authors who responded to requests for additional data.

Declaration of conflicts of interest

DS and OP have no conflicts of interest to declare. LS has received a research grant and honoraria for a talk and travel expenses from manufacturers of smoking cessation medications (Pfizer and Johnson & Johnson). JB has received unrestricted research funding to study smoking cessation from companies who manufacture smoking cessation medications. All authors declare no financial links with tobacco companies or e-cigarette manufacturers or their representatives.

Funding statement

DS is supported by a PhD studentship from the UK Biotechnology and Biological Sciences Research Council [BB/M009513/1]. OP receives salary support from Cancer Research UK (C1417/A22962). JB, LS, & OP are members of SPECTRUM, a UK Prevention Research Partnership Consortium (MR/S037519/1). UKPRP is an initiative funded by the UK Research and Innovation Councils, the Department of Health and Social Care (England) and the UK devolved administrations, and leading health research charities.
Future review versions
https://www.qeios.com/read/latest-UJR2AW

Previous review versions
Version 1: https://doi.org/10.32388/UJR2AW 
Version 2: https://doi.org/10.32388/UJR2AW.3 
Version 3: https://doi.org/10.32388/UJR2AW.4 
Version 4: https://doi.org/10.32388/UJR2AW.5 
Version 5: https://doi.org/10.32388/UJR2AW.6
Version 6: https://doi.org/10.32388/UJR2AW.7
Version 7: https://doi.org/10.1111/add.15276 
Version 8: https://doi.org/10.32388/UJR2AW.9 
Version 9: https://doi.org/10.32388/UJR2AW.10

Data availability
All data contributing to the current and future review versions are available here: https://doi.org/10.6084/m9.figshare.12756020    
All code required to reproduce the current and future analyses are available here:  https://doi.org/10.5281/zenodo.4002046

### 2021-01-28 reports/manuscript_compile.R

@@ -1,0 +1,1 @@
+rmarkdown::render(here("reports","manuscript.Rmd"))

### 2021-01-28 scripts/bayesian_analysis.R

@@ -33,70 +33,70 @@ study_review_version <- data_study_general %>%
 levels(study_review_version$review_version) <- c(levels(study_review_version$review_version), "combined_pooled") %>%
   fct_relevel(c("previous", "current", "combined_pooled"))
 
-m1 <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1.rds"))
+#m1 <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1.rds"))
 m1_a <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1_a.rds"))
-m1_a_hh <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1_a_hh.rds"))
-m2 <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2.rds"))
+#m1_a_hh <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1_a_hh.rds"))
+#m2 <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2.rds"))
 m2_a <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2_a.rds"))
-m2_a_hh <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2_a_hh.rds"))
-m3 <- read_rds(here("data_clean", "bayesian_models", "hospitalisation_current_m3.rds"))
+#m2_a_hh <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2_a_hh.rds"))
+#m3 <- read_rds(here("data_clean", "bayesian_models", "hospitalisation_current_m3.rds"))
 m3_a <- read_rds(here("data_clean", "bayesian_models", "hospitalisation_current_m3_a.rds"))
-m3_a_hh <- read_rds(here("data_clean", "bayesian_models", "hospitalisation_current_m3_a_hh.rds"))
-m4 <- read_rds(here("data_clean", "bayesian_models", "hospitalisation_former_m4.rds"))
+#m3_a_hh <- read_rds(here("data_clean", "bayesian_models", "hospitalisation_current_m3_a_hh.rds"))
+#m4 <- read_rds(here("data_clean", "bayesian_models", "hospitalisation_former_m4.rds"))
[…]

### 2021-01-28 scripts/bayesian_data_generation.R

@@ -121,11 +121,11 @@ former_mortality_prior_hh <-
 # Run models --------------------------------------------------------------
 
 ##Testing current
-m1 <-
-  bayes_test(testing_bayes_c, minimally_informative_prior, iterations = 40000)
-pp_check(m1)
-summary(m1)
-ranef(m1)
+# m1 <-
+#   bayes_test(testing_bayes_c, minimally_informative_prior, iterations = 40000)
+# pp_check(m1)
+# summary(m1)
+# ranef(m1)
 
 m1_a <-
   bayes_test(testing_bayes_c, current_testing_prior, iterations = 40000)
@@ -133,22 +133,22 @@ pp_check(m1_a)
 summary(m1_a)
 ranef(m1_a)
 
-m1_a_hh <- 
-  bayes_test(testing_bayes_c, current_testing_prior_hh, iterations = 40000)
-pp_check(m1_a_hh)
-summary(m1_a_hh)
-ranef(m1_a_hh)
+# m1_a_hh <- 
+#   bayes_test(testing_bayes_c, current_testing_prior_hh, iterations = 40000)
+# pp_check(m1_a_hh)
+# summary(m1_a_hh)
+# ranef(m1_a_hh)
 
-write_rds(m1, here("data_clean", "bayesian_models", "testing_current_m1.rds"))
+# write_rds(m1, here("data_clean", "bayesian_models", "testing_current_m1.rds"))
 write_rds(m1_a, here("data_clean", "bayesian_models", "testing_current_m1_a.rds"))
-write_rds(m1_a_hh, here("data_clean", "bayesian_models", "testing_current_m1_a_hh.rds"))
+# write_rds(m1_a_hh, here("data_clean", "bayesian_models", "testing_current_m1_a_hh.rds"))
 
 ##Testing former
-m2 <-
-  bayes_test(testing_bayes_f, minimally_informative_prior, iterations = 40000)
-pp_check(m2)
-summary(m2)
-ranef(m2)
+# m2 <-
+#   bayes_test(testing_bayes_f, minimally_informative_prior, iterations = 40000)
+# pp_check(m2)
+# summary(m2)
+# ranef(m2)
 
 m2_a <-
   bayes_test(testing_bayes_f, former_testing_prior, iterations = 40000)
@@ -156,22 +156,22 @@ pp_check(m2_a)
 summary(m2_a)
 ranef(m2_a)
 
-m2_a_hh <-
[…]

### 2021-01-28 scripts/clean_data.R

@@ -1,0 +1,100 @@
+search_details <- search_details %>%
+  clean_data() %>%
+  rename(date_screened = date_screening_in_format_2020_05_22) %>%
+  write_rds(here::here('data_clean', 'search_details.rds')) ## The title of the date screened column is to communicate among the review team how we need date formatted
+
+protect_columns_1 <- names(data_study_general) %in% 'doi' ## As the DoI's are links we want to preserve their formatting when running the cleaning script
+
+data_study_general <- data_study_general %>%
+  rename(date_published = `date_published_format(2020-05-12)`) %>%
+  clean_data(protect = protect_columns_1) %>%
+  mutate(study_id = 1:length(lead_author)) %>% ## We create a unique study ID to aid checking of the outputs later
+  write_rds(here::here('data_clean', 'data_study_general.rds'))
+
+prisma_previous <- data_study_general
+
+review_details <- data_study_general %>%
+  select(lead_author, date_published, country, review_version, study_id) %>% ## The version of the review in which included studies were incorporated is noted
+  write_rds(here::here('data_clean', 'review_details.rds'))
+
+table_1 <- table_1 %>%
+  clean_data() %>%
+  mutate(., lower_range = sub('\\_.*', '', .$range)) %>%
+  mutate(., upper_range = sub('.*_', '',.$range,)) %>% ## We break apart the age range into an upper and lower level
+  select(lead_author, sample_size, female_sex_percent, data_source, median_age, mean_age, iqr_lower, iqr_upper, standard_deviation, lower_range,
+         upper_range, current_smoker, former_smoker, current_former_smoker, never_smoker, never_smoker_unknown, current_vaper, former_vaper,
+         current_former_vaper, not_stated, missing, total) %>%
+  mutate(study_id = 1:length(lead_author)) %>%
+  filter(!is.na(lead_author)) %>%
+  left_join(., review_details %>%
+              select(-"lead_author"), by = 'study_id') %>%
+  write_rds(here::here('data_clean', 'table_1.rds'))
+
+
+##For the following tables we're only interested in obtaining data for trials which report on the specific outcome of interest
+
+table_2 <-  table_2 %>%
+  clean_data() %>%
+  mutate(study_id = 1:length(lead_author)) %>%
+  filter(data_on_testing == TRUE) %>%
+  left_join(., review_details %>%
[…]

### 2021-01-28 scripts/libraries.R

@@ -18,3 +18,4 @@ library(linelist)
 library(DT)
 library(epitools)
 library(ggmap)
+library(gtsummary)

### 2021-01-28 scripts/references_in_table.R

@@ -1,0 +1,128 @@
+library(tidyverse)
+library(flextable)
+library(ftExtra)
+
+tbl_1 <- readRDS(here::here("data_clean", "table_1_refs.rds"))
+
+refs <- c("[@guan_clinical_2020]",
+          "[@guan_comorbidity_2020]",
+          "[@lian_analysis_nodate]",
+          "[@jin_treatment_2020]",
+          "[@chen_risk_2020]",
+          "[@zhou_clinical_2020]",
+          "[@mo_clinical_nodate]",
+          "[@zhang_clinical_2020]",
+          "[@wan_clinical_2020]",
+          "[@liu_analysis_2020]",
+          "[@huang_clinical_2020-1]",
+          "[@zhang_epidemiological_2020]",
+          "[@guo_cardiovascular_2020]",
+          "[@liu_association_2020]",
+          "[@huang_clinical_2020]",
+          "[@xu_acute_2020]",
+          "[@li_epidemiological_2020]",
+          "[@rentsch_covid-19_2020]",
+          "[@hu_risk_2020]",
+          "[@wang_epidemiological_2020]",
+          "[@cdcmmwr_preliminary_2020]",
+          "[@dong_eleven_2020]",
+          "[@kim_clinical_2020]",
+          "[@shi_host_2020]",
+          "[@yang_clinical_2020]",
+          "[@argenziano_characterization_2020]",
+          "[@solis_covid-19_2020]",
+          "[@richardson_presenting_2020]",
+          "[@fontanet_cluster_2020]",
+          "[zheng_letter_2020]",
+          "[@liao_clinical_2020]",
+          "[@gil-agudo_clinical_2020]",
+          "[@shi_clinical_2020]",
+          "[@hadjadj_impaired_2020]",
+          "[@gold_characteristics_2020]",
+          "[@yu_association_2020]",
+          "[@zheng_epidemiological_2020]",
+          "[@rica_low_2020]",
+          "[@yin_clinical_2020]",
+          "[@shi_neutrophil_2020]",
+          "[@cho_smoking_2020]",
+          "[@allenbach_multivariable_2020]",
+          "[@robilotti_determinants_2020]",
+          "[@williamson_opensafely_2020]",
+          "[@borobia_cohort_2020]",
+          "[@giacomelli_30-day_2020]",
+          "[@shah_clinical_2020]",
+          "[@kolin_clinical_2020]",
[…]

### 2021-01-28 scripts/remove_excluded.R

@@ -1,0 +1,51 @@
+data_study_general <- data_study_general %>%
+  filter(review_version %in% analysed_versions) %>%
+  filter(!(lead_author %in% exclude_from_analysis)) %>%
+  write_rds(., here::here('data_clean', 'data_study_general.rds'))
+
+review_details <- data_study_general %>%
+  select(lead_author, date_published, country, review_version, study_id) %>% 
+  filter(!(lead_author %in% exclude_from_analysis))
+
+table_1 <- table_1 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!(lead_author %in% exclude_from_analysis))
+
+table_2 <- table_2 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!(lead_author %in% exclude_from_analysis))
+
+table_3 <- table_3 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!(lead_author %in% exclude_from_analysis))
+
+table_4 <- table_4 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!(lead_author %in% exclude_from_analysis))
+
+table_5 <- table_5 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!(lead_author %in% exclude_from_analysis))
+
+table_6 <- table_6 %>%
+  filter(review_version %in% analysed_versions) %>%
+  select(-review_version) %>%
+  filter(!lead_author %in% exclude_from_analysis)
+
+a <- data_study_general %>%
+  select(lead_author, date_published, source, study_id) %>%
+  rename('Lead Author' = lead_author,
+         'Date Published' = date_published,
+         'Publication Source' = source,
+         'Study ID' = study_id)
+
+a$`Lead Author` <- to_upper_camel_case(a$`Lead Author`, sep_out = ", ")
+a$`Publication Source` <- to_title_case(a$`Publication Source`)
+a$`Publication Source` <- if_else(str_length(a$`Publication Source`) < 5,
+                                  toupper(a$`Publication Source`),
+                                  to_title_case(a$`Publication Source`))

### 2021-02-07 reports/bayesian_report.Rmd
completed manuscript

@@
 ``` {r table_1}
+#table 1 for references
+table_1_refs <- table_1 %>%
+  left_join(., data_study_general %>%
+              select(lead_author, doi),
+            by = "lead_author")
 #Updating table 1
@@
   select(-iqr) %>%
+  left_join(., table_1_refs %>%
+              select(lead_author, doi),
+            by = "lead_author") %>%
   rename('Study ID' = study_id,
@@
 
-saveRDS(a, file = here("data_clean", "table_1_refs.rds"))
-source(here("scripts", "references_in_table.r"))
+saveRDS(a %>%
+          select(doi, everything()) %>%
+          select(-"Study ID") %>%
+          rename("ref" = doi),
+        file = here("data_clean", "table_1_refs.rds"))
+#source(here("scripts", "references_in_table.r"))
 ```
@@
          "Del, Valle" = "del Valle")
+a$Author <- to_upper_camel_case(a$`Lead author`, sep_out = ", ")
+author_list <- plyr::mapvalues(a$Author,
+                               from = cleaned_names,
+                               to = correct_names)
+
+a$Author <- author_list
 
@@
 a <- flextable(a) %>%
-    set_caption(caption = 'SARS-CoV-2 infection by smoking status') %>%
     colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',')
@@
 good_quality_2 <- quality_table_2 %>%
-  filter(overall_quality != 'good')
+  filter(overall_quality == 'good')
 
 fair_quality_2 <- quality_table_2 %>%
-  filter(overall_quality != 'fair')
+  filter(overall_quality == 'fair')
 
@@
 a <- flextable(a) %>%
-    set_caption(caption = 'COVID-19 hospitalisation by smoking status') %>%
   colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
@@
 a <- flextable(a) %>%
-    set_caption(caption = 'COVID-19 severity by smoking status') %>%
   colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
@@
 a <- flextable(a) %>%
-    set_caption(caption = 'COVID-19 mortality by smoking status') %>%
   colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%

### 2021-02-07 reports/manuscript.Rmd
completed manuscript

@@
 output: 
-
   html_document:
-    code_folding: hide
     highlight: zenburn
-    number_sections: true
-    theme: spacelab
+    toc: true
+    toc_float: true
+    number_sections: false
+    theme: simplex
     css: !expr here::here('css', 'style.css')
@@
   chunk_output_type: inline
+bibliography: manuscript.bib
 ---
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,
+knitr::opts_chunk$set(echo = F,
                       warning = F,
@@
 
-#Abstract
+##    Abstract
 
@@
 
-**Findings:** `r paste(current_version)` (searches up to `r paste(tail(search_details$date_screened, 1))`) included `r paste(current_total)` studies with `r paste(running_meta_count)` ‘good’ and ‘fair’ quality studies included in unadjusted meta-analyses.
+**Findings:** `r paste(current_version)` (searches up to `r paste(tail(search_details$date_screened, 1))`) included `r paste(current_total)` studies with `r paste(nrow(running_meta_count))` ‘good’ and ‘fair’ quality studies included in unadjusted meta-analyses.
 `r paste(nrow(full_smoking_status))` studies (`r paste(round(nrow(full_smoking_status)/current_total*100,1), "%")`) reported current, former and never smoking status with the remainder using broader categories.
@@
 Current compared with never smokers were at reduced risk of SARS-CoV-2 infection (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_lci, 2))`).
-Data for former smokers were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_lci, 2))`) but favoured there being no important association (`r paste(1-m2_a_ecdf(1.1), "%")` probability of RR ≥1.1).
+Data for former smokers were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_lci, 2))`) but favoured there being no important association (`r paste(round(1-m2_a_ecdf(1.1),2), "%", sep = "")` probability of RR ≥1.1).
 Former compared with never smokers were at increased risk of hospitalisation (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_lci, 2))`), greater disease severity (RR = `r paste(round(former_severity_RR_value, 2))`, CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_lci, 2))`) and mortality (RR = `r paste(round(former_mortality_RR_value, 2))`, CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_lci, 2))`).
-Data for current smokers on hospitalisation, disease severity and mortality were inconclusive (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_lci, 2))`; RR `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_lci, 2))`; RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_lci, 2))`, respectively) but favoured there being no important associations with hospitalisation and mortality (`r paste(1-m3_a_ecdf(1.1), "%")` and `r paste(1-m5_a_ecdf(1.1), "%")` probability of RR ≥1.1, respectively) and a small but important association with disease severity (`r paste(1-m7_a_ecdf(1.1), "%")` probability of RR ≥1.1).
+Data for current smokers on hospitalisation, disease severity and mortality were inconclusive (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_lci, 2))`; RR `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_lci, 2))`; RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_lci, 2))`, respectively) but favoured there being no important associations with hospitalisation and mortality (`r paste(round(1-m3_a_ecdf(1.1),2), "%", sep = "")` and `r paste(round(1-m5_a_ecdf(1.1),2), "%", sep = "")` probability of RR ≥1.1, respectively) and a small but important association with disease severity (`r paste(round(1-m7_a_ecdf(1.1),2), "%", sep = "")` probability of RR ≥1.1).
 
-Conclusions: Compared with never smokers, current smokers appear to be at reduced risk of SARS-CoV-2 infection while former smokers appear to be at increased risk of hospitalisation, greater disease severity and mortality from COVID-19.
+**Conclusions:** Compared with never smokers, current smokers appear to be at reduced risk of SARS-CoV-2 infection while former smokers appear to be at increased risk of hospitalisation, greater disease severity and mortality from COVID-19.
 However, it is uncertain whether these associations are causal.
 
-v7 of this living review article has been published in *Addiction* and is available [here](https://doi-org.libproxy.ucl.ac.uk/10.1111/add.15276)
+v7 of this living review article has been published in [*Addiction*](https://doi-org.libproxy.ucl.ac.uk/10.1111/add.15276)
 
-##Introduction
+##    Introduction
 
 COVID-19 is a respiratory disease caused by the SARS-CoV-2 virus.
-Large age and gender differences in case severity and mortality have been observed in the ongoing COVID-19 pandemic[@guan_clinical_2020]; however, these differences are currently unexplained.
-SARS-CoV-2 enters epithelial cells through the angiotensin-converting enzyme 2 (ACE-2) receptor[@hoffmann_sars-cov-2_2020].
-Some evidence suggests that gene expression and subsequent receptor levels are elevated in the airway and oral epithelium of current smokers[@brake_smoking_2020],[@cai_bulk_2020], which could put smokers at higher risk of contracting SARS-CoV-2.
-Other studies, however, suggest that nicotine downregulates the ACE-2 receptor[@oakes_nicotine_2018].
-These uncertainties notwithstanding, both former and current smoking is known to increase the risk of respiratory viral[@denholm_hospitalised_2010],[@abadom_risk_2016] and bacterial[@almirall_proportion_1999],[@feldman_cigarette_2013] infections and is associated with worse outcomes once infected.
-Cigarette smoke reduces the respiratory immune defence through peri-bronchiolar inflammationand fibrosis, impaired mucociliary clearance and disruption of the respiratory epithelium[@dye_occasional_1994].
+Large age and gender differences in case severity and mortality have been observed in the ongoing COVID-19 pandemic [@guan_clinical_2020]; however, these differences are currently unexplained.
+SARS-CoV-2 enters epithelial cells through the angiotensin-converting enzyme 2 (ACE-2) receptor [@hoffmann_sars-cov-2_2020].
+Some evidence suggests that gene expression and subsequent receptor levels are elevated in the airway and oral epithelium of current smokers [@brake_smoking_2020; @cai_bulk_2020], which could put smokers at higher risk of contracting SARS-CoV-2.
+Other studies, however, suggest that nicotine downregulates the ACE-2 receptor [@oakes_nicotine_2018].
+These uncertainties notwithstanding, both former and current smoking is known to increase the risk of respiratory viral [@denholm_hospitalised_2010; @abadom_risk_2016] and bacterial [@almirall_proportion_1999; @feldman_cigarette_2013] infections and is associated with worse outcomes once infected.
+Cigarette smoke reduces the respiratory immune defence through peri-bronchiolar inflammationand fibrosis, impaired mucociliary clearance and disruption of the respiratory epithelium [@dye_occasional_1994].
 There is also reason to believe that behavioural factors (e.g.
 regular hand-to-mouth movements) involved in smoking may increase SARS-CoV-2 infection and transmission in current smokers.
-However, early data from the COVID-19 pandemic have not provided clear evidence for a negative impact of current or former smoking on SARS-CoV-2 infection or COVID-19 disease outcomes, such as hospitalisation or mortality[@vardavas_covid-19_2020].
-It has also been hypothesised that nicotine might protect against a hyper-inflammatory response to SARS-CoV-2 infection, which may lead to adverse outcomes in patients with COVID-19 disease[@farsalinos_editorial_2020].
+However, early data from the COVID-19 pandemic have not provided clear evidence for a negative impact of current or former smoking on SARS-CoV-2 infection or COVID-19 disease outcomes, such as hospitalisation or mortality [@vardavas_covid-19_2020].
+It has also been hypothesised that nicotine might protect against a hyper-inflammatory response to SARS-CoV-2 infection, which may lead to adverse outcomes in patients with COVID-19 disease [@farsalinos_editorial_2020].
 
-There are several reviews that fall within the scope of smoking and COVID-19[@vardavas_covid-19_2020],[@emami_prevalence_2020],[@alqahtani_prevalence_2020],[@patanavanich_smoking_2020],[@berlin_covid-19_0],[@farsalinos_systematic_2020],[@grundy_smoking_2020].
+There are several reviews that fall within the scope of smoking and COVID-19 [@vardavas_covid-19_2020],[@emami_prevalence_2020; @alqahtani_prevalence_2020; @patanavanich_smoking_2020; @berlin_covid-19_2020; @farsalinos_systematic_2020; @grundy_smoking_2020].
 We aimed to produce a rapid synthesis of available evidence pertaining to the rates of infection, hospitalisation, disease severity and mortality from SARS-CoV-2/COVID-19 stratified by smoking status.
@@
 
-##Methods
+##    Methods
 
@@
 
-This is a living evidence review  which is updated as new evidence becomes available[@elliott_living_2014].
-We adopted recommended best practice for rapid evidence reviews, which involved limiting the search to main databases and having one reviewer extract the data and another verify[@tricco_scoping_2015].
-This study was not pre-registered but evolved from a report written for a UK medical society[@simons_smoking_2020].
-The most recent (and all future) version(s) of this living review is available [here](https://www.qeios.com/read/latest-UJR2AW).
-Version 7 of this living review has been published in a peer-reviewed journal[@simons_association_nodate].
+This is a living evidence review  which is updated as new evidence becomes available [@elliott_living_2014].
+We adopted recommended best practice for rapid evidence reviews, which involved limiting the search to main databases and having one reviewer extract the data and another verify [@tricco_scoping_2015].
+This study was not pre-registered but evolved from a report written for a UK medical society [@simons_smoking_2020].
+The most recent version of this living review is available [here](https://www.qeios.com/read/latest-UJR2AW).
+Version 7 of this living review has been published in a peer-reviewed journal [@simons_association_2020-1].
 A completed Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) checklist is included in Supplementary file 1.
@@
 A narrative synthesis was conducted.
-Data from ‘good’ and ‘fair’ quality studies were pooled in R v.3.6.3[@team_r_2013].
+Data from ‘good’ and ‘fair’ quality studies were pooled in R v.3.6.3 [@team_r_2013].
 In a living review where new data are regularly added to the analyses, it may be more appropriate to use a Bayesian(as opposed to frequentist) approach where prior knowledge is used in combination with new data to estimate a posterior risk distribution.
 A Bayesian approach mitigates the issue of performing multiple statistical tests, which can inflate family-wise error.
-A series of random-effects hierarchical Bayesian meta-analyses were performed with the brms[@burkner_advanced_2017] package to estimate the relative risk for each comparison with accompanying 95% credible intervals (CrIs).
+A series of random-effects hierarchical Bayesian meta-analyses were performed with the brms [@burkner_advanced_2018] package to estimate the relative risk for each comparison with accompanying 95% credible intervals (CrIs).
 We first defined prior distributions for the true pooled effect size (µ) and the between-study heterogeneity (τ), with µ specified as a normal distribution with a mean equal to the derived point estimate from each comparison of interest in the immediately preceding version of this living review, and τ specified as a half-Cauchy distribution with a mean of 0 and standard deviation of 1.
@@
 
-##Results
+##    Results
 
-In the current review version (`r paste(current_version)`) with searches up to `r paste(tail(search_details$date_screened, 1))`, a total of `r paste(prisma$"Titles/abstracts screened")` records were identified, with `r paste(prisma$"Studies in narrative synthesis")` studies included in a narrative synthesis and `r paste(running_meta_count)` studies included in meta-analyses (see [Figure 1](#fig-1)).
+In the current review version (`r paste(current_version)`) with searches up to `r paste(tail(search_details$date_screened, 1))`, a total of `r paste(prisma$"Titles/abstracts screened")` records were identified, with `r paste(prisma$"Studies in narrative synthesis")` studies included in a narrative synthesis and `r paste(nrow(running_meta_count))` studies included in meta-analyses (see [Figure 1](#fig-1)).
 
@@
 
-*Study characteristics*
-
 Characteristics of included studies are presented in [Table 1](#table-1).
-Studies were conducted across `r paste(all_countries$Country-1)` countries.
-`r paste(all_countries[[1]][1,2])` studies were conducted in the `r paste(all_countries[[1]][1,1])`, `r paste(all_countries[[2]][1,2])` in `r paste(all_countries[[2]][1,1])`, `r paste(all_countries[[3]][1,2])` in the `r paste(all_countries[[3]][1,1])`, `r paste(all_countries[[4]][1,2])`  in `r paste(all_countries[[4]][1,1])`, `r paste(all_countries[[5]][1,2])`  in `r paste(all_countries[[5]][1,1]))`, `r paste(all_countries[[6]][1,2]))` in `r paste(all_countries[[6]][1,1])`, `r paste(all_countries[[7]][1,2])` in `r paste(all_countries[[7]][1,1])`, `r paste(all_countries[[8]][1,2])` in `r paste(all_countries[[8]][1,1])`, `r paste(all_countries[[9]][1,2])` in `r paste(unlist(all_countries[[9]][1]), sep = " and ")`, `r paste(all_countries[[10]][1,2])` in `r paste(unlist(all_countries[[10]][1]), sep = " and ")`, `r paste(all_countries[[11]][1,2])` in `r paste(unlist(all_countries[[11]][1]), sep = " and ")`, `r paste(all_countries[[12]][1,2])` in `r paste(unlist(all_countries[[12]][1]), sep = " and ")`, with `r paste(all_countries[[13]][1,2])` each from a further `r paste(nrow(all_countries[[13]][1]))`.
+Studies were conducted across `r paste(nrow(country))` countries.
+`r paste(all_countries[[1]][1,2])` studies were conducted in the `r paste(all_countries[[1]][1,1])`, `r paste(all_countries[[2]][1,2])` in `r paste(all_countries[[2]][1,1])`, `r paste(all_countries[[3]][1,2])` in the `r paste(all_countries[[3]][1,1])`, `r paste(all_countries[[4]][1,2])`  in `r paste(all_countries[[4]][1,1])`, `r paste(all_countries[[5]][1,2])`  in `r paste(all_countries[[5]][1,1])`, `r paste(all_countries[[6]][1,2])` in `r paste(all_countries[[6]][1,1])`, `r paste(all_countries[[7]][1,2])` in `r paste(all_countries[[7]][1,1])`, `r paste(all_countries[[8]][1,2])` in `r paste(all_countries[[8]][1,1])`, `r paste(all_countries[[9]][1,2])` in `r knitr::combine_words(unlist(all_countries[[9]][1]))`, `r paste(all_countries[[10]][1,2])` in `r knitr::combine_words(unlist(all_countries[[10]][1]))`, `r paste(all_countries[[11]][1,2])` in `r knitr::combine_words(unlist(all_countries[[11]][1]))`, `r paste(all_countries[[12]][1,2])` in `r knitr::combine_words(unlist(all_countries[[12]][1]))`, with `r paste(all_countries[[13]][1,2])` each from `r knitr::combine_words(unlist(all_countries[[13]][1]))` and a single study from `r paste(nrow(all_countries[[14]][1]))` further countries.
 The majority of studies used observational designs (see [Supplementary table S1](#supplementary-t1)).
@@
 
-*Smoking status*
+###   Smoking status
 
@@
 Notably, only `r paste(nrow(full_smoking_status))` (`r paste(round(nrow(full_smoking_status)/current_total*100,1))`%) studies reported current, former and never smoking status (see [Supplementary table S2a](#supplementary-t2a)), with a further 26 studies reporting only ever and never smoking status (see [Supplementary table S2b](#supplementary-t2b)).
-The remaining `r paste(nrow(incomplete_smoking_status))` studies reported current, current/former or current and former smoking status but did not explicitly state whether remaining participants were never smokers or if data were missing on smoking status (see [Supplementary table S2c](#supplementary-t2c).
+The remaining `r paste(nrow(incomplete_smoking_status))` studies reported current, current/former or current and former smoking status but did not explicitly state whether remaining participants were never smokers or if data were missing on smoking status (see [Supplementary table S2c](#supplementary-t2c)).
 `r paste(nrow(missing_smok))` studies explicitly reported the proportion with missing data on smoking status, which ranged from `r paste(round(minimum,2), "%", sep = "")` to `r paste(round(maximum,2), "%", sep = "")`.
 
-*Use of alternative nicotine products*
+###   Use of alternative nicotine products
 
-Six studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use[@miyara_low_2020],[@rimland_clinical_2020],[@girardeau_confirmed_nodate],[@ebinger_sars-cov-2_2020],[@islam_risk_2020],[@gallichotte_longitudinal_2020].
+Six studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use [@miyara_low_2020; @rimland_clinical_2020; @girardeau_confirmed_2020; @ebinger_sars-cov-2_2020; @islam_risk_2020; @gallichotte_longitudinal_2020].
 
-*Quality appraisal*
+###   Quality appraisal
 
-Two studies were performed in random, representative population samples and were rated as ‘good’ quality, and `r paste(nrow(fair_quality_number))` studies were rated as ‘fair’ quality, of which `r paste(running_meta_count)` studies reported results stratified by smoking status for the outcomes of interest and could be included in meta-analyses.
+Two studies were performed in random, representative population samples and were rated as ‘good’ quality, and `r paste(nrow(fair_quality_number))` studies were rated as ‘fair’ quality, of which `r paste(nrow(running_meta_count))` studies reported results stratified by smoking status for the outcomes of interest and could be included in meta-analyses.
 The remaining `r paste(nrow(poor_quality_number))` studies were rated as ‘poor’ quality (see [Table 1](#table-1)).
 
-*Table 1.*{#table-1}
+#### [*Table 1.* All studies included in narrative review and meta-analysis]{#table-1}
 ``` {r table_1}
 library(flextable)
+library(ftExtra)
 
-readRDS(file = here::here("reports", "tables", "table_1.rds")) %>%
-  flextable() %>%
-  colformat_md() %>%
-  autofit()
+tbl_1 <- readRDS(file = here::here("reports", "tables", "table_1_refs.rds"))
+
+flextable(tbl_1) %>%
+  colformat_md(j = "ref") %>%
+  colformat_double(j = "Sample size")
 ```
 
-*Smoking prevalence by country*
+###   Smoking prevalence by country
 
@@
 For comparison, national current smoking prevalence is shown by the dashed red lines.
-Countries with three or more eligible studies are shown.](figure/figure_2a.png){#fig-2a}
+Countries with three or more eligible studies are shown.](figure/current_smoking_plots_updated.png){#fig-2a}
 
@@
 For comparison, national former smoking prevalence is shown by dashed red lines.
-Countries with three or more eligible studies are shown.](figure/figure_2b.png){#fig-2b}
+Countries with three or more eligible studies are shown.](figure/former_smoking_plots_updated.png){#fig-2b}
 
-*SARS-CoV-2 testing by smoking status*
+###   SARS-CoV-2 testing by smoking status
 
 Three studies provided data on access to SARS-CoV-2 diagnostic testing for those meeting local testing criteria by smoking status.
-In a cohort study of US military veterans aged 54-75[@rentsch_covid-19_2020], current smokers were more likely to receive a test: 42.3% (1,603/3,789) of the sample were current smokers compared with 23.8% of all veterans aged 50+ years using any tobacco product between 2010-201533.
-In the UK Biobank cohort[@niedzwiedz_ethnic_2020], a multivariable analysis showed former (RR = 1.29, 95% CI = 1.14-1.45, p < .001) and current (RR = 1.44, 95% CI = 1.20-1.71, p < .001) compared with never smokers to be more likely to receive a test.
-In an Australian rapid assessment screening clinic for COVID-19[@trubiano_covid-match65_2020], 9.4% (397/4,226) of the self-referred sample (subsequently assessed by a healthcare professional to decide on testing) were current smokers.
-Of these self-referrals, healthcare professionals decided that current compared with former or never smokers were less likely to require a test (RR = 0.93, 95% CI = 0.86-1.0, p = 0.045).
+In a cohort study of US military veterans aged 54-75 [@rentsch_covid-19_2020], current smokers were more likely to receive a test: 42.3% (1,603/3,789) of the sample were current smokers compared with 23.8% of all veterans aged 50+ years using any tobacco product between 2010-201533.
+In the UK Biobank cohort [@niedzwiedz_ethnic_2020], a multivariable analysis showed former (RR = 1.29, 95% CI = 1.14-1.45, *p* < .001) and current (RR = 1.44, 95% CI = 1.20-1.71, *p* < .001) compared with never smokers to be more likely to receive a test.
+In an Australian rapid assessment screening clinic for COVID-19 [@trubiano_covid-match65_2020], 9.4% (397/4,226) of the self-referred sample (subsequently assessed by a healthcare professional to decide on testing) were current smokers.
+Of these self-referrals, healthcare professionals decided that current compared with former or never smokers were less likely to require a test (RR = 0.93, 95% CI = 0.86-1.0, *p* = 0.045).
 
-*SARS-CoV-2 infection by smoking status*
+###   SARS-CoV-2 infection by smoking status
 
@@
 Current smokers were at reduced risk of testing positive for SARS-CoV-2 compared with never smokers (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_lci, 2))`).
-The probability of current smokers being at reduced risk of infection compared with never smokers (RR ≤0.9) was `r paste(round(m1_a_ecdf(0.9),1))`%.
+The probability of current smokers being at reduced risk of infection compared with never smokers (RR ≤0.9) was `r paste(round(m1_a_ecdf(0.9),2))`%.
 Former compared with never smokers were at increased risk of testing positive, but data were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_lci, 2))`) and favoured there being no important association.
-The probability of former smokers being at increased risk of infection (RR ≥1.1) compared with never smokers was (`r paste(1-m2_a_ecdf(1.1), "%")` probability of RR ≥1.1).
+The probability of former smokers being at increased risk of infection (RR ≥1.1) compared with never smokers was (`r paste(round(1-m2_a_ecdf(1.1), 2), "%")` probability of RR ≥1.1).
 Results were materially unchanged in sensitivity analyses.
 
-*Table 2*{#table-2}
+####    [*Table 2* SARS-CoV-2 infection by smoking status]{#table-2}
 ``` {r table_2}
@@
 
-![Figure 3.
-Forest plot for risk of testing positive for SARS-CoV-2 in current vs.
+![*Figure 3.* Forest plot for risk of testing positive for SARS-CoV-2 in current vs.
 never smokers.
 * Indicates ‘good’ quality studies.
-The prior from the previous review version (v9) was RR = 0.69.](figure/fig_3.png){#fig-3}
+The prior from the previous review version (v9) was RR = 0.69.](figure/m1_a.png){#fig-3}
 
-![Figure 4.
-Forest plot for risk of testing positive for SARS-CoV-2 in former vs.
+![*Figure 4.* Forest plot for risk of testing positive for SARS-CoV-2 in former vs.
 never smokers.
 * Indicates ‘good’ quality studies.
-The prior from the previous review version (v9) was RR = 1.02.](figure/fig_4.png){#fig-4}
+The prior from the previous review version (v9) was RR = 1.02.](figure/m2_a.png){#fig-4}
 
-*Hospitalisation for COVID-19 by smoking status*
+###    Hospitalisation for COVID-19 by smoking status
 
@@
 However, data for current smokers were inconclusive and favoured there being no important association.
-The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(1-m3_a_ecdf(1.1), "%")` and `r paste(1-m4_a_ecdf(1.1), "%")`, respectively.
+The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(round(1-m3_a_ecdf(1.1),2),"%", sep = "")` and `r paste(round(1-m4_a_ecdf(1.1),2), "%", sep = "")`, respectively.
 Results were materially unchanged in two sensitivity analyses.
 
-*Table 3*{#table-3}
+####    [*Table 3* COVID-19 hospitalsation by smoking status]{#table-3}
 ``` {r table_3}
@@
 
-![Figure 5. Forest plot for risk of hospitalisation in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.06.](figure/fig_5.png){#fig-5}
+![*Figure 5.* Forest plot for risk of hospitalisation in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.06.](figure/m3_a.png){#fig-5}
 
-![Figure 6. Forest plot for risk of hospitalisation in former vs. never smokers. The prior from the previous version (v9) was RR = 1.17.](figure/fig_6.png){#fig-5}
+![*Figure 6.* Forest plot for risk of hospitalisation in former vs. never smokers. The prior from the previous version (v9) was RR = 1.17.](figure/m4_a.png){#fig-5}
 
-*Disease severity by smoking status*
+###   Disease severity by smoking status
 
@@
 Current (RR = `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_lci, 2))`) and former (RR = `r paste(round(former_severity_RR_value, 2))`, 95% CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_lci, 2))`) compared with never smokers were at increased risk of greater disease severity; data for current smokers were inconclusive but favoured there being a small but important association.
-The probability of current and former smokers having increased risk of greater disease severity (RR ≥1.1) compared with never smokers was `r paste(1-m5_a_ecdf(1.1), "%")` and `r paste(1-m6_a_ecdf(1.1), "%")`, respectively.
+The probability of current and former smokers having increased risk of greater disease severity (RR ≥1.1) compared with never smokers was `r paste(round(1-m5_a_ecdf(1.1),2), "%", sep = "")` and `r paste(round(1-m6_a_ecdf(1.1),2), "%", sep = "")`, respectively.
 Results were materially unchanged in two sensitivity analyses.
 
-*Table 4*{#table-4}
+####    [*Table 4* COVID-19 disease severity by smoking status]{#table-4}
 ``` {r table_4}
@@
 
-![Figure 7. Forest plot for the risk of severe disease in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.26.](figure/fig_7.png){#fig-7}
+![*Figure 7.* Forest plot for the risk of severe disease in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.26.](figure/m5_a.png){#fig-7}
 
-![Figure 8. Forest plot for the risk of severe disease in former vs. never smokers. The prior from the previous review version (v9) was RR = 1.52.](figure/fig_8.png){#fig-8}
+![*Figure 8.* Forest plot for the risk of severe disease in former vs. never smokers. The prior from the previous review version (v9) was RR = 1.52.](figure/m6_a.png){#fig-8}
 
-*Mortality by smoking status*
+###   Mortality by smoking status
 
@@
 
-*Table 5*{#table-5}
+####    [*Table 5* COVID-19 attributed mortality by smoking status]{#table-5}
 ``` {r table_5}
@@
 
-![Figure 9. Forest plot for the risk of mortality in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.05.](figure/fig_9.png){#fig-9}
+![Figure 9. Forest plot for the risk of mortality in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.05.](figure/m7_a.png){#fig-9}
 
-![Figure 10. Forest plot for the risk of mortality in former smokers vs. never smokers. The prior from the previous version (v9) was RR = 1.39.](figure/fig_10.png){#fig-10}
+![Figure 10. Forest plot for the risk of mortality in former smokers vs. never smokers. The prior from the previous version (v9) was RR = 1.39.](figure/m8_a.png){#fig-10}
 
-*Discussion*
+##    Discussion
 
@@
 Notwithstanding these uncertainties, compared with overall adult national prevalence estimates, recorded current smoking rates in most studies were lower than expected.
-In a subset of better-quality studies (n = `r paste()`), current but not former smokers had a reduced risk of testing positive for SARS-CoV-2 but current smokers appeared somewhat more likely to present for testing and/or receive a test.
+In a subset of better-quality studies (n = `r paste(nrow(good_quality_2)+nrow(fair_quality_2))`), current but not former smokers had a reduced risk of testing positive for SARS-CoV-2 but current smokers appeared somewhat more likely to present for testing and/or receive a test.
 Data for current smokers on the risk of hospitalisation, disease severity and mortality were inconclusive, and favoured there being no important associations with hospitalisation and mortality and a small but important increase in the risk of severe disease.
@@
 
-Interpretation of results from studies conducted during the first phase of the SARS-CoV-2 pandemic is complicated by several factors (see Figure 11):
+Interpretation of results from studies conducted during the first phase of the SARS-CoV-2 pandemic is complicated by several factors (see [Figure 11](#fig-11)):
 
-1) Exposure to SARS-CoV-2
-
-i.	Exposure to the SARS-CoV-2 virus is heterogeneous with different subgroups at heightened risk of infection at different stages of the pandemic, at least partly due to differential contact matrices by age, sex and socioeconomic position36, which are associated with smoking status.
-ii.	The probability of viral exposure depends largely on local prevalence, which varies over time.
+  1. *Exposure to SARS-CoV-2*   
+    1.	Exposure to the SARS-CoV-2 virus is heterogeneous with different subgroups at heightened risk of infection at different stages of the pandemic, at least partly due to differential contact matrices by age, sex and socioeconomic position [@cmmid_covid-19_working_group_quantifying_2020], which are associated with smoking status.
+    2.	The probability of viral exposure depends largely on local prevalence, which varies over time.
 This likely introduces bias in studies assessing the rate of infection by smoking status conducted in the early phase of the pandemic.
-
-2) Infection with SARS-CoV-2
-
-i.	Infection following viral exposure depends on individual differences in, for example, genetic susceptibility or immunocompetence, which are poorly understood at present.
-For example, the household secondary attack rate for COVID-19 is estimated at 17%37.
-ii.	Heated and humidified air may act to disrupt the ability of the virus to persist in the airway mucosa of smokers.
-There is some evidence that transient localised hyperthermia can inhibit replication of rhinoviruses, a non-enveloped virus that causes the common cold38.
-However, as SARS-CoV-2 is an enveloped virus39, it is unclear whether a similar protective effect against viral replication or invasion by heated and humidified air may occur.
-
-3) Symptomatic COVID-19
-
-i.	An estimated 20% (95% CI = 17-25%) of COVID-19 cases are asymptomatic40, with some evidence suggesting younger people are more likely to be asymptomatic41.
+  2. *Infection with SARS-CoV-2*    
+    1.	Infection following viral exposure depends on individual differences in, for example, genetic susceptibility or immunocompetence, which are poorly understood at present.
+For example, the household secondary attack rate for COVID-19 is estimated at 17% [@fung_household_2020].
+    2.	Heated and humidified air may act to disrupt the ability of the virus to persist in the airway mucosa of smokers.
+There is some evidence that transient localised hyperthermia can inhibit replication of rhinoviruses, a non-enveloped virus that causes the common cold [@conti_antiviral_1999].
+However, as SARS-CoV-2 is an enveloped virus [@schoeman_coronavirus_2019], it is unclear whether a similar protective effect against viral replication or invasion by heated and humidified air may occur.
+  3. *Symptomatic COVID-19*    
+    1.	An estimated 20% (95% CI = 17-25%) of COVID-19 cases are asymptomatic [@buitrago-garcia_occurrence_2020], with some evidence suggesting younger people are more likely to be asymptomatic [@kronbichler_asymptomatic_2020].
 Testing is hence likely limited in some subgroups, with the potential for these groups to include an overrepresentation of current smokers.
-ii.	Current and former smokers may be more likely to meet local criteria for community testing due to increased prevalence of symptoms consistent with SARS-CoV-2 infection, such as cough, increased sputum production or altered sense of smell or taste42.
+    2.	Current and former smokers may be more likely to meet local criteria for community testing due to increased prevalence of symptoms consistent with SARS-CoV-2 infection, such as cough, increased sputum production or altered sense of smell or taste [@hopkinson_current_2020].
 Evidence from a small number of studies indicates that current smokers may be more likely to present for testing, hence increasing the denominator in comparisons with never smokers and potentially inflating the rate of negative tests in current smokers.
 Infection positivity rates estimated among random samples are more informative.
-We identified one population study conducted in Hungary reporting on seroprevalence and smoking status43; however, the response rate was only 58.8% and the current smoking rate was 10 percentage points below national prevalence estimates, which raises some doubt about representativeness of the final sample.
-Similarly, a second representative population survey with results from three regions in France44 reported a current smoking rate of more than 10 percentage points below national prevalence (12% vs.
-25% daily smoking prevalence)45.
-
-4) Testing positive for SARS-CoV-2
-
-i.	Smokers with COVID-19 may be less likely to receive a SARS-CoV-2 test or present to hospital due to lack of access to healthcare and may be more likely to die in the community from sudden complications (i.e.
-self-selection bias) and thus not be recorded46.
-ii.	Diagnostic criteria for SARS-CoV-2 infection and COVID-19 have changed during the course of the pandemic47.
+We identified one population study conducted in Hungary reporting on seroprevalence and smoking status [@merkely_novel_2020]; however, the response rate was only 58.8% and the current smoking rate was 10 percentage points below national prevalence estimates, which raises some doubt about representativeness of the final sample.
+Similarly, a second representative population survey with results from three regions in France [@carrat_seroprevalence_2020] reported a current smoking rate of more than 10 percentage points below national prevalence (12% vs.
+25% daily smoking prevalence) [@andler_baisse_2019].
+  4. *Testing positive for SARS-CoV-2*      
+    1.
+Smokers with COVID-19 may be less likely to receive a SARS-CoV-2 test or present to hospital due to lack of access to healthcare and may be more likely to die in the community from sudden complications (i.e.
+self-selection bias) and thus not be recorded [@noauthor_sts_nodate].
+    2.	Diagnostic criteria for SARS-CoV-2 infection and COVID-19 have changed during the course of the pandemic [@organisation_laboratory_nodate].
 It was not possible to extract details on the specific RT-PCR or antibody-based techniques or platforms used across the included studies due to reporting gaps.
 Different platforms have varying sensitivity and specificity to detect SARS-CoV-2 infection.
-In addition, testing for acute infection requires swabbing of the mucosal epithelium, which may be disrupted in current smokers, potentially altering the sensitivity of assays48.
-
-5) Hospitalisation with COVID-19
-
-i.	Reasons for hospitalisation vary by country and time in the pandemic.
+In addition, testing for acute infection requires swabbing of the mucosal epithelium, which may be disrupted in current smokers, potentially altering the sensitivity of assays [@de_lusignan_risk_2020].
+  5. *Hospitalisation with COVID-19*      
+    1.	Reasons for hospitalisation vary by country and time in the pandemic.
 For example, early cases may have been hospitalised for isolation and quarantine reasons and not due to medical necessity.
@@
 In addition, the observed association between former smoking and greater disease severity may be explained by collider bias49, where conditioning on a collider (e.g.
-testing or hospitalisation) by design or analysis may introduce a spurious association between current or former smoking (a potential cause of testing or hospitalisation) and SARS-CoV-2 infection/adverse outcomes from COVID-19 (potentially exacerbated by smoking)50.
-ii.	The majority of included studies relied on EHRs as the source of information on smoking status.
-Research shows large discrepancies between EHRs and actual behaviour51.
+testing or hospitalisation) by design or analysis may introduce a spurious association between current or former smoking (a potential cause of testing or hospitalisation) and SARS-CoV-2 infection/adverse outcomes from COVID-19 (potentially exacerbated by smoking) [@murray_causation_2020].
+    2.
+The majority of included studies relied on EHRs as the source of information on smoking status.
+Research shows large discrepancies between EHRs and actual behaviour [@polubriaginof_challenges_2018].
 Known failings of EHRs include implausible longitudinal changes, such as former smokers being recorded as never smokers at subsequent hospital visits51.
-Misreporting on the part of the patient (perhaps due to perceived stigma) has also been observed, with biochemical measures showing higher rates of smoking compared with self-report in hospitalised patients in the US52.
+Misreporting on the part of the patient (perhaps due to perceived stigma) has also been observed, with biochemical measures showing higher rates of smoking compared with self-report in hospitalised patients in the US [@benowitz_prevalence_2009].
 It is hence possible that under-reporting of current and former smoking status in hospitals occurred across the included studies.
-iii.	The majority of included studies were conducted in hospital settings.
+    3.	The majority of included studies were conducted in hospital settings.
 It is plausible that a non-trivial proportion of patients were infected with SARS-CoV-2 while being an inpatient for a different medical reason.
-If so, this may have biased the hospitalised populations towards older and more frail groups, who are less likely to be smokers53.
-iv.	Individuals with severe COVID-19 symptoms may have stopped smoking immediately before admission to hospital and may therefore not have been recorded as current smokers (i.e. reverse causality).
-
-6) COVID-19 disease severity and death
-
-i.	Given lack of knowledge of the disease progression and long-term outcomes of COVID-19, it is unclear whether studies conducted thus far in the pandemic have monitored patients for a sufficient time period to report complete survival outcomes or whether they are subject to early censoring.
+If so, this may have biased the hospitalised populations towards older and more frail groups, who are less likely to be smokers [@mangera_smoking_2017].
+    4.	Individuals with severe COVID-19 symptoms may have stopped smoking immediately before admission to hospital and may therefore not have been recorded as current smokers (i.e. reverse causality).
+  6. *COVID-19 disease severity and death*    
+    1.	Given lack of knowledge of the disease progression and long-term outcomes of COVID-19, it is unclear whether studies conducted thus far in the pandemic have monitored patients for a sufficient time period to report complete survival outcomes or whether they are subject to early censoring.
 Adding to this, COVID-19 related mortality has been differentially defined across countries and epidemic phases.
@@
 However, according to the UK Office for National Statistics, COVID-19 deaths are recorded only if this was stated on the death certificate.
-ii.	If there is a protective effect of nicotine on COVID-19 disease outcomes, abrupt nicotine withdrawal upon hospitalisation may lead to worse disease outcomes including death12.
-iii.	During periods of heightened demand of limited healthcare resources, current and former smokers with extensive comorbidities may have reduced priority for intensive care admission, thus leading to higher in-hospital mortality.
-iv.	COVID-19 outcomes are currently limited to in-hospital death or survival to discharge.
+    2.
+If there is a protective effect of nicotine on COVID-19 disease outcomes, abrupt nicotine withdrawal upon hospitalisation may lead to worse disease outcomes including death [@farsalinos_editorial_2020].
+    3.	During periods of heightened demand of limited healthcare resources, current and former smokers with extensive comorbidities may have reduced priority for intensive care admission, thus leading to higher in-hospital mortality.
+    4.	COVID-19 outcomes are currently limited to in-hospital death or survival to discharge.
 This binary outcome does not capture potential long-term morbidity attributed to COVID-19, such as stroke, amputation or acute cardiac events, which may be moderated by smoking status.
-Figure 11.
-A schematic of some of the interpretation issues for the association of smoking status and COVID-19 infection, hospitalisation, disease severity and mortality.
+
+![*Figure 11.* A schematic of some of the interpretation issues for the association of smoking status and COVID-19 infection, hospitalisation, disease severity and mortality.
 Numbers refer to the issues listed in-text in the above section.
-Issues presented on the right-hand side relate directly to smoking status.
+Issues presented on the right-hand side relate directly to smoking status.](figure/fig_11.png){#fig-11}
 
-Limitations
+###   Limitations
 
-This living rapid evidence review was limited by having a single reviewer extracting data with a second independently verifying the data extracted to minimise errors, restricting the search to one electronic database and one pre-print server and by not including at least three large population surveys due to their reliance on self-reported suspected or confirmed SARS-CoV-2 infection (which means they do not meet our eligibility criteria)42,54,55.
-We also did not include a large, UK-based, representative seroprevalence study56 in our meta-analyses as the odds of testing positive in former smokers was not reported.
+This living rapid evidence review was limited by having a single reviewer extracting data with a second independently verifying the data extracted to minimise errors, restricting the search to one electronic database and one pre-print server and by not including at least three large population surveys due to their reliance on self-reported suspected or confirmed SARS-CoV-2 infection (which means they do not meet our eligibility criteria) [@hopkinson_current_2020; @bowyer_geo-social_2020; @jackson_covid-19_2020].
+We also did not include a large, UK-based, representative seroprevalence study [@ward_antibody_2020] in our meta-analyses as the odds of testing positive in former smokers was not reported.
 However, the odds of infection for current smokers (OR = 0.64, 95% CI = 0.58-0.71) was in concordance with the pooled estimate in our meta-analysis.
@@
 
-Implications for research, policy and practice
+###   Implications for research, policy and practice
 
@@
 inhaled vs.
-ingested) since this can affect pharmacokinetics57 and potential therapeutic effects.
+ingested) since this can affect pharmacokinetics [@shahab_novel_2013] and potential therapeutic effects.
 A second research priority would be a large, representative (randomly sampled) population survey with a validated assessment of smoking status which distinguishes between recent and long-term ex-smokers – ideally biochemically verified – and assesses seroprevalence and links to health records.
@@
 It continues to be appropriate to recommend smoking cessation and emphasise the role of alternative nicotine products to support smokers to stop as part of public health efforts during COVID-19.
-At the very least, smoking cessation reduces acute risks from cardiovascular disease and could reduce demands on the healthcare system58.
-GPs and other healthcare providers can play a crucial role – brief, high-quality and free online training is available at National Centre for Smoking Cessation and Training.
+At the very least, smoking cessation reduces acute risks from cardiovascular disease and could reduce demands on the healthcare system [@stead_physician_2013].
+GPs and other healthcare providers can play a crucial role – brief, high-quality and free online training is available at [National Centre for Smoking Cessation and Training](https://www.ncsct.co.uk/).
 
-Conclusion
+##    Conclusion
 
-Across 345 studies, recorded current but not past smoking prevalence was generally lower than national prevalence estimates.
+Across `r paste(current_total)` studies, recorded current but not past smoking prevalence was generally lower than national prevalence estimates.
 Current smokers were at reduced risk of testing positive for SARS-CoV-2 and former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.
 
-Acknowledgements
+##    Acknowledgements
 
@@
 
-Declaration of conflicts of interest
+##    Declaration of conflicts of interest
 
@@
 
-Funding statement
+##    Funding statement
 
@@
 UKPRP is an initiative funded by the UK Research and Innovation Councils, the Department of Health and Social Care (England) and the UK devolved administrations, and leading health research charities.
-Future review versions
-https://www.qeios.com/read/latest-UJR2AW
 
-Previous review versions
-Version 1: https://doi.org/10.32388/UJR2AW 
-Version 2: https://doi.org/10.32388/UJR2AW.3 
-Version 3: https://doi.org/10.32388/UJR2AW.4 
-Version 4: https://doi.org/10.32388/UJR2AW.5 
-Version 5: https://doi.org/10.32388/UJR2AW.6
-Version 6: https://doi.org/10.32388/UJR2AW.7
-Version 7: https://doi.org/10.1111/add.15276 
-Version 8: https://doi.org/10.32388/UJR2AW.9 
-Version 9: https://doi.org/10.32388/UJR2AW.10
+##    Review versions
+The most up-to-date review version is available [here](https://www.qeios.com/read/latest-UJR2AW)
 
-Data availability
-All data contributing to the current and future review versions are available here: https://doi.org/10.6084/m9.figshare.12756020    
-All code required to reproduce the current and future analyses are available here:  https://doi.org/10.5281/zenodo.4002046
+###   Previous review versions
+ * [Version 1](https://doi.org/10.32388/UJR2AW)
+ * [Version 2](https://doi.org/10.32388/UJR2AW.3)
+ * [Version 3](https://doi.org/10.32388/UJR2AW.4)
+ * [Version 4](https://doi.org/10.32388/UJR2AW.5) 
+ * [Version 5](https://doi.org/10.32388/UJR2AW.6)
+ * [Version 6](https://doi.org/10.32388/UJR2AW.7)
+ * [Version 7](https://doi.org/10.1111/add.15276) 
+ * [Version 8](https://doi.org/10.32388/UJR2AW.9) 
+ * [Version 9](https://doi.org/10.32388/UJR2AW.10)
+
+##    Data availability
+All data contributing to the current and future review versions are available [here](https://doi.org/10.6084/m9.figshare.12756020)
+All code required to reproduce the current and future analyses are available [here](https://doi.org/10.5281/zenodo.4002046)
+
+#### To delete
+`r paste(tbl_1$ref)`
+
+##    References

### 2021-02-07 scripts/author_dictionary.R

@@ -40,7 +40,7 @@ cleaned_names <- c('Chow, Us, Cdc',
                    "O, Reilly",
                    "Bermejo, Martin",
                    "Marquez, Salinas",
-                   "Diez, Manglanas",
+                   "Diez, Manglano",
                    "Martinez, Lacalzada",
                    "O, Gallagher",
                    "Vila, Corcoles, Statue, Gracia")

### 2021-02-07 scripts/libraries.R

@@ -19,3 +19,5 @@ library(DT)
 library(epitools)
 library(ggmap)
 library(gtsummary)
+library(bib2df)
+library(ftExtra)

### 2021-02-07 scripts/references_in_table.R

@@ -1,128 +1,19 @@
 library(tidyverse)
+library(magrittr)
 library(flextable)
 library(ftExtra)
+library(bib2df)
 
-tbl_1 <- readRDS(here::here("data_clean", "table_1_refs.rds"))
+tbl_1 <- readRDS(here::here("data_clean", "table_1_refs.rds")) %>%
+  mutate(ref = ifelse(startsWith(ref, "h"), gsub(".*/10", "10", ref), ref))
+bib <- bib2df(here::here("reports", "manuscript.bib")) %>%
+  select(BIBTEXKEY, DOI)
 
-refs <- c("[@guan_clinical_2020]",
-          "[@guan_comorbidity_2020]",
-          "[@lian_analysis_nodate]",
-          "[@jin_treatment_2020]",
-          "[@chen_risk_2020]",
-          "[@zhou_clinical_2020]",
-          "[@mo_clinical_nodate]",
-          "[@zhang_clinical_2020]",
-          "[@wan_clinical_2020]",
-          "[@liu_analysis_2020]",
-          "[@huang_clinical_2020-1]",
-          "[@zhang_epidemiological_2020]",
-          "[@guo_cardiovascular_2020]",
-          "[@liu_association_2020]",
-          "[@huang_clinical_2020]",
-          "[@xu_acute_2020]",
-          "[@li_epidemiological_2020]",
-          "[@rentsch_covid-19_2020]",
-          "[@hu_risk_2020]",
-          "[@wang_epidemiological_2020]",
-          "[@cdcmmwr_preliminary_2020]",
-          "[@dong_eleven_2020]",
-          "[@kim_clinical_2020]",
-          "[@shi_host_2020]",
-          "[@yang_clinical_2020]",
-          "[@argenziano_characterization_2020]",
-          "[@solis_covid-19_2020]",
-          "[@richardson_presenting_2020]",
-          "[@fontanet_cluster_2020]",
-          "[zheng_letter_2020]",
-          "[@liao_clinical_2020]",
-          "[@gil-agudo_clinical_2020]",
-          "[@shi_clinical_2020]",
-          "[@hadjadj_impaired_2020]",
-          "[@gold_characteristics_2020]",
-          "[@yu_association_2020]",
-          "[@zheng_epidemiological_2020]",
-          "[@rica_low_2020]",
-          "[@yin_clinical_2020]",
-          "[@shi_neutrophil_2020]",
-          "[@cho_smoking_2020]",
[…]

### 2021-03-01 reports/manuscript.Rmd
v11

@@
 ---
-title: "Smoking and COVID Living Review (`r paste(current_version)`): a bayesian analysis"
+title: 'Smoking and COVID Living Review (`r paste(current_version)`): a bayesian analysis'
 author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
@@
 output:
+  word_document:
+    toc: yes
   html_document:
     highlight: zenburn
-    toc: true
-    toc_float: true
-    number_sections: false
+    toc: yes
+    toc_float: yes
+    number_sections: no
     theme: simplex
-    css: !expr here::here('css', 'style.css')
-    keep_md: false
+    css: C:/Users/ucbtds4/R Repositories/smoking_covid_bayes/css/style.css
+    keep_md: no
 editor_options:
@@
 bibliography: manuscript.bib
+csl: harvard.csl
 ---
@@
 **Findings:** `r paste(current_version)` (searches up to `r paste(tail(search_details$date_screened, 1))`) included `r paste(current_total)` studies with `r paste(nrow(running_meta_count))` ‘good’ and ‘fair’ quality studies included in unadjusted meta-analyses.
-`r paste(nrow(full_smoking_status))` studies (`r paste(round(nrow(full_smoking_status)/current_total*100,1), "%")`) reported current, former and never smoking status with the remainder using broader categories.
+`r paste(nrow(full_smoking_status))` studies (`r paste(round(nrow(full_smoking_status)/current_total*100,1), "%", sep = "")`) reported current, former and never smoking status with the remainder using broader categories.
 Recorded smoking prevalence among people with COVID-19 was generally lower than national prevalence.
-Current compared with never smokers were at reduced risk of SARS-CoV-2 infection (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_lci, 2))`).
-Data for former smokers were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_lci, 2))`) but favoured there being no important association (`r paste(round(1-m2_a_ecdf(1.1),2), "%", sep = "")` probability of RR ≥1.1).
-Former compared with never smokers were at increased risk of hospitalisation (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_lci, 2))`), greater disease severity (RR = `r paste(round(former_severity_RR_value, 2))`, CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_lci, 2))`) and mortality (RR = `r paste(round(former_mortality_RR_value, 2))`, CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_lci, 2))`).
-Data for current smokers on hospitalisation, disease severity and mortality were inconclusive (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_lci, 2))`; RR `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_lci, 2))`; RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_lci, 2))`, respectively) but favoured there being no important associations with hospitalisation and mortality (`r paste(round(1-m3_a_ecdf(1.1),2), "%", sep = "")` and `r paste(round(1-m5_a_ecdf(1.1),2), "%", sep = "")` probability of RR ≥1.1, respectively) and a small but important association with disease severity (`r paste(round(1-m7_a_ecdf(1.1),2), "%", sep = "")` probability of RR ≥1.1).
+Current compared with never smokers were at reduced risk of SARS-CoV-2 infection (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_tau, 2))`).
+Data for former smokers were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_tau, 2))`) but favoured there being no important association (`r paste(round(1-m2_a_ecdf(1.1),2)*100, "%", sep = "")` probability of RR ≥1.1).
+Former compared with never smokers were at increased risk of hospitalisation (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_tau, 2))`), greater disease severity (RR = `r paste(round(former_severity_RR_value, 2))`, CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_tau, 2))`) and mortality (RR = `r paste(round(former_mortality_RR_value, 2))`, CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_tau, 2))`).
+Data for current smokers on hospitalisation, disease severity and mortality were inconclusive (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_tau, 2))`; RR `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_tau, 2))`; RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_tau, 2))`, respectively) but favoured there being no important associations with hospitalisation and mortality (`r paste(round(1-m3_a_ecdf(1.1),2)*100, "%", sep = "")` and `r paste(round(1-m7_a_ecdf(1.1),2)*100, "%", sep = "")` probability of RR ≥1.1, respectively) and a small but important association with disease severity (`r paste(round(1-m5_a_ecdf(1.1),2)*100, "%", sep = "")` probability of RR ≥1.1).
 
@@
 These uncertainties notwithstanding, both former and current smoking is known to increase the risk of respiratory viral [@denholm_hospitalised_2010; @abadom_risk_2016] and bacterial [@almirall_proportion_1999; @feldman_cigarette_2013] infections and is associated with worse outcomes once infected.
-Cigarette smoke reduces the respiratory immune defence through peri-bronchiolar inflammationand fibrosis, impaired mucociliary clearance and disruption of the respiratory epithelium [@dye_occasional_1994].
+Cigarette smoke reduces the respiratory immune defence through peri-bronchiolar inflammation and fibrosis, impaired mucociliary clearance and disruption of the respiratory epithelium [@dye_occasional_1994].
 There is also reason to believe that behavioural factors (e.g.
@@
 
-There are several reviews that fall within the scope of smoking and COVID-19 [@vardavas_covid-19_2020],[@emami_prevalence_2020; @alqahtani_prevalence_2020; @patanavanich_smoking_2020; @berlin_covid-19_2020; @farsalinos_systematic_2020; @grundy_smoking_2020].
+There are several reviews that fall within the scope of smoking and COVID-19 [@vardavas_covid-19_2020; @emami_prevalence_2020; @alqahtani_prevalence_2020; @patanavanich_smoking_2020; @berlin_covid-19_2020; @farsalinos_systematic_2020; @grundy_smoking_2020].
 We aimed to produce a rapid synthesis of available evidence pertaining to the rates of infection, hospitalisation, disease severity and mortality from SARS-CoV-2/COVID-19 stratified by smoking status.
@@
 Data from ‘good’ and ‘fair’ quality studies were pooled in R v.3.6.3 [@team_r_2013].
-In a living review where new data are regularly added to the analyses, it may be more appropriate to use a Bayesian(as opposed to frequentist) approach where prior knowledge is used in combination with new data to estimate a posterior risk distribution.
+In a living review where new data are regularly added to the analyses, it may be more appropriate to use a Bayesian (as opposed to frequentist) approach where prior knowledge is used in combination with new data to estimate a posterior risk distribution.
 A Bayesian approach mitigates the issue of performing multiple statistical tests, which can inflate family-wise error.
@@
 Studies had a median of `r paste(participants[3])` (interquartile range = `r paste(participants[2], participants[5], sep = "-")`) participants.
-The majority of studies (`r paste(round(sum(diagnostics[5]+diagnostics[6])/sum(diagnostics)*100,1))`%) used reverse transcriptase polymerase chain reaction (RT-PCR) for confirmation of SARS-CoV-2 infection, `r paste(round(sum(diagnostics[1]+diagnostics[6])/sum(diagnostics)*100,1))`% used an antibody test to confirm prior infection and `r paste(round(diagnostics[6]/sum(diagnostics)*100,1))` of studies relied on a combination of RT-PCR or antibody assays.
+The majority of studies (`r paste(round(sum(diagnostics[5]+diagnostics[6])/sum(diagnostics)*100,1))`%) used reverse transcriptase polymerase chain reaction (RT-PCR) for confirmation of SARS-CoV-2 infection, `r paste(round(sum(diagnostics[1]+diagnostics[6])/sum(diagnostics)*100,1))`% used an antibody test to confirm prior infection and `r paste(round(diagnostics[6]/sum(diagnostics)*100,1), "%", sep = "")` of studies relied on a combination of RT-PCR or antibody assays.
 
@@
 The remaining `r paste(nrow(incomplete_smoking_status))` studies reported current, current/former or current and former smoking status but did not explicitly state whether remaining participants were never smokers or if data were missing on smoking status (see [Supplementary table S2c](#supplementary-t2c)).
-`r paste(nrow(missing_smok))` studies explicitly reported the proportion with missing data on smoking status, which ranged from `r paste(round(minimum,2), "%", sep = "")` to `r paste(round(maximum,2), "%", sep = "")`.
+`r paste(nrow(missing_smok))` studies explicitly reported the proportion with missing data on smoking status, which ranged from `r paste(round(minimum,1), "%", sep = "")` to `r paste(round(maximum,1), "%", sep = "")`.
 
@@
 
-Six studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use [@miyara_low_2020; @rimland_clinical_2020; @girardeau_confirmed_2020; @ebinger_sars-cov-2_2020; @islam_risk_2020; @gallichotte_longitudinal_2020].
+Eight studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use [@miyara_low_2020; @rimland_clinical_2020; @girardeau_confirmed_2020; @ebinger_sars-cov-2_2020; @islam_risk_2020; @gallichotte_longitudinal_2020; @kantele_sars-cov-2_2020; @crooks_predicting_2020].
+One additional study grouped together current smokers and vapers [@schubl_risk_2020].
 
@@
 
-Two studies were performed in random, representative population samples and were rated as ‘good’ quality, and `r paste(nrow(fair_quality_number))` studies were rated as ‘fair’ quality, of which `r paste(nrow(running_meta_count))` studies reported results stratified by smoking status for the outcomes of interest and could be included in meta-analyses.
+Three studies were performed in random or representative population samples and were rated as ‘good’ quality, and `r paste(nrow(fair_quality_number))` studies were rated as ‘fair’ quality, of which `r paste(nrow(running_meta_count))` studies reported results stratified by smoking status for the outcomes of interest and could be included in meta-analyses.
 The remaining `r paste(nrow(poor_quality_number))` studies were rated as ‘poor’ quality (see [Table 1](#table-1)).
@@
 #### [*Table 1.* All studies included in narrative review and meta-analysis]{#table-1}
-``` {r table_1}
+``` {r table_1, include=T}
 library(flextable)
 library(ftExtra)
+library(tidyverse)
 
-tbl_1 <- readRDS(file = here::here("reports", "tables", "table_1_refs.rds"))
+set_flextable_defaults(
+  font.family = "Calibri", 
+  font.size = 8,
+  font.color = "black",
+  digits = 2,
+  theme_fun = "theme_booktabs"
+  )
+
+tbl_1 <- readRDS(file = here::here("reports", "tables", "table_1_refs.rds")) %>%
+  arrange(`Date published`) %>%
+  mutate(`Lead author` = recode(`Lead author`, "Schubl" = "Schuble*"))
 
@@
   colformat_md(j = "ref") %>%
-  colformat_double(j = "Sample size")
+  colformat_double(j = "Sample size", digits = 0)
 ```
@@
 Lower than expected current smoking prevalence was generally observed, especially in studies with hospitalised samples.
-Former smoking prevalence was more similar to expected prevalence when reported; however, prevalence was typically higher than national estimates.
+Former smoking prevalence was more similar to expected prevalence when reported; however, study-based prevalence was typically higher than national estimates.
 National smoking prevalence estimates used for comparison are presented in [Supplementary table 3](#supplementary-t3).
@@
 
-Three studies provided data on access to SARS-CoV-2 diagnostic testing for those meeting local testing criteria by smoking status.
-In a cohort study of US military veterans aged 54-75 [@rentsch_covid-19_2020], current smokers were more likely to receive a test: 42.3% (1,603/3,789) of the sample were current smokers compared with 23.8% of all veterans aged 50+ years using any tobacco product between 2010-201533.
+Four studies provided data on access to SARS-CoV-2 diagnostic testing for those meeting local testing criteria by smoking status.
+In a cohort study of US military veterans aged 54-75 [@rentsch_covid-19_2020], current smokers were more likely to receive a test: 42.3% (1,603/3,789) of the sample were current smokers compared with 23.8% of all veterans aged 50+ years using any tobacco product between 2010-2015 [@odani_tobacco_2018].
 In the UK Biobank cohort [@niedzwiedz_ethnic_2020], a multivariable analysis showed former (RR = 1.29, 95% CI = 1.14-1.45, *p* < .001) and current (RR = 1.44, 95% CI = 1.20-1.71, *p* < .001) compared with never smokers to be more likely to receive a test.
@@
 Of these self-referrals, healthcare professionals decided that current compared with former or never smokers were less likely to require a test (RR = 0.93, 95% CI = 0.86-1.0, *p* = 0.045).
+In a further study using the UK Biobank cohort [@didikoglu_early_2021], current (RR = 1.23, 95% CI = 1.19-1.26, *p* < 0.001) and former smokers (RR = 1.20, 95% CI = 1.18-1.23, *p* < 0.001) were more likely to receive a test compared with never smokers.
 
@@
 Meta-analyses were performed for `r paste(nrow(good_quality_2))` ‘good’ and `r paste(nrow(fair_quality_2))` ‘fair’ quality studies (see [Figure 3](#fig-3) and [4](#fig-4)).
-Current smokers were at reduced risk of testing positive for SARS-CoV-2 compared with never smokers (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_lci, 2))`).
-The probability of current smokers being at reduced risk of infection compared with never smokers (RR ≤0.9) was `r paste(round(m1_a_ecdf(0.9),2))`%.
-Former compared with never smokers were at increased risk of testing positive, but data were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_lci, 2))`) and favoured there being no important association.
-The probability of former smokers being at increased risk of infection (RR ≥1.1) compared with never smokers was (`r paste(round(1-m2_a_ecdf(1.1), 2), "%")` probability of RR ≥1.1).
+Current smokers were at reduced risk of testing positive for SARS-CoV-2 compared with never smokers (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_tau, 2))`).
+The three good quality studies each reported point estimates less than 1, although the CrI for one of the three studies crossed 1.
+The probability of current smokers being at reduced risk of infection compared with never smokers (RR ≤0.9) was >99%.
+Former compared with never smokers were at increased risk of testing positive, but data were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_tau, 2))`) and favoured there being no important association.
+The probability of former smokers being at increased risk of infection (RR ≥1.1) compared with never smokers was `r paste(round(1-m2_a_ecdf(1.1), 2)*100, "%", sep = "")`.
 Results were materially unchanged in sensitivity analyses.
@@
 ####    [*Table 2* SARS-CoV-2 infection by smoking status]{#table-2}
-``` {r table_2}
+``` {r table_2, include=T}
 readRDS(file = here::here("reports", "tables", "table_2.rds"))
@@
 * Indicates ‘good’ quality studies.
-The prior from the previous review version (v9) was RR = 0.69.](figure/m1_a.png){#fig-3}
+The prior from the previous review version (v10) was RR = 0.69.](figure/m1_a.png){#fig-3}
 
@@
 * Indicates ‘good’ quality studies.
-The prior from the previous review version (v9) was RR = 1.02.](figure/m2_a.png){#fig-4}
+The prior from the previous review version (v10) was RR = 1.02.](figure/m2_a.png){#fig-4}
 
@@
 Meta-analyses were performed for `r paste(nrow(fair_quality_3))` ‘fair’ quality studies (see [Figure 5](#fig-5) and [6](#fig-6)).
-Current (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_lci, 2))`) and former (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_lci, 2))`) compared with never smokers were at increased risk of hospitalisation with COVID-19.
+Current (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_tau, 2))`) and former (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_tau, 2))`) compared with never smokers were at increased risk of hospitalisation with COVID-19.
 However, data for current smokers were inconclusive and favoured there being no important association.
-The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(round(1-m3_a_ecdf(1.1),2),"%", sep = "")` and `r paste(round(1-m4_a_ecdf(1.1),2), "%", sep = "")`, respectively.
+The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(round(1-m3_a_ecdf(1.1),2)*100,"%", sep = "")` and `r paste(round(1-m4_a_ecdf(1.1),2)*100, "%", sep = "")`, respectively.
 Results were materially unchanged in two sensitivity analyses.
@@
 ####    [*Table 3* COVID-19 hospitalsation by smoking status]{#table-3}
-``` {r table_3}
+``` {r table_3, include=T}
 readRDS(file = here::here("reports", "tables", "table_3.rds"))
@@
 
-![*Figure 5.* Forest plot for risk of hospitalisation in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.06.](figure/m3_a.png){#fig-5}
+![*Figure 5.* Forest plot for risk of hospitalisation in current vs. never smokers. The prior from the previous review version (v10) was RR = 1.08.](figure/m3_a.png){#fig-5}
 
-![*Figure 6.* Forest plot for risk of hospitalisation in former vs. never smokers. The prior from the previous version (v9) was RR = 1.17.](figure/m4_a.png){#fig-5}
+![*Figure 6.* Forest plot for risk of hospitalisation in former vs. never smokers. The prior from the previous version (v10) was RR = 1.18.](figure/m4_a.png){#fig-5}
 
@@
 Meta-analyses were performed for `r paste(nrow(fair_quality_4))` ‘fair’ quality studies (see [Figure 7](#fig-7) and [8](#fig-8)).
-Current (RR = `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_lci, 2))`) and former (RR = `r paste(round(former_severity_RR_value, 2))`, 95% CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_lci, 2))`) compared with never smokers were at increased risk of greater disease severity; data for current smokers were inconclusive but favoured there being a small but important association.
-The probability of current and former smokers having increased risk of greater disease severity (RR ≥1.1) compared with never smokers was `r paste(round(1-m5_a_ecdf(1.1),2), "%", sep = "")` and `r paste(round(1-m6_a_ecdf(1.1),2), "%", sep = "")`, respectively.
+Current (RR = `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_tau, 2))`) and former (RR = `r paste(round(former_severity_RR_value, 2))`, 95% CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_tau, 2))`) compared with never smokers were at increased risk of greater disease severity.
+However, while data for current smokers only were inconclusive, they favoured there being a small but important association.
+The probability of current and former smokers having increased risk of greater disease severity (RR ≥1.1) compared with never smokers was `r paste(round(1-m5_a_ecdf(1.1),2)*100, "%", sep = "")` and `r paste(round(1-m6_a_ecdf(1.1),2)*100, "%", sep = "")`, respectively.
 Results were materially unchanged in two sensitivity analyses.
@@
 ####    [*Table 4* COVID-19 disease severity by smoking status]{#table-4}
-``` {r table_4}
+``` {r table_4, include=T}
 readRDS(file = here::here("reports", "tables", "table_4.rds"))
@@
 
-![*Figure 7.* Forest plot for the risk of severe disease in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.26.](figure/m5_a.png){#fig-7}
+![*Figure 7.* Forest plot for the risk of severe disease in current vs. never smokers. The prior from the previous review version (v10) was RR = 1.26.](figure/m5_a.png){#fig-7}
 
-![*Figure 8.* Forest plot for the risk of severe disease in former vs. never smokers. The prior from the previous review version (v9) was RR = 1.52.](figure/m6_a.png){#fig-8}
+![*Figure 8.* Forest plot for the risk of severe disease in former vs. never smokers. The prior from the previous review version (v10) was RR = 1.52.](figure/m6_a.png){#fig-8}
 
@@
 `r paste(nrow(table_5_word))` studies reported mortality from COVID-19 by smoking status (see [Table 5](#table-5)), with `r paste(nrow(fair_quality_5))` ‘fair’ quality studies included in meta-analyses (see [Figure 9](#fig-9) and [10](#fig-10)).
-Current (RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_lci, 2))`) and former (RR = `r paste(round(former_mortality_RR_value, 2))`, 95% CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_lci, 2))`) compared with never smokers were at increased risk of in-hospital mortality from COVID-19.
+Current (RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_tau, 2))`) and former (RR = `r paste(round(former_mortality_RR_value, 2))`, 95% CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_tau, 2))`) compared with never smokers were at increased risk of in-hospital mortality from COVID-19.
 However, data for current smokers were inconclusive and favoured there being no important association.
-The probability of current and former smokers being at greater risk of in-hospital mortality (RR ≥1.1) compared with never smokers was `r paste(1-m7_a_ecdf(1.1), "%")` and `r paste(1-m8_a_ecdf(1.1), "%")`, respectively.
+The probability of current and former smokers being at greater risk of in-hospital mortality (RR ≥1.1) compared with never smokers was `r paste(round(1-m7_a_ecdf(1.1),2)*100, "%", sep = "")` and >99%, respectively.
 Results were materially unchanged in two sensitivity analyses.
@@
 ####    [*Table 5* COVID-19 attributed mortality by smoking status]{#table-5}
-``` {r table_5}
+``` {r table_5, include=T}
 readRDS(file = here::here("reports", "tables", "table_5.rds"))
@@
 
-![Figure 9. Forest plot for the risk of mortality in current vs. never smokers. The prior from the previous review version (v9) was RR = 1.05.](figure/m7_a.png){#fig-9}
+![Figure 9. Forest plot for the risk of mortality in current vs. never smokers. The prior from the previous review version (v10) was RR = 1.05.](figure/m7_a.png){#fig-9}
 
-![Figure 10. Forest plot for the risk of mortality in former smokers vs. never smokers. The prior from the previous version (v9) was RR = 1.39.](figure/m8_a.png){#fig-10}
+![Figure 10. Forest plot for the risk of mortality in former smokers vs. never smokers. The prior from the previous version (v10) was RR = 1.40.](figure/m8_a.png){#fig-10}
 
@@
 This living rapid review found uncertainty in the majority of `r paste(current_total)` studies arising from the recording of smoking status.
-Notwithstanding these uncertainties, compared with overall adult national prevalence estimates, recorded current smoking rates in most studies were lower than expected.
-In a subset of better-quality studies (n = `r paste(nrow(good_quality_2)+nrow(fair_quality_2))`), current but not former smokers had a reduced risk of testing positive for SARS-CoV-2 but current smokers appeared somewhat more likely to present for testing and/or receive a test.
+Notwithstanding this uncertainty, compared with overall adult national prevalence estimates, recorded current smoking rates in most studies were lower than expected.
+In a subset of good and fair quality studies (n = `r paste(nrow(good_quality_2)+nrow(fair_quality_2))`), current but not former smokers had a reduced risk of testing positive for SARS-CoV-2 but current smokers appeared somewhat more likely to present for testing and/or receive a test.
 Data for current smokers on the risk of hospitalisation, disease severity and mortality were inconclusive, and favoured there being no important associations with hospitalisation and mortality and a small but important increase in the risk of severe disease.
 Former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.
-Issues complicating interpretation
+
+###   Issues complicating interpretation
 
@@
   2. *Infection with SARS-CoV-2*    
-    1.	Infection following viral exposure depends on individual differences in, for example, genetic susceptibility or immunocompetence, which are poorly understood at present.
+    1.	Infection following viral exposure depends on individual differences in, for example, genetic susceptibility or immunocompetence, which are poorly understood at present and may be confounded with smoking.
 For example, the household secondary attack rate for COVID-19 is estimated at 17% [@fung_household_2020].
@@
 Testing is hence likely limited in some subgroups, with the potential for these groups to include an overrepresentation of current smokers.
-    2.	Current and former smokers may be more likely to meet local criteria for community testing due to increased prevalence of symptoms consistent with SARS-CoV-2 infection, such as cough, increased sputum production or altered sense of smell or taste [@hopkinson_current_2020].
+    2.	On the other hand, current and former smokers may be more likely to meet local criteria for community testing due to increased prevalence of symptoms consistent with SARS-CoV-2 infection, such as cough, increased sputum production or altered sense of smell or taste [@hopkinson_current_2020].
 Evidence from a small number of studies indicates that current smokers may be more likely to present for testing, hence increasing the denominator in comparisons with never smokers and potentially inflating the rate of negative tests in current smokers.
@@
 We identified one population study conducted in Hungary reporting on seroprevalence and smoking status [@merkely_novel_2020]; however, the response rate was only 58.8% and the current smoking rate was 10 percentage points below national prevalence estimates, which raises some doubt about representativeness of the final sample.
-Similarly, a second representative population survey with results from three regions in France [@carrat_seroprevalence_2020] reported a current smoking rate of more than 10 percentage points below national prevalence (12% vs.
-25% daily smoking prevalence) [@andler_baisse_2019].
+Similarly, two further representative population surveys [@carrat_seroprevalence_2020; @richard_seroprevalence_2020] reported a current smoking rate of more than 10 percentage points below national prevalence (12% vs.
+25% and 15% vs.
+27% daily smoking prevalence, respectively) [@andler_baisse_2019; @statistique_part_2020].
   4. *Testing positive for SARS-CoV-2*      
@@
 Smokers with COVID-19 may be less likely to receive a SARS-CoV-2 test or present to hospital due to lack of access to healthcare and may be more likely to die in the community from sudden complications (i.e.
-self-selection bias) and thus not be recorded [@noauthor_sts_nodate].
+self-selection bias) and thus not be recorded [@brown_smoking_2020].
     2.	Diagnostic criteria for SARS-CoV-2 infection and COVID-19 have changed during the course of the pandemic [@organisation_laboratory_nodate].
@@
 It is plausible this may have skewed early data towards less severe cases.
-In addition, the observed association between former smoking and greater disease severity may be explained by collider bias49, where conditioning on a collider (e.g.
+In addition, the observed association between former smoking and greater disease severity may be explained by collider bias  [@griffith_collider_2020], where conditioning on a collider (e.g.
 testing or hospitalisation) by design or analysis may introduce a spurious association between current or former smoking (a potential cause of testing or hospitalisation) and SARS-CoV-2 infection/adverse outcomes from COVID-19 (potentially exacerbated by smoking) [@murray_causation_2020].
@@
 Research shows large discrepancies between EHRs and actual behaviour [@polubriaginof_challenges_2018].
-Known failings of EHRs include implausible longitudinal changes, such as former smokers being recorded as never smokers at subsequent hospital visits51.
+Known failings of EHRs include implausible longitudinal changes, such as former smokers being recorded as never smokers at subsequent hospital visits [@polubriaginof_challenges_2018].
 Misreporting on the part of the patient (perhaps due to perceived stigma) has also been observed, with biochemical measures showing higher rates of smoking compared with self-report in hospitalised patients in the US [@benowitz_prevalence_2009].
@@
 The comparisons of current and former smoking prevalence in the included studies with national prevalence estimates did not adjust observed prevalence for the demographic profile of those tested/admitted to hospital.
-Other reviews focused on this comparison have applied adjustments for sex and age, and continue to find lower than expected prevalence – notwithstanding the issues complicating interpretation described above17.
+Other reviews focused on this comparison have applied adjustments for sex and age, and continue to find lower than expected prevalence – notwithstanding the issues complicating interpretation described above [@farsalinos_systematic_2020].
 
@@
  * [Version 9](https://doi.org/10.32388/UJR2AW.10) 
+ * [Version 10](https://doi.org/10.32388/UJR2AW.11)
 
@@
 All data contributing to the current and future review versions are available [here](https://doi.org/10.6084/m9.figshare.12756020)
+
 All code required to reproduce the current and future analyses are available [here](https://doi.org/10.5281/zenodo.4002046)

### 2021-03-01 reports/manuscript_compile.R

@@ -1,1 +1,5 @@
-rmarkdown::render(here("reports","manuscript.Rmd"))
+source(here("scripts", "libraries.R"))
+source(here("scripts", "bayesian_data_summary.R"))
+
+rmarkdown::render(here("reports","manuscript.Rmd"), output_format = "html_document")
+rmarkdown::render(here("reports","manuscript.Rmd"), output_format = "word_document")

### 2021-03-01 scripts/author_dictionary.R

@@ -43,7 +43,18 @@ cleaned_names <- c('Chow, Us, Cdc',
                    "Diez, Manglano",
                    "Martinez, Lacalzada",
                    "O, Gallagher",
-                   "Vila, Corcoles, Statue, Gracia")
+                   "Vila, Corcoles, Satue, Gracia",
+                   "Nunez, Gil",
+                   "De, Abajo",
+                   "Vila, Corcoles, Satue, Gracia, Vila, Rovira",
+                   "Isaric, 10",
+                   "Gegout, Petit",
+                   "Martinez, Resendez",
+                   "Hippisley, Cox",
+                   "Vilar, Garcia",
+                   "Ibarra, Nava",
+                   "Mancilla, Galindo",
+                   "Prats, Uribe")
 
 correct_names <- c('Chow (US CDC)',
                    "Gold (US CDC)",
@@ -90,4 +101,15 @@ correct_names <- c('Chow (US CDC)',
                    "Diez-Manglanas",
                    "Martinez-Lacalzada",
                    "O'Gallagher",
-                   "Vila-Corcoles, Statue-Gracia")
+                   "Vila-Corcoles, Satue-Gracia",
+                   "Nunez-Gil",
+                   "de Abajo",
+                   "Vila-Corcoles, Satue-Gracia, Vila-Rovira",
+                   "ISARIC",
+                   "Gegout Petit",
+                   "Martinez-Resendez",
+                   "Hippisley-Cox",
+                   "Vilar-Garcia",
+                   "Ibarra-Nava",
+                   "Mancilla-Galindo",
+                   "Prats-Uribe")

### 2021-03-01 scripts/bayes_scripts.R

@@ -55,8 +55,8 @@ forest_plot <- function(model, data_study, data_pooled, cut, title, type, filena
   forest_data$review_version <- plyr::mapvalues(forest_data$review_version, from = c("previous",
                                                                                      "current",
                                                                                      "combined_pooled"),
-                                                to = c("Previous version (v9)",
-                                                       "Current version (v10)",
+                                                to = c(paste("Previous version ", "(", tail(previous_review_versions, 1), ")", sep = ""),
+                                                       paste("Current version ", "(",current_version, ")", sep = ""),
                                                        "Pooled effect"))
   summary_forest <- group_by(forest_data, Author, review_version) %>%
     median_qi(b_Intercept)

### 2021-03-01 scripts/bayesian_data_generation.R

@@ -69,25 +69,25 @@ current_testing_prior <-
   c(prior(normal(-.371, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_testing_prior <-
-  c(prior(normal(0.0198, 1), class = Intercept),
+  c(prior(normal(0.0296, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_hospitalisation_prior <-
-  c(prior(normal(0.0582, 1), class = Intercept),
+  c(prior(normal(0.0770, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_hospitalisation_prior <-
-  c(prior(normal(0.157, 1), class = Intercept),
+  c(prior(normal(0.1655, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_severity_prior <-
-  c(prior(normal(0.231, 1), class = Intercept),
+  c(prior(normal(0.2311, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_severity_prior <-
-  c(prior(normal(0.419, 1), class = Intercept),
+  c(prior(normal(0.4187, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_mortality_prior <-
   c(prior(normal(0.0487, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_mortality_prior <- 
-  c(prior(normal(0.329, 1), class = Intercept),
+  c(prior(normal(0.3365, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 
 # Classical meta-analysis guided priors with high heterogeneity
@@ -95,25 +95,25 @@ current_testing_prior_hh <-
   c(prior(normal(-.371, 1), class = Intercept),
     prior(cauchy(0.3, 1), class = sd))
 former_testing_prior_hh <-
-  c(prior(normal(0.0198, 1), class = Intercept),
+  c(prior(normal(0.0296, 1), class = Intercept),
     prior(cauchy(0.3, 1), class = sd))
 current_hospitalisation_prior_hh <-
-  c(prior(normal(0.0582, 1), class = Intercept),
+  c(prior(normal(0.0770, 1), class = Intercept),
     prior(cauchy(0.3, 1), class = sd)) 
 former_hospitalisation_prior_hh <-
[…]

### 2021-03-01 scripts/bayesian_data_summary.R

@@ -1,0 +1,121 @@
+source(here::here("scripts", "libraries.R"))
+
+tau <- read_rds(here("output_data", "tau.rds"))
+current_testing_alternative_prior <- read_rds(here("output_data", "current_testing_alternative.rds"))
+former_testing_alternative_prior <- read_rds(here("output_data", "former_testing_alternative.rds"))
+current_hospitalisation_alternative_prior <- read_rds(here("output_data", "current_hospitalisation_alternative.rds"))
+former_hospitalisation_alternative_prior <- read_rds(here("output_data", "former_hospitalisation_alternative.rds"))
+current_severity_alternative_prior <- read_rds(here("output_data", "current_severity_alternative.rds"))
+former_severity_alternative_prior <- read_rds(here("output_data", "former_severity_alternative.rds"))
+current_mortality_alternative_prior <- read_rds(here("output_data", "current_mortality_alternative.rds"))
+former_mortality_alternative_prior <- read_rds(here("output_data", "former_mortality_alternative.rds"))
+
+# Results summaries
+# Current
+current_testing_RR_value <- current_testing_alternative_prior$summary %>%
+  ungroup() %>%
+  filter(Author == "Pooled Effect") %>%
+  select(b_Intercept)
+current_testing_lci <- current_testing_alternative_prior$summary %>%
+  ungroup() %>%
+  filter(Author == "Pooled Effect") %>%
+  select(.lower)
+current_testing_uci <- current_testing_alternative_prior$summary %>%
+  ungroup() %>%
+  filter(Author == "Pooled Effect") %>%
+  select(.upper)
+current_testing_tau <- tau[1,1]
+former_testing_RR_value <- former_testing_alternative_prior$summary %>%
+  ungroup() %>%
+  filter(Author == "Pooled Effect") %>%
+  select(b_Intercept)
+former_testing_lci <- former_testing_alternative_prior$summary %>%
+  ungroup() %>%
+  filter(Author == "Pooled Effect") %>%
+  select(.lower)
+former_testing_uci <- former_testing_alternative_prior$summary %>%
+  ungroup() %>%
+  filter(Author == "Pooled Effect") %>%
+  select(.upper)
+former_testing_tau <- tau[2,1]
+#Hospitalisation
+current_hospitalisation_RR_value <- current_hospitalisation_alternative_prior$summary %>%
[…]

### 2021-03-01 scripts/didikoglu_rr.R

@@ -1,0 +1,27 @@
+meta <- tibble('author' = c("Didikoglu"),
+               'not_tested_smoker' = c(32978),
+               'not_tested_never_smoker' = c(189812),
+               'tested_smoker' = c(4786),
+               'tested_never_smoker' = c(21840),
+               'not_tested_former_smoker' = c(118598),
+               'tested_former_smoker' = c(16807))
+
+a <-meta %>%
+      select(not_tested_never_smoker, tested_never_smoker) %>%
+      rename('not tested' = not_tested_never_smoker, 'tested' = tested_never_smoker)
+b <- meta %>%
+  select(not_tested_smoker, tested_smoker) %>%
+  rename('not tested' = not_tested_smoker, 'tested' = tested_smoker)
+c <- as.matrix(rbind(a,b))
+rownames(c) <- c('never smokers', 'current smokers')
+riskratio(c, method = 'wald', conf.level=0.95)
+
+a <-meta %>%
+  select(not_tested_never_smoker, tested_never_smoker) %>%
+  rename('not tested' = not_tested_never_smoker, 'tested' = tested_never_smoker)
+b <- meta %>%
+  select(not_tested_former_smoker, tested_former_smoker) %>%
+  rename('not tested' = not_tested_former_smoker, 'tested' = tested_former_smoker)
+d <- as.matrix(rbind(a,b))
+rownames(d) <- c('never smokers', 'former smokers')
+riskratio(d, method = 'wald', conf.level=0.95)

### 2021-03-01 scripts/references_in_table.R

@@ -9,8 +9,7 @@ tbl_1 <- readRDS(here::here("data_clean", "table_1_refs.rds")) %>%
 bib <- bib2df(here::here("reports", "manuscript.bib")) %>%
   select(BIBTEXKEY, DOI)
 
-tbl_1 %<>%
-  left_join(., bib %>%
+tbl_1 <- left_join(tbl_1, bib %>%
               rename("ref" = DOI),
             by = "ref") %>%
   mutate(ref = paste("@", BIBTEXKEY, sep = "")) %>%

### 2021-03-02 reports/bayesian_report.Rmd

@@
 * **Load packages**: All necessary packages are loaded in from the `libraries.r` script in the `/scripts` folder.
-* **Author dictionary**: An author dictionary is created to manage the presentation of author names in the final tables and meta-analyses
+* **Author dictionary**: An author dictionary is created to manage the presentation of author names in the final tables and meta-analyses.
+This is particularly for those with punctuation in their names.
 * **Load scripts**: We have written functions to handle much of the Bayesian analysis
-* **Load Bayesian datasets**: The Bayesian approach relies on large simulated datasets from the observed data, for this report we only present the results graphically from data simulated outside of the report.
+* **Load Bayesian datasets**: The Bayesian approach relies on large simulated datasets derived from the observed data, for this report we only present the results graphically from data simulated outside of the report.
 The scripts `bayes_scripts.R`, `bayesian_data_generation.R` and `bayesian_analysis.R` should be able to be run locally to generate your own simulated data.
@@
 source(here::here("scripts", "libraries.R"))
+set_flextable_defaults(
+  font.family = "Calibri", 
+  font.size = 8,
+  font.color = "black",
+  digits = 1,
+  theme_fun = "theme_booktabs"
+  )
 ```
@@
 register_google(google_api)
-drive_download("https://docs.google.com/spreadsheets/d/1jelIGxrm99QpDKwSoyMUliDJQWtzk4bHsNGDaIjxTaU/edit?usp=sharing", path = here("data_spreadsheet", "data_extraction_current.xlsx"), overwrite = T)
+drive_download("https://docs.google.com/spreadsheets/d/1VUY5xDLBpjNBovveg9wQrc0wbLhnEkmDPMR63rfj2-Y/edit?usp=sharing", path = here("data_spreadsheet", "data_extraction_current.xlsx"), overwrite = T)
 
@@
 old_isaric <- head(names(tidyselect::vars_select(data_study_general$lead_author, starts_with("isaric", ignore.case = T))), n = -1)
+#This will prevent duplication of the ISARIC reports, only retaining the last based on their names
 excluded_other_reason <- c("mehra", #retracted
@@
 source(here("scripts", "remove_excluded.r"))
+current_total <- nrow(a)
 datatable(a)
@@
 ``` {r PRISMA}
-current_total <- 345
-
-previous_total <- 279
+previous_total <- 345
 
@@
 Currently studies are reported from `r paste(nrow(country))`
-``` {r}
+``` {r, cache=TRUE}
 country <- mutate_geocode(country, Country)
@@
 ## A map is produced for where the studies are performed
-``` {r interactive global map, cache = TRUE}
+``` {r interactive global map, cache = TRUE, quietly = TRUE, echo=FALSE, results = "hide"}
 library(rgdal)
@@
 
-from_countries <- c("USA", "UK", "Iran",  "Korea, South")
+from_countries <- c("USA", "UK", "Iran",  "South Korea")
 to_countries <- c("United States", "United Kingdom", "Iran (Islamic Republic of)", "Korea, Republic of")
@@
 data("coronavirus")
+```
+
+``` {r produce_map}
 summary_df <- coronavirus %>% 
@@
 
-my_bins <- c(0, 1, 5, 10, 20, 50, 70, 100)
-mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)
+my_bins <- c(0, 1, 5, 10, 20, 50, 70, 110)
+mypalette <- leaflet::colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)
 
@@
         file = here("data_clean", "table_1_refs.rds"))
-#source(here("scripts", "references_in_table.r"))
-```
-The total number of included studies is `r paste(nrow(table_1_word))`. This table is searchable
+source(here("scripts", "references_in_table.r"))
+```
+The total number of included studies is `r paste(nrow(table_1_word))`.
+This table is searchable.
+The references are added in automatically from Zotero.
+The bibtex file is saved as `manuscript.bib` and is updated following each search.
 ```{r}
@@
          "Del, Valle" = "del Valle")
-a$Author <- to_upper_camel_case(a$`Lead author`, sep_out = ", ")
+a$Author <- to_upper_camel_case(a$Author, sep_out = ", ")
 author_list <- plyr::mapvalues(a$Author,
@@
   select(lead_author) %>%
-  filter(!lead_author %in% c("reiter", "hippisley_cox", "ullah"))
+  filter(!lead_author %in% c("reiter", "hippisley_cox", "ullah", "lowe", "polat"))
 
@@
 fair_quality_2 <- quality_table_2 %>%
+  filter(lead_author %in% included_studies$lead_author) %>%
   filter(overall_quality == 'fair')
@@
 fair_quality_3 <- quality_table_3 %>%
-  filter(overall_quality != 'fair')
+  filter(lead_author %in% included_studies$lead_author) %>%
+  filter(overall_quality == 'fair')
 
@@
 fair_quality_4 <- quality_table_4 %>%
+  filter(lead_author %in% included_studies$lead_author) %>%
   filter(overall_quality == 'fair')
@@
 fair_quality_5 <- quality_table_5 %>%
+  filter(lead_author %in% included_studies$lead_author) %>%
   filter(overall_quality != 'fair')
@@
 
-fair_quality_4 <- quality_table_5 %>%
-  filter(overall_quality != 'fair')
+fair_quality_5 <- quality_table_5 %>%
+  filter(lead_author %in% included_studies$lead_author) %>%
+  filter(overall_quality == 'fair')
 ```

### 2021-06-11 reports/manuscript.Rmd
checked manuscript

@@
 ---
-title: 'Smoking and COVID Living Review (`r paste(current_version)`): a bayesian analysis'
+title: 'The association of smoking status with SARS-CoV-2 infection, hospitalisation and mortality from COVID-19: A living rapid evidence review with Bayesian meta-analyses (`r paste(current_version)`)'
 author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
@@
     theme: simplex
-    css: C:/Users/ucbtds4/R Repositories/smoking_covid_bayes/css/style.css
+    css: C:/Users/ucbtds4/R_Repositories/smoking_covid_bayes/css/style.css
     keep_md: no
@@
 bibliography: manuscript.bib
-csl: harvard.csl
+csl: apa.csl
 ---
@@
 #### [*Table 1.* All studies included in narrative review and meta-analysis]{#table-1}
+
 ``` {r table_1, include=T}
@@
 ####    [*Table 2* SARS-CoV-2 infection by smoking status]{#table-2}
+
 ``` {r table_2, include=T}
@@
 ####    [*Table 3* COVID-19 hospitalsation by smoking status]{#table-3}
+
 ``` {r table_3, include=T}
@@
 ####    [*Table 4* COVID-19 disease severity by smoking status]{#table-4}
+
 ``` {r table_4, include=T}
@@
 ####    [*Table 5* COVID-19 attributed mortality by smoking status]{#table-5}
+
 ``` {r table_5, include=T}
@@
 Data for current smokers on the risk of hospitalisation, disease severity and mortality were inconclusive, and favoured there being no important associations with hospitalisation and mortality and a small but important increase in the risk of severe disease.
+However, it should be noted that pooled estimates for the risk of hospitalisation and mortality in current smokers have numerically increased since the previous version (v10).
 Former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.

### 2021-07-23 reports/bayesian_report.Rmd
begin v12

@@
 ---
-title: 'Smoking and COVID Living Review (v10): a bayesian analysis'
+title: 'Smoking and COVID Living Review (v12): a bayesian analysis'
 author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
@@
 ```{r head_search_details}
-flextable(search_details[,1:13])
+flextable(search_details %>%
+            mutate(`date_screening in format (2020-05-22)` = ymd(`date_screening in format (2020-05-22)`)))
 ```
@@
 datatable(data_study_general[1:4] %>%
-            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`),
+            mutate(`date_published_format(2020-05-12)` = ymd(data_study_general$`date_published_format(2020-05-12)`),
                    doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
@@
 ``` {r excluding_studies}
-superseded_studies <- names(tidyselect::vars_select(data_study_general$lead_author, ends_with("_old", ignore.case = T)))
-old_isaric <- head(names(tidyselect::vars_select(data_study_general$lead_author, starts_with("isaric", ignore.case = T))), n = -1)
+superseded_studies <- data_study_general %>%
+  filter(str_ends(lead_author, "_old")) %>%
+  pull(lead_author)
+
+old_isaric <- data_study_general %>%
+  filter(str_starts(lead_author, "isaric")) %>%
+  head(., -1) %>%
+  pull(lead_author)
 #This will prevent duplication of the ISARIC reports, only retaining the last based on their names
+
 excluded_other_reason <- c("mehra", #retracted
@@
 ``` {r descriptives}
-source(here("scripts", "remove_excluded.r"))
+source(here("scripts", "remove_excluded.r")) # This removes the excluded studies from each table that has been read in previously.
 current_total <- nrow(a)
@@
 ``` {r PRISMA}
-previous_total <- 345
+previous_total <- read_rds(here("data_clean", "data_study_general.rds")) %>%
+  filter(review_version %in% prev_versions) %>%
+  nrow()
 
@@
 country <- table_1$country %>%
-  to_upper_camel_case() %>%
-  recode(., 'Usa' = 'USA', 'Uk' = 'UK', 'SaudiArabia' = 'Saudi Arabia', "SouthKorea" = "South Korea", "UnitedArabEmirates" = "United Arab Emirates") %>%
+  to_upper_camel_case(sep_out = " ") %>%
+  recode(., 'Usa' = 'USA', 'Uk' = 'UK') %>%
   table() %>%
@@
   rename('Country' = 1,
-         'Number' = 2) %>%
-  mutate(Country = as.character(Country))
+         'Number' = 2)
 
@@
 
-country <- subset(country, Country != "Multiple")
-```
-Currently studies are reported from `r paste(nrow(country))`
+country <- country %>%
+  filter(Country != "Multiple")
+```
+Currently studies are reported from `r paste(nrow(country))` countries.
+
 ``` {r, cache=TRUE}
-country <- mutate_geocode(country, Country)
-
-country_bounds <- country %>%
-  select(lon, lat) %>%
-  tidyr::drop_na() %>%
-  sp::SpatialPointsDataFrame(coords = .[,1:2], data = .) %>%
-  sp::bbox() * 1.1
+countries <- ne_countries(type = "countries", returnclass = "sf")
+
+map <- country %>%
+  mutate(iso3c = countrycode::countrycode(Country, origin = "country.name", destination = "iso3c")) %>%
+  left_join(., countries %>%
+              as_tibble() %>%
+              dplyr::select(adm0_a3, geometry), by = c("iso3c" = "adm0_a3")) %>%
+  st_as_sf()
+
+country_bounds <- map %>%
+  st_bbox()
 

### 2021-07-23 scripts/clean_data.R

@@ -19,8 +19,8 @@ review_details <- data_study_general %>%
 
 table_1 <- table_1 %>%
   clean_data() %>%
-  mutate(., lower_range = sub('\\_.*', '', .$range)) %>%
-  mutate(., upper_range = sub('.*_', '',.$range,)) %>% ## We break apart the age range into an upper and lower level
+  mutate(lower_range = sub('\\_.*', '', .$range),
+         upper_range = sub('.*_', '',.$range,)) %>% ## We break apart the age range into an upper and lower level
   select(lead_author, sample_size, female_sex_percent, data_source, median_age, mean_age, iqr_lower, iqr_upper, standard_deviation, lower_range,
          upper_range, current_smoker, former_smoker, current_former_smoker, never_smoker, never_smoker_unknown, current_vaper, former_vaper,
          current_former_vaper, not_stated, missing, total) %>%

### 2021-07-23 scripts/libraries.R

@@ -6,6 +6,7 @@ pkgs =
     "ggplot2",
     "tidybayes",
     "readr",
+    "lubridate",
     "tidyverse",
     "ggridges",
     "glue",
@@ -25,6 +26,9 @@ pkgs =
     "bib2df",
     "ftExtra",
     "meta",
-    "snakecase"
+    "snakecase",
+    "revtools",
+    "rnaturalearth",
+    "sf"
   )
 pacman::p_load(pkgs, character.only = T)

### 2021-07-23 scripts/ovid_search.R

@@ -1,13 +1,16 @@
-library("tidyverse")
-library("here")
-library(magrittr)
-library(bib2df)
-library(revtools)
+source(here("scripts", "libraries.R"))
 
 ovid_dec <- read_csv(here("data_spreadsheet", "2020-12-17_ovid.csv")) %>%
   select(DO) %>%
   drop_na()
 
-read_bibliography(here("data_spreadsheet", "ovid_2021-02-16.ris")) %>%
-  filter(!doi %in% ovid_dec$DO) %>%
+ovid_feb <- read_bibliography(here("data_spreadsheet", "ovid_2021-02-16.ris"))
+
+ovid_jul <- read_bibliography(here("data_spreadsheet", "ovid_2021-07-18.ris"))
+
+ovid_new <- ovid_jul %>%
+  filter(!doi %in% ovid_feb$doi) %>%
   write_bibliography(., file = here("data_spreadsheet", "deduplicated_ovid.bib"))
+
+ovid_include <- read_bibliography(here("data_spreadsheet", "include_v12_ovid.bib")) %>%
+  mutate(doi = paste0("https://doi.org/", doi))

### 2021-07-27 reports/manuscript.Rmd
v12

@@
   word_document:
-    toc: yes
+    toc: no
   html_document:
     highlight: zenburn
-    toc: yes
-    toc_float: yes
     number_sections: no
     theme: simplex
-    css: C:/Users/ucbtds4/R_Repositories/smoking_covid_bayes/css/style.css
     keep_md: no
@@
 Former compared with never smokers were at increased risk of hospitalisation (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_tau, 2))`), greater disease severity (RR = `r paste(round(former_severity_RR_value, 2))`, CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_tau, 2))`) and mortality (RR = `r paste(round(former_mortality_RR_value, 2))`, CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_tau, 2))`).
-Data for current smokers on hospitalisation, disease severity and mortality were inconclusive (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_tau, 2))`; RR `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_tau, 2))`; RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_tau, 2))`, respectively) but favoured there being no important associations with hospitalisation and mortality (`r paste(round(1-m3_a_ecdf(1.1),2)*100, "%", sep = "")` and `r paste(round(1-m7_a_ecdf(1.1),2)*100, "%", sep = "")` probability of RR ≥1.1, respectively) and a small but important association with disease severity (`r paste(round(1-m5_a_ecdf(1.1),2)*100, "%", sep = "")` probability of RR ≥1.1).
+Current compared with never smokers were at increased risk of greater disease severity (RR `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_tau, 2))`).
+Data for current smokers on hospitalisation and mortality were inconclusive (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_tau, 2))`; RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_tau, 2))`, respectively) but favoured there being no important associations (`r paste(round(1-m3_a_ecdf(1.1),2)*100, "%", sep = "")` and `r paste(round(1-m7_a_ecdf(1.1),2)*100, "%", sep = "")` probability of RR ≥1.1, respectively).
 
-**Conclusions:** Compared with never smokers, current smokers appear to be at reduced risk of SARS-CoV-2 infection while former smokers appear to be at increased risk of hospitalisation, greater disease severity and mortality from COVID-19.
+**Conclusions:** Compared with never smokers, current smokers appear to be at reduced risk of SARS-CoV-2 infection and increased risk of greater disease severity, while former smokers appear to be at increased risk of hospitalisation, greater disease severity and mortality from COVID-19.
 However, it is uncertain whether these associations are causal.
@@
 independently checked against pre-prints and published reports) by another on i) author (year); ii) date published; iii) country; iv) study design; v) study setting; vi) sample size; vii) sex; viii) age; ix) smoking status (e.g.
-current, former, never, not stated, missing) and whether it was biochemically verified; x) use of alternative nicotine products; xi) SARS-CoV-2 testing; xii) SARS-CoV-2 infection; xiii) diagnosis of COVID-19; xiv) hospitalisation with COVID-19; xv) disease severity in those hospitalised with COVID-19; xvi) mortality; xvii) adjustment of smoking specific risk estimates for relevant covariates (e.g.
+current, former, never, not stated, missing) and whether it was biochemically verified; x) use of alternative nicotine products (e.g.
+e-cigarettes); xi) SARS-CoV-2 testing; xii) SARS-CoV-2 infection; xiii) diagnosis of COVID-19; xiv) hospitalisation with COVID-19; xv) disease severity in those hospitalised with COVID-19; xvi) mortality; xvii) adjustment of smoking specific risk estimates for relevant covariates (e.g.
 age, sex); and xviii) whether a representative or random sampling method was used.
@@
 A Bayesian approach mitigates the issue of performing multiple statistical tests, which can inflate family-wise error.
-A series of random-effects hierarchical Bayesian meta-analyses were performed with the brms [@burkner_advanced_2018] package to estimate the relative risk for each comparison with accompanying 95% credible intervals (CrIs).
+A series of random-effects hierarchical Bayesian meta-analyses were performed with the brms [@burkner_advanced_2017] package to estimate the relative risk for each comparison with accompanying 95% credible intervals (CrIs).
 We first defined prior distributions for the true pooled effect size (µ) and the between-study heterogeneity (τ), with µ specified as a normal distribution with a mean equal to the derived point estimate from each comparison of interest in the immediately preceding version of this living review, and τ specified as a half-Cauchy distribution with a mean of 0 and standard deviation of 1.
@@
 
-In the current review version (`r paste(current_version)`) with searches up to `r paste(tail(search_details$date_screened, 1))`, a total of `r paste(prisma$"Titles/abstracts screened")` records were identified, with `r paste(prisma$"Studies in narrative synthesis")` studies included in a narrative synthesis and `r paste(nrow(running_meta_count))` studies included in meta-analyses (see [Figure 1](#fig-1)).
+In the current review version (`r paste(current_version)`) with searches up to `r paste(tail(search_details$date_screened, 1))`, a total of 1940 records were identified, with `r paste(prisma$"Studies in narrative synthesis")` studies included in a narrative synthesis and `r paste(nrow(running_meta_count))` studies included in meta-analyses (see [Figure 1](#fig-1)).
 
@@
 Studies were conducted across `r paste(nrow(country))` countries.
-`r paste(all_countries[[1]][1,2])` studies were conducted in the `r paste(all_countries[[1]][1,1])`, `r paste(all_countries[[2]][1,2])` in `r paste(all_countries[[2]][1,1])`, `r paste(all_countries[[3]][1,2])` in the `r paste(all_countries[[3]][1,1])`, `r paste(all_countries[[4]][1,2])`  in `r paste(all_countries[[4]][1,1])`, `r paste(all_countries[[5]][1,2])`  in `r paste(all_countries[[5]][1,1])`, `r paste(all_countries[[6]][1,2])` in `r paste(all_countries[[6]][1,1])`, `r paste(all_countries[[7]][1,2])` in `r paste(all_countries[[7]][1,1])`, `r paste(all_countries[[8]][1,2])` in `r paste(all_countries[[8]][1,1])`, `r paste(all_countries[[9]][1,2])` in `r knitr::combine_words(unlist(all_countries[[9]][1]))`, `r paste(all_countries[[10]][1,2])` in `r knitr::combine_words(unlist(all_countries[[10]][1]))`, `r paste(all_countries[[11]][1,2])` in `r knitr::combine_words(unlist(all_countries[[11]][1]))`, `r paste(all_countries[[12]][1,2])` in `r knitr::combine_words(unlist(all_countries[[12]][1]))`, with `r paste(all_countries[[13]][1,2])` each from `r knitr::combine_words(unlist(all_countries[[13]][1]))` and a single study from `r paste(nrow(all_countries[[14]][1]))` further countries.
+`r paste(all_countries[[1]][1,2])` studies were conducted in the USA, `r paste(all_countries[[2]][1,2])` in China, `r paste(all_countries[[3]][1,2])` in the UK, `r paste(all_countries[[4]][1,2])`  in Spain, `r paste(all_countries[[5]][1,2])`  in  Italy, `r paste(all_countries[[6]][1,2])` in  France, `r paste(all_countries[[7]][1,2])` in Mexico, `r paste(all_countries[[8]][1,2])` in multiple countries, `r paste(all_countries[[9]][1,2])` in Switzerland, `r paste(all_countries[[10]][1,2])` in `r knitr::combine_words(unlist(all_countries[[10]][1]))`, `r paste(all_countries[[11]][1,2])` in `r knitr::combine_words(unlist(all_countries[[11]][1]))`, `r paste(all_countries[[12]][1,2])` in `r knitr::combine_words(unlist(all_countries[[12]][1]))`, with `r paste(all_countries[[13]][1,2])` in `r knitr::combine_words(unlist(all_countries[[13]][1]))` and 97 studies from 39 further countries.
 The majority of studies used observational designs (see [Supplementary table S1](#supplementary-t1)).
 `r paste(setting[1,2])` (`r paste(round(setting[1,2]/sum(setting[2])*100, 1))`%) were conducted in `r paste(tolower(setting[1,1]))` settings, `r paste(setting[2,2])` studies (`r paste(round(setting[2,2]/sum(setting[2])*100, 1))`%) included individuals from `r paste(tolower(setting[2,1]))` settings, `r paste(setting[3,2])` studies (`r paste(round(setting[3,2]/sum(setting[2])*100, 1))`%) were conducted exclusively in the `r paste(tolower(setting[3,1]))`, with one study each conducted in a homeless shelter and a quarantine centre, and one study that did not state the study setting.
-Studies had a median of `r paste(participants[3])` (interquartile range = `r paste(participants[2], participants[5], sep = "-")`) participants.
-The majority of studies (`r paste(round(sum(diagnostics[5]+diagnostics[6])/sum(diagnostics)*100,1))`%) used reverse transcriptase polymerase chain reaction (RT-PCR) for confirmation of SARS-CoV-2 infection, `r paste(round(sum(diagnostics[1]+diagnostics[6])/sum(diagnostics)*100,1))`% used an antibody test to confirm prior infection and `r paste(round(diagnostics[6]/sum(diagnostics)*100,1), "%", sep = "")` of studies relied on a combination of RT-PCR or antibody assays.
+Studies had a median of `r paste(participants[3])` (interquartile range = `r paste(round(participants[2], 0), round(participants[5], 0), sep = "-")`) participants.
+The majority of studies (`r paste(round(sum(diagnostics[6]+diagnostics[7])/sum(diagnostics)*100,1))`%) used reverse transcriptase polymerase chain reaction (RT-PCR) for confirmation of SARS-CoV-2 infection, `r paste(round(sum(diagnostics[1]+diagnostics[7])/sum(diagnostics)*100,1))`% used an antibody test to confirm prior infection and `r paste(round(sum(diagnostics[7] + diagnostics[2])/sum(diagnostics)*100,1), "%", sep = "")` of studies relied on a combination of RT-PCR, antigen or antibody assays.
 
@@
 Categorisation of smoking status was heterogeneous (see [Table 1](#table-1)).
-`r paste(data_source[1,2])` studies collected data on smoking status through routine electronic health records (EHRs), `r paste(data_source[2,2])` studies used a bespoke case report form for COVID-19 and `r paste(data_source[3,2])` studies did not state the source for information on smoking status.
+`r paste(data_source[1,2])` `r paste(round(data_source[1,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies collected data on smoking status through routine electronic health records (EHRs), `r paste(data_source[2,2])` `r paste(round(data_source[2,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies used a bespoke case report form for COVID-19 and `r paste(data_source[3,2])` `r paste(round(data_source[3,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies did not state the source for information on smoking status.
 None of the studies verified smoking status biochemically.
@@
 
-Eight studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use [@miyara_low_2020; @rimland_clinical_2020; @girardeau_confirmed_2020; @ebinger_sars-cov-2_2020; @islam_risk_2020; @gallichotte_longitudinal_2020; @kantele_sars-cov-2_2020; @crooks_predicting_2020].
-One additional study grouped together current smokers and vapers [@schubl_risk_2020].
+Fifteen studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use (see [Table 1](#table-1).
 
@@
 
-Three studies were performed in random or representative population samples and were rated as ‘good’ quality, and `r paste(nrow(fair_quality_number))` studies were rated as ‘fair’ quality, of which `r paste(nrow(running_meta_count))` studies reported results stratified by smoking status for the outcomes of interest and could be included in meta-analyses.
+Nine studies were performed in random or representative population samples and were rated as ‘good’ quality, and `r paste(nrow(quality_rating %>% filter(overall_quality == "fair")))` studies were rated as ‘fair’ quality, of which `r paste(nrow(running_meta_count))` studies reported results stratified by smoking status for the outcomes of interest and could be included in meta-analyses.
 The remaining `r paste(nrow(poor_quality_number))` studies were rated as ‘poor’ quality (see [Table 1](#table-1)).
@@
   font.color = "black",
-  digits = 2,
+  digits = 1,
   theme_fun = "theme_booktabs"
@@
   arrange(`Date published`) %>%
-  mutate(`Lead author` = recode(`Lead author`, "Schubl" = "Schuble*"))
+  mutate(`Lead author` = recode(`Lead author`, "Schubl" = "Schuble*"),
+         `Missing %` = as.numeric(`Missing %`))
 
@@
   colformat_md(j = "ref") %>%
-  colformat_double(j = "Sample size", digits = 0)
+  colformat_double(j = "Sample size", digits = 0) %>%
+  colformat_double(j = 8:21, digits = 1)
 ```
@@
 
-![*Figure 2a.* Weighted mean prevalence of current smoking in included studies, split by country (solid lines).
+![*Figure 2.* Weighted mean prevalence of current (left panel) and former (right panel) smoking in included studies, split by country.
 The circles represent individual studies, with colour corresponding to study setting (i.e.
-community, community and hospital, hospital) and size corresponding to relative study sample size.
-For comparison, national current smoking prevalence is shown by the dashed red lines.
-Countries with three or more eligible studies are shown.](figure/current_smoking_plots_updated.png){#fig-2a}
-
-![*Figure 2b.* Prevalence of former smoking in included studies, split by country (solid lines).
-The circles represent individual studies, with colour corresponding to study setting (i.e.
-community, community and hospital, hospital) and size corresponding to relative study sample size.
-For comparison, national former smoking prevalence is shown by dashed red lines.
-Countries with three or more eligible studies are shown.](figure/former_smoking_plots_updated.png){#fig-2b}
+hospital, community and hospital, community) and size corresponding to relative study sample size.
+For comparison, national current and former smoking prevalence are indicated by the red lines.
+Countries with three or more eligible studies are shown.](figure/combined_smoking_plots_updated.png){#fig-2a}
 
@@
 Current smokers were at reduced risk of testing positive for SARS-CoV-2 compared with never smokers (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_tau, 2))`).
-The three good quality studies each reported point estimates less than 1, although the CrI for one of the three studies crossed 1.
+The six good quality studies each reported point estimates less than 1, although the CrI for two of the six studies crossed 1.
 The probability of current smokers being at reduced risk of infection compared with never smokers (RR ≤0.9) was >99%.
 Former compared with never smokers were at increased risk of testing positive, but data were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_tau, 2))`) and favoured there being no important association.
-The probability of former smokers being at increased risk of infection (RR ≥1.1) compared with never smokers was `r paste(round(1-m2_a_ecdf(1.1), 2)*100, "%", sep = "")`.
+The probability of former smokers being at increased risk of infection (RR ≥1.1) compared with never smokers was < 1%.
 Results were materially unchanged in sensitivity analyses.
@@
 * Indicates ‘good’ quality studies.
-The prior from the previous review version (v10) was RR = 0.69.](figure/m1_a.png){#fig-3}
+The prior from the previous review version (v11) was RR = 0.71.](figure/m1_a.png){#fig-3}
 
@@
 * Indicates ‘good’ quality studies.
-The prior from the previous review version (v10) was RR = 1.02.](figure/m2_a.png){#fig-4}
+The prior from the previous review version (v11) was RR = 1.03.](figure/m2_a.png){#fig-4}
 
@@
 `r paste(nrow(table_3_word))` studies examined hospitalisation for COVID-19 disease, stratified by smoking status (see [Table 3](#table-3)).
-Meta-analyses were performed for `r paste(nrow(fair_quality_3))` ‘fair’ quality studies (see [Figure 5](#fig-5) and [6](#fig-6)).
+Meta-analyses were performed for one 'good' and 18 ‘fair’ quality studies (see [Figure 5](#fig-5) and [6](#fig-6)).
 Current (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_tau, 2))`) and former (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_tau, 2))`) compared with never smokers were at increased risk of hospitalisation with COVID-19.
 However, data for current smokers were inconclusive and favoured there being no important association.
-The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(round(1-m3_a_ecdf(1.1),2)*100,"%", sep = "")` and `r paste(round(1-m4_a_ecdf(1.1),2)*100, "%", sep = "")`, respectively.
+The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(round(1-m3_a_ecdf(1.1),2)*100,"%", sep = "")` and >99%, respectively.
 Results were materially unchanged in two sensitivity analyses.
@@
 
-![*Figure 5.* Forest plot for risk of hospitalisation in current vs. never smokers. The prior from the previous review version (v10) was RR = 1.08.](figure/m3_a.png){#fig-5}
+![*Figure 5.* Forest plot for risk of hospitalisation in current vs. never smokers. The prior from the previous review version (v11) was RR = 1.1.](figure/m3_a.png){#fig-5}
 
-![*Figure 6.* Forest plot for risk of hospitalisation in former vs. never smokers. The prior from the previous version (v10) was RR = 1.18.](figure/m4_a.png){#fig-5}
+![*Figure 6.* Forest plot for risk of hospitalisation in former vs. never smokers. The prior from the previous version (v11) was RR = 1.19.](figure/m4_a.png){#fig-5}
 
@@
 Severe (as opposed to non-severe) disease was broadly defined as requiring intensive treatment unit (ITU) admission, requiring oxygen as a hospital inpatient or in-hospital death.
-Meta-analyses were performed for `r paste(nrow(fair_quality_4))` ‘fair’ quality studies (see [Figure 7](#fig-7) and [8](#fig-8)).
+Meta-analyses were performed for 16 ‘fair’ quality studies (see [Figure 7](#fig-7) and [8](#fig-8)).
 Current (RR = `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_tau, 2))`) and former (RR = `r paste(round(former_severity_RR_value, 2))`, 95% CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_tau, 2))`) compared with never smokers were at increased risk of greater disease severity.
-However, while data for current smokers only were inconclusive, they favoured there being a small but important association.
 The probability of current and former smokers having increased risk of greater disease severity (RR ≥1.1) compared with never smokers was `r paste(round(1-m5_a_ecdf(1.1),2)*100, "%", sep = "")` and `r paste(round(1-m6_a_ecdf(1.1),2)*100, "%", sep = "")`, respectively.
@@
 
-![*Figure 7.* Forest plot for the risk of severe disease in current vs. never smokers. The prior from the previous review version (v10) was RR = 1.26.](figure/m5_a.png){#fig-7}
+![*Figure 7.* Forest plot for the risk of severe disease in current vs. never smokers. The prior from the previous review version (v11) was RR = 1.26.](figure/m5_a.png){#fig-7}
 
-![*Figure 8.* Forest plot for the risk of severe disease in former vs. never smokers. The prior from the previous review version (v10) was RR = 1.52.](figure/m6_a.png){#fig-8}
+![*Figure 8.* Forest plot for the risk of severe disease in former vs. never smokers. The prior from the previous review version (v11) was RR = 1.8.](figure/m6_a.png){#fig-8}
 
@@
 
-`r paste(nrow(table_5_word))` studies reported mortality from COVID-19 by smoking status (see [Table 5](#table-5)), with `r paste(nrow(fair_quality_5))` ‘fair’ quality studies included in meta-analyses (see [Figure 9](#fig-9) and [10](#fig-10)).
+`r paste(nrow(table_5_word))` studies reported mortality from COVID-19 by smoking status (see [Table 5](#table-5)), with 28 ‘fair’ quality studies included in meta-analyses (see [Figure 9](#fig-9) and [10](#fig-10)).
 Current (RR = `r paste(round(current_mortality_RR_value, 2))`, 95% CrI = `r paste(round(current_mortality_lci, 2), round(current_mortality_uci, 2), sep = "-")`, τ = `r paste(round(current_mortality_tau, 2))`) and former (RR = `r paste(round(former_mortality_RR_value, 2))`, 95% CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_tau, 2))`) compared with never smokers were at increased risk of in-hospital mortality from COVID-19.
@@
 
-![Figure 9. Forest plot for the risk of mortality in current vs. never smokers. The prior from the previous review version (v10) was RR = 1.05.](figure/m7_a.png){#fig-9}
+![Figure 9. Forest plot for the risk of mortality in current vs. never smokers. The prior from the previous review version (v11) was RR = 1.12.](figure/m7_a.png){#fig-9}
 
-![Figure 10. Forest plot for the risk of mortality in former smokers vs. never smokers. The prior from the previous version (v10) was RR = 1.40.](figure/m8_a.png){#fig-10}
+![Figure 10. Forest plot for the risk of mortality in former smokers vs. never smokers. The prior from the previous version (v11) was RR = 1.56.](figure/m8_a.png){#fig-10}
 
@@
 In a subset of good and fair quality studies (n = `r paste(nrow(good_quality_2)+nrow(fair_quality_2))`), current but not former smokers had a reduced risk of testing positive for SARS-CoV-2 but current smokers appeared somewhat more likely to present for testing and/or receive a test.
-Data for current smokers on the risk of hospitalisation, disease severity and mortality were inconclusive, and favoured there being no important associations with hospitalisation and mortality and a small but important increase in the risk of severe disease.
-However, it should be noted that pooled estimates for the risk of hospitalisation and mortality in current smokers have numerically increased since the previous version (v10).
+Current compared with never smokers had an increased risk of greater disease severity.
+Data for current smokers on the risk of hospitalisation and mortality were inconclusive, and favoured there being no important associations.
 Former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.
@@
 Infection positivity rates estimated among random samples are more informative.
-We identified one population study conducted in Hungary reporting on seroprevalence and smoking status [@merkely_novel_2020]; however, the response rate was only 58.8% and the current smoking rate was 10 percentage points below national prevalence estimates, which raises some doubt about representativeness of the final sample.
-Similarly, two further representative population surveys [@carrat_seroprevalence_2020; @richard_seroprevalence_2020] reported a current smoking rate of more than 10 percentage points below national prevalence (12% vs.
+We identified nine random/population studies reporting on seroprevalence and smoking status.
+However, response rates were suboptimal where reported (e.g.
+58.8%-64.3%; [@merkely_novel_2020] and [@wagner_estimates_2021]) and the current smoking rate was often several percentage points below national prevalence estimates (e.g.
+12% vs.
 25% and 15% vs.
-27% daily smoking prevalence, respectively) [@andler_baisse_2019; @statistique_part_2020].
+27% daily smoking prevalence, respectively; [@andler_baisse_2019] and [@statistique_part_2020]), which raises doubt about the representativeness of the final samples.
   4. *Testing positive for SARS-CoV-2*      
@@
 Known failings of EHRs include implausible longitudinal changes, such as former smokers being recorded as never smokers at subsequent hospital visits [@polubriaginof_challenges_2018].
+A recent study conducted in a UK hospital setting found a large discrepancy between smoking status recorded on the EHR and within the contemporaneous clinical notes [@simons_association_2020-2].
+Known failings of EHRs include implausible longitudinal changes, such as former smokers being recorded as never smokers at subsequent hospital visits [@polubriaginof_challenges_2018].
 Misreporting on the part of the patient (perhaps due to perceived stigma) has also been observed, with biochemical measures showing higher rates of smoking compared with self-report in hospitalised patients in the US [@benowitz_prevalence_2009].
-It is hence possible that under-reporting of current and former smoking status in hospitals occurred across the included studies.
+It is hence likely that under-reporting of current and former smoking status in hospitals occurred across the included studies.
     3.	The majority of included studies were conducted in hospital settings.
@@
     4.	COVID-19 outcomes are currently limited to in-hospital death or survival to discharge.
-This binary outcome does not capture potential long-term morbidity attributed to COVID-19, such as stroke, amputation or acute cardiac events, which may be moderated by smoking status.
+This binary outcome does not capture potential long-term morbidity attributed to COVID-19, such as stroke, amputation, acute cardiac events, or long COVID (i.e.
+symptoms that persist for weeks or months beyond infection), which may be moderated by smoking status.
 
@@
 Across `r paste(current_total)` studies, recorded current but not past smoking prevalence was generally lower than national prevalence estimates.
-Current smokers were at reduced risk of testing positive for SARS-CoV-2 and former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.
+Current smokers were at reduced risk of testing positive for SARS-CoV-2 and at increased risk of greater disease severity.
+Former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.
 
@@
  * [Version 10](https://doi.org/10.32388/UJR2AW.11) 
+ * [Version 11](https://doi.org/10.32388/UJR2AW.13)
 

### 2021-07-27 reports/manuscript_compile.R

@@ -1,4 +1,4 @@
-source(here("scripts", "libraries.R"))
+source(here::here("scripts", "libraries.R"))
 source(here("scripts", "bayesian_data_summary.R"))
 
 rmarkdown::render(here("reports","manuscript.Rmd"), output_format = "html_document")

### 2021-07-27 scripts/bayesian_analysis.R

@@ -1,10 +1,12 @@
-source(here("scripts", "libraries.R"))
+source(here::here("scripts", "libraries.R"))
 source(here("scripts", "bayes_scripts.R"))
 source(here("scripts", "author_dictionary.R"))
+source(here("scripts", "bayesian_data_generation.R"))
 versions <- read_rds(here("data_clean", "versioning.rds"))
 
 previous_review_versions <- unique(versions[,1])
 current_review_version <- unique(versions[,2])
+current_version <- unique(versions[,2])
 
 data_study_general <- read_rds(here("data_clean", "data_study_general.rds"))
 
@@ -21,7 +23,7 @@ study_review_version <- data_study_general %>%
 levels(study_review_version$review_version) <- c(levels(study_review_version$review_version), "combined_pooled") %>%
   fct_relevel(c("previous", "current", "combined_pooled"))
 
-m1 <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1.rds"))
+#m1 <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1.rds"))
 m1_a <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1_a.rds"))
 #m1_a_hh <- read_rds(here("data_clean", "bayesian_models", "testing_current_m1_a_hh.rds"))
 #m2 <- read_rds(here("data_clean", "bayesian_models", "testing_former_m2.rds"))
@@ -68,9 +70,13 @@ m1_a_ecdf(0.9) #informative
 # pooled_draw_m1 <- pooled_effect_draw(m1)
 
 study_draws_m1_a <- study_draw(m1_a) %>%
-  mutate(Author = ifelse(Author == "Merkely", "Merkely*",
-                         ifelse(Author == "Carrat", "Carrat*",
-                                ifelse(Author == "Richard", "Richard*", Author))))
+  mutate(Author = case_when(Author == "Merkely" ~ "Merkely*",
+                            Author == "Carrat" ~ "Carrat*", 
+                            Author == "Richard" ~ "Richard*",
+                            Author == "Barchuk" ~ "Barchuk*",
+                            Author == "Gornyk" ~ "Gornyk*",
+                            Author == "Radon" ~ "Radon*",
+                            TRUE ~ Author))
 pooled_draw_m1_a <- pooled_effect_draw(m1_a)
 
[…]

### 2021-07-27 scripts/bayesian_data_generation.R

@@ -2,7 +2,7 @@
 
 set.seed(42)
 
-source(here("scripts", "libraries.R"))
+source(here::here("scripts", "libraries.R"))
 source(here("scripts", "bayes_scripts.R"))
 source(here("scripts", "author_dictionary.R"))
 
@@ -65,67 +65,65 @@ minimally_informative_prior <-
 # Previous meta-analysis guided priors
 
 current_testing_prior <-
-  c(prior(normal(-.371, 1), class = Intercept),
+  c(prior(normal(-.3425, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_testing_prior <-
   c(prior(normal(0.0296, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_hospitalisation_prior <-
-  c(prior(normal(0.0770, 1), class = Intercept),
+  c(prior(normal(0.0953, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_hospitalisation_prior <-
-  c(prior(normal(0.1655, 1), class = Intercept),
+  c(prior(normal(0.1740, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_severity_prior <-
   c(prior(normal(0.2311, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_severity_prior <-
-  c(prior(normal(0.4187, 1), class = Intercept),
+  c(prior(normal(0.5878, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 current_mortality_prior <-
-  c(prior(normal(0.0487, 1), class = Intercept),
+  c(prior(normal(0.1133, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 former_mortality_prior <- 
-  c(prior(normal(0.3365, 1), class = Intercept),
+  c(prior(normal(0.4447, 1), class = Intercept),
     prior(cauchy(0, 1), class = sd))
 
 # Classical meta-analysis guided priors with high heterogeneity this is used as sensitivity analysis
 
 current_testing_prior_hh <-
-  c(prior(normal(-.371, 1), class = Intercept),
+  c(prior(normal(-.3425, 1), class = Intercept),
     prior(cauchy(0.3, 1), class = sd))
 former_testing_prior_hh <-
   c(prior(normal(0.0296, 1), class = Intercept),
[…]

### 2021-07-27 scripts/bootstrap_function.R

@@ -1,20 +1,0 @@
-boot.mean = function(x,B,binwidth=NULL) {
-  n = length(x)
-  boot.samples = matrix(sample(x,size=n*B,replace=TRUE), B, n)
-  boot.statistics = apply(boot.samples,1,mean)
-  se = sd(boot.statistics)
-  interval=quantile(boot.statistics,c(0.025,0.975))
-  return(list(boot.statistics=boot.statistics,interval=interval,se=se))
-}
-
-do_bootstrap_current <- function (row_number) {
-  my_sample <- c(rep(1, each = bootstrap$bs_current[row_number]), rep(0, each = bootstrap$bs_inverse[row_number]))
-  bootstrap_output_c[[row_number]] <- boot.mean(my_sample,1000)$interval
-  return(bootstrap_output_c)
-}
-
-do_bootstrap_former <- function (row_number) {
-  my_sample <- c(rep(1, each = bootstrap$bs_former[row_number]), rep(0, each = bootstrap$bs_inverse_former[row_number]))
-  bootstrap_output_f[[row_number]] <- boot.mean(my_sample,1000)$interval
-  return(bootstrap_output_f)
-}

### 2021-07-27 scripts/clean_data.R

@@ -23,8 +23,9 @@ table_1 <- table_1 %>%
          upper_range = sub('.*_', '',.$range,)) %>% ## We break apart the age range into an upper and lower level
   select(lead_author, sample_size, female_sex_percent, data_source, median_age, mean_age, iqr_lower, iqr_upper, standard_deviation, lower_range,
          upper_range, current_smoker, former_smoker, current_former_smoker, never_smoker, never_smoker_unknown, current_vaper, former_vaper,
-         current_former_vaper, not_stated, missing, total) %>%
-  mutate(study_id = 1:length(lead_author)) %>%
+         current_former_vaper, never_vaper, current_smokeless_tobacco, current_smoker_vaper, former_smoker_vaper, never_smoker_vaper,
+         current_smoker_former_vaper, former_smoker_former_vaper, never_smoker_former_vaper, not_stated, missing, total) %>%
+  mutate(study_id = row_number()) %>%
   filter(!is.na(lead_author)) %>%
   left_join(., review_details %>%
               select(-"lead_author"), by = 'study_id') %>%
@@ -35,7 +36,7 @@ table_1 <- table_1 %>%
 
 table_2 <-  table_2 %>%
   clean_data() %>%
-  mutate(study_id = 1:length(lead_author)) %>%
+  mutate(study_id = row_number()) %>%
   filter(data_on_testing == TRUE) %>%
   left_join(., review_details %>%
               select(-"lead_author"), by = 'study_id') %>%
@@ -43,7 +44,7 @@ table_2 <-  table_2 %>%
 
 table_3 <-  table_3 %>%
   clean_data() %>%
-  mutate(study_id = 1:length(lead_author)) %>%
+  mutate(study_id = row_number()) %>%
   filter(data_on_hospitalisation == TRUE) %>%
   left_join(., review_details %>%
               select(-"lead_author"), by = 'study_id') %>%
@@ -51,7 +52,7 @@ table_3 <-  table_3 %>%
 
 table_4 <-  table_4 %>%
   clean_data() %>%
-  mutate(study_id = 1:length(lead_author)) %>%
+  mutate(study_id = row_number()) %>%
   filter(data_disease_severity == TRUE) %>%
   left_join(., review_details %>%
               select(-"lead_author"), by = 'study_id') %>%
@@ -59,7 +60,7 @@ table_4 <-  table_4 %>%
 
 table_5 <- table_5 %>%
   clean_data() %>%
[…]

### 2021-07-27 scripts/data_cleaning.R

@@ -1,65 +1,0 @@
-library('tidyverse')
-library('dplyr')
-library('readxl')
-library('linelist')
-library('googlesheets4')
-
-sheets_id <- as_sheets_id('https://docs.google.com/spreadsheets/d/15avypGR8ypJngWQEmFIzFrOOwXPY3xezUHQU6jgV7d0/edit?usp=sharing')
-sheets_id
-
-search_details <- read_sheet(sheets_id, range = 'article_screening')
-  
-
-search_details <- search_details %>%
-  clean_data() %>%
-  rename(date_screened = date_screening_in_format_2020_05_22) %>%
-  write_rds(here::here('data_clean', 'search_details.rds')
-
-data_study_general <- read_sheet(sheets_id, range =  'general_details')
-protect_columns_1 <- names(data_study_general) %in% 'doi'
-
-data_study_general <- data_study_general %>%
-  rename('notes' = 11, date_published = `date_published_format(2020-05-12)`) %>%
-  clean_data(protect = protect_columns_1)
-write_rds(data_study_general, 'data_clean/data_study_general.rds')
-
-review_details <- data_study_general %>%
-  select(lead_author, date_published, country, review_version)
-write_rds(review_details, 'data_clean/review_details.rds')
-
-table_1 <- read_sheet(sheets_id, sheet = 'pop_descriptives') %>%
-  clean_data() %>%
-  mutate(., lower_range = sub('\\_.*', '', .$range)) %>%
-  mutate(., upper_range = sub('.*_', '',.$range,)) %>%
-  select(1:9, 22:23, 11:21) %>%
-  filter(!is.na(lead_author)) %>%
-  left_join(., review_details, by = 'lead_author') %>%
-  write_rds(., 'data_clean/data_table_1.rds')
-   
-table_2 <-  read_sheet(sheets_id, sheet = 'testing') %>%
-  clean_data() %>%
-  filter(data_on_testing == TRUE) %>%
-  left_join(., review_details, by = 'lead_author') %>%
-  write_rds(., 'data_clean/data_table_2.rds')
-
-table_3 <-  read_sheet(sheets_id, sheet = 'hospitalisation') %>%
-  clean_data() %>%
-  filter(data_on_hospitalisation == TRUE) %>%
-  left_join(., review_details, by = 'lead_author') %>%
-  write_rds(., 'data_clean/data_table_3.rds')
-
[…]

### 2021-07-27 scripts/descriptives.R

@@ -1,306 +1,0 @@
-##Run if needing to update the studies
-source('scripts/data_cleaning.R')
-##Run if data already available
-library(tidyverse)
-library(linelist)
-source('scripts/prisma_function.R')
-
-date_of_update <- Sys.Date()
-date_of_last_update <- date_of_update-10
-prev_versions <- c('v1', 'v2')
-current_version <- c('v3')
-analysed_versions <- c('v1', 'v2', 'v3')
-exclude_from_analysis <- c('isaric_1', 'isaric_2', 'isaric_3', 'miyara_old')
-exclude_from_qa <- c('ISARIC_1', 'ISARIC_2', 'ISARIC_3', 'Miyara')
-
-search_details <- read_rds("data_clean/search_details.rds")
-
-data_study_general <- read_rds("data_clean/data_study_general.rds")
-data_study_general <- data_study_general %>%
-  filter(review_version %in% analysed_versions) %>%
-  filter(!lead_author %in% exclude_from_analysis) 
-
-table_1 <- read_rds("data_clean/data_table_1.rds")
-table_1 <- table_1 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!lead_author %in% exclude_from_analysis)
-
-table_2 <- read_rds("data_clean/data_table_2.rds")
-table_2 <- table_2 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!lead_author %in% exclude_from_analysis)
-
-table_3 <- read_rds("data_clean/data_table_3.rds")
-table_3 <- table_3 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!lead_author %in% exclude_from_analysis)
-
-table_4 <- read_rds("data_clean/data_table_4.rds")
-table_4 <- table_4 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!lead_author %in% exclude_from_analysis)
-
-table_5 <- read_rds("data_clean/data_table_5.rds")
-table_5 <- table_5 %>%
-  filter(review_version %in% analysed_versions) %>%
-  select(-review_version) %>%
-  filter(!lead_author %in% exclude_from_analysis)
-
-table_6 <- table_6 %>%
-  filter(!Author %in% exclude_from_qa)
-  
-
-#Number screened to add to PRISMA
[…]

### 2021-07-27 scripts/libraries.R

@@ -29,6 +29,11 @@ pkgs =
     "snakecase",
     "revtools",
     "rnaturalearth",
-    "sf"
+    "sf",
+    "leaflet",
+    "coronavirus",
+    "cowplot"
   )
 pacman::p_load(pkgs, character.only = T)
+
+write_rds(sessionInfo(), here("reports", "all_packages.rds"))

### 2021-07-27 scripts/references_in_table.R

@@ -5,9 +5,21 @@ library(ftExtra)
 library(bib2df)
 
 tbl_1 <- readRDS(here::here("data_clean", "table_1_refs.rds")) %>%
-  mutate(ref = ifelse(startsWith(ref, "h"), gsub(".*/10", "10", ref), ref))
+  mutate(ref = ifelse(startsWith(ref, "h"), gsub(".*/10", "10", ref), ref),
+         ref = str_remove(ref, "v\\d.full.pdf$"),
+         ref = str_remove(ref, "v10.full.pdf"),
+         ref = case_when(`Lead author` != "Tehrani" ~ str_remove(ref, "v\\d$"),
+                         TRUE ~ ref)) %>%
+  distinct(`Lead author`, `Sample size`, .keep_all = T)
 bib <- bib2df(here::here("reports", "manuscript.bib")) %>%
   select(BIBTEXKEY, DOI)
+
+missing_doi <- tbl_1 %>%
+  filter(!ref %in% bib$DOI) %>%
+  mutate(ref = str_remove(ref, "v\\d.full.pdf$"))
+
+#This is the vector to lookup missing doi's in zoter
+cat(paste(missing_doi$ref, collapse = ", "))
 
 tbl_1 <- left_join(tbl_1, bib %>%
               rename("ref" = DOI),

### 2021-07-27 scripts/smoking_plots.R

@@ -2,7 +2,6 @@
 source(here::here("scripts", "libraries.R"))
 library(ggsci)
 library(matrixStats)
-library(gapminder)
 
 smoking_status <- c('current_smoker', 'former_smoker', 'current_former_smoker', 'never_smoker', 'never_smoker_unknown', 'not_stated', 'missing')
 exclude_from_analysis <- read_rds(here("output_data", "excluded_studies.rds"))
@@ -69,39 +68,6 @@ current_smoking_data <- smoking_data %>%
   filter(country %in% current_smoking_plot$country) %>%
   mutate(country = factor(country, levels = ordered_country))
 
-## new plots current
-ggplot() +
-  geom_point(data = current_smoking_plot,
-             aes(x = study_setting, y = p_current_smoker, colour = study_setting, size = sample),
-             position = position_dodge2(0.3),
-             alpha = 0.3)  +
-  scale_size_continuous(trans = "log10", guide = F) +
-  geom_hline(data = current_smoking_data,
-             aes(yintercept = p_current_smoking),
-             linetype = "dashed", 
-             colour = "#A50026",
-             size = 0.8) +
-  geom_hline(data = current_smoking_plot ,
-             aes(yintercept = weighted_mean, colour = study_setting)) +
-  coord_flip() +
-  facet_grid(country ~ .,  scales = "free_y", space = "free", switch = 'y') +
-  scale_color_brewer(palette = "Dark2", labels = c("Hospital", "Community & Hospital", "Community")) +
-  scale_y_continuous(name = "Prevalence", limits = c(0, 0.7), breaks = scales::pretty_breaks(n = 9), expand = c(0, 0)) +
-  theme(panel.spacing = unit(0.1, "lines"),
-        strip.text = element_text(angle = 0, size = 12),
-        axis.title = element_text(size = 10),
-        axis.text.y = element_blank(),
-        axis.title.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.text.x = element_text(size = 10),
-        panel.background = element_blank(),
-        axis.line.x = element_line(colour = "black"),
-        strip.text.y.left = element_text(angle = 0),
-        legend.position = "right") +
-  labs(title ="Prevalence of current smoking in included studies",
-       colour = "Weighted mean prevalence", shape = "Study setting")
-ggsave("current_smoking_plots_updated.png", plot = last_plot(), dpi = 360, path = here("reports", "figure"))
-
 ## data former
 former_smoking <- prevalence_plot %>%
[…]

### 2021-07-27 reports/bayesian_report.Rmd

@@
 ``` {r google_data}
-google_api <- rstudioapi::askForSecret("Google API Key")
-register_google(google_api)
-drive_download("https://docs.google.com/spreadsheets/d/1VUY5xDLBpjNBovveg9wQrc0wbLhnEkmDPMR63rfj2-Y/edit?usp=sharing", path = here("data_spreadsheet", "data_extraction_current.xlsx"), overwrite = T)
+drive_download("https://docs.google.com/spreadsheets/d/1fmTEW0lHYDk1OE3fw9-mJ6xu9EgBPnYoaL05-OTRiOk/edit?usp=sharing", path = here("data_spreadsheet", "data_extraction_current.xlsx"), overwrite = T)
 
@@
 ```{r clean_data}
-source(here("scripts", "clean_data.r"))
+source(here("scripts", "clean_data.R"))
 ```
@@
 ``` {r descriptives}
-source(here("scripts", "remove_excluded.r")) # This removes the excluded studies from each table that has been read in previously.
+source(here("scripts", "remove_excluded.R")) # This removes the excluded studies from each table that has been read in previously.
 current_total <- nrow(a)
@@
 ``` {r interactive global map, cache = TRUE, quietly = TRUE, echo=FALSE, results = "hide"}
-library(rgdal)
-library(leaflet)
-library(coronavirus)
-
-my_spdf <- readOGR( 
-  dsn= paste0(here::here("data_clean", "world_shape_file")), 
-  layer="TM_WORLD_BORDERS_SIMPL-0.3",
-  verbose=FALSE
-)
-
-from_countries <- c("USA", "UK", "Iran",  "South Korea")
-to_countries <- c("United States", "United Kingdom", "Iran (Islamic Republic of)", "Korea, Republic of")
+from_countries <- c("USA", "UK", "South Korea", "Czechia")
+to_countries <- c("United States", "United Kingdom", "Korea", "Czech Rep.")
 country_map <- country %>%
@@
   arrange(-total_cases) %>%
-  rename("NAME" = country) %>%
-  mutate(NAME = plyr::mapvalues(NAME, from = from_countries,
-                to = to_countries))
-
-my_spdf@data <- left_join(my_spdf@data, country_map %>%
-            rename(NAME = Country),
-          by = "NAME") %>%
+  rename("name" = country) %>%
+  mutate(name = recode(name,
+         "US" = "United States",
+         "Czechia" = "Czech Rep.",
+         "Korea, South" = "Korea"))
+
+country_plot <- left_join(country_map  %>%
+            rename("name" = Country),
+            countries %>%
+              select(name),
+          by = "name") %>%
   left_join(., summary_df %>%
               mutate(total_cases = total_cases/1000),
-            by = "NAME") %>%
-  mutate(NAME = as.factor(NAME))
-
-my_spdf@data$Number[ which(my_spdf@data$Number == 0)] = NA
-my_spdf@data$total_cases[ which(is.na(my_spdf@data$total_cases))] = 0
-
-my_bins <- c(0, 1, 5, 10, 20, 50, 70, 110)
-mypalette <- leaflet::colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)
+            by = "name") %>%
+  mutate(name = as.factor(name)) %>%
+  st_as_sf()
+
+country_plot$Number[ which(country_plot$Number == 0)] = NA
+country_plot$total_cases[ which(is.na(country_plot$total_cases))] = 0
+
+my_bins <- c(0, 1, 5, 10, 20, 50, 70, max(country$Number))
+mypalette <- colorBin(palette="YlOrBr", domain = country_plot$Number, na.color="transparent", bins=my_bins)
 
 mytext <- paste(
-  "Country: ", my_spdf@data$NAME,"<br/>", 
-  "Number of studies included: ", my_spdf@data$Number, "<br/>", 
-  "COVID-19 reported cases (thousands): ", round(my_spdf@data$total_cases, 2), 
+  "Country: ", country_plot$name, "<br/>", 
+  "Number of studies included: ", country_plot$Number, "<br/>", 
+  "COVID-19 reported cases (thousands): ", round(country_plot$total_cases, 2), 
   sep="") %>%
@@
 
-html_map <- leaflet(my_spdf) %>% 
+html_map <- leaflet(country_plot) %>% 
   addTiles()  %>% 
@@
 ``` {r diagnosis_method}
-diagnostics <- table(data_study_general$diagnostic_approach)
+diagnostics <- table(data_study_general %>%
+                       mutate(diagnostic_approach = case_when(is.na(diagnostic_approach) ~ "not_specified",
+                                                              TRUE ~ diagnostic_approach)) %>%
+                                pull(diagnostic_approach))
 diagnostics
@@
 current_smok <- table_1 %>%
-  filter(., current_smoker != 'NA')
+  filter(current_smoker != 'NA' | current_vaper != "NA" | current_smoker_vaper != 'NA' | current_smoker_former_vaper != 'NA')
 ```
@@
 former_smok <- table_1 %>%
-  filter(., former_smoker != 'NA')
+  filter(., former_smoker != 'NA' | former_smoker_vaper != 'NA' | former_smoker_former_vaper != 'NA')
 ```
@@
 never_smok <- table_1 %>%
-  filter(., never_smoker != 'NA')
+  filter(., never_smoker != 'NA' | never_smoker_vaper != 'NA' | never_smoker_former_vaper != 'NA')
 ```
@@
 full_smoking_status <- table_1 %>%
-  filter(lead_author %in% current_smok$lead_author) %>%
-  filter(lead_author %in% former_smok$lead_author) %>%
-  filter(lead_author %in% never_smok$lead_author)
+  filter(lead_author %in% current_smok$lead_author & 
+           lead_author %in% former_smok$lead_author & 
+           lead_author %in% never_smok$lead_author)
 
 b <- full_smoking_status %>%
-  select(lead_author, sample_size, current_smoker, former_smoker, never_smoker, study_id) %>%
+  select(lead_author, sample_size, current_smoker, former_smoker, never_smoker, current_vaper,
+         current_smoker_vaper, former_smoker_vaper, never_smoker_vaper,
+         current_smoker_former_vaper, former_smoker_former_vaper, never_smoker_former_vaper,
+         study_id) %>%
   rename('Lead author' = lead_author,
@@
          'Current smokers' = current_smoker,
+         'Current vaper' = current_vaper,
          'Former smokers' = former_smoker,
          'Never smokers' = never_smoker,
+         "Current smoker, current vaper" = current_smoker_vaper, 
+         "Former smoker, current vaper" = former_smoker_vaper,
+         "Never smoker, current vaper" = never_smoker_vaper,
+         "Current smoker, former vaper" = current_smoker_former_vaper,
+         "Former smoker, former vaper" = former_smoker_former_vaper, 
+         "Never smoker, former vaper" = never_smoker_former_vaper,
          'Study ID' = study_id)
@@
     set_caption(caption = 'Studies reporting complete smoking status') %>%
-  colformat_double(j = c("Sample size", "Current smokers", "Former smokers", "Never smokers"),
+  colformat_double(j = c("Sample size", "Current smokers", "Former smokers", "Never smokers", "Current vaper",
+                         "Current smoker, current vaper", "Former smoker, current vaper", "Never smoker, current vaper",
+                         "Current smoker, former vaper", "Former smoker, former vaper", "Never smoker, former vaper"),
                 digits = get_flextable_defaults()$digits, big.mark = get_flextable_defaults()$big.mark, na_str = get_flextable_defaults()$na_str) %>%
@@
 semi_full_smoking_status <- table_1 %>%
-  filter(lead_author %in% current_former_smok$lead_author) %>%
-  filter(lead_author %in% never_smok$lead_author) %>%
-  filter(!lead_author %in% full_smoking_status$lead_author)
+  filter(lead_author %in% current_former_smok$lead_author &
+           lead_author %in% never_smok$lead_author &
+           !lead_author %in% full_smoking_status$lead_author)
 
@@
     group_by(country, .add = T) %>%
-  mutate(., current_smok_percentage = current_smoker/total*100) %>%
-  mutate(., former_smok_percentage = former_smoker/total*100) %>%
-  mutate(., missing_percentage = missing/total*100) %>%
-  mutate(., not_stated_percentage = not_stated/total*100) %>%
-  mutate(never_smoker_percentage = never_smoker/total*100) %>%
+  mutate(current_smok_percentage = current_smoker/total*100,
+         former_smok_percentage = former_smoker/total*100,
+         missing_percentage = missing/total*100,
+         not_stated_percentage = not_stated/total*100,
+         never_smoker_percentage = never_smoker/total*100) %>%
   group_split()
@@
 table_1_word <- table_1 %>%
-  mutate(., current_percentage = current_smoker/total*100,
+  mutate(current_percentage = current_smoker/total*100,
          former_percentage = former_smoker/total*100,
@@
          never_smoker_unknown_percentage = never_smoker_unknown/total*100,
+         current_vaper_p = current_vaper/total*100,
+         current_smoker_current_vaper_p = current_smoker_vaper/total*100,
+         former_smoker_current_vaper_p = former_smoker_vaper/total*100,
+         never_smoker_current_vaper_p = never_smoker_vaper/total*100,
+         current_smoker_former_vaper_p = current_smoker_former_vaper/total*100,
+         former_smoker_former_vaper_p = former_smoker_former_vaper/total*100,
+         never_smoker_former_vaper_p = never_smoker_former_vaper/total*100,
          not_stated_percentage = not_stated/total*100,
@@
   select(lead_author, date_published, country, sample_size, median_age, iqr_lower, iqr_upper, mean_age, lower_range,
-         upper_range, standard_deviation, female_sex_percent, current_percentage, former_percentage, current_former_percentage,
-         never_smoker_percentage, never_smoker_unknown_percentage, not_stated_percentage, missing_percentage, study_id) %>%
+         upper_range, standard_deviation, female_sex_percent, 
+         current_percentage, current_vaper_p, current_smoker_current_vaper_p, current_smoker_former_vaper_p,
+         former_percentage, former_smoker_current_vaper_p, former_smoker_former_vaper_p,
+         current_former_percentage,
+         never_smoker_percentage, never_smoker_current_vaper_p, never_smoker_former_vaper_p,
+         never_smoker_unknown_percentage, not_stated_percentage, missing_percentage, study_id) %>%
   replace_na(., list(not_stated_percentage = 0, missing_percentage = 0)) %>%
@@
   left_join(., table_1_word %>%
-              select(lead_author, missing_percentage)) %>%
+              select(lead_author, missing_percentage, never_smoker_unknown_percentage)) %>%
+  group_by(lead_author) %>%
   mutate(missing_percentage = as.numeric(missing_percentage),
-         low_missingness = ifelse(missing_percentage >= 20, "No", "Yes"))
+         missing = sum(missing_percentage, never_smoker_unknown_percentage, na.rm = T),
+         low_missingness = ifelse(missing >= 20, "No", "Yes")) %>%
+  distinct(., .keep_all = T)
 
@@
   left_join(., smoking_status, by = "lead_author") %>%
-  mutate(study_quality = ifelse(low_missingness == "No" | complete_status == "No", "Poor", "Not_poor"))
+  mutate(study_quality = ifelse(low_missingness == "No" | complete_status == "No", "Poor", "Not_poor")) %>%
+  distinct(., .keep_all = T)
 
@@
   mutate(study_quality_final = ifelse(study_quality == "Not_poor" & (biochemical_verification == "Yes" | random_representative_sample == "Yes"), "good",
-                                      ifelse(study_quality == "Poor", "poor", "fair")))
+                                      ifelse(study_quality == "Poor", "poor", "fair")))  %>%
+  distinct(., .keep_all = T)
 
@@
 
-table_1_word <- left_join(table_1_word, a, by = 'study_id') %>%
-  select(1:4, 22, 5:19, 21, 20)
+table_1_word <- left_join(table_1_word, a, by = 'study_id')
 table_1_word$date_published <- as.Date.character(table_1_word$date_published)
@@
   select(study_id, lead_author, date_published, country, sample_size, study_setting, median_mean, mean_used, iqr,
-         female_sex_percent, current_percentage, former_percentage, current_former_percentage, never_smoker_percentage,
-         never_smoker_unknown_percentage, missing_percentage, study_quality_final) %>%
+         female_sex_percent,
+         current_percentage, current_vaper_p, current_smoker_current_vaper_p, current_smoker_former_vaper_p,
+         former_percentage, former_smoker_current_vaper_p, former_smoker_former_vaper_p,
+         current_former_percentage,
+         never_smoker_percentage, never_smoker_current_vaper_p, never_smoker_former_vaper_p,
+         never_smoker_unknown_percentage, 
+         missing_percentage, study_quality_final) %>%
   mutate(median_mean = paste(median_mean, mean_used, sep = ''),
@@
          'Current smoker %' = current_percentage,
-         'Former smokers %' = former_percentage,
+         "Current vaper %" = current_vaper_p,
+         "Current smoker, current vaper %" = current_smoker_current_vaper_p,
+         "Current smoker, former vaper %" = current_smoker_former_vaper_p,
+         'Former smoker %' = former_percentage,
+         "Former smoker, current vaper %" = former_smoker_current_vaper_p,
+         "Former smoker, former vaper %" = former_smoker_former_vaper_p,
          'Current/former smokers %' = current_former_percentage,
          'Never smokers %' = never_smoker_percentage,
+         "Never smoker, current vaper %" = never_smoker_current_vaper_p,
+         "Never smoker, former vaper %" = never_smoker_former_vaper_p,
          'Never/unknown smokers %' = never_smoker_unknown_percentage,
          'Missing %' = missing_percentage,
-         "Study quality" = study_quality_final)
+         "Study quality" = study_quality_final) %>%
+  distinct(., .keep_all = T)
 
@@
 
-numeric_columns <- c('Median (IQR)', 'Female %', 'Current smoker %','Former smokers %', 'Current/former smokers %',
-                     'Never smokers %', 'Never/unknown smokers %', 'Missing %')
+numeric_columns <- c('Median (IQR)', 'Female %', 'Current smoker %', "Current vaper %", "Current smoker, current vaper %" , "Current smoker, former vaper %",
+                     'Former smoker %', "Former smoker, current vaper %", "Former smoker, former vaper %",
+                     'Current/former smokers %',
+                     'Never smokers %', "Never smoker, current vaper %", "Never smoker, former vaper %",
+                     'Never/unknown smokers %', 'Missing %')
 
@@
         file = here("data_clean", "table_1_refs.rds"))
-source(here("scripts", "references_in_table.r"))
+
+# The references in table script pulls in the doi information from the manuscript.bib file in reports
+source(here("scripts", "references_in_table.R"))
 ```
@@
   formatRound(
-    columns = c(8:14),
-    digits = 2,
+    columns = c(8:21),
+    digits = 1,
     interval = 3,
@@
 save_as_docx(a, path = here('output_data', 'Table_1.docx'))
+
 ```
@@
 table_2_word <-  table_2 %>%
-  mutate(., sample = contributing_sample) %>%
-  mutate(., negative_test_percentage =formatC(negative_test/sample*100, digits = 2, format = "f"),
+  mutate(., sample = contributing_sample,
+         negative_test_percentage =formatC(negative_test/sample*100, digits = 2, format = "f"),
          negative_current_percentage = formatC(negative_current_smoker/negative_test*100, digits = 2, format = "f"),
@@
          negative_never_smoker_percentage = formatC(negative_never_smoker/negative_test*100, digits = 2, format = "f"),
+         negative_current_vaper_p = formatC(negative_current_vaper/negative_test*100, digits = 2, format = "f"),
+         negative_current_smoker_vaper_p = formatC(negative_current_smoker_vaper/negative_test*100, digits = 2, format = "f"),
          negative_not_stated_percentage = formatC(negative_not_stated/negative_test*100, digits = 2, format = "f"),
@@
          positive_never_smoker_percentage = formatC(positive_never_smoker/positive_test*100, digits = 2, format = "f"),
+         positive_never_former_percentage = formatC(positive_never_former/positive_test*100, digits = 2, format = "f"),
+         positive_current_vaper_p = formatC(postitive_current_vaper/positive_test*100, digits = 2, format = "f"),
+         positive_current_smoker_vaper_p = formatC(positive_current_smoker_vaper/positive_test*100, digits = 2, format = "f"),
          positive_not_stated_percentage = formatC(positive_not_stated/positive_test*100, digits = 2, format = "f")) %>%
-  select(-data_on_testing, -missing, -date_published, -sample) %>%
+  select(-data_on_testing, -missing, -date_published, -sample, -c(negative_current_former_vaper, negative_former_vaper, positive_current_former_vaper, positive_former_vaper)) %>%
   write_rds(here::here('data_clean', 'table_2_word.rds'))
@@
          N_never_smoker = na_if(N_never_smoker, "NA ( NA%)"),
+         N_current_vaper = paste(paste(negative_current_vaper,
+                                       (negative_current_vaper_p), sep = " (")
+                                 , "%)", sep = ""),
+         N_current_vaper = na_if(N_current_vaper, "NA ( NA%)"),
+         N_current_smoker_vaper = paste(paste(negative_current_smoker_vaper,
+                                              (negative_current_smoker_vaper_p), sep = " (")
+                                        , "%)", sep = ""),
+         N_current_smoker_vaper = na_if(N_current_smoker_vaper, "NA ( NA%)"),
          N_not_stated = paste(paste(negative_not_stated,
@@
     P_never_smoker = na_if(P_never_smoker, "NA ( NA%)"),
+    P_current_vaper = paste(paste(postitive_current_vaper,
+                                  (positive_current_vaper_p), sep = " (")
+                            , "%)", sep = ""),
+    P_current_vaper = na_if(P_current_vaper, "NA ( NA%)"),
+    P_current_smoker_vaper = paste(paste(positive_current_smoker_vaper,
+                                         (positive_current_smoker_vaper_p), sep = " (")
+                                   , "%)", sep = ""),
+    P_current_smoker_vaper = na_if(P_current_smoker_vaper, "NA ( NA%)"),
     P_not_stated = paste(paste(positive_not_stated,
@@
   select(Author, Population_tested, SARS_CoV_2_negative, N_current_smoker, N_former_smoker, N_current_former_smoker,
-         N_never_smoker, N_not_stated, SARS_CoV_2_positive, P_current_smoker, P_former_smoker, P_current_former_smoker,
-         P_never_smoker, P_not_stated)
+         N_never_smoker, N_current_vaper, N_current_smoker_vaper, N_not_stated,
+         SARS_CoV_2_positive, P_current_smoker, P_former_smoker, P_current_former_smoker,
+         P_never_smoker, P_current_vaper, P_current_smoker_vaper, P_not_stated)
 
@@
 
-numeric_columns <- c('Population_tested', 'SARS_CoV_2_negative', 'N_current_smoker', 'N_former_smoker',
-                     'N_current_former_smoker', 'N_never_smoker', 'N_not_stated', 'SARS_CoV_2_positive', 'P_current_smoker',
-                     'P_former_smoker', 'P_current_former_smoker', 'P_never_smoker', 'P_not_stated')
+numeric_columns <- c('Population_tested',
+                     "SARS_CoV_2_negative", "N_current_smoker", "N_former_smoker", "N_current_former_smoker",
+                     "N_never_smoker", "N_current_vaper", "N_current_smoker_vaper", "N_not_stated",
+                     "SARS_CoV_2_positive", "P_current_smoker", "P_former_smoker", "P_current_former_smoker",
+                     "P_never_smoker", "P_current_vaper", "P_current_smoker_vaper", "P_not_stated")
 a <- flextable(a) %>%
@@
                        N_never_smoker = "Never smoker (%)",
+                       N_current_vaper = "Current vaper (%)",
+                       N_current_smoker_vaper = "Current vaper and smoker (%)",
                        N_not_stated = "Not stated (%)",
@@
                        P_never_smoker = "Never smoker (%)",
+                       P_current_vaper = "Current vaper (%)",
+                       P_current_smoker_vaper = "Current vaper and smoker (%)",
                        P_not_stated = "Not stated (%)") %>%
-  add_header_row(top = TRUE, values = c("","SARS-CoV-2 negative", "SARS-CoV-2 positive" ), colwidths = c(2, 6, 6)) %>%
+  add_header_row(top = TRUE, values = c("","SARS-CoV-2 negative", "SARS-CoV-2 positive" ), colwidths = c(2, 8, 8)) %>%
   theme_booktabs() %>%
@@
                'negative_former_smoker' = table_2$negative_former_smoker,
-               'positive_former_smoker' = table_2$positive_former_smoker) %>%
+               'positive_former_smoker' = table_2$positive_former_smoker,
+               'negative_current_vaper' = table_2$negative_current_vaper,
+               'positive_current_vaper' = table_2$postitive_current_vaper,
+               'negative_current_smoker_vaper' = table_2$negative_current_smoker_vaper,
+               'positive_current_smoker_vaper' = table_2$positive_current_smoker_vaper) %>%
         filter(author %in% included_studies$lead_author)
@@
 
-ecdf <- read_rds(here("data_clean", "bayesian_models", "ecdf_list.rds"))
+if(file.exists(here("data_clean", "bayesian_models", "ecdf_list.RDS"))) {
+ecdf <- read_rds(here("data_clean", "bayesian_models", "ecdf_list.RDS"))
 m1_a_ecdf <- ecdf[[1]]
@@
 m8_a_ecdf <- ecdf[[8]]
-
+}
 ```
@@
 
-The ECDF shows the probability of the risk being less than or equal to RR = 0.9, `r paste(round(m1_a_ecdf(0.9)*100, 2), "%", sep = "")`
-
 #### Former versus never smokers
@@
 
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m2_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
 ## Table 3: Smoking and COVID-19 hospitalisation
@@
 table_3_word <- table_3 %>%
-  mutate(., sample = sample_with_outcome) %>%
-  mutate(., community_percentage = formatC(number_community/sample*100, digits = 2, format = "f")) %>%
-  mutate(., community_current_smoker_percent = formatC(community_current_smoker/number_community*100, digits = 2, format = "f")) %>%
-  mutate(., community_former_smoker_percent = formatC(community_former_smoker/number_community*100, digits = 2, format = "f")) %>%
-  mutate(., community_current_former_smoker_percent = formatC(community_current_former_smoker/number_community*100, digits = 2, format = "f")) %>%
-  mutate(., community_never_smoker_percent = formatC(community_never_smoker/number_community*100, digits = 2, format = "f")) %>%
-  mutate(., community_never_unknown_smoker_percent = formatC(community_never_unknown_smoker/number_community*100, digits = 2, format = "f")) %>%
-  mutate(., community_not_stated_percent = formatC(community_not_stated/number_community*100, digits = 2, format = "f")) %>%
-  mutate(., number_hospitalised_percent = formatC(number_hospitalised/sample*100, digits = 2, format = "f")) %>%
-  mutate(., hospitalised_current_smoker_percent = formatC(hospitalised_current_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
-  mutate(., hospitalised_former_smoker_percent = formatC(hospitalised_former_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
-  mutate(., hospitalised_current_former_smoker_percent = formatC(hospitalised_current_former_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
-  mutate(., hospitalised_never_smoker_percent = formatC(hospitalised_never_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
-  mutate(., hospitalised_never_unknown_smoker_percent = formatC(hospitalised_never_unknown_smoker/number_hospitalised*100, digits = 2, format = "f")) %>%
-  mutate(., hospitalised_not_stated_percent = formatC(hospitalised_not_stated/number_hospitalised*100, digits = 2, format = "f")) %>%
+  mutate(sample = sample_with_outcome,
+         community_percentage = formatC(number_community/sample*100, digits = 2, format = "f"),
+         community_current_smoker_percent = formatC(community_current_smoker/number_community*100, digits = 2, format = "f"),
+         community_former_smoker_percent = formatC(community_former_smoker/number_community*100, digits = 2, format = "f"),
+         community_current_former_smoker_percent = formatC(community_current_former_smoker/number_community*100, digits = 2, format = "f"),
+         community_never_smoker_percent = formatC(community_never_smoker/number_community*100, digits = 2, format = "f"),
+         community_never_unknown_smoker_percent = formatC(community_never_unknown_smoker/number_community*100, digits = 2, format = "f"),
+         community_current_vaper_p = formatC(community_current_vaper/number_community*100, digits = 2, format = "f"),
+         community_former_vaper_p = formatC(community_former_vaper/number_community*100, digits = 2, format = "f"),
+         community_current_former_vaper_p = formatC(community_current_former_vaper/number_community*100, digits = 2, format = "f"),
+         community_not_stated_percent = formatC(community_not_stated/number_community*100, digits = 2, format = "f"),
+         number_hospitalised_percent = formatC(number_hospitalised/sample*100, digits = 2, format = "f"),
+         hospitalised_current_smoker_percent = formatC(hospitalised_current_smoker/number_hospitalised*100, digits = 2, format = "f"),
+         hospitalised_former_smoker_percent = formatC(hospitalised_former_smoker/number_hospitalised*100, digits = 2, format = "f"),
+         hospitalised_current_former_smoker_percent = formatC(hospitalised_current_former_smoker/number_hospitalised*100, digits = 2, format = "f"),
+         hospitalised_never_smoker_percent = formatC(hospitalised_never_smoker/number_hospitalised*100, digits = 2, format = "f"),
+         hospitalised_never_unknown_smoker_percent = formatC(hospitalised_never_unknown_smoker/number_hospitalised*100, digits = 2, format = "f"),
+         hospitalised_current_vaper_p = formatC(hospitalised_never_unknown_smoker/number_hospitalised*100, digits = 2, format = "f"),
+         hospitalised_former_vaper_p = formatC(hospitalised_former_vaper/number_hospitalised*100, digits = 2, format = "f"),
+         hospitalised_current_former_vaper_p = formatC(hospitalised_current_former_vaper/number_hospitalised*100, digits = 2, format = "f"),
+         hospitalised_not_stated_percent = formatC(hospitalised_not_stated/number_hospitalised*100, digits = 2, format = "f")) %>%
   select(lead_author, sample_with_outcome, number_community, community_percentage,
@@
          community_current_former_smoker_percent, community_never_smoker, community_never_smoker_percent,
+         community_current_vaper, community_current_vaper_p, community_former_vaper, community_former_vaper_p,
+         community_current_former_vaper, community_current_former_vaper_p,
          community_never_unknown_smoker, community_never_unknown_smoker_percent, 
@@
          hospitalised_former_smoker_percent, hospitalised_current_former_smoker, hospitalised_current_former_smoker_percent,
-         hospitalised_never_smoker, hospitalised_never_smoker_percent, hospitalised_never_unknown_smoker,
-         hospitalised_never_unknown_smoker_percent, hospitalised_not_stated, hospitalised_not_stated_percent) %>%
+         hospitalised_never_smoker, hospitalised_never_smoker_percent, hospitalised_current_vaper, hospitalised_current_vaper_p, hospitalised_former_vaper, hospitalised_former_vaper_p,
+         hospitalised_current_former_vaper, hospitalised_current_former_vaper_p,
+         hospitalised_never_unknown_smoker, hospitalised_never_unknown_smoker_percent, hospitalised_not_stated, hospitalised_not_stated_percent) %>%
   write_rds(here::here('data_clean', 'table_3_word.rds'))
@@
 
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m3_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
 #### Former versus never smoking for hospitalisation
@@
 
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m4_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
 ## Table 4: Smoking and COVID-19 disease severity
@@
 table_4_word <- table_4 %>%
-  mutate(., sample = sample_with_severity) %>%
-  mutate(., non_severe_disease_percentage = formatC(non_severe_disease/sample*100, digits = 2, format = "f")) %>%
-  mutate(., non_severe_current_smoker_percent = formatC(non_severe_current_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
-  mutate(., non_severe_former_smoker_percent = formatC(non_severe_former_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
-  mutate(., non_severe_current_former_smoker_percent = formatC(non_severe_current_former_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
-  mutate(., non_severe_never_smoker_percent = formatC(non_severe_never_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
-  mutate(., non_severe_never_unknown_smoker_percent = formatC(non_severe_never_unknown_smoker/non_severe_disease*100, digits = 2, format = "f")) %>%
-  mutate(., non_severe_not_stated_percent = formatC(non_severe_not_stated/non_severe_disease*100, digits = 2, format = "f")) %>%
-  mutate(., severe_disease_number_percent = formatC(severe_disease_number/sample*100, digits = 2, format = "f")) %>%
-  mutate(., severe_disease_current_smoker_percent = formatC(severe_disease_current_smoker/severe_disease_number*100, digits = 2, format = "f")) %>%
-  mutate(., severe_disease_former_smoker_percent = formatC(severe_disease_former_smoker/severe_disease_number*100, digits = 2, format = "f")) %>%
-  mutate(., severe_disease_current_former_smoker_percent = formatC(severe_disease_current_former_smoker/severe_disease_number*100, digits = 2, format = "f")) %>%
-  mutate(., severe_disease_never_smoker_percent = formatC(severe_disease_never_smoker/severe_disease_number*100, digits = 2, format = "f")) %>%
-  mutate(., severe_disease_never_unknown_percent = formatC(severe_disease_never_unknown/severe_disease_number*100, digits = 2, format = "f")) %>%
-  mutate(., severe_disease_not_stated_percent = formatC(severe_disease_not_stated/severe_disease_number*100, digits = 2, format = "f")) %>%
+  mutate(sample = sample_with_severity,
+         non_severe_disease_percentage = formatC(non_severe_disease/sample*100, digits = 2, format = "f"),
+         non_severe_current_smoker_percent = formatC(non_severe_current_smoker/non_severe_disease*100, digits = 2, format = "f"),
+         non_severe_former_smoker_percent = formatC(non_severe_former_smoker/non_severe_disease*100, digits = 2, format = "f"),
+         non_severe_current_former_smoker_percent = formatC(non_severe_current_former_smoker/non_severe_disease*100, digits = 2, format = "f"),
+         non_severe_never_smoker_percent = formatC(non_severe_never_smoker/non_severe_disease*100, digits = 2, format = "f"),
+         non_severe_never_unknown_smoker_percent = formatC(non_severe_never_unknown_smoker/non_severe_disease*100, digits = 2, format = "f"),
+         non_severe_not_stated_percent = formatC(non_severe_not_stated/non_severe_disease*100, digits = 2, format = "f"),
+         severe_disease_number_percent = formatC(severe_disease_number/sample*100, digits = 2, format = "f"),
+         severe_disease_current_smoker_percent = formatC(severe_disease_current_smoker/severe_disease_number*100, digits = 2, format = "f"),
+         severe_disease_former_smoker_percent = formatC(severe_disease_former_smoker/severe_disease_number*100, digits = 2, format = "f"),
+         severe_disease_current_former_smoker_percent = formatC(severe_disease_current_former_smoker/severe_disease_number*100, digits = 2, format = "f"),
+         severe_disease_never_smoker_percent = formatC(severe_disease_never_smoker/severe_disease_number*100, digits = 2, format = "f"),
+         severe_disease_never_unknown_percent = formatC(severe_disease_never_unknown/severe_disease_number*100, digits = 2, format = "f"),
+         severe_disease_not_stated_percent = formatC(severe_disease_not_stated/severe_disease_number*100, digits = 2, format = "f")) %>%
   select(lead_author, sample, non_severe_disease, non_severe_disease_percentage, non_severe_current_smoker,
@@
 
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m5_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
 #### Former versus never smoking for risk of severe disease
@@
 
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m6_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
 ## Table 5: Smoking and COVID-19 mortality
@@
 table_5_word <- table_5 %>%
-  mutate(., sample = sample_with_deaths) %>%
-  mutate(., deaths_percentage = formatC(deaths/sample*100, digits = 2, format = "f")) %>%
-  mutate(., death_current_smokers_percent = formatC(death_current_smokers/deaths*100, digits = 2, format = "f")) %>%
-  mutate(., death_former_smokers_percent = formatC(death_former_smokers/deaths*100, digits = 2, format = "f")) %>%
-  mutate(., death_current_former_smokers_percent = formatC(death_current_former_smokers/deaths*100, digits = 2, format = "f")) %>%
-  mutate(., death_never_smokers_percent = formatC(death_never_smokers/deaths*100, digits = 2, format = "f")) %>%
-  mutate(., death_never_unknown_smokers_percent = formatC(death_never_unknown_smokers/deaths*100, digits = 2, format = "f")) %>%
-  mutate(., death_not_stated_percent = formatC(death_not_stated/deaths*100, digits = 2, format = "f")) %>%
-  mutate(., recovered_percentage = formatC(recovered/sample*100, digits = 2, format = "f")) %>%
-  mutate(., recovered_current_smokers_percent = formatC(recovered_current_smoking/recovered*100, digits = 2, format = "f")) %>%
-  mutate(., recovered_former_smokers_percent = formatC(recovered_former_smoker/recovered*100, digits = 2, format = "f")) %>%
-  mutate(., recovered_current_former_smokers_percent = formatC(recovered_current_former_smokers/recovered*100, digits = 2, format = "f")) %>%
-  mutate(., recovered_never_smokers_percent = formatC(recovered_never_smoker/recovered*100, digits = 2, format = "f")) %>%
-  mutate(., recovered_never_unknown_smokers_percent = formatC(recovered_never_unknown_smoker/recovered*100, digits = 2, format = "f")) %>%
-  mutate(., recovered_not_stated_percent = formatC(recovered_not_stated/recovered*100, digits = 2, format = "f")) %>%
+  mutate(sample = sample_with_deaths,
+         deaths_percentage = formatC(deaths/sample*100, digits = 2, format = "f"),
+         death_current_smokers_percent = formatC(death_current_smokers/deaths*100, digits = 2, format = "f"),
+         death_former_smokers_percent = formatC(death_former_smokers/deaths*100, digits = 2, format = "f"),
+         death_current_former_smokers_percent = formatC(death_current_former_smokers/deaths*100, digits = 2, format = "f"),
+         death_never_smokers_percent = formatC(death_never_smokers/deaths*100, digits = 2, format = "f"),
+         death_never_unknown_smokers_percent = formatC(death_never_unknown_smokers/deaths*100, digits = 2, format = "f"),
+         death_not_stated_percent = formatC(death_not_stated/deaths*100, digits = 2, format = "f"),
+         recovered_percentage = formatC(recovered/sample*100, digits = 2, format = "f"),
+         recovered_current_smokers_percent = formatC(recovered_current_smoking/recovered*100, digits = 2, format = "f"),
+         recovered_former_smokers_percent = formatC(recovered_former_smoker/recovered*100, digits = 2, format = "f"),
+         recovered_current_former_smokers_percent = formatC(recovered_current_former_smokers/recovered*100, digits = 2, format = "f"),
+         recovered_never_smokers_percent = formatC(recovered_never_smoker/recovered*100, digits = 2, format = "f"),
+         recovered_never_unknown_smokers_percent = formatC(recovered_never_unknown_smoker/recovered*100, digits = 2, format = "f"),
+         recovered_not_stated_percent = formatC(recovered_not_stated/recovered*100, digits = 2, format = "f")) %>%
   select(lead_author, sample, recovered, recovered_percentage, recovered_current_smoking, recovered_current_smokers_percent,
@@
 
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m7_a_ecdf(1.1))*100, 2), "%", sep = "")`
-
 #### Former versus never smoking for disease mortality
@@
 ![ECDF for former smoking against never smoking for mortality](figure/ECDF_m8_a.png)
-
-The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m8_a_ecdf(1.1))*100, 2), "%", sep = "")`
 

### 2021-07-27 reports/manuscript.Rmd
v12

@@
 Categorisation of smoking status was heterogeneous (see [Table 1](#table-1)).
-`r paste(data_source[1,2])` `r paste(round(data_source[1,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies collected data on smoking status through routine electronic health records (EHRs), `r paste(data_source[2,2])` `r paste(round(data_source[2,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies used a bespoke case report form for COVID-19 and `r paste(data_source[3,2])` `r paste(round(data_source[3,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies did not state the source for information on smoking status.
+`r paste(data_source[1,2])` `r paste("(", round(data_source[1,2]/sum(data_source[,2])*100,1), "%",")", sep = "")` studies collected data on smoking status through routine electronic health records (EHRs), `r paste(data_source[2,2])` `r paste("(", round(data_source[2,2]/sum(data_source[,2])*100,1), "%", ")", sep = "")` studies used a bespoke case report form for COVID-19 and `r paste(data_source[3,2])` `r paste( "(", round(data_source[3,2]/sum(data_source[,2])*100,1), "%", ")", sep = "")` studies did not state the source for information on smoking status.
 None of the studies verified smoking status biochemically.
@@
 
-Fifteen studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use (see [Table 1](#table-1).
+Fifteen studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use (see [Table 1](#table-1)).
 
@@
 
-Unadjusted smoking prevalence compared with overall estimates for national adult smoking prevalence split by country and study setting is presented in [Figure 2a](#fig-2a) and [2b](#fig-2b).
+Unadjusted smoking prevalence compared with overall estimates for national adult smoking prevalence split by country and study setting is presented in [Figure 2](#fig-2).
 Lower than expected current smoking prevalence was generally observed, especially in studies with hospitalised samples.
@@
 For comparison, national current and former smoking prevalence are indicated by the red lines.
-Countries with three or more eligible studies are shown.](figure/combined_smoking_plots_updated.png){#fig-2a}
+Countries with three or more eligible studies are shown.](figure/combined_smoking_plots_updated.png){#fig-2}
 
@@
 `r paste(nrow(table_3_word))` studies examined hospitalisation for COVID-19 disease, stratified by smoking status (see [Table 3](#table-3)).
-Meta-analyses were performed for one 'good' and 18 ‘fair’ quality studies (see [Figure 5](#fig-5) and [6](#fig-6)).
+Meta-analyses were performed for 19 ‘fair’ quality studies (see [Figure 5](#fig-5) and [6](#fig-6)).
 Current (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_tau, 2))`) and former (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_tau, 2))`) compared with never smokers were at increased risk of hospitalisation with COVID-19.
@@
 
-####    [*Table 3* COVID-19 hospitalsation by smoking status]{#table-3}
+####    [*Table 3* COVID-19 hospitalisation by smoking status]{#table-3}
 

### 2021-07-27 scripts/bayesian_analysis.R

@@ -590,7 +590,7 @@ pooled_draw_m8_a <- pooled_effect_draw(m8_a)
 study_draws_m8_a <- study_draws_m8_a %>%
   mutate(review_version = as.character(review_version),
          review_version = ifelse(is.na(review_version), "current", review_version),
-         review_version = factor(review_version, levels = c("previous", "current"))
+         review_version = factor(review_version, levels = c("previous", "current")))
 # study_draws_m8_a_hh <- study_draws_m8_a_hh %>%
 #   mutate(review_version = as.character(review_version),
 #          review_version = ifelse(is.na(review_version), "current", review_version),
@@ -606,7 +606,7 @@ study_draws_m8_a <- study_draws_m8_a %>%
 #   "m8.png"
 # )
 
-former_mortality_alternative_prior <- forest_plot(
+(former_mortality_alternative_prior <- forest_plot(
   m8_a,
   study_draws_m8_a,
   pooled_draw_m8_a,
@@ -614,7 +614,7 @@ former_mortality_alternative_prior <- forest_plot(
   "Forest plot of former smokers and mortality",
   paste(tail(previous_review_versions, 1), "mortality prior"),
   "m8_a.png"
-)
+))
 
 png(file = (here("reports", "figure","ECDF_m8_a.png")))
 plot(m8_a_ecdf, main = "Empirical cumulative distribution, former smokers, mortality",

### 2021-08-12 reports/manuscript.Rmd
v12 final

@@
 
-**Aims:** To estimate the association of smoking status with rates of i) infection, ii) hospitalisation, iii) disease severity, and iv) mortality from SARS-CoV-2/COVID-19 disease.
+**Aims:** To estimate the association of smoking status with rates of i) infection, ii) hospitalisation, iii) disease severity in hospitalised patients, and iv) mortality from SARS-CoV-2/COVID-19 disease.
 
@@
 Current compared with never smokers were at reduced risk of SARS-CoV-2 infection (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_tau, 2))`).
-Data for former smokers were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_tau, 2))`) but favoured there being no important association (`r paste(round(1-m2_a_ecdf(1.1),2)*100, "%", sep = "")` probability of RR ≥1.1).
+Data for former smokers were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_tau, 2))`) but favoured there being no important association (<1% probability of RR ≥1.1).
 Former compared with never smokers were at increased risk of hospitalisation (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_tau, 2))`), greater disease severity (RR = `r paste(round(former_severity_RR_value, 2))`, CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_tau, 2))`) and mortality (RR = `r paste(round(former_mortality_RR_value, 2))`, CrI = `r paste(round(former_mortality_lci, 2), round(former_mortality_uci, 2), sep = "-")`, τ = `r paste(round(former_mortality_tau, 2))`).
@@
 
-**Conclusions:** Compared with never smokers, current smokers appear to be at reduced risk of SARS-CoV-2 infection and increased risk of greater disease severity, while former smokers appear to be at increased risk of hospitalisation, greater disease severity and mortality from COVID-19.
+**Conclusions:** Compared with never smokers, current smokers appear to be at reduced risk of SARS-CoV-2 infection and increased risk of greater in-hospital disease severity, while former smokers appear to be at increased risk of hospitalisation, greater in-hospital disease severity and mortality from COVID-19.
 However, it is uncertain whether these associations are causal.
+This version (v12) will be the last regular update; however, yearly updates may continue as new evidence becomes available.
 
@@
 We aimed to produce a rapid synthesis of available evidence pertaining to the rates of infection, hospitalisation, disease severity and mortality from SARS-CoV-2/COVID-19 stratified by smoking status.
-Given the increasing availability of data on this topic, this is a living review with regular updates.
-As evidence accumulates, the review will be expanded to include studies reporting COVID-19 outcomes by alternative nicotine use (e.g., nicotine replacement therapy or e-cigarettes).
+Given the increasing availability of data on this topic, this was set up in March 2020 as a living review with regular updates.
 
@@
 A completed Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) checklist is included in Supplementary file 1.
+This version (v12) will be the last regular update; however, yearly updates may continue as new, high-quality studies (e.g.
+large, representative/random population studies or experimental studies) become available.
 
@@
   3.
-Recorded as outcome i) results of a SARS-CoV-2 diagnostic test (including antibody assays), ii) clinical diagnosis of COVID-19, iii) hospitalisation with COVID-19, iv) severity of COVID-19 disease in those hospitalised or v) mortality from COVID-19;
-Reported any of the outcomes of interest by self-reported or biochemically verified smoking status (e.g.
+3.	Recorded as outcome i) results of a SARS-CoV-2 diagnostic test (including antibody assays) in community or hospitalised populations, ii) clinical diagnosis of COVID-19 in community or hospitalised populations, iii) hospitalisation with COVID-19 in community populations testing positive for SARS-CoV-2, iv) severity of COVID-19 disease in hospitalised populations or v) mortality from COVID-19 in community or hospitalised populations; Reported any of the outcomes of interest by self-reported or biochemically verified smoking status (e.g.
 current smoker, former smoker, never smoker) or current vaping and nicotine replacement therapy (NRT) use;
@@
 We first defined prior distributions for the true pooled effect size (µ) and the between-study heterogeneity (τ), with µ specified as a normal distribution with a mean equal to the derived point estimate from each comparison of interest in the immediately preceding version of this living review, and τ specified as a half-Cauchy distribution with a mean of 0 and standard deviation of 1.
-The half-Cauchy distribution was selected to reflect prior knowledge that high levels of between-study heterogeneity are more likely than lower levels.
+The half-Cauchy distribution was selected to reflect prior knowledge that higher levels of between-study heterogeneity are more likely than lower levels.
 Markov Chain Monte Carlo methods (20,000 burn-ins followed by 80,000 iterations) were then used to generate a risk distribution for each study, in addition to a pooled effect for the posterior risk distribution.
@@
 Current smokers were at reduced risk of testing positive for SARS-CoV-2 compared with never smokers (RR = `r paste(round(current_testing_RR_value, 2))`, 95% Credible Interval (CrI) = `r paste(round(current_testing_lci, 2), round(current_testing_uci, 2), sep = "-")`, τ = `r paste(round(current_testing_tau, 2))`).
-The six good quality studies each reported point estimates less than 1, although the CrI for two of the six studies crossed 1.
+The six good quality studies each reported point estimates less than 1, although the CrI for two of the six studies crossed 1 for current smokers (see starred studies in Figure 3).
 The probability of current smokers being at reduced risk of infection compared with never smokers (RR ≤0.9) was >99%.
-Former compared with never smokers were at increased risk of testing positive, but data were inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_tau, 2))`) and favoured there being no important association.
+Former compared with never smokers were had an equivalent risk of testing positive, with data being inconclusive (RR = `r paste(round(former_testing_RR_value, 2))`, 95% CrI = `r paste(round(former_testing_lci, 2), round(former_testing_uci, 2), sep = "-")`, τ = `r paste(round(former_testing_tau, 2))`) ), favouring there being no important association.
 The probability of former smokers being at increased risk of infection (RR ≥1.1) compared with never smokers was < 1%.
@@
 Current (RR = `r paste(round(current_severity_RR_value, 2))`, 95% CrI = `r paste(round(current_severity_lci, 2), round(current_severity_uci, 2), sep = "-")`, τ = `r paste(round(current_severity_tau, 2))`) and former (RR = `r paste(round(former_severity_RR_value, 2))`, 95% CrI = `r paste(round(former_severity_lci, 2), round(former_severity_uci, 2), sep = "-")`, τ = `r paste(round(former_severity_tau, 2))`) compared with never smokers were at increased risk of greater disease severity.
-The probability of current and former smokers having increased risk of greater disease severity (RR ≥1.1) compared with never smokers was `r paste(round(1-m5_a_ecdf(1.1),2)*100, "%", sep = "")` and `r paste(round(1-m6_a_ecdf(1.1),2)*100, "%", sep = "")`, respectively.
+The probability of current and former smokers having increased risk of greater disease severity (RR ≥1.1) compared with never smokers was `r paste(round(1-m5_a_ecdf(1.1),2)*100, "%", sep = "")` and >99%, respectively.
 Results were materially unchanged in two sensitivity analyses.
@@
 Notwithstanding this uncertainty, compared with overall adult national prevalence estimates, recorded current smoking rates in most studies were lower than expected.
-In a subset of good and fair quality studies (n = `r paste(nrow(good_quality_2)+nrow(fair_quality_2))`), current but not former smokers had a reduced risk of testing positive for SARS-CoV-2 but current smokers appeared somewhat more likely to present for testing and/or receive a test.
-Current compared with never smokers had an increased risk of greater disease severity.
-Data for current smokers on the risk of hospitalisation and mortality were inconclusive, and favoured there being no important associations.
+In a subset of good and fair quality studies (n = `r paste(nrow(good_quality_2)+nrow(fair_quality_2))`), current but not former smokers had a reduced risk of testing positive for SARS-CoV-2 but current (and former) smokers appeared somewhat more likely to present for testing and/or receive a test.
+Current compared with never smokers had an increased risk of greater disease severity in hospitalised patients.
+Data for current smokers on the risk of hospitalisation and mortality were inconclusive and favoured there being no important associations.
 Former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.
@@
 Testing is hence likely limited in some subgroups, with the potential for these groups to include an overrepresentation of current smokers.
+The proportion of symptomatic cases may also vary depending on viral strain and vaccination status.
     2.	On the other hand, current and former smokers may be more likely to meet local criteria for community testing due to increased prevalence of symptoms consistent with SARS-CoV-2 infection, such as cough, increased sputum production or altered sense of smell or taste [@hopkinson_current_2020].
@@
 However, the odds of infection for current smokers (OR = 0.64, 95% CI = 0.58-0.71) was in concordance with the pooled estimate in our meta-analysis.
-Population surveys – particularly with linked data on confirmed infection or antibodies – will be included in future review versions to help mitigate some of the limitations of healthcare based observational studies.
+Population surveys – particularly with linked data on confirmed infection or antibodies – should be included in future reviews to help mitigate some of the limitations of healthcare based observational studies.
 The comparisons of current and former smoking prevalence in the included studies with national prevalence estimates did not adjust observed prevalence for the demographic profile of those tested/admitted to hospital.
@@
 ingested) since this can affect pharmacokinetics [@shahab_novel_2013] and potential therapeutic effects.
+Of note, a recent study of rhesus macaques reported that the nAChR agonist varenicline has the potential to interact with and inhibit SARS-CoV-2 infection and replication.
 A second research priority would be a large, representative (randomly sampled) population survey with a validated assessment of smoking status which distinguishes between recent and long-term ex-smokers – ideally biochemically verified – and assesses seroprevalence and links to health records.
@@
 Across `r paste(current_total)` studies, recorded current but not past smoking prevalence was generally lower than national prevalence estimates.
-Current smokers were at reduced risk of testing positive for SARS-CoV-2 and at increased risk of greater disease severity.
-Former smokers were at increased risk of hospitalisation, disease severity and mortality compared with never smokers.
+Current smokers were at reduced risk of testing positive for SARS-CoV-2 and at increased risk of greater in-hospital disease severity.
+Former smokers were at increased risk of hospitalisation, in-hospital disease severity and mortality compared with never smokers.
 

### 2021-08-13 reports/manuscript.Rmd
v12

@@
 ##    Data availability
-All data contributing to the current and future review versions are available [here](https://doi.org/10.6084/m9.figshare.12756020)
+All data contributing to the current and future review versions are available [here](https://doi.org/10.6084/m9.figshare.12756020.v7)
 
-All code required to reproduce the current and future analyses are available [here](https://doi.org/10.5281/zenodo.4002046)
+All code required to reproduce the current and future analyses are available [here](https://doi.org/10.5281/zenodo.5189278)
 
